---
title: "Bayesian Model - All sites"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
knitr::opts_knit$set(root.dir = "~/Project Files/Analysis/Cleaning and Compilation/Cleaning")
```

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(lme4) # to use glmer
library(data.table)
library(R2jags)

```

# Clean out environment

```{r}
rm(list = ls())
```

# Load data

At time of creation, I want to load in original data, combine visits, then add point data, just to make this as clean as possible(ie. not having duplicate columns with multiple distances and targets and stuff)

```{r, warning=FALSE, echo=FALSE}

data <- read_xlsx("data_with_bearings.xlsx")

```

# Infestation Sum

```{r}

# Visit 1
# data <- data %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")), 
#             funs(if_else(startsWith(., "1-"), "1", .))) %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")), as.numeric)
# Change all "1-" ratings to just a singular 1 value and make numeric

data <- data %>%
  mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")),~coalesce(., 0)) # change all NA's to 0

data$infestation_sum_v1_focal <- data$NL_v1 + data$NM_v1 + data$NH_v1 + data$SL_v1 + data$SM_v1 + data$SH_v1

# Visit 2
# data <- data %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")), 
#             funs(if_else(startsWith(., "1-"), "1", .))) %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")), as.numeric)
# # Change all "1-" ratings to just a singular 1 valueand make numeric
# 
data <- data %>%
  mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")),~coalesce(., 0)) # change all NA's to 0

data$infestation_sum_v2_focal <- data$NL_v2 + data$NM_v2 + data$NH_v2 + data$SL_v2 + data$SM_v2 + data$SH_v2

data <- data %>%
  mutate_at(vars(c("NL_v3", "NM_v3", "NH_v3", "SL_v3", "SM_v3", "SH_v3")),~coalesce(., 0)) # change all NA's to 0

data$infestation_sum_v3_focal <- data$NL_v3 + data$NM_v3 + data$NH_v3 + data$SL_v3 + data$SM_v3 + data$SH_v3

data <- data %>%
  mutate_at(vars(c("NL_v4", "NM_v4", "NH_v4", "SL_v4", "SM_v4", "SH_v4")),~coalesce(., 0)) # change all NA's to 0

data$infestation_sum_v4_focal <- data$NL_v4 + data$NM_v4 + data$NH_v4 + data$SL_v4 + data$SM_v4 + data$SH_v4

# Subset target trees to get infesation sums
targets <- as.character(data$TARGET)
infested_subset_inf_sum <- data %>% filter(unique_ID %in% targets) %>% dplyr::select(c("unique_ID", "infestation_sum_v1_focal", "infestation_sum_v2_focal", "infestation_sum_v3_focal", "infestation_sum_v4_focal"))
names(infested_subset_inf_sum)[1] <- "TARGET"
names(infested_subset_inf_sum)[2] <- "inf_sum_v1_target"
names(infested_subset_inf_sum)[3] <- "inf_sum_v2_target"
names(infested_subset_inf_sum)[4] <- "inf_sum_v3_target"
names(infested_subset_inf_sum)[5] <- "inf_sum_v4_target"


# append sums to our main dataframe
# Append target coordinates to data frame
data <- inner_join(data, infested_subset_inf_sum, by = "TARGET")
str(data)

data$species <- as.factor(data$species)

# Make things!

# BA in cubed meters or something
data$dbh <- data$dbh * 2.54 # convert DBH to cm
data <- mutate(data, BA = (0.00007854 * (dbh ^ 2)))

# Crown Height in meters
data$crown_height <- (data$height - data$canopy_ht) * 0.3048

# Convert to sf
data_sf <- st_as_sf(data, coords = c("Longitude.focal", "Latitude.focal"), crs = 4326)

# Define the target UTM zone
target_crs <- 32612  # EPSG code for UTM zone 12

# Transform the coordinates
data_utm <- st_transform(data_sf, crs = target_crs)
all_coords <- st_coordinates(data_utm)
all_coords <- as.data.frame(all_coords)

# Put back into original datframe

data_utm <- data
data_utm$Longitude.focal <- all_coords$X
data_utm$Latitude.focal <- all_coords$Y

# data <- data[!duplicated(data), ] #- there shouldn't be duplicates now
rownames(data_utm) <- data_utm$unique_ID

# Function to get distances for every tree 10m and under
TEMP_DIST_10 = vegdist(data_utm[,c("Longitude.focal","Latitude.focal")], method = "euclid")
TEMP_DIST_10.df = as.data.frame(as.matrix(TEMP_DIST_10))
rownames(TEMP_DIST_10.df) = (data$unique_ID)
colnames(TEMP_DIST_10.df) = (data$unique_ID)
TEMP_DIST_10.df$`Tag #` = rownames(TEMP_DIST_10.df)
TEMP_DIST_MELT_10 = reshape2::melt(TEMP_DIST_10.df, id.vars = c("Tag #"))
colnames(TEMP_DIST_MELT_10) = c("FROM", "TARGET", "DISTANCE")

Holder_10 <- TEMP_DIST_MELT_10

Holder_new_10 <- merge(Holder_10, data, by.x = "TARGET", by.y = "unique_ID")
data.dt <- as.data.table(Holder_new_10)

# Get all trees under 10m from focal tree
under10_holder <- data.dt[,.(TARGET, DISTANCE.x), by=.(FROM)][,.SD[which(DISTANCE.x <= 10 & DISTANCE.x != 0)], by=.(FROM)]

filtered_distances <- under10_holder %>%
  # Extract the site info (assuming the site info is before the "_")
  mutate(site1 = sub("_.*", "", FROM), 
         site2 = sub("_.*", "", TARGET)) %>%
  # Filter rows where site1 matches site2
  filter(site1 == site2) %>%
  # Optionally, remove the helper columns site1 and site2
  select(FROM, TARGET, DISTANCE.x)

# View the filtered data
df <- filtered_distances

### Different approach to accelerate processing time

# Create a lookup table from your `data`
lookup <- data %>% select(unique_ID, infested_v1, infestation_sum_v1_focal, dbh, brooms, Gouting_v1, species, BA, crown_height, `Dead / Alive_v1`)

# Join based on TARGET to bring in these variables
df <- df %>%
  left_join(lookup, by = c("TARGET" = "unique_ID")) %>%
  rename(infestation_under10 = infested_v1,
         infestation.sum.1_under10 = infestation_sum_v1_focal,
         dbh_under10 = dbh,
         brooms_under10 = brooms,
         gouting_under10 = Gouting_v1,
         species_under10 = species,
         BA_under10 = BA,
         crown_ht__under10 = crown_height,
         dead_alive = `Dead / Alive_v1`)

# Host and non-host BA
df <- df %>%
  mutate(host_non = if_else(species_under10 == "ABLA", 1, 0))

# Now calculate local stats
local.stats <- df %>%
  group_by(FROM) %>%
  summarise(
    n = n(),
    number.infested_under10 = sum(infestation_under10, na.rm = TRUE),  # number of infested
    percent.infested_under10 = (sum(infestation_under10, na.rm = TRUE) / n) * 100,  # percent infested
    infestation.sum_under10 = sum(infestation.sum.1_under10, na.rm = TRUE),  # infestation load
    average.dbh_under10 = mean(dbh_under10, na.rm = TRUE),  # average dbh
    brooms_under10 = sum(brooms_under10, na.rm = TRUE),  # number of brooms
    gouting.sum_under = sum(gouting_under10, na.rm = TRUE),  # gouting sum
    basal_area_non_host = sum(BA_under10[host_non == 0], na.rm = TRUE),  # non-host BA
    basal_area_host = sum(BA_under10[host_non == 1], na.rm = TRUE),  # host BA
    percent_basal_area_non_host = basal_area_non_host / (basal_area_non_host + basal_area_host) * 100,  # percent non-host BA
    crown_ht_sum = sum(crown_ht__under10, na.rm = TRUE),
    mean_distance_infested = ifelse(sum(infestation_under10) == 0, NA, mean(DISTANCE.x[infestation_under10 == 1], na.rm = TRUE))  # mean distance of infested trees
  )

###

# Create a function to extract the variable based on the linked tree ID
# extractVariable <- function(TARGET, variable) {
#   linkedTreeRow <- data[data$unique_ID == TARGET, ]
#   if (nrow(linkedTreeRow) > 0) {
#     return(linkedTreeRow[[variable]])
#   } else {
#     return(NA)  # or any other value to indicate missing data
#   }
# }
# 
# str(data)
# 
# # Apply the function to create a new column with the extracted variable - need to change function for each extraction !!
# df$infestation_under10 <- sapply(df$TARGET, extractVariable, variable = "infested_v1")
# df$infestation.sum.1_under10 <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v1_focal")
# df$dbh_under10 <- sapply(df$TARGET, extractVariable, variable = "dbh")
# df$brooms_under10 <- sapply(df$TARGET, extractVariable, variable = "brooms")
# df$gouting_under10 <- sapply(df$TARGET, extractVariable, variable = "Gouting_v1")
# df$species_under10 <- sapply(df$TARGET, extractVariable, variable = "Species_v1")
# df$BA_under10 <- sapply(df$TARGET, extractVariable, variable = "BA")
# df$crown_ht__under10 <- sapply(df$TARGET, extractVariable, variable = "crown_height")
# df$dead_alive <- sapply(df$TARGET, extractVariable, variable = "dead_alive")
# 
# # Host and non-host BA
# df$host_non <- if_else(df$species_under10 == "ABLA", 1, 0)
# 
# # create a summary table of local stats for each tree
# local.stats <- df %>% 
#   dplyr::group_by(FROM) %>% 
#   dplyr::summarise(n = n(),
#                    number.infested_under10 = sum(infestation_under10), # number of infested
#                    percent.infested_under10 = (sum(infestation_under10)/n) * 100, # percent infested
#                    infestation.sum_under10 = sum(infestation.sum.1_under10), # infestation load or propagule pressure
#                    average.dbh_under10 = mean(dbh_under10, na.rm = TRUE), # average dbh
#                    brooms_under10 = sum(brooms_under10), # number of brooms (this may not be anything)
#                    # avg.gouting_under10 = mean(gouting_under10, na.rm = TRUE), # Since we only have 1's at EMIG, average isn't useful
#                    gouting.sum_under = sum(gouting_under10, na.rm = TRUE), # gouting sum
#                    basal_area_non_host = sum(BA_under10[!is.na(host_non) & host_non != 1], na.rm = TRUE), # non-host BA
#                    basal_area_host = sum(BA_under10[host_non == 1], na.rm = TRUE), # host BA
#                    percent_basal_area_non_host = basal_area_non_host / (basal_area_non_host + basal_area_host) * 100, # percent BA of non-host
#                    crown_ht_sum = sum(crown_ht__under10, na.rm = TRUE),
#                    mean_distance_infested = ifelse(sum(infestation_under10) == 0, NA, mean(DISTANCE[infestation_under10 == 1], na.rm = TRUE))) # If there's no infestation within 10m, it'll come out as NA 

names(local.stats)[1] <- "unique_ID"

# combine with data
data_with_local_stats <- data %>% inner_join(local.stats, by = "unique_ID")


```

# Visit 2 - Local Stats

```{r}

# Get all trees under 10m from focal tree
under10_holder2 <- data.dt[,.(TARGET2, DISTANCE.x), by=.(FROM)][,.SD[which(DISTANCE.x <= 10 & DISTANCE.x != 0)], by=.(FROM)]

filtered_distances2 <- under10_holder2 %>%
  # Extract the site info (assuming the site info is before the "_")
  mutate(site1 = sub("_.*", "", FROM), 
         site2 = sub("_.*", "", TARGET2)) %>%
  # Filter rows where site1 matches site2
  filter(site1 == site2) %>%
  # Optionally, remove the helper columns site1 and site2
  select(FROM, TARGET2, DISTANCE.x)

# View the filtered data
df2 <- filtered_distances2

### Different approach to accelerate processing time

# Create a lookup table from your `data`
lookup2 <- data %>% select(unique_ID, infested_v2, infestation_sum_v2_focal, dbh, brooms, Gouting_v2, species, BA, crown_height, `Dead / Alive_v2`)

# Join based on TARGET to bring in these variables
df2 <- df2 %>%
  left_join(lookup2, by = c("TARGET2" = "unique_ID")) %>%
  rename(infestation_under10_2 = infested_v2,
         infestation.sum.1_under10_2 = infestation_sum_v2_focal,
         dbh_under10_2 = dbh,
         brooms_under10_2 = brooms,
         gouting_under10_2 = Gouting_v2,
         species_under10_2 = species,
         BA_under10_2 = BA,
         crown_ht__under10_2 = crown_height,
         dead_alive_2 = `Dead / Alive_v2`)

# Host and non-host BA
df2 <- df2 %>%
  mutate(host_non = if_else(species_under10_2 == "ABLA", 1, 0))

# Now calculate local stats
local.stats_2 <- df2 %>%
  group_by(FROM) %>%
  summarise(
    n_2 = n(),
    number.infested_under10_2 = sum(infestation_under10_2, na.rm = TRUE),  # number of infested
    percent.infested_under10_2 = (sum(infestation_under10_2, na.rm = TRUE) / n_2) * 100,  # percent infested
    infestation.sum_under10_2 = sum(infestation.sum.1_under10_2, na.rm = TRUE),  # infestation load
    average.dbh_under10_2 = mean(dbh_under10_2, na.rm = TRUE),  # average dbh
    brooms_under10_2 = sum(brooms_under10_2, na.rm = TRUE),  # number of brooms
    gouting.sum_under_2 = sum(gouting_under10_2, na.rm = TRUE),  # gouting sum
    basal_area_non_host_2 = sum(BA_under10_2[host_non == 0], na.rm = TRUE),  # non-host BA
    basal_area_host_2 = sum(BA_under10_2[host_non == 1], na.rm = TRUE),  # host BA
    percent_basal_area_non_host_2 = basal_area_non_host_2 / (basal_area_non_host_2 + basal_area_host_2) * 100,  # percent non-host BA
    crown_ht_sum_2 = sum(crown_ht__under10_2, na.rm = TRUE),
    mean_distance_infested_2 = ifelse(sum(infestation_under10_2) == 0, NA, mean(DISTANCE.x[infestation_under10_2 == 1], na.rm = TRUE))  # mean distance of infested trees
  )

names(local.stats_2)[1] <- "unique_ID"

# combine with data
data_with_local_stats <- data_with_local_stats %>% inner_join(local.stats_2, by = "unique_ID")

```

# Visit 3 - Local Stats

```{r}

# Get all trees under 10m from focal tree
under10_holder3 <- data.dt[,.(TARGET3, DISTANCE.x), by=.(FROM)][,.SD[which(DISTANCE.x <= 10 & DISTANCE.x != 0)], by=.(FROM)]

filtered_distances3 <- under10_holder3 %>%
  # Extract the site info (assuming the site info is before the "_")
  mutate(site1 = sub("_.*", "", FROM), 
         site2 = sub("_.*", "", TARGET3)) %>%
  # Filter rows where site1 matches site2
  filter(site1 == site2) %>%
  # Optionally, remove the helper columns site1 and site2
  select(FROM, TARGET3, DISTANCE.x)

# View the filtered data
df3 <- filtered_distances3

### Different approach to accelerate processing time

# Create a lookup table from your `data`
lookup3 <- data %>% select(unique_ID, infested_v3, infestation_sum_v3_focal, dbh, brooms, Gouting_v3, species, BA, crown_height, `Dead / Alive_v3`)

# Join based on TARGET to bring in these variables
df3 <- df3 %>%
  left_join(lookup3, by = c("TARGET3" = "unique_ID")) %>%
  rename(infestation_under10_3 = infested_v3,
         infestation.sum.1_under10_3 = infestation_sum_v3_focal,
         dbh_under10_3 = dbh,
         brooms_under10_3 = brooms,
         gouting_under10_3 = Gouting_v3,
         species_under10_3 = species,
         BA_under10_3 = BA,
         crown_ht__under10_3 = crown_height,
         dead_alive_3 = `Dead / Alive_v3`)

# Host and non-host BA
df3 <- df3 %>%
  mutate(host_non = if_else(species_under10_3 == "ABLA", 1, 0))

# Now calculate local stats
local.stats_3 <- df3 %>%
  group_by(FROM) %>%
  summarise(
    n_3 = n(),
    number.infested_under10_3 = sum(infestation_under10_3, na.rm = TRUE),  # number of infested
    percent.infested_under10_3 = (sum(infestation_under10_3, na.rm = TRUE) / n_3) * 100,  # percent infested
    infestation.sum_under10_3 = sum(infestation.sum.1_under10_3, na.rm = TRUE),  # infestation load
    average.dbh_under10_3 = mean(dbh_under10_3, na.rm = TRUE),  # average dbh
    brooms_under10_3 = sum(brooms_under10_3, na.rm = TRUE),  # number of brooms
    gouting.sum_under_3 = sum(gouting_under10_3, na.rm = TRUE),  # gouting sum
    basal_area_non_host_3 = sum(BA_under10_3[host_non == 0], na.rm = TRUE),  # non-host BA
    basal_area_host_3 = sum(BA_under10_3[host_non == 1], na.rm = TRUE),  # host BA
    percent_basal_area_non_host_3 = basal_area_non_host_3 / (basal_area_non_host_3 + basal_area_host_3) * 100,  # percent non-host BA
    crown_ht_sum_3 = sum(crown_ht__under10_3, na.rm = TRUE),
    mean_distance_infested_3 = ifelse(sum(infestation_under10_3) == 0, NA, mean(DISTANCE.x[infestation_under10_3 == 1], na.rm = TRUE))  # mean distance of infested trees
  )

names(local.stats_3)[1] <- "unique_ID"

# combine with data
data_with_local_stats <- data_with_local_stats %>% inner_join(local.stats_3, by = "unique_ID")

```

# Visit 4 - Local Stats

```{r}

# Get all trees under 10m from focal tree
under10_holder4 <- data.dt[,.(TARGET4, DISTANCE.x), by=.(FROM)][,.SD[which(DISTANCE.x <= 10 & DISTANCE.x != 0)], by=.(FROM)]

filtered_distances4 <- under10_holder4 %>%
  # Extract the site info (assuming the site info is before the "_")
  mutate(site1 = sub("_.*", "", FROM), 
         site2 = sub("_.*", "", TARGET4)) %>%
  # Filter rows where site1 matches site2
  filter(site1 == site2) %>%
  # Optionally, remove the helper columns site1 and site2
  select(FROM, TARGET4, DISTANCE.x)

# View the filtered data
df4 <- filtered_distances4

### Different approach to accelerate processing time

# Create a lookup table from your `data`
lookup4 <- data %>% select(unique_ID, infested_v4, infestation_sum_v4_focal, dbh, brooms, Gouting_v4, species, BA, crown_height, `Dead / Alive_v4`)

# Join based on TARGET to bring in these variables
df4 <- df4 %>%
  left_join(lookup4, by = c("TARGET4" = "unique_ID")) %>%
  rename(infestation_under10_4 = infested_v4,
         infestation.sum.1_under10_4 = infestation_sum_v4_focal,
         dbh_under10_4 = dbh,
         brooms_under10_4 = brooms,
         gouting_under10_4 = Gouting_v4,
         species_under10_4 = species,
         BA_under10_4 = BA,
         crown_ht__under10_4 = crown_height,
         dead_alive_4 = `Dead / Alive_v4`)

# Host and non-host BA
df4 <- df4 %>%
  mutate(host_non = if_else(species_under10_4 == "ABLA", 1, 0))

# Now calculate local stats
local.stats_4 <- df4 %>%
  group_by(FROM) %>%
  summarise(
    n_4 = n(),
    number.infested_under10_4 = sum(infestation_under10_4, na.rm = TRUE),  # number of infested
    percent.infested_under10_4 = (sum(infestation_under10_4, na.rm = TRUE) / n_4) * 100,  # percent infested
    infestation.sum_under10_4 = sum(infestation.sum.1_under10_4, na.rm = TRUE),  # infestation load
    average.dbh_under10_4 = mean(dbh_under10_4, na.rm = TRUE),  # average dbh
    brooms_under10_4 = sum(brooms_under10_4, na.rm = TRUE),  # number of brooms
    gouting.sum_under_4 = sum(gouting_under10_4, na.rm = TRUE),  # gouting sum
    basal_area_non_host_4 = sum(BA_under10_4[host_non == 0], na.rm = TRUE),  # non-host BA
    basal_area_host_4 = sum(BA_under10_4[host_non == 1], na.rm = TRUE),  # host BA
    percent_basal_area_non_host_4 = basal_area_non_host_4 / (basal_area_non_host_4 + basal_area_host_4) * 100,  # percent non-host BA
    crown_ht_sum_4 = sum(crown_ht__under10_4, na.rm = TRUE),
    mean_distance_infested_4 = ifelse(sum(infestation_under10_4) == 0, NA, mean(DISTANCE.x[infestation_under10_4 == 1], na.rm = TRUE))  # mean distance of infested trees
  )

names(local.stats_4)[1] <- "unique_ID"

# combine with data
data_with_local_stats <- data_with_local_stats %>% inner_join(local.stats_4, by = "unique_ID")

#library(writexl)
#write_xlsx(data_with_local_stats, "all_sites_all_visits_under10_data.xlsx")

```

# Bayesian Model

```{r}

data_with_local_stats <- read_xlsx("all_sites_all_visits_under10_data.xlsx")

# Filter out hosts to not skew infested metric
host_local <- data_with_local_stats %>% filter(species == "ABLA" | species == "ABCO")

# Change NA's in mean distance infested to 0 - this is just to get it to run!! - NEED TO CHANGE FOR FINAL MODEL - 0 might be right since I think we're getting NA's because there's no infested within 10 meters
host_local$mean_distance_infested[is.na(host_local$mean_distance_infested)] <- 0

# Check for NA's

host_local$dbh[is.na(host_local$dbh)] <- 2.54/2
host_local$Gouting_v1[is.na(host_local$Gouting_v1)] <- 0
host_local$infested_v1[is.na(host_local$infested_v1)] <- 0

na_subset <- host_local[, c("infested_v1", "DISTANCE", "dbh", "Gouting_v1", "brooms", "infestation.sum_under10", "mean_distance_infested", "percent_basal_area_non_host", "basal_area_host", "average.dbh_under10", "number.infested_under10", "percent.infested_under10", "crown_ht_sum", "Name"), drop = FALSE]

anyNA(na_subset)

host_local$crown_ht_sum_std <- (host_local$crown_ht_sum - mean(host_local$crown_ht_sum))/sd(host_local$crown_ht_sum)
host_local$infest_sum_10_std <- (host_local$infestation.sum_under10 - mean(host_local$infestation.sum_under10))/sd(host_local$infestation.sum_under10)
host_local$basal_area_host_std <- (host_local$basal_area_host - mean(host_local$basal_area_host))/sd(host_local$basal_area_host)

pow <- function(base, exponent) {
  return(base^exponent)
}

brooms_recode <- ifelse(host_local$brooms > 0, 0, 1)
distance_to_nearest_recode <- ifelse(host_local$mean_distance_infested == 0, 11, host_local$mean_distance_infested)

infested_jags_data_re <- list(infested = host_local$infested_v1, 
                            logDISTANCE = log(host_local$DISTANCE),
                            DBH_focal = log(host_local$dbh), # DBH is zero-inflated
                            Gouting_focal = host_local$Gouting_v1,
#                            Brooms_focal = log(host_local$brooms + 1), # might want to consider binning brooms
                            Brooms_focal = brooms_recode, # might want to consider binning brooms
                            infest_sum_10 = host_local$infest_sum_10_std,
                            mean_dist_infest_10 = distance_to_nearest_recode,
                            # non_host_ba_perc = host_local$percent_basal_area_non_host,
                            host_ba = host_local$basal_area_host_std,
                            # avg_dbh_10 = host_local$average.dbh_under10,
                            n_infest_10 = log(host_local$number.infested_under10 + 1),
                            percent_infested_10 = host_local$percent.infested_under10,
                            # crown_ht_sum = host_local$crown_ht_sum_std,
                            tree = as.numeric(as.factor(host_local$unique_ID)),
                            N = nrow(host_local), 
                            N_tree = length(levels(factor(host_local$unique_ID)))
                                                              )
# Model
infested_jags_naive_re <- function(){
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)
  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
  # Likelihood:
  for (i in 1:N){
    lin_comb[i] <- beta_0 + (beta_distance * logDISTANCE[i]) + (beta_dbh * DBH_focal[i]) +
      (beta_gouting * Gouting_focal[i]) + (beta_brooms * Brooms_focal[i]) + (beta_infestsum * infest_sum_10[i]) + (beta_mean_dist_infest * mean_dist_infest_10[i]) + (beta_host_ba * host_ba[i])+ beta_n_infest * n_infest_10[i] + beta_percent_infest * percent_infested_10[i]
    #+ beta_non_host_ba_perc * non_host_ba_perc[i] + beta_host_ba * host_ba[i] + beta_avg_dbh * avg_dbh_10[i]  + beta_crown_ht * crown_ht_sum[i] + (beta_non_host_ba_perc * non_host_ba_perc[i])
    + beta_tree[tree[i]]
    
    # here's our linear combination of predictors, just like in a standard regression
    
     p[i] <- exp(lin_comb[i])/(1 + exp(lin_comb[i]))
    infested[i] ~ dbinom(p[i], 1) # Here's the Binomial likelihood,
  }
}

infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, 1), 
       beta_distance = rnorm(1, 0, 1),
       beta_dbh = rnorm(1, 0, 1),
       beta_gouting = rnorm(1, 0, 1),
       beta_brooms = rnorm(1, 0, 1),
       beta_infestsum = rnorm(1, 0, 1),
       beta_mean_dist_infest = rnorm(1, 0, 1),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, 1),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, 1),
       beta_percent_infest = rnorm(1, 0, 1),
       # beta_crown_ht = rnorm(1, 0, 1))
    #,
       beta_tree = rep(0, length(levels(factor(host_local$unique_ID)))))
       #p = runif(1, 0.1, 0.9))
}

 infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", # "beta_non_host_ba_perc", "beta_avg_dbh",  "beta_crown_ht",
   "sigma_sq_tree",
   "p")

infested_fit_naive_re <- jags(data = infested_jags_data_re, 
                            inits = infested_inits_re, 
                            parameters.to.save = infested_params_re, 
                            model.file = infested_jags_naive_re,
                            n.chains = 3,
                            n.iter = 12000, 
                            n.burnin = 2000, 
                            n.thin = 10, 
                            DIC = F)

mcmcplots::mcmcplot(infested_fit_naive_re)


```

# Nimble Bayesian model

This model contains our basic GLMM model structure with tree as a random effect, our local statistics under 10m, and focal tree attributes.

```{r}
library(nimble)


# Filter out hosts to not skew infested metric
host_local <- data_with_local_stats %>% filter(species == "ABLA" | species == "ABCO")

# Change NA's in mean distance infested to 0 - this is just to get it to run!! - NEED TO CHANGE FOR FINAL MODEL - 0 might be right since I think we're getting NA's because there's no infested within 10 meters
host_local$mean_distance_infested[is.na(host_local$mean_distance_infested)] <- 0

# Check for NA's

host_local$dbh[is.na(host_local$dbh)] <- 2.54/2
host_local$Gouting_v1[is.na(host_local$Gouting_v1)] <- 0
host_local$infested_v1[is.na(host_local$infested_v1)] <- 0

na_subset <- host_local[, c("infested_v1", "DISTANCE", "dbh", "Gouting_v1", "brooms", "infestation.sum_under10", "mean_distance_infested", "percent_basal_area_non_host", "basal_area_host", "average.dbh_under10", "number.infested_under10", "percent.infested_under10", "crown_ht_sum", "Name"), drop = FALSE]

anyNA(na_subset)

host_local$crown_ht_sum_std <- (host_local$crown_ht_sum - mean(host_local$crown_ht_sum))/sd(host_local$crown_ht_sum)
host_local$infest_sum_10_std <- (host_local$infestation.sum_under10 - mean(host_local$infestation.sum_under10))/sd(host_local$infestation.sum_under10)
host_local$basal_area_host_std <- (host_local$basal_area_host - mean(host_local$basal_area_host))/sd(host_local$basal_area_host)

pow <- function(base, exponent) {
  return(base^exponent)
}

brooms_recode <- ifelse(host_local$brooms > 0, 0, 1)
distance_to_nearest_recode <- ifelse(host_local$mean_distance_infested == 0, 11, host_local$mean_distance_infested)

infested_jags_constants_re <- list(infested = host_local$infested_v1, 
                            logDISTANCE = log(host_local$DISTANCE),
                            DBH_focal = log(host_local$dbh), # DBH is zero-inflated
                            Gouting_focal = host_local$Gouting_v1,
#                            Brooms_focal = log(host_local$brooms + 1), # might want to consider binning brooms
                            Brooms_focal = brooms_recode, # might want to consider binning brooms
                            infest_sum_10 = host_local$infest_sum_10_std,
                            mean_dist_infest_10 = distance_to_nearest_recode,
                            # non_host_ba_perc = host_local$percent_basal_area_non_host,
                            host_ba = host_local$basal_area_host_std,
                            # avg_dbh_10 = host_local$average.dbh_under10,
                            n_infest_10 = log(host_local$number.infested_under10 + 1),
                            percent_infested_10 = host_local$percent.infested_under10,
                            # crown_ht_sum = host_local$crown_ht_sum_std,
                            tree = as.numeric(as.factor(host_local$unique_ID)),
                            N = nrow(host_local), 
                            N_tree = length(levels(factor(host_local$unique_ID)))
                                                              )
# Model
infested_jags_naive_re_nimble <- nimbleCode({
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)
  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
  # Likelihood:
  for (i in 1:N){
    lin_comb[i] <- beta_0 + (beta_distance * logDISTANCE[i]) + (beta_dbh * DBH_focal[i]) + (beta_gouting * Gouting_focal[i]) + (beta_brooms * Brooms_focal[i]) + (beta_infestsum * infest_sum_10[i]) + (beta_mean_dist_infest * mean_dist_infest_10[i]) + (beta_host_ba * host_ba[i])+ beta_n_infest * n_infest_10[i] + beta_percent_infest * percent_infested_10[i] + beta_tree[tree[i]]
    #+ beta_non_host_ba_perc * non_host_ba_perc[i] + beta_host_ba * host_ba[i] + beta_avg_dbh * avg_dbh_10[i]  + beta_crown_ht * crown_ht_sum[i] + (beta_non_host_ba_perc * non_host_ba_perc[i])

    
    # here's our linear combination of predictors, just like in a standard regression
    
     p[i] <- exp(lin_comb[i])/(1 + exp(lin_comb[i]))
    infested[i] ~ dbinom(p[i], 1) # Here's the Binomial likelihood,
  }
}) # closes out nimbleCode

input_sd <- .5
infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, input_sd), 
       beta_distance = rnorm(1, 0, input_sd),
       beta_dbh = rnorm(1, 0, input_sd),
       beta_gouting = rnorm(1, 0, input_sd),
       beta_brooms = rnorm(1, 0, input_sd),
       beta_infestsum = rnorm(1, 0, input_sd),
       beta_mean_dist_infest = rnorm(1, 0, input_sd),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, input_sd),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, input_sd),
       beta_percent_infest = rnorm(1, 0, input_sd),
       # beta_crown_ht = rnorm(1, 0, 1))
    #,
       beta_tree = rep(0, length(levels(factor(host_local$unique_ID)))))
       #p = runif(1, 0.1, 0.9))
}

 infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", 
   "sigma_sq_tree",
   "p")

## iii. Build Nimble Model ----
## (this can be a medium-slow step) (specifies the graph associated with the model)
nimbleModel_vect <- nimbleModel(infested_jags_naive_re_nimble, # nimbleCode model
                                constants = infested_jags_constants_re,
                                inits = infested_inits_re())

 # assign default samplers to nodes
mcmcConf <- configureMCMC(nimbleModel_vect, monitors = infested_params_re)

# 3. build MCMC object: can customize thinning, monitors, etc.
modelMCMC <- buildMCMC(mcmcConf) # update mcmc configuration based on graph

# compile the code to run the samplers to C
cModel <- compileNimble(nimbleModel_vect) # compiled model
Cmcmc <- compileNimble(modelMCMC, project = cModel, resetFunctions = T) # compiled mcmc

M1_out <- runMCMC(Cmcmc,
                  niter = 100000, # 1000000,
                  nburnin = 70000, # 500000,
                  thin = 10, # 500,
                  nchains = 3,
                  samplesAsCodaMCMC = T,
                  summary = T)

mcmcplots::mcmcplot(M1_out$samples)

```

# Model structure with lag t built in - option 2

```{r}

# Filter out hosts to not skew infested metric
host_local <- data_with_local_stats %>% filter(species == "ABLA" | species == "ABCO")

# Change NA's in mean distance infested to 0 - this is just to get it to run!! - NEED TO CHANGE FOR FINAL MODEL - 0 might be right since I think we're getting NA's because there's no infested within 10 meters
host_local$mean_distance_infested[is.na(host_local$mean_distance_infested)] <- 0

# Check for NA's

host_local$dbh[is.na(host_local$dbh)] <- 2.54/2
host_local$Gouting_v1[is.na(host_local$Gouting_v1)] <- 0
host_local$Gouting_v2[is.na(host_local$Gouting_v2)] <- 0
host_local$Gouting_v3[is.na(host_local$Gouting_v3)] <- 0
host_local$Gouting_v4[is.na(host_local$Gouting_v4)] <- 0
host_local$infested_v1[is.na(host_local$infested_v1)] <- 0
host_local$infested_v2[is.na(host_local$infested_v2)] <- 0
host_local$infested_v3[is.na(host_local$infested_v3)] <- 0
host_local$infested_v4[is.na(host_local$infested_v4)] <- 0

# Log distance for model
host_local$DISTANCE <- log(host_local$DISTANCE)
host_local$DISTANCE <- log(host_local$DISTANCE2)
host_local$DISTANCE3 <- log(host_local$DISTANCE3)
host_local$DISTANCE4 <- log(host_local$DISTANCE4)

# Log number of infested for indexing
host_local$number.infested_under10 <- log(host_local$number.infested_under10 + 1)
host_local$number.infested_under10_2 <- log(host_local$number.infested_under10_2 + 1)
host_local$number.infested_under10_3<- log(host_local$number.infested_under10_3 + 1)
host_local$number.infested_under10_4 <- log(host_local$number.infested_under10_4 + 1)


na_subset <- host_local[, c("infested_v1", "infested_v2", "infested_v3", "infested_v4", "DISTANCE", "dbh", "Gouting_v1", "Gouting_v2", "Gouting_v3", "Gouting_v4", "brooms", "infestation.sum_under10", "mean_distance_infested", "percent_basal_area_non_host", "basal_area_host", "average.dbh_under10", "number.infested_under10", "percent.infested_under10", "crown_ht_sum", "Name", "infestation.sum_under10_2", "mean_distance_infested_2", "number.infested_under10_2", "percent.infested_under10_2", "infestation.sum_under10_3", "mean_distance_infested_3", "number.infested_under10_3", "percent.infested_under10_3", "infestation.sum_under10_4", "mean_distance_infested_4", "number.infested_under10_4", "percent.infested_under10_4"), drop = FALSE]

anyNA(na_subset$infested_v2)

# Standardizing variables
host_local$crown_ht_sum_std <- (host_local$crown_ht_sum - mean(host_local$crown_ht_sum))/sd(host_local$crown_ht_sum)
host_local$infest_sum_10_std <- (host_local$infestation.sum_under10 - mean(host_local$infestation.sum_under10))/sd(host_local$infestation.sum_under10)
host_local$basal_area_host_std <- (host_local$basal_area_host - mean(host_local$basal_area_host))/sd(host_local$basal_area_host)
brooms_recode <- ifelse(host_local$brooms > 0, 0, 1)

distance_to_nearest_recode <- ifelse(host_local$mean_distance_infested == 0, 11, host_local$mean_distance_infested) # this changes zeros to 11 so they're not included in local stats
distance_to_nearest_recode_2 <- ifelse(host_local$mean_distance_infested_2 == 0, 11, host_local$mean_distance_infested_2) 
distance_to_nearest_recode_3 <- ifelse(host_local$mean_distance_infested_3 == 0, 11, host_local$mean_distance_infested_3) 
distance_to_nearest_recode_4 <- ifelse(host_local$mean_distance_infested_4 == 0, 11, host_local$mean_distance_infested_4) 



infested_jags_constants_re <- list(infested = as.matrix(host_local[, c("infested_v1", "infested_v2", "infested_v3", "infested_v4")]), 
                            logDISTANCE = as.matrix(host_local[, c("DISTANCE", "DISTANCE2", "DISTANCE3", "DISTANCE4")]),
                            DBH_focal = log(as.matrix(host_local[, c("dbh", "dbh", "dbh", "dbh")])), # DBH is zero-inflated
                            Gouting_focal = as.matrix(host_local[, c("Gouting_v1", "Gouting_v2", "Gouting_v3", "Gouting_v4")]),
                            # Brooms_focal = log(host_local$brooms + 1), # might want to consider binning brooms
                            Brooms_focal = as.matrix(cbind(brooms_recode, brooms_recode, brooms_recode, brooms_recode)), # might want to consider binning brooms
                            infest_sum_10 = as.matrix(host_local[, c("infestation.sum_under10", "infestation.sum_under10_2", "infestation.sum_under10_3", "infestation.sum_under10_4")]),
                            mean_dist_infest_10 = as.matrix(cbind(distance_to_nearest_recode, distance_to_nearest_recode_2, distance_to_nearest_recode_3, distance_to_nearest_recode_4)),
                            # non_host_ba_perc = host_local$percent_basal_area_non_host,
                            host_ba = as.matrix(host_local[, c("basal_area_host_std", "basal_area_host_std", "basal_area_host_std", "basal_area_host_std")]),
                            # avg_dbh_10 = host_local$average.dbh_under10,
                            n_infest_10 = as.matrix(host_local[, c("number.infested_under10", "number.infested_under10_2", "number.infested_under10_3", "number.infested_under10_4")]),
                            percent_infested_10 = as.matrix(host_local[, c("percent.infested_under10", "percent.infested_under10_2", "percent.infested_under10_3", "percent.infested_under10_4")]),
                            # crown_ht_sum = host_local$crown_ht_sum_std,
                            tree = as.numeric(as.factor(host_local$unique_ID)),
                            N = nrow(host_local), 
                            N_tree = length(levels(factor(host_local$unique_ID))),
                            visit = 4
                                                              )
# Model
infested_jags_naive_re_nimble <- nimbleCode({
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)
  beta_distance_infest ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh_infest ~ dnorm(0, 0.5)
  beta_gouting_infest ~ dnorm(0, 0.5)
  beta_brooms_infest ~ dnorm(0, 0.5)
  beta_infestsum_infest ~ dnorm(0, 0.5)
  beta_mean_dist_infest_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba_infest ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest_infest ~ dnorm(0, 0.5)
  beta_percent_infest_infest ~ dnorm(0, 0.5)

  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
# Likelihood:
  for (i in 1:N) {
    # Visit 1 (no prior information from previous visits)
    # lin_comb[i, 1] <- beta_0 + (beta_distance * logDISTANCE[i, 1]) + (beta_dbh * DBH_focal[i, 1]) +
    #                   (beta_gouting * Gouting_focal[i, 1]) + (beta_brooms * Brooms_focal[i, 1]) +
    #                   (beta_infestsum * infest_sum_10[i, 1]) + (beta_mean_dist_infest * mean_dist_infest_10[i, 1]) +
    #                   (beta_host_ba * host_ba[i, 1]) + (beta_n_infest * n_infest_10[i, 1]) +
    #                   (beta_percent_infest * percent_infested_10[i, 1]) + beta_tree[tree[i]]
    # # 
    # # Logistic regression for visit 1:
    # p[i, 1] <- exp(lin_comb[i, 1]) / (1 + exp(lin_comb[i, 1]))
    # infested[i, 1] ~ dbinom(p[i, 1], 1)  # Likelihood for visit 1
    # # 
    # # # Visits 2 to 4: Including information from previous visit (infested[i, v-1])
    for (v in 2:visit) {
    # #   # option 1: completely remove effects associated with propagule pressure once you're infested
    #   lin_comb[i, v] <- (beta_0 + (beta_distance * logDISTANCE[i, v]) + (beta_dbh * DBH_focal[i, v]) +
    #                     (beta_gouting * Gouting_focal[i, v]) + (beta_brooms * Brooms_focal[i, v]) +
    #                     (beta_infestsum * infest_sum_10[i, v]) + (beta_mean_dist_infest * mean_dist_infest_10[i, v]) +
    #                     (beta_host_ba * host_ba[i, v]) + (beta_n_infest * n_infest_10[i, v]) +
    #                     (beta_percent_infest * percent_infested_10[i, v])) * (1 - infested[i, v-1]) +
    #                     beta_tree[tree[i]] +
    #                     beta_previous * infested[i, v-1]  # Infestation status from the previous visit - absolute value to make the linear combo drop out if it was infested last time
      
      # option 2: put all propagule pressure effects in interaction with infestation status
       lin_comb[i, v] <- beta_0 + (beta_distance * logDISTANCE[i, v]) + (beta_dbh * DBH_focal[i, v]) +
                        (beta_gouting * Gouting_focal[i, v]) + (beta_brooms * Brooms_focal[i, v]) +
                        (beta_infestsum * infest_sum_10[i, v]) + (beta_mean_dist_infest * mean_dist_infest_10[i, v]) +
                        (beta_host_ba * host_ba[i, v]) + (beta_n_infest * n_infest_10[i, v]) +
                        (beta_percent_infest * percent_infested_10[i, v]) +
         (beta_distance_infest * logDISTANCE[i, v] * infested[i, v-1]) +
         (beta_dbh_infest * DBH_focal[i, v] * infested[i, v-1]) +
         (beta_gouting_infest * Gouting_focal[i, v] * infested[i, v-1]) +
         (beta_brooms * Brooms_focal[i, v] * infested[i, v-1]) +
         (beta_infestsum_infest * infest_sum_10[i, v] * infested[i, v-1]) +
         (beta_mean_dist_infest_infest * mean_dist_infest_10[i, v] * infested[i, v-1]) +
         (beta_host_ba_infest * host_ba[i, v] * infested[i, v-1]) +
         (beta_n_infest_infest * n_infest_10[i, v] * infested[i, v-1]) +
         (beta_percent_infest_infest * percent_infested_10[i, v] * infested[i, v-1]) +
       beta_tree[tree[i]] +
                        beta_previous * infested[i, v-1]  # Infestation status from the previous visit
     
      # Logistic regression for visits 2 to 4:
       # Constrained logistic regression for visits 2 to visit
        # p[i, v] <- max(1e-8, min(1 - 1e-8, exp(lin_comb[i, v]) / (1 + exp(lin_comb[i, v]))))
       p[i, v] <- exp(lin_comb[i, v]) / (1 + exp(lin_comb[i, v]))
      infested[i, v] ~ dbinom(p[i, v], 1)  # Likelihood for visit v
    }
  }
}) # closes out nimbleCode

input_sd <- .5
infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, input_sd), 
       beta_distance = rnorm(1, 0, input_sd),
       beta_dbh = rnorm(1, 0, input_sd),
       beta_gouting = rnorm(1, 0, input_sd),
       beta_brooms = rnorm(1, 0, input_sd),
       beta_infestsum = rnorm(1, 0, input_sd),
       beta_mean_dist_infest = rnorm(1, 0, input_sd),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, input_sd),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, input_sd),
       beta_percent_infest = rnorm(1, 0, input_sd),
       beta_distance_infest = rnorm(1, 0, input_sd),
       beta_dbh_infest = rnorm(1, 0, input_sd),
       beta_gouting_infest = rnorm(1, 0, input_sd),
       beta_brooms_infest = rnorm(1, 0, input_sd),
       beta_infestsum_infest = rnorm(1, 0, input_sd),
       beta_mean_dist_infest_infest = rnorm(1, 0, input_sd),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba_infest = rnorm(1, 0, input_sd),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest_infest = rnorm(1, 0, input_sd),
       beta_percent_infest_infest = rnorm(1, 0, input_sd),

       beta_previous = rnorm(1, 0, input_sd),
       # beta_crown_ht = rnorm(1, 0, 1))
    #,
       beta_tree = rep(0, length(levels(factor(host_local$unique_ID)))))
       #p = runif(1, 0.1, 0.9))
}

 infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest",
   "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
   "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest",
   "sigma_sq_tree",
   "p")

## iii. Build Nimble Model ----
## (this can be a medium-slow step) (specifies the graph associated with the model)
nimbleModel_vect <- nimbleModel(infested_jags_naive_re_nimble, # nimbleCode model
                                constants = infested_jags_constants_re,
                                inits = infested_inits_re())

 # assign default samplers to nodes
mcmcConf <- configureMCMC(nimbleModel_vect, monitors = infested_params_re)

# 3. build MCMC object: can customize thinning, monitors, etc.
modelMCMC <- buildMCMC(mcmcConf) # update mcmc configuration based on graph

# compile the code to run the samplers to C
cModel <- compileNimble(nimbleModel_vect) # compiled model
Cmcmc <- compileNimble(modelMCMC, project = cModel, resetFunctions = T) # compiled mcmc

## update samples to multivariate normals
### vi. customize sampler ----
# customize sampler to block betas for svbs
# try blocking all the baseline types together
# mcmcConf$removeSamplers("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
#    "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
#    "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest")
# 
# mcmcConf$addSampler(target = c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
#    "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
#    "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest"
# ), type = 'RW_block')
# 
# updatedMcmc <- buildMCMC(mcmcConf)
# CupdatedMcmc <- compileNimble(updatedMcmc)

# started: 9:24 for Moj
mcmc_rev_out <- runMCMC(Cmcmc,
                        niter = 5000,
                        nburnin = 1000,
                        nchains = 3,
                        samplesAsCodaMCMC = T,
                        summary = T)

# M1_out <- runMCMC(Cmcmc,
#                   niter = 100000, # 1000000,
#                   nburnin = 50000, # 500000,
#                   thin = 50, # 500,
#                   nchains = 3,
#                   samplesAsCodaMCMC = T,
#                   summary = T)

mcmcplots::mcmcplot(mcmc_rev_out$samples)

```

# Model structure with lag t built in - option 1

```{r}

infested_jags_constants_re <- list(infested = as.matrix(host_local[, c("infested_v1", "infested_v2", "infested_v3", "infested_v4")]), 
                            logDISTANCE = as.matrix(host_local[, c("DISTANCE", "DISTANCE2", "DISTANCE3", "DISTANCE4")]),
                            DBH_focal = log(as.matrix(host_local[, c("dbh", "dbh", "dbh", "dbh")])), # DBH is zero-inflated
                            Gouting_focal = as.matrix(host_local[, c("Gouting_v1", "Gouting_v2", "Gouting_v3", "Gouting_v4")]),
                            # Brooms_focal = log(host_local$brooms + 1), # might want to consider binning brooms
                            Brooms_focal = as.matrix(cbind(brooms_recode, brooms_recode, brooms_recode, brooms_recode)), # might want to consider binning brooms
                            infest_sum_10 = as.matrix(host_local[, c("infestation.sum_under10", "infestation.sum_under10_2", "infestation.sum_under10_3", "infestation.sum_under10_4")]),
                            mean_dist_infest_10 = as.matrix(cbind(distance_to_nearest_recode, distance_to_nearest_recode_2, distance_to_nearest_recode_3, distance_to_nearest_recode_4)),
                            # non_host_ba_perc = host_local$percent_basal_area_non_host,
                            host_ba = as.matrix(host_local[, c("basal_area_host_std", "basal_area_host_std", "basal_area_host_std", "basal_area_host_std")]),
                            # avg_dbh_10 = host_local$average.dbh_under10,
                            n_infest_10 = as.matrix(host_local[, c("number.infested_under10", "number.infested_under10_2", "number.infested_under10_3", "number.infested_under10_4")]),
                            percent_infested_10 = as.matrix(host_local[, c("percent.infested_under10", "percent.infested_under10_2", "percent.infested_under10_3", "percent.infested_under10_4")]),
                            # crown_ht_sum = host_local$crown_ht_sum_std,
                            tree = as.numeric(as.factor(host_local$unique_ID)),
                            N = nrow(host_local), 
                            N_tree = length(levels(factor(host_local$unique_ID))),
                            visit = 4
                                                              )
# Model
infested_jags_naive_re_nimble <- nimbleCode({
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)
  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
# Likelihood:
  for (i in 1:N) {
    
    # Visit 1 (no prior information from previous visits)
    lin_comb[i, 1] <- beta_0 + (beta_distance * logDISTANCE[i, 1]) + (beta_dbh * DBH_focal[i, 1]) +
                      (beta_gouting * Gouting_focal[i, 1]) + (beta_brooms * Brooms_focal[i, 1]) +
                      (beta_infestsum * infest_sum_10[i, 1]) + (beta_mean_dist_infest * mean_dist_infest_10[i, 1]) +
                      (beta_host_ba * host_ba[i, 1]) + (beta_n_infest * n_infest_10[i, 1]) +
                      (beta_percent_infest * percent_infested_10[i, 1]) + beta_tree[tree[i]]
    
    # Logistic regression for visit 1:
    p[i, 1] <- exp(lin_comb[i, 1]) / (1 + exp(lin_comb[i, 1]))
    infested[i, 1] ~ dbinom(p[i, 1], 1)  # Likelihood for visit 1
    
    # # Visits 2 to 4: Including information from previous visit (infested[i, v-1])
    for (v in 2:visit) {
    #   # option 1: completely remove effects associated with propagule pressure once you're infested
      lin_comb[i, v] <- (beta_0 + (beta_distance * logDISTANCE[i, v]) + (beta_dbh * DBH_focal[i, v]) +
                        (beta_gouting * Gouting_focal[i, v]) + (beta_brooms * Brooms_focal[i, v]) +
                        (beta_infestsum * infest_sum_10[i, v]) + (beta_mean_dist_infest * mean_dist_infest_10[i, v]) +
                        (beta_host_ba * host_ba[i, v]) + (beta_n_infest * n_infest_10[i, v]) +
                        (beta_percent_infest * percent_infested_10[i, v])) * (1 - infested[i, v-1]) +
                        beta_tree[tree[i]] +
                        beta_previous * infested[i, v-1]  # Infestation status from the previous visit - absolute value to make the linear combo drop out if it was infested last time
      
     
      p[i, v] <- exp(lin_comb[i, v]) / (1 + exp(lin_comb[i, v]))
      infested[i, v] ~ dbinom(p[i, v], 1)  # Likelihood for visit v
    }
  }
}) # closes out nimbleCode

input_sd <- .25
infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, input_sd), 
       beta_distance = rnorm(1, 0, input_sd),
       beta_dbh = rnorm(1, 0, input_sd),
       beta_gouting = rnorm(1, 0, input_sd),
       beta_brooms = rnorm(1, 0, input_sd),
       beta_infestsum = rnorm(1, 0, input_sd),
       beta_mean_dist_infest = rnorm(1, 0, input_sd),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, input_sd),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, input_sd),
       beta_percent_infest = rnorm(1, 0, input_sd),
       beta_previous = rnorm(1, 0, input_sd),
       beta_tree = rep(0, length(levels(factor(host_local$unique_ID)))),
       sigma_sq_tree = .2,
       p = runif(1, 0.1, 0.9))
}

infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_previous", "sigma_sq_tree"
  # , "p"
  )

## iii. Build Nimble Model ----
## (this can be a medium-slow step) (specifies the graph associated with the model)
nimbleModel_vect <- nimbleModel(infested_jags_naive_re_nimble, # nimbleCode model
                                constants = infested_jags_constants_re,
                                inits = infested_inits_re())

 # assign default samplers to nodes
mcmcConf <- configureMCMC(nimbleModel_vect, monitors = infested_params_re)

# 3. build MCMC object: can customize thinning, monitors, etc.
modelMCMC <- buildMCMC(mcmcConf) # update mcmc configuration based on graph

# compile the code to run the samplers to C
cModel <- compileNimble(nimbleModel_vect) # compiled model
Cmcmc <- compileNimble(modelMCMC, project = cModel, resetFunctions = T) # compiled mcmc

## update samples to multivariate normals
### vi. customize sampler ----
# customize sampler to block betas for svbs
# try blocking all the baseline types together
# mcmcConf$removeSamplers("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
#    "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
#    "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest")
# 
# mcmcConf$addSampler(target = c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
#    "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
#    "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest"
# ), type = 'RW_block')
# 
# updatedMcmc <- buildMCMC(mcmcConf)
# CupdatedMcmc <- compileNimble(updatedMcmc)

# started: 9:24 for Moj
mcmc_rev_out_opt1 <- runMCMC(Cmcmc,
                        niter = 5000,
                        nburnin = 1000,
                        nchains = 3,
                        samplesAsCodaMCMC = T,
                        summary = T)

mcmcplots::mcmcplot(mcmc_rev_out_opt1$samples)

```

# Bayesian Model - Infestation sum and change

```{r}


infested_jags_constants_re <- list(
                            infest_sum = as.matrix(host_local[, c("infestation_sum_v1_focal", "infestation_sum_v2_focal", "infestation_sum_v3_focal", "infestation_sum_v4_focal")]),
                            # infested = as.matrix(host_local[, c("infested_v1", "infested_v2", "infested_v3", "infested_v4")]),
                            logDISTANCE = as.matrix(host_local[, c("DISTANCE", "DISTANCE2", "DISTANCE3", "DISTANCE4")]),
                            DBH_focal = log(as.matrix(host_local[, c("dbh", "dbh", "dbh", "dbh")])), # DBH is zero-inflated
                            Gouting_focal = as.matrix(host_local[, c("Gouting_v1", "Gouting_v2", "Gouting_v3",                                           "Gouting_v4")]),
                            # Brooms_focal = log(host_local$brooms + 1), # might want to consider binning brooms
                            Brooms_focal = as.matrix(cbind(brooms_recode, brooms_recode, brooms_recode, brooms_recode)), # might want to consider binning brooms
                            infest_sum_10 = as.matrix(host_local[, c("infestation.sum_under10", "infestation.sum_under10_2",                              "infestation.sum_under10_3", "infestation.sum_under10_4")]),
                            mean_dist_infest_10 = as.matrix(cbind(distance_to_nearest_recode, distance_to_nearest_recode_2,                               distance_to_nearest_recode_3, distance_to_nearest_recode_4)),
                            # non_host_ba_perc = host_local$percent_basal_area_non_host,
                            host_ba = as.matrix(host_local[, c("basal_area_host_std", "basal_area_host_std",                                              "basal_area_host_std", "basal_area_host_std")]),
                            # avg_dbh_10 = host_local$average.dbh_under10,
                            n_infest_10 = as.matrix(host_local[, c("number.infested_under10", "number.infested_under10_2",                                "number.infested_under10_3", "number.infested_under10_4")]),
                            percent_infested_10 = as.matrix(host_local[, c("percent.infested_under10",                                                    "percent.infested_under10_2", "percent.infested_under10_3", "percent.infested_under10_4")]),
                            # crown_ht_sum = host_local$crown_ht_sum_std,
                            tree = as.numeric(as.factor(host_local$unique_ID)),
                            N = nrow(host_local), 
                            N_tree = length(levels(factor(host_local$unique_ID))),
                            visit = 4
                                                              )
# Model
infested_jags_naive_re_nimble <- nimbleCode({
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  # beta_infested ~ dnorm(0, 0.5)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)

  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
# Likelihood:
  for (i in 1:N) {
    # Visit 1 (no prior information from previous visits)
    lin_comb[i, 1] <- beta_0 + (beta_infested * infested[i, 1]) + (beta_distance * logDISTANCE[i, 1]) + (beta_dbh *                                  DBH_focal[i, 1]) + (beta_gouting * Gouting_focal[i, 1]) + (beta_brooms * Brooms_focal[i, 1]) +
                      (beta_infestsum * infest_sum_10[i, 1]) + (beta_mean_dist_infest * mean_dist_infest_10[i, 1]) +
                      (beta_host_ba * host_ba[i, 1]) + (beta_n_infest * n_infest_10[i, 1]) +
                      (beta_percent_infest * percent_infested_10[i, 1]) + beta_tree[tree[i]]
    #
    # Logistic regression for visit 1:
    lambda[i, 1] <- exp(lin_comb[i, 1])
    infest_sum[i, 1] ~ dpois(lambda[i, 1])  # Poisson likelihood for visit 1
    # #
    # # # Visits 2 to 4: Including information from previous visit (infested[i, v-1])
    for (v in 2:visit) {
    #   # option 1: completely remove effects associated with propagule pressure once you're infested
      lin_comb[i, v] <- (beta_0 + (beta_infested * infested[i, 1]) + (beta_distance * logDISTANCE[i, v]) +                                            (beta_dbh * DBH_focal[i, v]) + (beta_gouting * Gouting_focal[i, v]) +                                                         (beta_brooms * Brooms_focal[i,  v]) + (beta_infestsum * infest_sum_10[i, v]) +                                                (beta_mean_dist_infest * mean_dist_infest_10[i, v]) +
                        (beta_host_ba * host_ba[i, v]) + (beta_n_infest * n_infest_10[i, v]) +
                        (beta_percent_infest * percent_infested_10[i, v])) * (1 - infested[i, v-1]) +
                        beta_tree[tree[i]] +
                        beta_previous * infested[i, v-1]
     
      # Logistic regression for visits 2 to 4:
      lambda[i, v] <- exp(lin_comb[i, v])
      infest_sum[i, v] ~ dpois(lambda[i, v])  # Poisson likelihood for visits 2 to 4
    }
  }
}) # closes out nimbleCode

input_sd <- .5
infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, input_sd), 
       # beta_infested = rnorm(1, 0, input_sd),
       beta_distance = rnorm(1, 0, input_sd),
       beta_dbh = rnorm(1, 0, input_sd),
       beta_gouting = rnorm(1, 0, input_sd),
       beta_brooms = rnorm(1, 0, input_sd),
       beta_infestsum = rnorm(1, 0, input_sd),
       beta_mean_dist_infest = rnorm(1, 0, input_sd),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, input_sd),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, input_sd),
       beta_percent_infest = rnorm(1, 0, input_sd),
       # beta_distance_infest = rnorm(1, 0, input_sd),
       # beta_dbh_infest = rnorm(1, 0, input_sd),
       # beta_gouting_infest = rnorm(1, 0, input_sd),
       # beta_brooms_infest = rnorm(1, 0, input_sd),
       # beta_infestsum_infest = rnorm(1, 0, input_sd),
       # beta_mean_dist_infest_infest = rnorm(1, 0, input_sd),
       # #beta_non_host_ba_perc = rnorm(1, 0, 1),
       # beta_host_ba_infest = rnorm(1, 0, input_sd),
       # # beta_avg_dbh = rnorm(1, 0, 1),
       # beta_n_infest_infest = rnorm(1, 0, input_sd),
       # beta_percent_infest_infest = rnorm(1, 0, input_sd),

       beta_previous = rnorm(1, 0, input_sd),
       # beta_crown_ht = rnorm(1, 0, 1))
    #,
       beta_tree = rep(0, length(levels(factor(host_local$unique_ID)))))
       lambda = runif(1, 0.1, 0.9)
}

 infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest",
   # "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
   # "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest",
   "sigma_sq_tree",
   "lambda")

## iii. Build Nimble Model ----
## (this can be a medium-slow step) (specifies the graph associated with the model)
nimbleModel_vect <- nimbleModel(infested_jags_naive_re_nimble, # nimbleCode model
                                constants = infested_jags_constants_re,
                                inits = infested_inits_re())

 # assign default samplers to nodes
mcmcConf <- configureMCMC(nimbleModel_vect, monitors = infested_params_re)

# 3. build MCMC object: can customize thinning, monitors, etc.
modelMCMC <- buildMCMC(mcmcConf) # update mcmc configuration based on graph

# compile the code to run the samplers to C
cModel <- compileNimble(nimbleModel_vect) # compiled model
Cmcmc <- compileNimble(modelMCMC, project = cModel, resetFunctions = T) # compiled mcmc

## update samples to multivariate normals
### vi. customize sampler ----
# customize sampler to block betas for svbs
# try blocking all the baseline types together
# mcmcConf$removeSamplers("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
#    "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
#    "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest")
# 
# mcmcConf$addSampler(target = c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
#    "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest", "beta_distance_infest", "beta_dbh_infest", "beta_gouting_infest", "beta_brooms_infest", "beta_infestsum_infest",
#    "beta_mean_dist_infest_infest", "beta_host_ba_infest", "beta_n_infest_infest", "beta_percent_infest_infest"
# ), type = 'RW_block')
# 
# updatedMcmc <- buildMCMC(mcmcConf)
# CupdatedMcmc <- compileNimble(updatedMcmc)

# started: 9:24 for Moj
mcmc_rev_out <- runMCMC(Cmcmc,
                        niter = 5000,
                        nburnin = 1000,
                        nchains = 3,
                        samplesAsCodaMCMC = T,
                        summary = T)

```