---
title: "Cox Spring Local"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Local Densities
2/8/24
Michael Wayman

The purpose of this script and the following seven scripts is to break each site down into local densities and look at how the patterns we've seen may or may not appear on the site level. The first pattern we want to look at is this infestation sum and prevalence idea that we've seen across sites. If it holds for individual sites, then we may have some insight into a site-level density-dependence relationship.

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

# clean out environment

```{r}
rm(list = ls())

```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(Matrix) # to use glmer
library(raster) # for moving window analysis

```

## Load data

```{r}

trees <- read_xlsx("data_with_nearest_info.xlsx")
cxsp <- trees %>% dplyr::filter(trees$`Site ID_Visit1` == "CXSP")
cxsp_points <- read_xlsx("cxsp_points.xlsx")
combined <- read_xlsx("combined_cleaned.xlsx")

```

## Convert coordinates to use meters

```{r}

xy = data.frame(cxsp$Longitude.focal, cxsp$Latitude.focal)
coordinates(xy) <- c("cxsp.Longitude.focal", "cxsp.Latitude.focal")
proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")
NE <- spTransform(xy, CRS("+proj=utm +zone=12 ellps=WGS84"))
NE <- as.data.frame(NE)

cxsp$Longitude.focal <- NE$coords.x1
cxsp$Latitude.focal <- NE$coords.x2

```

# Get all trees within 10m of focal tree

```{r}

Holder = as.data.frame(matrix(nrow=0,ncol=4)) 
colnames(Holder) =c("PLOT", "FROM", "TARGET", "DISTANCE")

#Assign rownames as Tree-ID, dataset also needs UTM_X, UTM_Y, and PLOT
TEMP_DIST = vegdist(cxsp[,c("Longitude.focal","Latitude.focal")], method = "euclid")
TEMP_DIST.df = as.data.frame(as.matrix(TEMP_DIST))
rownames(TEMP_DIST.df) = (cxsp$unique_ID)
colnames(TEMP_DIST.df) = (cxsp$unique_ID)
TEMP_DIST.df$`Tag #` = rownames(TEMP_DIST.df)
TEMP_DIST_MELT = reshape2::melt(TEMP_DIST.df, id.vars = c("Tag #"))
colnames(TEMP_DIST_MELT) = c("FROM", "TARGET", "DISTANCE")

Holder = rbind(Holder, TEMP_DIST_MELT) #This is your exit


#Merge with infestation data
names(Holder)[1] <- "unique_ID"

Holder_new <- merge(Holder, cxsp, by.x = "TARGET", by.y = "unique_ID")
holder_new.dt <- as.data.table(Holder_new)
length(unique(holder_new.dt$unique_ID))
nrow(holder_new.dt)
names(holder_new.dt)[11] <- "infested_1"
names(holder_new.dt)[38] <- "infested_2"

#min_dist_holder <- holder_new.dt[`infested` == 1 & (DISTANCE == min(DISTANCE[(DISTANCE != 0)])) ,.(DISTANCE, TARGET),by=.("unique_ID")]

# distance to all infested less than 10m away
ten_m_dist_holder <- holder_new.dt[infested_1 == 1 ,.(TARGET, DISTANCE.x),by=.(unique_ID)][, .SD[which(DISTANCE.x <= 10 & (DISTANCE.x[which(DISTANCE.x!=0)]))],by=.(unique_ID)]

# eliminate zeros
ten_m_dist_holder <- ten_m_dist_holder[ten_m_dist_holder[, DISTANCE.x] != 0, ]


# distance to all trees less than 10m away
ten_m_all_dist_holder <- holder_new.dt[,.(TARGET, DISTANCE.x),by=.(unique_ID)][, .SD[which(DISTANCE.x <= 10)],by=.(unique_ID)]

ten_m_all_dist_holder <- ten_m_all_dist_holder[ten_m_all_dist_holder[, DISTANCE.x] != 0, ]

```

# Non-host Point Pattern

We're missing GPS points for many non-host in Cox Spring so we'll do Emigration in the next chink instead

```{r}

ppp.cxsp.nh <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window, marks = cxsp$Species_Visit1)

cxsp_points <- cxsp_points %>% dplyr::select(-c("Elevation", "distance"))
cxsp_points <- cxsp_points[-10,] # get rid of host trees
cxsp_points <- cxsp_points[-10,]
cxsp_points <- cxsp_points[1:255,] # just keep non-host for combination
names(cxsp_points)[2] <- "Latitude.focal"
names(cxsp_points)[3] <- "Longitude.focal"

# change coordinates to UTM

xy.nh = data.frame(cxsp_points$Longitude.focal, cxsp_points$Latitude.focal)
coordinates(xy.nh) <- c("cxsp_points.Longitude.focal", "cxsp_points.Latitude.focal")
proj4string(xy.nh) <- CRS("+proj=longlat +datum=WGS84")
NE <- spTransform(xy.nh, CRS("+proj=utm +zone=12 ellps=WGS84"))
NE <- as.data.frame(NE)

cxsp_points$Longitude.focal <- NE$coords.x1
cxsp_points$Latitude.focal <- NE$coords.x2

cxsp_points.nh <- cxsp_points

# Append non-host data
# Get all site data
cxsp.nh.data <- combined %>% dplyr::filter(combined$`Site ID` == "CXSP")
# filter for non-host
cxsp.nh.data <- cxsp.nh.data %>% dplyr::filter(Species == "POTR"| Species == "ACGR" | Species == "JUSP" | Species == "PSMEG")
cxsp.nh.data <- cxsp.nh.data %>% dplyr::filter(Visit == 1)

#change case of species for combination
cxsp.nh.data$Species <- cxsp.nh.data$Species %>% str_to_lower()
cxsp.nh.data$Name <- paste(cxsp.nh.data$Species, cxsp.nh.data$`Tag #`, sep = "")

cxsp.nh.combined <- inner_join(cxsp.nh.data, cxsp_points.nh, by = "Name")



# Get host frame ready for combo
names(cxsp)[2] <- "tag"
names(cxsp)[3] <- "Species"

```

Since we missed a bunch of non-host in Cox Spring, I'm gonna quickly check other sites
```{r}
emig_points <- read_xlsx("emig_points.xlsx")
emig.nh.data <- combined %>% dplyr::filter(combined$`Site ID` == "EMIG")
unique(emig.nh.data$Species)
# filter for non-host
emig.nh.data <- emig.nh.data %>% dplyr::filter(Species == "POTR"| Species == "PICO" | Species == "PSMEG")
emig.nh.data <- emig.nh.data %>% dplyr::filter(Visit == 1)
emig_points <- emig_points[1:150,]
# only missing one which is pretty good - fixed in combined_cleaned data and original site data
```

# EMIG Point pattern with non-host

```{r}

emig_points <- emig_points %>% dplyr::select(-c("Elevation", "distance"))
names(emig_points)[2] <- "Latitude.focal"
names(emig_points)[3] <- "Longitude.focal"

# change coordinates to UTM

#xy.nh = data.frame(emig_points$Longitude.focal, emig_points$Latitude.focal)
#coordinates(xy.nh) <- c("emig_points.Longitude.focal", "emig_points.Latitude.focal")
#proj4string(xy.nh) <- CRS("+proj=longlat +datum=WGS84")
#NE <- spTransform(xy.nh, CRS("+proj=utm +zone=12 ellps=WGS84"))
#NE <- as.data.frame(NE)

#emig_points$Longitude.focal <- NE$coords.x1
#emig_points$Latitude.focal <- NE$coords.x2

emig_points.nh <- emig_points
# Append non-host data

#change case of species for combination
emig.nh.data$Species <- emig.nh.data$Species %>% str_to_lower()
emig.nh.data$Name <- paste(emig.nh.data$Species, emig.nh.data$`Tag #`, sep = "")

emig.nh.combined <- inner_join(emig.nh.data, emig_points.nh, by = "Name")

# Error fixes
emig_points.nh <- emig_points.nh[-which(emig_points.nh$Name == "pico24"),]# Get rid of Pico24, it was deleted, but point was taken
emig_points.nh <- emig_points.nh[-which(emig_points.nh$Name == "potr37"),]
emig_points.nh <- emig_points.nh[-which(emig_points.nh$Name == "potr38"),]
emig_points.nh <- emig_points.nh[-which(emig_points.nh$Name == "potr39"),]
# These trees were included with others

# so we're missing points for POTR 45 and 57, which are under 2in totaling 7 trees, I think it's okay to leave out
emig.nh.combined <- inner_join(emig.nh.data, emig_points.nh, by = "Name")

# Now we need to combine this data with host data
emig.h <- trees %>% dplyr::filter(trees$`Site ID_Visit1` == "EMIG" & trees$Species_Visit1 == "ABLA")

# We need to mess with non-host dataframe to be able to combine
names(emig.nh.combined) <- paste(names(emig.nh.combined), "_Visit1", sep = "")
names(emig.nh.combined)[27] <- "tag_ID"
emig.nh.combined$tag_ID <- str_to_upper(emig.nh.combined$tag_ID)
names(emig.nh.combined)[28] <- "Latitude.focal"
names(emig.nh.combined)[29] <- "Longitude.focal"

str(emig.nh.combined)
str(emig.h)
emig.nh.combined$`Inf Location_Visit1` <- as.numeric(emig.nh.combined$`Inf Location_Visit1`)
emig.nh.combined <- emig.nh.combined %>% mutate_at(vars(NL_Visit1, NM_Visit1, NH_Visit1, SL_Visit1, SM_Visit1, SH_Visit1), as.numeric)

# combine
emig.all <- bind_rows(emig.nh.combined, emig.h, .id = "source")

# change classes for ArcGIS compatibility
emig.all$`Inf Location_Visit1` <- as.numeric(emig.all$`Inf Location_Visit1`)
emig.all$infestation_sum_Visit1_focal <- as.numeric(emig.all$infestation_sum_Visit1_focal)
emig.all$DISTANCE <- as.numeric(emig.all$DISTANCE)
# replace NA's for ArcGIS
emig.all$`Inf Location_Visit1` <- emig.all$`Inf Location_Visit1` %>% replace_na(0)
emig.all$infestation_sum_Visit1_focal <- emig.all$infestation_sum_Visit1_focal %>% replace_na(0)
emig.all$DBH_Visit1 <- emig.all$DBH_Visit1 %>% replace_na(0)

write_csv(emig.all, "emig_host_nonhost_with_points_WSG.csv")

# convert coordinates to UTM
xy = data.frame(emig.all$Longitude.focal, emig.all$Latitude.focal)
coordinates(xy) <- c("emig.all.Longitude.focal", "emig.all.Latitude.focal")
proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")
NE <- spTransform(xy, CRS("+proj=utm +zone=12 ellps=WGS84"))
NE <- as.data.frame(NE)

emig.all$Longitude.focal <- NE$coords.x1
emig.all$Latitude.focal <- NE$coords.x2

# Make Species a Factor
emig.all$Species_Visit1 <- as.factor(emig.all$Species_Visit1)

# make ppp object

ppp.window.emig.all <- owin(xrange = c(min(emig.all$Latitude.focal), max(emig.all$Latitude.focal)), yrange = c(min(emig.all$Longitude.focal), max(emig.all$Longitude.focal)))

ppp.emig.nh <- ppp(emig.all$Latitude.focal, emig.all$Longitude.focal, window = ppp.window.emig.all, marks = emig.all$Species_Visit1)

plot(split(ppp.emig.nh))


```

# Gridded Data

```{r}
# Following data table made in ArcGIS Pro
grid_data <- read_xlsx("emigration_grid_data.xlsx")
str(grid_data)
grid_data$TARGET_FID <- as.factor(grid_data$TARGET_FID)
prev_intensity_table <- grid_data %>% dplyr::group_by(TARGET_FID) %>%
                    dplyr::summarize(prevalence = sum(Inf_Location_Visit1)/n() * 100,
                               intensity = mean(infestation_sum_Visit1_focal))

ggplot(prev_intensity_table, aes(x = prevalence, y = intensity)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Percent infestation by tree count", y = "Infestation mean", title = "Infestation percent by tree count vs infestation mean for grid cells")


```