---
title: "Neighborhood Effect"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
knitr::opts_knit$set(root.dir = "~/Project Files/Analysis/Cleaning and Compilation/Cleaning")
```

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

# Clean out environment

```{r}
rm(list = ls())
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(lme4) # to use glmer
library(data.table)
library(glmmTMB)
library(DHARMa)
library(MuMIn)
library(spectralGP)
library(mgcv)
library(broom)
library(broom.helpers)
library(broom.mixed)
library(dotwhisker)
```

# Load data

```{r}

data <- read_xlsx("all_merged_all_spp.xlsx")
#df <- read_xlsx("under20_pairwise_distances_summary_live_crown.xlsx")
#df <- read_xlsx("under8_pairwise_distances_summary_live_crown_new.xlsx") - this has mistakes for emigration

## Final data used in modelling - you can import and skip the first few chunks that make this

#data <- read_xlsx("modelling_1_16_data.xlsx")

```



```{r}

# So we need to convert our decimel degrees to UTM for calculation
all_merged <- data

all_merged$host_non <- ifelse(all_merged$species == "ABLA" | all_merged$species == "ABCO",1,0)
all_merged$infested_v1[is.na(all_merged$infested_v1)] <- 0
all_merged$infested_v2[is.na(all_merged$infested_v2)] <- 0
all_merged$infested_v3[is.na(all_merged$infested_v3)] <- 0
all_merged$infested_v4[is.na(all_merged$infested_v4)] <- 0

unique(all_merged$infested_v4)

all_merged <- all_merged[!is.na(all_merged$Latitude),] # We lose 1.8% of data here unfortunately. We don't have coords for those trees

# Remove TMPF outlier

all_merged <- filter(all_merged, tagID != "TMPFABLA224")

# Convert to sf
all_merged_sf <- st_as_sf(all_merged, coords = c("Longitude", "Latitude"), crs = 4326)

# Define the target UTM zone
target_crs <- 32612  # EPSG code for UTM zone 12

# Transform the coordinates
all_merged_utm <- st_transform(all_merged_sf, crs = target_crs)
all_coords <- st_coordinates(all_merged_utm)
all_coords <- as.data.frame(all_coords)

# Put back into original datframe


all_merged_utm$Longitude <- all_coords$X
all_merged_utm$Latitude <- all_coords$Y


all_merged_utm$unique_ID <- paste(all_merged_utm$unique_ID, rownames(all_merged_utm), sep = "_")

all_infested <- all_merged_utm[all_merged_utm$infested_v1 == 1, ] # Infested trees with UTM coordinates


#write_clip(all_merged_utm) # copy dataframe without making file
#all_merged <- all_merged_utm[!duplicated(all_merged_utm$unique_ID), ] #- there shouldn't be duplicates now


row.names(all_merged_utm) <- all_merged_utm$unique_ID

all_merged_utm <- as.data.frame(all_merged_utm)

# change crown height NA's to zero
all_merged_utm$canopy_ht[is.na(all_merged_utm$canopy_ht)] <- 0

# Infestation sum because its not here for some reason
all_merged_utm$infestation_sum_v1 <- all_merged_utm$NL_v1 + all_merged_utm$NM_v1 + all_merged_utm$NH_v1 + all_merged_utm$SL_v1 + all_merged_utm$SM_v1 + all_merged_utm$SH_v1

all_merged_utm$infestation_sum_v2 <- all_merged_utm$NL_v2 + all_merged_utm$NM_v2 + all_merged_utm$NH_v2 + all_merged_utm$SL_v2 + all_merged_utm$SM_v2 + all_merged_utm$SH_v2

all_merged_utm$infestation_sum_v3 <- all_merged_utm$NL_v3 + all_merged_utm$NM_v3 + all_merged_utm$NH_v3 + all_merged_utm$SL_v3 + all_merged_utm$SM_v3 + all_merged_utm$SH_v3

all_merged_utm$infestation_sum_v4 <- all_merged_utm$NL_v4 + all_merged_utm$NM_v4 + all_merged_utm$NH_v4 + all_merged_utm$SL_v4 + all_merged_utm$SM_v4 + all_merged_utm$SH_v4

# New way to do this
all_merged_utm$dbh <- all_merged_utm$dbh * 2.54 # convert DBH to cm
all_merged_utm <- mutate(all_merged_utm, BA = (0.00007854 * (dbh ^ 2)))

# Add a percent live crown measurement

all_merged_utm$perc_live_crown <- ((all_merged_utm$height - all_merged_utm$canopy_ht) / all_merged_utm$height) * 100

```

```{r}

rm(Holder)
Holder = as.data.frame(matrix(nrow=0,ncol=4)) 
colnames(Holder) =c("PLOT", "FROM", "TARGET", "DISTANCE")

#Assign rownames as Tree-ID, dataset also needs UTM_X, UTM_Y, and PLOT
for( i in levels(as.factor(all_merged_utm$site))){
  subset_frame = subset(all_merged_utm, site == i)
TEMP_DIST = vegdist(subset_frame[,c("Longitude","Latitude")], method = "euclid")
TEMP_DIST.df = as.data.frame(as.matrix(TEMP_DIST))
rownames(TEMP_DIST.df) = (subset_frame$unique_ID)
colnames(TEMP_DIST.df) = (subset_frame$unique_ID)
TEMP_DIST.df$`Tag #` = rownames(TEMP_DIST.df)
TEMP_DIST_MELT = reshape2::melt(TEMP_DIST.df, id.vars = c("Tag #"))
colnames(TEMP_DIST_MELT) = c("FROM", "TARGET", "DISTANCE")
#Make a new column with plot ID
TEMP_DIST_MELT$PLOT = i
Holder = rbind(Holder, TEMP_DIST_MELT) #This is your exit
}

#Merge with infestation data
names(Holder)[1] <- "unique_ID"

Holder_new <- merge(Holder, all_merged_utm, by.x = "TARGET", by.y = "unique_ID")
holder_new.dt <- as.data.table(Holder_new)
length(unique(holder_new.dt$unique_ID))

# names(holder_new.dt)[85] <- "infested_1"
# names(holder_new.dt)[86] <- "infested_2"
#min_dist_holder <- holder_new.dt[`infested` == 1 & (DISTANCE == min(DISTANCE[(DISTANCE != 0)])) ,.(DISTANCE, TARGET),by=.("unique_ID")]

# Get all trees under xm from focal tree

holder_8 <- holder_new.dt[,.(TARGET, DISTANCE), by=.(unique_ID)][,.SD[which(DISTANCE <= 8 & DISTANCE != 0)], by=.(unique_ID)]

holder_20 <- holder_new.dt[,.(TARGET, DISTANCE), by=.(unique_ID)][,.SD[which(DISTANCE <= 20 & DISTANCE != 0)], by=.(unique_ID)]


# add distances to data
# visit 1 distances
#data_with_nearest_infested_dis <- all_merged_utm %>% inner_join(min_dist_holder, by = "unique_ID")


```

# Function to extract info from target trees

```{r}


# Create a function to extract the variable based on the linked tree ID
extractVariable <- function(TARGET, variable) {
  linkedTreeRow <- all_merged_utm[all_merged_utm$unique_ID == TARGET, ]
  if (nrow(linkedTreeRow) > 0) {
    return(linkedTreeRow[[variable]])
  } else {
    return(NA)  # or any other value to indicate missing data
  }
}

```


# Nearest tree and infested tree calculations

```{r}

#str(all_merged_utm)

# Apply the function to create a new column with the extracted variable - need to change function for each extraction !!



# distance to nearest infested at time 1
 min_dist_holder <- holder_new.dt[infested_v1 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]

 min_dist_holder <- min_dist_holder[!duplicated(min_dist_holder$unique_ID)]
 
 # Distance to nearest infested that's not a clump
 
 min_dist_holder_no_clump <- holder_new.dt[infested_v1 == 1 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_clump <- min_dist_holder_no_clump[!duplicated(min_dist_holder_no_clump$unique_ID)]
 
 # Distance to nearest non-infested and not a clump
 
 min_dist_holder_no_infest <- holder_new.dt[infested_v1 == 0 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_infest <- min_dist_holder_no_infest[!duplicated(min_dist_holder_no_infest$unique_ID)]
 
 min_dist_holder_no_infest <- min_dist_holder_no_infest %>% select(-4)
 
 # Distance to nearest tree
 
 min_dist_holder_tree <- holder_new.dt[,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]

 min_dist_holder_tree <- min_dist_holder_tree[!duplicated(min_dist_holder_tree$unique_ID)]
 
# Get rid of not needed dbh column
min_dist_holder_no_clump <- min_dist_holder_no_clump %>% select(1:3)

 # distance to nearest infested at time 2
 min_dist_holder2 <- holder_new.dt[infested_v2 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]

  min_dist_holder2 <- min_dist_holder2[!duplicated(min_dist_holder2$unique_ID)]
  
# Distance to nearest infested that's not a clump
  
  min_dist_holder_no_clump_v2 <- holder_new.dt[infested_v2 == 1 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_clump_v2 <- min_dist_holder_no_clump_v2[!duplicated(min_dist_holder_no_clump_v2$unique_ID)]
 
 min_dist_holder_no_clump_v2 <- min_dist_holder_no_clump_v2 %>% select(-4)
 
 # Distance to nearest non-infested and not a clump - time 2
 
 min_dist_holder_no_infest_v2 <- holder_new.dt[infested_v2 == 0 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_infest_v2 <- min_dist_holder_no_infest_v2[!duplicated(min_dist_holder_no_infest_v2$unique_ID)]
 
 min_dist_holder_no_infest_v2 <- min_dist_holder_no_infest_v2 %>% select(-4)
 
 # distance to nearest infested at time 3
 min_dist_holder3 <- holder_new.dt[infested_v3 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]

  min_dist_holder3 <- min_dist_holder3[!duplicated(min_dist_holder3$unique_ID)]
  
# Distance to nearest infested that's not a clump
  
  min_dist_holder_no_clump_v3 <- holder_new.dt[infested_v3 == 1 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_clump_v3 <- min_dist_holder_no_clump_v3[!duplicated(min_dist_holder_no_clump_v3$unique_ID)]
 
 min_dist_holder_no_clump_v3 <- min_dist_holder_no_clump_v3 %>% select(-4)
 
  # Distance to nearest non-infested and not a clump - time 3
 
 min_dist_holder_no_infest_v3 <- holder_new.dt[infested_v3 == 0 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_infest_v3 <- min_dist_holder_no_infest_v3[!duplicated(min_dist_holder_no_infest_v3$unique_ID)]
 
 min_dist_holder_no_infest_v3 <- min_dist_holder_no_infest_v3 %>% select(-4)
 
 # distance to nearest infested at time 4
 min_dist_holder4 <- holder_new.dt[infested_v4 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]
 
  min_dist_holder4 <- min_dist_holder4[!duplicated(min_dist_holder4$unique_ID)]
  
  # Distance to nearest infested that's not a clump
  
  min_dist_holder_no_clump_v4 <- holder_new.dt[infested_v4 == 1 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_clump_v4 <- min_dist_holder_no_clump_v4[!duplicated(min_dist_holder_no_clump_v4$unique_ID)]
 
 min_dist_holder_no_clump_v4 <- min_dist_holder_no_clump_v4 %>% select(-4)
 
  # Distance to nearest non-infested and not a clump - time 4
 
 min_dist_holder_no_infest_v4 <- holder_new.dt[infested_v4 == 0 ,.(TARGET, DISTANCE, dbh),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0 & dbh > 2.54)]))],by=.(unique_ID)]

 min_dist_holder_no_infest_v4 <- min_dist_holder_no_infest_v4[!duplicated(min_dist_holder_no_infest_v4$unique_ID)]
 
 min_dist_holder_no_infest_v4 <- min_dist_holder_no_infest_v4 %>% select(-4)


 names(min_dist_holder2)[2] <- "TARGET2"
 names(min_dist_holder2)[3] <- "DISTANCE2"
  names(min_dist_holder3)[2] <- "TARGET3"
 names(min_dist_holder3)[3] <- "DISTANCE3"
  names(min_dist_holder4)[2] <- "TARGET4"
 names(min_dist_holder4)[3] <- "DISTANCE4"
 names(min_dist_holder_no_clump)[2] <- "TARGET_v1_noclump"
  names(min_dist_holder_no_clump)[3] <- "DISTANCE_v1_noclump"
   names(min_dist_holder_no_clump_v2)[2] <- "TARGET_v2_noclump"
  names(min_dist_holder_no_clump_v2)[3] <- "DISTANCE_v2_noclump"
   names(min_dist_holder_no_clump_v3)[2] <- "TARGET_v3_noclump"
  names(min_dist_holder_no_clump_v3)[3] <- "DISTANCE_v3_noclump"
   names(min_dist_holder_no_clump_v4)[2] <- "TARGET_v4_noclump"
  names(min_dist_holder_no_clump_v4)[3] <- "DISTANCE_v4_noclump"
  names(min_dist_holder_no_infest)[2] <- "TARGET_v1_noinfest"
  names(min_dist_holder_no_infest)[3] <- "DISTANCE_v1_noinfest"
   names(min_dist_holder_no_infest_v2)[2] <- "TARGET_v2_noinfest"
  names(min_dist_holder_no_infest_v2)[3] <- "DISTANCE_v2_noinfest"
   names(min_dist_holder_no_infest_v3)[2] <- "TARGET_v3_noinfest"
  names(min_dist_holder_no_infest_v3)[3] <- "DISTANCE_v3_noinfest"
   names(min_dist_holder_no_infest_v4)[2] <- "TARGET_v4_noinfest"
  names(min_dist_holder_no_infest_v4)[3] <- "DISTANCE_v4_noinfest"
  names(min_dist_holder_tree)[2] <- "nearest_tree"
  names(min_dist_holder_tree)[3] <- "dist_nearest_tree"

 # add distances to data
merge_data <- all_merged_utm
 # visit 1 distances
 merge_data <- merge_data %>% inner_join(min_dist_holder, by = "unique_ID")
 merge_data <- merge_data %>% inner_join(min_dist_holder_no_clump, by = "unique_ID")
 merge_data <- merge_data %>% inner_join(min_dist_holder_no_infest, by = "unique_ID")

 # visit 2 distances
 merge_data <- merge_data %>% inner_join(min_dist_holder2, by = "unique_ID")
 merge_data <- merge_data %>% inner_join(min_dist_holder_no_clump_v2, by = "unique_ID")
 merge_data <- merge_data %>% inner_join(min_dist_holder_no_infest_v2, by = "unique_ID")
  
 # visit 3 distances
 merge_data <- merge_data %>% left_join(min_dist_holder3, by = "unique_ID")
 merge_data <- merge_data %>% left_join(min_dist_holder_no_clump_v3, by = "unique_ID")
 merge_data <- merge_data %>% inner_join(min_dist_holder_no_infest_v3, by = "unique_ID")
 
 # visit 4 distances
 merge_data <- merge_data %>% left_join(min_dist_holder4, by = "unique_ID")
 merge_data <- merge_data %>% left_join(min_dist_holder_no_clump_v4, by = "unique_ID")
 merge_data <- merge_data %>% inner_join(min_dist_holder_no_infest_v4, by = "unique_ID")
 
 # nearest tree distance
merge_data <- merge_data %>% inner_join(min_dist_holder_tree, by = "unique_ID")

#df <- holder_20
  # extract_local <- function(df){

# df$infestation_under <- sapply(df$TARGET, extractVariable, variable = "infested_v1")
# df$infestation_2_under <- sapply(df$TARGET, extractVariable, variable = "infested_v2")
# df$infestation_3_under <- sapply(df$TARGET, extractVariable, variable = "infested_v3")
# df$infestation_4_under <- sapply(df$TARGET, extractVariable, variable = "infested_v4")
# df$infestation.sum.1_under <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v1")
# df$infestation.sum.2_under <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v2")
# df$infestation.sum.3_under <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v3")
# df$infestation.sum.4_under <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v4")
# df$dbh_under <- sapply(df$TARGET, extractVariable, variable = "dbh")
# df$brooms_under <- sapply(df$TARGET, extractVariable, variable = "brooms")
# df$gouting_under <- sapply(df$TARGET, extractVariable, variable = "Gouting_v1")
# df$species_under <- sapply(df$TARGET, extractVariable, variable = "species")
# df$BA_under <- sapply(df$TARGET, extractVariable, variable = "BA")
# df$crown_ht__under <- sapply(df$TARGET, extractVariable, variable = "canopy_ht")
# df$dead_alive <- sapply(df$TARGET, extractVariable, variable = "Dead / Alive_v1")
# df$host_non <- sapply(df$TARGET, extractVariable, variable = "host_non")
# df$percent_live_crown <- sapply(df$TARGET, extractVariable, variable = "perc_live_crown")
#df$height_under <- pbsapply(df$TARGET, extractVariable, variable = "height")
#  # }
  
# vectorized solution to speed up df computation
df <- df %>%
  dplyr::left_join(all_merged_utm %>% dplyr::select(unique_ID, height), 
                   by = c("TARGET" = "unique_ID")) %>%
  dplyr::rename(height_under = height)


 #writexl::write_xlsx(df, "under8_pairwise_distances_summary_1_15.xlsx")
 
#df <- read_xlsx("under20_pairwise_distances_summary_live_crown.xlsx")


#writexl::write_xlsx(df, "under20_pairwise_distances_summary_2_7.xlsx")

#df <- read_xlsx("under8_pairwise_distances_summary_live_crown_new.xlsx")

# create a summary table of local stats for each tree

#df <- read_xlsx("under20_pairwise_distances_summary_1_18.xlsx")

# Fill in NA heights before competition index calculation
# Fix NA heights - We're missing some heights, so we'll take the average ratio of dbh-height from each site and apply that to fill in our NA's
hts <- merge_data %>% filter(is.na(height)) # Emig,FRNK,MOCR,MTTP,SOAP

emig_ratio <- merge_data %>% filter(site == "EMIG") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))

frnk_ratio <- merge_data %>% filter(site == "FRNK") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))

mocr_ratio <- merge_data %>% filter(site == "MOCR") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))

mttp_ratio <- merge_data %>% filter(site == "MTTP") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))

soap_ratio <- merge_data %>% filter(site == "SOAP") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))

all_merged_utm$height[all_merged_utm$site=="EMIG" & is.na(all_merged_utm$height)] <-   all_merged_utm$dbh[is.na(all_merged_utm$height) & all_merged_utm$site == "EMIG"] * emig_ratio
all_merged_utm$height <- as.numeric(all_merged_utm$height)
all_merged_utm$height[all_merged_utm$site=="FRNK" & is.na(all_merged_utm$height)] <-   all_merged_utm$dbh[is.na(all_merged_utm$height) & all_merged_utm$site == "FRNK"] * frnk_ratio
all_merged_utm$height <- as.numeric(all_merged_utm$height)
all_merged_utm$height[all_merged_utm$site=="MOCR" & is.na(all_merged_utm$height)] <-   all_merged_utm$dbh[is.na(all_merged_utm$height) & all_merged_utm$site == "MOCR"] * mocr_ratio
all_merged_utm$height <- as.numeric(all_merged_utm$height)
all_merged_utm$height[all_merged_utm$site=="MTTP" & is.na(all_merged_utm$height)] <-   all_merged_utm$dbh[is.na(all_merged_utm$height) & all_merged_utm$site == "MTTP"] * mttp_ratio
all_merged_utm$height <- as.numeric(all_merged_utm$height)
all_merged_utm$height[all_merged_utm$site=="SOAP" & is.na(all_merged_utm$height)] <-   all_merged_utm$dbh[is.na(all_merged_utm$height) & all_merged_utm$site == "SOAP"] * soap_ratio
all_merged_utm$height <- as.numeric(all_merged_utm$height)
all_merged_utm$logheight <- log(all_merged_utm$height + 1)

local_stats <- function(df){
  df %>% 
  dplyr::group_by(unique_ID) %>% 
  dplyr::summarise(n = n(),
                   n.infested_v1 = sum(infestation_under, na.rm = TRUE),
                   n.infested_v2 = sum(infestation_2_under, na.rm = TRUE),
                   n.infested_v3 = sum(infestation_3_under, na.rm = TRUE),
                   n.infested_v4 = sum(infestation_4_under, na.rm = TRUE),
                   perc.infested_v1 = (sum(infestation_under, na.rm = TRUE)/n) * 100,
                   perc.infested_v2 = (sum(infestation_2_under, na.rm = TRUE)/n) * 100,
                   perc.infested_v3 = (sum(infestation_3_under, na.rm = TRUE)/n) * 100,
                   perc.infested_v4 = (sum(infestation_4_under, na.rm = TRUE)/n) * 100,
                   inf.sum_nh_v1 = sum(infestation.sum.1_under, na.rm = TRUE),
                   inf.sum_nh_v2 = sum(infestation.sum.2_under, na.rm = TRUE),
                   inf.sum_nh_v3 = sum(infestation.sum.3_under, na.rm = TRUE),
                   inf.sum_nh_v4 = sum(infestation.sum.4_under, na.rm = TRUE),
                   avg.dbh = mean(dbh_under, na.rm = TRUE),
                   brooms_nh = sum(brooms_under, na.rm = TRUE),
                   avg.gouting = mean(gouting_under, na.rm = TRUE),
                   gouting.sum = sum(gouting_under, na.rm = TRUE),
                   avg_crown_ht = mean(crown_ht__under, na.rm = TRUE),
                   mean_dist_infested_v1 = ifelse(sum(infestation_under, na.rm = TRUE) == 0, NA, mean(DISTANCE[infestation_under == 1], na.rm = TRUE)),
                   mean_dist_infested_v2 = ifelse(sum(infestation_2_under, na.rm = TRUE) == 0, NA, mean(DISTANCE[infestation_2_under == 1], na.rm = TRUE)),
                   mean_dist_infested_v3 = ifelse(sum(infestation_3_under, na.rm = TRUE) == 0, NA, mean(DISTANCE[infestation_3_under == 1], na.rm = TRUE)),
                   mean_dist_infested_v4 = ifelse(sum(infestation_4_under, na.rm = TRUE) == 0, NA, mean(DISTANCE[infestation_4_under == 1], na.rm = TRUE)),
    basal_area_non_host = sum(BA_under[host_non == 0], na.rm = TRUE),  # non-host BA
    basal_area_host = sum(BA_under[host_non == 1], na.rm = TRUE),  # host BA
    basal_area_nh = basal_area_non_host + basal_area_host, # all tree BA
    percent_basal_area_non_host = basal_area_non_host / (basal_area_non_host + basal_area_host) * 100,# percent non-host BA
    mean_percent_live_crown = mean(percent_live_crown, na.rm = TRUE),
    mean_ABLA_ht = mean(height_under[species_under == "ABLA"], na.rm = TRUE),
    comp_index = 100 * (1/dplyr::n_distinct(species_under[species_under != "ABLA"])) * sum((mean_ABLA_ht - height_under[species_under != "ABLA"]) / mean_ABLA_ht, na.rm = TRUE)
    )
}



# Alternate way - Do this out to 20 meters, subset that by distance and then run local metrics

local_20 <- local_stats(df)
#local_19 <- df %>% filter(DISTANCE <= 19) %>% local_stats()
local_18 <- df %>% filter(DISTANCE <= 18) %>% local_stats()
#local_17 <- df %>% filter(DISTANCE <= 17) %>% local_stats()
local_16 <- df %>% filter(DISTANCE <= 16) %>% local_stats()
#local_15 <- df %>% filter(DISTANCE <= 15) %>% local_stats()
local_14 <- df %>% filter(DISTANCE <= 14) %>% local_stats()
#local_13 <- df %>% filter(DISTANCE <= 13) %>% local_stats()
local_12 <- df %>% filter(DISTANCE <= 12) %>% local_stats()
#local_11 <- df %>% filter(DISTANCE <= 11) %>% local_stats()
local_10 <- df %>% filter(DISTANCE <= 10) %>% local_stats()
#local_9 <- df %>% filter(DISTANCE <= 9) %>% local_stats()
local_8 <- df %>% filter(DISTANCE <= 8) %>% local_stats()
#local_7 <- df %>% filter(DISTANCE <= 7) %>% local_stats()
local_6 <- df %>% filter(DISTANCE <= 6) %>% local_stats()
#local_5 <- df %>% filter(DISTANCE <= 5) %>% local_stats()
local_4 <- df %>% filter(DISTANCE <= 4) %>% local_stats()
#local_3 <- df %>% filter(DISTANCE <= 3) %>% local_stats()
local_2 <- df %>% filter(DISTANCE <= 2) %>% local_stats()
local_1 <- df %>% filter(DISTANCE <= 1) %>% local_stats()
#local_.5 <- df %>% filter(DISTANCE <= 0.5) %>% local_stats()

# Doing this to get mean live crown in local stats and closest live infested that's not a clump
# local_8 <- df %>% local_stats()

# writexl::write_xlsx(local_19, "local_19.xlsx")
# writexl::write_xlsx(local_18, "local_18.xlsx")
# writexl::write_xlsx(local_17, "local_17.xlsx")
# writexl::write_xlsx(local_16, "local_16.xlsx")
# writexl::write_xlsx(local_15, "local_15.xlsx")
# writexl::write_xlsx(local_14, "local_14.xlsx")
# writexl::write_xlsx(local_13, "local_13.xlsx")
# writexl::write_xlsx(local_12, "local_12.xlsx")
# writexl::write_xlsx(local_11, "local_11.xlsx")
# writexl::write_xlsx(local_9, "local_9.xlsx")
# writexl::write_xlsx(local_8, "local_8.xlsx")
# writexl::write_xlsx(local_7, "local_7.xlsx")
# writexl::write_xlsx(local_6, "local_6.xlsx")
# writexl::write_xlsx(local_5, "local_5.xlsx")
# writexl::write_xlsx(local_4, "local_4.xlsx")
# writexl::write_xlsx(local_3, "local_3.xlsx")
# writexl::write_xlsx(local_2, "local_2.xlsx")
# writexl::write_xlsx(local_1, "local_1.xlsx")
# writexl::write_xlsx(local_.5, "local_.5.xlsx")
```

# Combine with original data

```{r}

 #data$unique_ID <- paste(data$site, data$unique_ID, sep = "")
# data <- data %>%
#   dplyr::mutate(unique_ID = paste(unique_ID, row_number(), sep = "_"))
# data <- data %>% filter(species == "ABLA" | species == "ABCO")

local_1$unique_ID <- sub("_.*", "", local_1$unique_ID)
local_2$unique_ID <- sub("_.*", "", local_2$unique_ID)
local_4$unique_ID <- sub("_.*", "", local_4$unique_ID)
local_6$unique_ID <- sub("_.*", "", local_6$unique_ID)
local_8$unique_ID <- sub("_.*", "", local_8$unique_ID)
local_10$unique_ID <- sub("_.*", "", local_10$unique_ID)
local_12$unique_ID <- sub("_.*", "", local_12$unique_ID)
local_14$unique_ID <- sub("_.*", "", local_14$unique_ID)
local_16$unique_ID <- sub("_.*", "", local_16$unique_ID)
local_18$unique_ID <- sub("_.*", "", local_18$unique_ID)
local_20$unique_ID <- sub("_.*", "", local_20$unique_ID)

names(merge_data)[94] <- "unique_ID"
names(merge_data)[87] <- "tagID"
merge_data <- merge_data[, -94]
merge_data$unique_ID <- sub("_.*", "", merge_data$unique_ID)# remove row number for combination


 local_data_20 <- merge_data %>% left_join(local_20, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_19 <- data %>% left_join(local_19, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_18 <- merge_data %>% left_join(local_18, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_17 <- data %>% left_join(local_17, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_16 <- merge_data %>% left_join(local_16, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_15 <- data %>% left_join(local_15, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_14 <- merge_data %>% left_join(local_14, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_13 <- data %>% left_join(local_13, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_12 <- merge_data %>% left_join(local_12, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_11 <- data %>% left_join(local_11, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_10 <- merge_data %>% left_join(local_10, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_9 <- data %>% left_join(local_9, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
local_data_8 <- merge_data %>% left_join(local_8,by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
#writexl::write_xlsx(local_data_8, "local_data_8_new.xlsx")
# local_data_7 <- data %>% left_join(local_7, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_6 <- merge_data %>% left_join(local_6, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_5 <- data %>% left_join(local_5, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_4 <- merge_data %>% left_join(local_4, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_3 <- data %>% left_join(local_3, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_2 <- merge_data %>% left_join(local_2, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_1 <- merge_data %>% left_join(local_1, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
# local_data_.5 <- data %>% left_join(local_.5, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")

```

# Run models to get coefficients

We need to transform some variables so we'll do that for everything and then run models, pull out coefficients and then plot them

```{r}

#object_names <- c("local_data_.5", paste0("local_data_", 1:20))

object_names <- c(paste0("local_data_", c(1,2,4,6,8,10,12,14,16,18,20)))

data_list <- mget(object_names)

data_list <- lapply(data_list, function(df) {
  df$dbh[is.na(df$dbh)] <- 2.54/2
  df$Gouting_v1[is.na(df$Gouting_v1)] <- 0
df$infested_v1[is.na(df$infested_v1)] <- 0
df$infested_v2[is.na(df$infested_v2)] <- 0
df$infested_v3[is.na(df$infested_v3)] <- 0
df$infested_v4[is.na(df$infested_v4)] <- 0
df$logdbh <- log(df$dbh)
df$n.infested_v1[is.na(df$n.infested_v1)] <- 0
df$n.infested_v2[is.na(df$n.infested_v2)] <- 0
df$n.infested_v3[is.na(df$n.infested_v3)] <- 0
df$n.infested_v4[is.na(df$n.infested_v4)] <- 0
df$log_number_infested_1 <- log(df$n.infested_v1)
df$log_number_infested_2 <- log(df$n.infested_v2)
df$log_number_infested_3 <- log(df$n.infested_v3)
df$log_number_infested_4 <- log(df$n.infested_v4)
names(df)[104] <- "TARGET_v1"
names(df)[105] <- "DISTANCE_v1"
names(df)[106] <- "TARGET_noclump_v1"
names(df)[107] <- "DISTANCE_noclump_v1"
names(df)[108] <- "TARGET_noinfest_v1"
names(df)[109] <- "DISTANCE_noinfest_v1"
names(df)[110] <- "TARGET_v2"
names(df)[111] <- "DISTANCE_v2"
names(df)[112] <- "TARGET_noclump_v2"
names(df)[113] <- "DISTANCE_noclump_v2"
names(df)[114] <- "TARGET_noinfest_v2"
names(df)[115] <- "DISTANCE_noinfest_v2"
names(df)[116] <- "TARGET_v3"
names(df)[117] <- "DISTANCE_v3"
names(df)[118] <- "TARGET_noclump_v3"
names(df)[119] <- "DISTANCE_noclump_v3"
names(df)[120] <- "TARGET_noinfest_v3"
names(df)[121] <- "DISTANCE_noinfest_v3"
names(df)[122] <- "TARGET_v4"
names(df)[123] <- "DISTANCE_v4"
names(df)[124] <- "TARGET_noclump_v4"
names(df)[125] <- "DISTANCE_noclump_v4"
names(df)[126] <- "TARGET_noinfest_v4"
names(df)[127] <- "DISTANCE_noinfest_v4"
df$logDISTANCE_v1 <- log(df$DISTANCE_v1)
df$logDISTANCE_v2 <- log(df$DISTANCE_v2)
df$logDISTANCE_v3 <- log(df$DISTANCE_v3)
df$logDISTANCE_v4 <- log(df$DISTANCE_v4)
df$avg.gouting[is.nan(df$avg.gouting)] <- 0
df$Gouting_v1 <- as.factor(df$Gouting_v1)
df$Gouting_v1[is.na(df$Gouting_v1)] <- 0
unique(df$Gouting_v1)
df$Gouting_v2 <- as.factor(df$Gouting_v2)
df$Gouting_v2[is.na(df$Gouting_v2)] <- 0
unique(df$Gouting_v2)
df$Gouting_v3 <- as.factor(df$Gouting_v3)
df$Gouting_v3[is.na(df$Gouting_v3)] <- 0
unique(df$Gouting_v3)
df$Gouting_v4 <- as.factor(df$Gouting_v4)
df$Gouting_v4[is.na(df$Gouting_v4)] <- 0
df$gouting_recode_v1 <- ifelse(df$Gouting_v1 == 0, 0, 1)
df$gouting_recode_v2 <- ifelse(df$Gouting_v2 == 0, 0, 1)
df$gouting_recode_v3 <- ifelse(df$Gouting_v3 == 0, 0, 1)
df$gouting_recode_v4 <- ifelse(df$Gouting_v4 == 0, 0, 1)

# Standardize variables
df$infest_sum_std_v1 <- (df$inf.sum_nh_v1 - mean(df$inf.sum_nh_v1, na.rm = TRUE))/sd(df$inf.sum_nh_v1, na.rm = TRUE)
df$infest_sum_std_v2 <- (df$inf.sum_nh_v2 - mean(df$inf.sum_nh_v2, na.rm = TRUE))/sd(df$inf.sum_nh_v2, na.rm = TRUE)
df$infest_sum_std_v3 <- (df$inf.sum_nh_v3 - mean(df$inf.sum_nh_v3, na.rm = TRUE))/sd(df$inf.sum_nh_v3, na.rm = TRUE)
df$infest_sum_std_v4 <- (df$inf.sum_nh_v4 - mean(df$inf.sum_nh_v4, na.rm = TRUE))/sd(df$inf.sum_nh_v4, na.rm = TRUE)

df$basal_area_host_std <- (df$basal_area_host - mean(df$basal_area_host, na.rm = TRUE))/sd(df$basal_area_host, na.rm = TRUE)
df$basal_area_nh_std <- (df$basal_area_nh - mean(df$basal_area_nh, na.rm = TRUE))/sd(df$basal_area_nh, na.rm = TRUE)

df$brooms_recode <- ifelse(df$brooms > 0, 0, 1)

df$mean_percent_live_crown_std <- (df$mean_percent_live_crown - mean(df$mean_percent_live_crown))/sd(df$mean_percent_live_crown)

df$new_infested_v2 <- ifelse(df$infested_v1 == 0 & df$infested_v2 == 1, 1, 0)
df$new_infested_v3 <- ifelse(df$infested_v1 == 0 & df$infested_v2 == 0 & df$infested_v3 == 1, 1, 0)
df$new_infested_v4 <- ifelse(df$infested_v1 == 0 & df$infested_v2 == 0 & df$infested_v3 == 0 & df$infested_v4 == 1, 1, 0)

df$Beetle_v1[is.na(df$Beetle_v1)] <- 0
df$Beetle_v1 <- ifelse(grepl("[a-zA-Z]", df$Beetle_v1), 1, 0)
df$Beetle_v2[is.na(df$Beetle_v2)] <- 0
df$Beetle_v2 <- ifelse(grepl("[a-zA-Z]", df$Beetle_v2), 1, 0)
df$Beetle_v3[is.na(df$Beetle_v3)] <- 0
df$Beetle_v3 <- ifelse(grepl("[a-zA-Z]", df$Beetle_v3), 1, 0)
df$Beetle_v4[is.na(df$Beetle_v4)] <- 0
df$Beetle_v4 <- ifelse(grepl("[a-zA-Z]", df$Beetle_v4), 1, 0)

df$beetle_recode <- ifelse(rowSums(df[, c("Beetle_v1", "Beetle_v2", "Beetle_v3", "Beetle_v4")] != 0) > 0, 1, 0)

df$mean_dist_infested_v1[is.na(df$mean_dist_infested_v1)] <- 0
df$mean_dist_infested_recode_v1 <- ifelse(df$mean_dist_infested_v1 == 0, 11, df$mean_dist_infested_v1)

df$mean_dist_infested_v2[is.na(df$mean_dist_infested_v2)] <- 0
df$mean_dist_infested_recode_v2 <- ifelse(df$mean_dist_infested_v2 == 0, 11, df$mean_dist_infested_v2)

df$mean_dist_infested_v3[is.na(df$mean_dist_infested_v3)] <- 0
df$mean_dist_infested_recode_v3 <- ifelse(df$mean_dist_infested_v3 == 0, 11, df$mean_dist_infested_v3)

df$mean_dist_infested_v4[is.na(df$mean_dist_infested_v4)] <- 0
df$mean_dist_infested_recode_v4 <- ifelse(df$mean_dist_infested_v4 == 0, 11, df$mean_dist_infested_v4)

df %>% mutate(new_infestation = ifelse(infested_v1 == 0 & (infested_v2 == 1 | infested_v3 == 1 | infested_v4 == 1), 1, 0))

df$infest_sum_std_new_v1 <- df$infest_sum_std_v1 + 2 # Get rid of negatives and zeros
df$infest_sum_std_new_v2 <- df$infest_sum_std_v2 + 2
df$infest_sum_std_new_v3 <- df$infest_sum_std_v3 + 2
df$infest_sum_std_new_v4 <- df$infest_sum_std_v4 + 2
# df$infestation.sum_under10_new <- df$infestation.sum_under10 + 1 # try with non-standardized variable
  return(df)
})

list2env(data_list, .GlobalEnv)


# cor(data[, c("logdbh", "DISTANCE", "height", "Gouting_v1", "crown_height", "brooms_recode", "infestation.sum_under10", "distance_to_nearest_recode", "DISTANCE_nearest_tree", "basal_area_host_std", "percent.infested_under10", "number.infested_under10", "average_crown_ht_under_10", "average.dbh_under10", "qmd_under10")], use = "complete.obs")

```

Model to run

```{r}
data_list <- mget(object_names)

glmm_results <- lapply(data_list, function(df) {
  glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE + logdbh + Gouting_v1 + infest_sum_std_new + (1|unique_ID) + (1|site), data = df, family = binomial(link = "logit"))
})

coefficients_list <- lapply(glmm_results, coef)


dbh_coefficients <- sapply(glmm_results, function(model) {
    return(fixef(model)$cond["logdbh"])  # Extract the fixed-effect coefficient for "dbh"
})

distance_coefficients <- sapply(glmm_results, function(model) {
    return(fixef(model)$cond["logDISTANCE"])  # Extract the fixed-effect coefficient for "dbh"
})

gouting_coefficients <- sapply(glmm_results, function(model) {
    return(fixef(model)$cond["Gouting_v1"])  # Extract the fixed-effect coefficient for "dbh"
})

infest_sum_coefficients <- sapply(glmm_results, function(model) {
    return(fixef(model)$cond["infest_sum_std_new"])  # Extract the fixed-effect coefficient for "dbh"
})

# Combine into a data frame
dbh_coeff_data <- data.frame(
  Distance = c(0.5,1:20),
  DBH_Coefficient = dbh_coefficients)

distance_coeff_data <- data.frame(
  Distance = c(0.5,1:20),
  Distance_Coefficient = distance_coefficients)

gouting_coeff_data <- data.frame(
  Distance = c(0.5,1:20),
  Gouting_Coefficient = gouting_coefficients)

infest_sum_coeff_data <- data.frame(
  Distance = c(0.5,1:20),
  Infestation_Sum_Coefficient = infest_sum_coefficients)


dbh_coefficeient_plot <- ggplot(dbh_coeff_data, aes(x = Distance, y = DBH_Coefficient)) +
  geom_line(color = "blue", size = 1) +  # Line connecting points
  geom_point(color = "red", size = 2) +  # Add points for clarity
  labs(
    title = "Coefficient of DBH vs. Distance",
    x = "Distance",
    y = "DBH Coefficient"
  ) +
  theme_minimal()

distance_coefficient_plot <- ggplot(distance_coeff_data, aes(x = Distance, y = Distance_Coefficient)) +
  geom_line(color = "blue", size = 1) +  # Line connecting points
  geom_point(color = "red", size = 2) +  # Add points for clarity
  labs(
    title = "Coefficient of distance to nearest infested vs. Distance",
    x = "Distance",
    y = "Distance to nearest infested Coefficient"
  ) +
  theme_minimal()

gouting_coefficient_plot <- ggplot(gouting_coeff_data, aes(x = Distance, y = Gouting_Coefficient)) +
  geom_line(color = "blue", size = 1) +  # Line connecting points
  geom_point(color = "red", size = 2) +  # Add points for clarity
  labs(
    title = "Coefficient of Gouting vs. Distance",
    x = "Distance",
    y = "Gouting Coefficient"
  ) +
  theme_minimal()

infest_sum_coefficient_plot <- ggplot(infest_sum_coeff_data, aes(x = Distance, y = Infestation_Sum_Coefficient)) +
  geom_line(color = "blue", size = 1) +  # Line connecting points
  geom_point(color = "red", size = 2) +  # Add points for clarity
  labs(
    title = "Coefficient of Infestation Sum vs. Distance",
    x = "Distance",
    y = "Infestation Sum Coefficient"
  ) +
  theme_minimal()

print(dbh_coefficeient_plot)
print(gouting_coefficient_plot)
print(infest_sum_coefficient_plot)
print(distance_coefficient_plot)
```

# Question 1: Who's infested??

We have this answer in our GLM - tree-level markdown. We copy that model into here with an 8m neighborhood and then do:

Question 2: Are the risk factors the same for different sized trees?

- Looks like we should use a neighborhood of 8m to pick up the largest effects of our coefficients

```{r}

## data manipulation chunk



# Standardize infestation sum and take care of NA's
# local_data_8$dbh[is.na(local_data_8$dbh)] <- 2.54/2
# local_data_8$height <- local_data_8$height * 0.3048
# local_data_8$Gouting_v1[is.na(local_data_8$Gouting_v1)] <- 0
# local_data_8$infest_sum_std_v1 <- (local_data_8$inf.sum_nh_v1 - mean(local_data_8$inf.sum_nh_v1))/sd(local_data_8$inf.sum_nh_v1)
# local_data_8$infest_sum_std_v2 <- (local_data_8$inf.sum_nh_v2 - mean(local_data_8$inf.sum_nh_v2))/sd(local_data_8$inf.sum_nh_v2)
# local_data_8$infest_sum_std_v3 <- (local_data_8$inf.sum_nh_v3 - mean(local_data_8$inf.sum_nh_v3))/sd(local_data_8$inf.sum_nh_v3)
# local_data_8$infest_sum_std_v4 <- (local_data_8$inf.sum_nh_v4 - mean(local_data_8$inf.sum_nh_v4))/sd(local_data_8$inf.sum_nh_v4)
# local_data_8$basal_area_host_std <- (local_data_8$basal_area_host - mean(local_data_8$basal_area_host))/sd(local_data_8$basal_area_host)
# local_data_8$mean_percent_live_crown_std <- (local_data_8$mean_percent_live_crown - mean(local_data_8$mean_percent_live_crown))/sd(local_data_8$mean_percent_live_crown)


# Fix beetles before modelling
# data$Beetle_v1[is.na(data$Beetle_v1)] <- 0
# data$Beetle_v1 <- ifelse(grepl("[a-zA-Z]", data$Beetle_v1), 1, 0)
# data$Beetle_v2[is.na(data$Beetle_v2)] <- 0
# data$Beetle_v2 <- ifelse(grepl("[a-zA-Z]", data$Beetle_v2), 1, 0)
# data$Beetle_v3[is.na(data$Beetle_v3)] <- 0
# data$Beetle_v3 <- ifelse(grepl("[a-zA-Z]", data$Beetle_v3), 1, 0)
# data$Beetle_v4[is.na(data$Beetle_v4)] <- 0
# data$Beetle_v4 <- ifelse(grepl("[a-zA-Z]", data$Beetle_v4), 1, 0)
# 
# data$beetle_recode <- ifelse(rowSums(data[, c("Beetle_v1", "Beetle_v2", "Beetle_v3", "Beetle_v4")] != 0) > 0, 1, 0)

# Make new infestation variable

# data <- data %>%
#   mutate(new_infestation = ifelse(infested_v1 == 0 & (infested_v2 == 1 | infested_v3 == 1 | infested_v4 == 1), 1, 0))


#data$brooms_recode <- ifelse(data$brooms > 0, 1, 0)

# filter out dead



# log/tweak things
#data$height[data$height == 0] <- 0.1

# data$logdbh <- log(data$dbh)
# data$logDISTANCE_v1 <- log(data$DISTANCE_v1)
# data$logDISTANCE_v2 <- log(data$DISTANCE_v2)
# data$logDISTANCE_v3 <- log(data$DISTANCE_v3)
# data$logDISTANCE_v4 <- log(data$DISTANCE_v4)
# data$Gouting_v1 <- as.factor(data$Gouting_v1)
# data$Gouting_v1[is.na(data$Gouting_v1)] <- 0
# unique(data$Gouting_v1)
# data$Gouting_v2 <- as.factor(data$Gouting_v2)
# data$Gouting_v2[is.na(data$Gouting_v2)] <- 0
# unique(data$Gouting_v2)
# data$Gouting_v3 <- as.factor(data$Gouting_v3)
# data$Gouting_v3[is.na(data$Gouting_v3)] <- 0
# unique(data$Gouting_v3)
# data$Gouting_v4 <- as.factor(data$Gouting_v4)
# data$Gouting_v4[is.na(data$Gouting_v4)] <- 0
# unique(data$Gouting_v4)

# data$mean_dist_infested_v1[is.na(data$mean_dist_infested_v1)] <- 0
# data$mean_dist_infested_recode_v1 <- ifelse(data$mean_dist_infested_v1 == 0, 11, data$mean_dist_infested_v1)
# data$mean_dist_infested_v2[is.na(data$mean_dist_infested_v2)] <- 0
# data$mean_dist_infested_recode_v2 <- ifelse(data$mean_dist_infested_v2 == 0, 11, data$mean_dist_infested_v2)
# data$mean_dist_infested_v3[is.na(data$mean_dist_infested_v3)] <- 0
# data$mean_dist_infested_recode_v3 <- ifelse(data$mean_dist_infested_v3 == 0, 11, data$mean_dist_infested_v3)
# data$mean_dist_infested_v4[is.na(data$mean_dist_infested_v4)] <- 0
# data$mean_dist_infested_recode_v4 <- ifelse(data$mean_dist_infested_v4 == 0, 11, data$mean_dist_infested_v4)
# 
# data$avg.gouting[is.nan(data$avg.gouting)] <- 0 # Was taking mean of 0 to get NaN - replace with zero

# Fix NA heights - We're missing some heights, so we'll take the average ratio of dbh-height from each site and apply that to fill in our NA's
# hts <- data %>% filter(is.na(height)) # Emig,FRNK,MOCR,MTTP,SOAP
# 
# emig_ratio <- data %>% filter(site == "EMIG") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))
# data$height[data$site=="EMIG" & is.na(data$height)] <-   data$dbh[is.na(data$height) & data$site == "EMIG"] * emig_ratio
# 
# data$height <- as.numeric(data$height)
# 
# frnk_ratio <- data %>% filter(site == "FRNK") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))
# data$height[data$site=="FRNK" & is.na(data$height)] <-   data$dbh[is.na(data$height) & data$site == "FRNK"] * frnk_ratio
# 
# data$height <- as.numeric(data$height)
# 
# mocr_ratio <- data %>% filter(site == "MOCR") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))
# data$height[data$site=="MOCR" & is.na(data$height)] <-   data$dbh[is.na(data$height) & data$site == "MOCR"] * mocr_ratio
# 
# data$height <- as.numeric(data$height)
# 
# mttp_ratio <- data %>% filter(site == "MTTP") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))
# data$height[data$site=="MTTP" & is.na(data$height)] <-   data$dbh[is.na(data$height) & data$site == "MTTP"] * mttp_ratio
# 
# data$height <- as.numeric(data$height)
# 
# soap_ratio <- data %>% filter(site == "SOAP") %>% dplyr::summarize(ratios = (height/dbh)) %>% dplyr::summarize(avg_ratio = mean(ratios, na.rm = TRUE))
# data$height[data$site=="SOAP" & is.na(data$height)] <-   data$dbh[is.na(data$height) & data$site == "SOAP"] * soap_ratio
# 
# data$height <- as.numeric(data$height)

# data$logheight <- log(data$height + 1)

data <- local_data_10

data <- data %>% filter(`Dead / Alive_v1`==1)

glmm <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + logdbh + avg.dbh + Gouting_v1 + infest_sum_std_new_v1 + avg.gouting + (1|site), data = data, family = binomial(link = "logit"))

glmm.sims <- simulateResiduals(glmm)
plot(glmm.sims)
summary(glmm)

# Make data long

data_long <- data %>%
  pivot_longer(
    cols = ends_with("_v1") | ends_with("_v2") | ends_with("_v3") | ends_with("_v4"),  # columns with visit data
    names_to = c(".value", "visit"),       # use ".value" to keep the original variable name
    names_pattern = "(.*)_(v[0-9])"        # regex to separate variable name and visit
  ) %>%
  mutate(visit = as.numeric(gsub("v", "", visit)))  # Convert "v1", "v2", etc. to 1, 2, etc.

names(data_long)[49] <- "t"

# make logged and lagged variables

data_long <- data_long %>%
  arrange(unique_ID, t) %>%
  group_by(unique_ID) %>%
  mutate(
    logDISTANCE = log(DISTANCE),
    logDISTANCE_noclump = log(DISTANCE_noclump),
    logdbh = log(dbh),
    lag_logDISTANCE = lag(logDISTANCE),
    lag_infest_sum_std_new = lag(infest_sum_std_new),
    lag_infest_sum = lag(infestation_sum),
    lag_Gouting = lag(Gouting),
    lag_infested = lag(infested),
    lag_perc.infested = lag(perc.infested),
    lag_infestation_sum = lag(infestation_sum)
  ) %>%
  ungroup()

data_long$qmd_under <- sqrt(((data_long$basal_area_non_host + data_long$basal_area_host)/(data_long$n*31.83))/0.00007854)

## End of data manipulation

## Load this to skip manipulation steps - this is all live hosts in long format with lagged variables, ready to be subset into last three visits and put into models

#writexl::write_xlsx(data_long, "local_data_8_long_new.xlsx") - 1/13/24 

#data_long <- read_xlsx("local_data_8_long.xlsx")


sm <- subset(data_long, t %in% c(2, 3, 4))

sm$site <- as.factor(sm$site)
sm$unique_ID <- as.factor(sm$unique_ID)

```

# Lagged infestation model

```{r}

time.glmm.withlag <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + lag_logDISTANCE + avg.dbh + (infest_sum_std_new * lag_infest_sum_std_new) + Gouting + logdbh + avg.gouting + (1|unique_ID) + (1|site), data = sm, family = binomial(link = "logit"))

time.glmm.withlag.wosite <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + lag_logDISTANCE + avg.dbh + (infest_sum_std_new * lag_infest_sum_std_new) + Gouting + logdbh + avg.gouting + (1|unique_ID), data = sm, family = binomial(link = "logit"))

time.sims <- simulateResiduals(time.glmm.withlag)
plot(time.sims)

summary(time.glmm.withlag)

# Looking at interactions
time.glmm.lag.interact <- glmmTMB(cbind(infested, 1 - infested) ~ (logDISTANCE * logdbh) + lag_logDISTANCE + dist_nearest_tree + (infest_sum_std_new * lag_infest_sum_std_new) + Gouting + logheight + (1|unique_ID) + (1|site), data = sm, family = binomial(link = "logit"))

time.sims.interact <- simulateResiduals(time.glmm.lag.interact)
plot(time.sims.interact)

summary(time.glmm.lag.interact)

# Checking effect of random effect

summary(time.glmm.withlag.wosite)
anova(time.glmm.withlag, time.glmm.withlag.wosite)

library(sjPlot)
re_plot <- sjPlot::plot_model(time.glmm.withlag, type = "re")

library(ggstats)

coef_plot <- ggcoef_model(time.glmm.withlag)



cor(sm[, c("logheight", "logDISTANCE", "lag_logDISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new", "lag_infest_sum_std_new", "mean_dist_infested", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "n.infested", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])

infest.final <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logdbh + logDISTANCE_v1 + avg.dbh + infest_sum_std_new_v1 + Gouting_v1 + logdbh + avg.gouting + (1|unique_ID) + (1|site), data = data, family = binomial(link = "logit"))

final.sims <- simulateResiduals(infest.final)
plot(final.sims)

summary(infest.final)

# Final lagged model
time.glmm.withlag <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + avg.dbh + gouting_recode + logdbh + avg.dbh + lag_infested + perc.infested + dist_nearest_tree + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = sm, family = binomial(link = "logit"))

time.sims <- simulateResiduals(time.glmm.withlag)
plot(time.sims)

summary(time.glmm.withlag)

tidy_main <- broom::tidy(time.glmm.withlag) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_main$model <- "main" 

library(dotwhisker)
dw_main <- dwplot(tidy_main) +
  scale_y_discrete(
    labels = c("logDISTANCE" = "Distance to nearest infested",
               "lag_infested" = "Lagged infestation status",
               "perc.infested" = "Percent infested (neighborhood)",
                "dist_nearest_tree" = "Distance to nearest tree",
                "gouting_recode" = "Gouting",
                "avg.dbh" = "Average DBH in neighborhood",
                "logdbh" = "DBH",
                "mean_percent_live_crown_std" = "Mean percent live crown (Neighborhood)"),
    limits = c( "logDISTANCE",
                "mean_percent_live_crown_std",
                "avg.dbh",
                "perc.infested",
                "dist_nearest_tree",
                "logdbh",
                "gouting_recode", 
                "lag_infested") # Specify the desired order
  )
print(dw_main+geom_vline(xintercept=0,lty=2))


```

# Full-factorial infestation status model

```{r}
# MAke indicator variable
local_data_1$avg.dbh_rad <- 1
local_data_1$perc.infested_rad <- 1
local_data_2$avg.dbh_rad <- 2
local_data_2$perc.infested_rad <- 2
local_data_4$avg.dbh_rad <- 4
local_data_4$perc.infested_rad <- 4
local_data_6$avg.dbh_rad <- 6
local_data_6$perc.infested_rad <- 6
local_data_8$avg.dbh_rad <- 8
local_data_8$perc.infested_rad <- 8
local_data_10$avg.dbh_rad <- 10
local_data_10$perc.infested_rad <- 10
local_data_12$avg.dbh_rad <- 12
local_data_12$perc.infested_rad <- 12
local_data_14$avg.dbh_rad <- 14
local_data_14$perc.infested_rad <- 14
local_data_16$avg.dbh_rad <- 16
local_data_16$perc.infested_rad <- 16
local_data_18$avg.dbh_rad <- 18
local_data_18$perc.infested_rad <- 18
local_data_20$avg.dbh_rad <- 20
local_data_20$perc.infested_rad <- 20

# combine all data
all_local_data <- local_data_1 %>% bind_rows(local_data_2, local_data_4, local_data_6, local_data_8, local_data_10, local_data_12, local_data_14, local_data_16, local_data_18, local_data_20)

# make long
# Make long versions 
all_local_data <- all_local_data %>% filter(`Dead / Alive_v1`==1)

# Make data long

all_local_long <- all_local_data %>%
  pivot_longer(
    cols = ends_with("_v1") | ends_with("_v2") | ends_with("_v3") | ends_with("_v4"),  # columns with visit data
    names_to = c(".value", "visit"),       # use ".value" to keep the original variable name
    names_pattern = "(.*)_(v[0-9])"        # regex to separate variable name and visit
  ) %>%
  mutate(visit = as.numeric(gsub("v", "", visit)))  # Convert "v1", "v2", etc. to 1, 2, etc.

names(all_local_long)[48] <- "t"

# make logged and lagged variables

all_local_long <- all_local_long %>%
  arrange(unique_ID, t) %>%
  group_by(unique_ID) %>%
  mutate(
    logDISTANCE = log(DISTANCE),
    logDISTANCE_noclump = log(DISTANCE_noclump),
    logdbh = log(dbh),
    lag_logDISTANCE = lag(logDISTANCE),
    lag_infest_sum_std_new = lag(infest_sum_std_new),
    lag_infest_sum = lag(infestation_sum),
    lag_Gouting = lag(Gouting),
    lag_infested = lag(infested),
    lag_perc.infested = lag(perc.infested)
  ) %>%
  ungroup()

all_local_long$qmd_under <- sqrt(((all_local_long$basal_area_non_host + all_local_long$basal_area_host)/(all_local_long$n*31.83))/0.00007854)


all_sm <- subset(all_local_long, t %in% c(2, 3, 4))

all_sm$site <- as.factor(all_sm$site)
all_sm$unique_ID <- as.factor(all_sm$unique_ID)


radii = c(1,2,4,6,8,10,12,14,16,18,20)
combos = as.data.frame(t(combn(radii,2)))

output_list <- vector("list", nrow(combos))

for (i in 1:nrow(combos)){
     model_data = all_sm %>% 
          dplyr::filter(avg.dbh_rad == combos$V1[i] | perc.infested_rad == combos$V2[i])
     
     model_data <- model_data %>%
    dplyr::mutate(
      avg_dbh_current = ifelse(avg.dbh_rad == combos$V1[i], avg.dbh, NA),
      perc_infested_current = ifelse(perc.infested_rad == combos$V2[i], perc.infested, NA)
    )
     
     model <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + logdbh + avg_dbh_current + lag_infested + perc_infested_current + dist_nearest_tree, data = model_data, family = binomial(link = "logit"))
     
     coeffs = model$coeffs
     aic = AICc(model)
     
     output_list[i] = list(avg.dbh_rad = combos$V1[i], 
                           perc.infested_rad = combos$V2[i],
                           coeffs = coeffs,
                           aic = aic)
}


```

# Visualize predictor variables and effect on infestation probability

- This is just for infestation sum in neighborhood, but we could do these plots for all if needed

```{r}
# Specify your independent variable of interest and typical values for others
independent_var <- "infest_sum_std_new"  # Replace with your variable of interest
other_vars <- c("logDISTANCE", "lag_logDISTANCE", "lag_infest_sum_std_new", "Gouting", "lag_Gouting", "logdbh")  # Your other predictors

# Create a sequence for the variable of interest
new_data <- data.frame(
  infest_sum_std_new = seq(min(sm[[independent_var]]), max(sm[[independent_var]]), length.out = 100)
)

# Add other predictors with their typical values
for (var in other_vars) {
  new_data[[var]] <- mean(sm[[var]], na.rm = TRUE)  # Replace with median or other value if preferred
}

# Predict probabilities
new_data$predicted_y <- predict(time.glmm.withlag, newdata = new_data, type = "response", re.form = NA)

# Plot
plot1 <- ggplot() +
  geom_jitter(data = sm, aes(x = infest_sum_std_new, y = infested), height = 0.05, width = 0, alpha = 0.5, color = "red") +
  geom_line(data = new_data, aes(x = infest_sum_std_new, y = predicted_y), size = 1) +
  labs(
    x = "Infestation sum in the neighborhood (8m)",
    y = "Probability",
    title = "Observed Binary Response vs. Predicted Probabilities"
  ) +
  theme_minimal()

print(plot1)

```

# Clump Model!

We'll do current and lagged variables for our last three visits

List of variables:
- Gouting
- brooms
- beetle
- infest sum in neighborhood
- mean distance to infested
- number infested in nh
- percent infested in nh
- number of trees in nh (competition)
- lagged variables after we subset just times with lagged variables available
- mean percent live crown

## Build model

```{r}

clumps <- filter(data, dbh == 1.27)

cor(clumps[, c("logheight", "DISTANCE_v1", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new_v1", "dist_nearest_tree", "basal_area_host_std", "perc.infested_v1", "avg.dbh", "avg.gouting", "mean_percent_live_crown", "n")])

## Building model below

 clump.base <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.base.sims <- simulateResiduals(clump.base)
plot(clump.base.sims)

summary(clump.base)

# Add brooms

clump.broom <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + brooms_recode + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.broom.sims <- simulateResiduals(clump.broom)
plot(clump.broom.sims)

summary(clump.broom)

clump.broom.nh <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + brooms_nh + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.broom.sims <- simulateResiduals(clump.broom.nh)
plot(clump.broom.sims)

summary(clump.broom.nh)

# Add beetle

clump.beetle <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + beetle_recode + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.beetle.sims <- simulateResiduals(clump.beetle)
plot(clump.beetle.sims)

summary(clump.beetle)

# Add infest sum

clump.sum <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.sum.sims <- simulateResiduals(clump.sum)
plot(clump.sum.sims)

summary(clump.sum)

# percent infested

clump.pinfest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + perc.infested_v1 + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.pinfest.sims <- simulateResiduals(clump.pinfest)
plot(clump.pinfest.sims)

summary(clump.pinfest)

# distance to nearest tree

clump.dist.nearinfest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + dist_nearest_tree + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.dist.nearinfest.sims <- simulateResiduals(clump.dist.nearinfest)
plot(clump.dist.nearinfest.sims)

summary(clump.dist.nearinfest)

# avg. dbh in neighborhood

clump.dbh.infest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + avg.dbh + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.dbh.infest.sims <- simulateResiduals(clump.dbh.infest)
plot(clump.dbh.infest.sims)

summary(clump.dbh.infest)

# mean percent live crown
clump.pcrown <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + avg.dbh + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.pcrown.sims <- simulateResiduals(clump.pcrown)
plot(clump.pcrown.sims)

summary(clump.pcrown)


# Building model above
clump.final <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + avg.dbh + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = clumps, family = binomial(link = "logit"))

clump.final.sims <- simulateResiduals(clump.final)
plot(clump.final.sims)

summary(clump.final)

# lagged variables - avg. dbh drops

clump.sm <- sm %>% filter(dbh == 1.27)

cor(clump.sm[, c("logDISTANCE", "lag_logDISTANCE", "infest_sum_std_new","lag_infested", "lag_perc.infested",  "lag_infest_sum_std_new", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "avg.dbh", "avg.gouting", "mean_percent_live_crown", "n")])

clump.lag <- glmmTMB(cbind(infested, 1 - infested) ~ lag_infested + perc.infested + gouting_recode + perc.infested + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = clump.sm, family = binomial(link = "logit"))

clump.lag.sims <- simulateResiduals(clump.lag)
plot(clump.lag.sims)

summary(clump.lag)

testZeroInflation(clump.lag.sims)

# adding lag infestation status drops logDISTANCE and avg dbh


# Long dataset

clump_long <- data_long %>% filter(dbh == 1.27)

clump.all <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.all.sims <- simulateResiduals(clump.all)
plot(clump.all.sims)

summary(clump.all)

```


# Large model!!

This includes all trees besides clumps

```{r}
large <- filter(data, dbh != 1.27 & `Dead / Alive_v1`==1)

cor(large[, c("logdbh","logheight", "DISTANCE_v1", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new_v1", "mean_dist_infested_v1", "dist_nearest_tree", "basal_area_host_std", "perc.infested_v1", "n.infested_v1", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])

large.base <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.base.sims <- simulateResiduals(large.base)
plot(large.base.sims)

summary(large.base)

# Add brooms

large.broom <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + brooms_recode + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.broom.sims <- simulateResiduals(large.broom)
plot(large.broom.sims)

summary(large.broom)

# brooms in neighborhood

large.broom.nh <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + brooms_nh + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.broom.sims <- simulateResiduals(large.broom.nh)
plot(large.broom.sims)

summary(large.broom.nh)

# Add beetle
large.beetle <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + beetle_recode + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.beetle.sims <- simulateResiduals(large.beetle)
plot(large.beetle.sims)

summary(large.beetle)

# Add infest sum

large.sum <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1  + infest_sum_std_new_v1 + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.sum.sims <- simulateResiduals(large.sum)
plot(large.sum.sims)

summary(large.sum)

# mean distance

large.mean.dis <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + mean_dist_infested_v1 + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.mean.dis.sims <- simulateResiduals(large.mean.dis)
plot(large.mean.dis.sims)

summary(large.mean.dis)


# percent infested - 0.75 correlation with infest sum so questionable
#large$logperc.infested_v1 <- log(large$perc.infested_v1 + 1)

large.pinfest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + perc.infested_v1 + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.pinfest.sims <- simulateResiduals(large.pinfest)
plot(large.pinfest.sims)

summary(large.pinfest)

# number in neighborhood

large.n.nhinfest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + n + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.n.nhinfest.sims <- simulateResiduals(large.n.nhinfest)
plot(large.n.nhinfest.sims)

summary(large.n.nhinfest)

# distance to nearest tree

large.dist.nearinfest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + dist_nearest_tree + n + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.dist.nearinfest.sims <- simulateResiduals(large.dist.nearinfest)
plot(large.dist.nearinfest.sims)

summary(large.dist.nearinfest)

# avg. dbh in neighborhood

large.dbh.infest <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + dist_nearest_tree + avg.dbh + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.dbh.infest.sims <- simulateResiduals(large.dbh.infest)
plot(large.dbh.infest.sims)

summary(large.dbh.infest)

## Building model above

# Final Model!

large.final <- glmmTMB(cbind(infested_v1, 1 - infested_v1) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + avg.dbh + (1|unique_ID) + (1|site), data = large, family = binomial(link = "logit"))

large.final.sims <- simulateResiduals(large.final)
plot(large.final.sims)

summary(large.final)

# lagged variables

large.sm <- sm %>% filter(dbh != 1.27)

large_long <- data_long %>% filter(dbh != 1.27)

large.lag <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + lag_infested + avg.dbh + gouting_recode + dist_nearest_tree + perc.infested + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = large.sm, family = binomial(link = "logit"))

large.lag.sims <- simulateResiduals(large.lag)
plot(large.lag.sims)

summary(large.lag)
summary(clump.lag)

large.all <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.all.sims <- simulateResiduals(large.all)
plot(large.all.sims)

summary(large.all)

```


```{r}
tidy_large <- broom::tidy(large.lag) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_large$model <- "large.lag" 

dw_large <- dwplot(tidy_large)

tidy_small <- broom::tidy(clump.lag)

tidy_small$model <- "clump.lag"

dw_small <- dwplot(tidy_small)

tidy_all <- rbind(tidy_large, tidy_small)

dw_all <- dwplot(tidy_all) +
  scale_y_discrete(
    labels = c("logDISTANCE" = "Distance to nearest infested",
               "lag_infested" = "Lagged infestation status",
               "dist_nearest_tree" = "Distance to nearest tree",
               "perc.infested" = "Percent infested (neighborhood)",
                "gouting_recode" = "Gouting",
                "avg.dbh" = "Average dbh in neighborhood",
                "logdbh" = "DBH",
                "mean_percent_live_crown_std" = "Mean percent live crown (Neighborhood)"
      ),
    limits = c("logDISTANCE",
                "mean_percent_live_crown_std",
                "avg.dbh",
                "perc.infested",
                "dist_nearest_tree",
                "logdbh",
                "gouting_recode", 
                "lag_infested") # Specify the desired order
  )

print(dw_large+geom_vline(xintercept=0,lty=2))
print(dw_small+geom_vline(xintercept=0,lty=2))
print(dw_all+geom_vline(xintercept=0,lty=2))
   
```

# Testing lagged dataset

```{r}

time.glmm.withlag <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + brooms_recode + lag_logDISTANCE + dist_nearest_tree + infest_sum_std_new + lag_infest_sum_std_new + Gouting + logdbh + (1|unique_ID) + (1|site), data = sm, family = binomial(link = "logit"))

large.sm <- sm %>% filter(dbh != 1.27)

large.lag <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE +  lag_logDISTANCE + lag_infest_sum + Gouting + brooms_recode + dist_nearest_tree + infest_sum_std_new + (1|unique_ID) + (1|site), data = large.sm, family = binomial(link = "logit"))

clump.sm <- sm %>% filter(dbh == 1.27)

# change n to number in neighborhood
names(clump.sm)[25] <- "number_nh"

clump.lag <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE +  lag_logDISTANCE + lag_infest_sum + Gouting + brooms_recode + infest_sum_std_new + number_nh + (1|unique_ID) + (1|site), data = clump.sm, family = binomial(link = "logit"))

summary(time.glmm.withlag)
summary(large.lag)
summary(clump.lag)

```

# Question 3: What are the risk factors for new infestation?

We need to make a variable for trees for new infestations and compare with trees that never got infested

# Get data prepped
```{r}

# Filter out so we have just new infested and never infested

new <- local_data_10 %>% filter(infested_v1 == 0) %>% filter(species == "ABLA" | species == "ABCO")

new_check <- data %>% select(c("infested_v1", "infested_v2", "infested_v3", "infested_v4", "new_infested_v2", "new_infested_v3", "new_infested_v4"))


cor(new[, c("logdbh","logheight", "DISTANCE_v1", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new_v1", "mean_dist_infested_v1", "dist_nearest_tree", "basal_area_host_std", "perc.infested_v1", "n.infested_v1", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])

# lagged variables

new_long <- new %>%
  pivot_longer(
    cols = ends_with("_v1") | ends_with("_v2") | ends_with("_v3") | ends_with("_v4"),  # columns with visit data
    names_to = c(".value", "visit"),       # use ".value" to keep the original variable name
    names_pattern = "(.*)_(v[0-9])"        # regex to separate variable name and visit
  ) %>%
  mutate(visit = as.numeric(gsub("v", "", visit)))  # Convert "v1", "v2", etc. to 1, 2, etc.

names(new_long)[49] <- "t"

new_long <- new_long %>%
  dplyr::arrange(unique_ID, t) %>%
  dplyr::group_by(unique_ID) %>%
  dplyr::mutate(
    lag_logDISTANCE = lag(logDISTANCE),
    lag_infest_sum_std_new = lag(infest_sum_std_new),
    lag_Gouting = lag(Gouting),
    lag_infested = lag(infested),
    lag_mean_dis = lag(mean_dist_infested),
    lag_perc_inf = lag(perc.infested)
  ) %>%
  ungroup()

new_three_visits <- subset(new_long, t %in% c(2, 3, 4))

new_three_visits$gouting_recode <- ifelse(new_three_visits$Gouting == 0, 0, 1)

new_long_check <- new_three_visits %>% select(c("infested", "new_infested"))

```

# Building new infestation model

```{r}
cor(new_three_visits[, c("logdbh","logheight", "logDISTANCE", "lag_logDISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new", "lag_infest_sum_std_new", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "n.infested", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])

## Building model below

# Start a-modelling

new.base <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.base.sims <- simulateResiduals(new.base)
plot(new.base.sims)

summary(new.base)

# Add gouting

new.gout <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.gout.sims <- simulateResiduals(new.gout)
plot(new.gout.sims)

summary(new.gout)

# Add brooms

new.broom <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.broom.sims <- simulateResiduals(new.broom)
plot(new.broom.sims)

summary(new.broom)

# Brooms - nh
new.broom.nh <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_nh + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.broom.sims <- simulateResiduals(new.broom.nh)
plot(new.broom.sims)

summary(new.broom.nh)


# Add beetle

new.beetle <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.beetle.sims <- simulateResiduals(new.beetle)
plot(new.beetle.sims)

summary(new.beetle)

# Add infest sum

new.sum <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + infest_sum_std_new + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.sum.sims <- simulateResiduals(new.sum)
plot(new.sum.sims)

summary(new.sum)

# percent infested

new.pinfest <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + infest_sum_std_new + perc.infested + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.pinfest.sims <- simulateResiduals(new.pinfest)
plot(new.pinfest.sims)

summary(new.pinfest)

# number in neighborhood

new.n.nhinfest <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + n + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.n.nhinfest.sims <- simulateResiduals(new.n.nhinfest)
plot(new.n.nhinfest.sims)

summary(new.n.nhinfest)

# add basal area host

new.ba_host <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode  + infest_sum_std_new + perc.infested + n + basal_area_host_std + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.ba_host.sims <- simulateResiduals(new.ba_host)
plot(new.ba_host.sims)

summary(new.ba_host)

# Tried avg.gouting, avg.dbh, n, and basal area host

# Final Model!

new.final <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + mean_percent_live_crown + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))
 
new.final.sims <- simulateResiduals(new.final)
plot(new.final.sims)
 
summary(new.final)


new.lag <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + lag_logDISTANCE + gouting_recode + dist_nearest_tree + mean_percent_live_crown + (1|unique_ID) + (1|site), data = new_three_visits, family = binomial(link = "logit"))

new.lag.sims <- simulateResiduals(new.lag)
plot(new.lag.sims)
testDispersion(new.lag.sims)

summary(new.lag)
summary(new.final)

new.all <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode + infest_sum_std_new + avg.dbh + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))
 
new.all.sims <- simulateResiduals(new.all)
plot(new.all.sims)
 
summary(new.all)

tidy_new <- broom::tidy(new.final) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_new$model <- "new" 

#new.final <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + beetle_recode + infest_sum_std_new_v1 + avg.dbh + (1|unique_ID) + (1|site), data = new, family = binomial(link = "logit"))


dw_new <- dwplot(tidy_new) +
  scale_y_discrete(
    labels = c("logDISTANCE" = "Distance to nearest infested",
                "gouting_recode" = "Gouting",
                "perc.infested" = "Percent infested (neighborhood)",
                "mean_percent_live_crown" = "Mean percent live crown (neighborhood)",
                "dist_nearest_tree" = "Distance to nearest tree",
                "logdbh" = "DBH"),
    limits = c( "logDISTANCE",
                "mean_percent_live_crown",
                "perc.infested",
                "dist_nearest_tree",
                "gouting_recode",
                "logdbh") # Specify the desired order
  )
print(dw_new+geom_vline(xintercept=0,lty=2))


#data %>% group_by(site) %>% count(data$new_infestation)

# Build new infestation model with all data 

# Start a-modelling

new.base <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.base.sims <- simulateResiduals(new.base)
plot(new.base.sims)

summary(new.base)

# Add gouting

new.gout <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.gout.sims <- simulateResiduals(new.gout)
plot(new.gout.sims)

summary(new.gout)

# Add brooms

new.broom <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + brooms_recode + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.broom.sims <- simulateResiduals(new.broom)
plot(new.broom.sims)

summary(new.broom)

# Brooms - nh
new.broom.nh <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + brooms_nh + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.broom.sims <- simulateResiduals(new.broom.nh)
plot(new.broom.sims)

summary(new.broom.nh)


# Add beetle

new.beetle <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode + brooms_nh + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.beetle.sims <- simulateResiduals(new.beetle)
plot(new.beetle.sims)

summary(new.beetle)

# Add infest sum

new.sum <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode + infest_sum_std_new + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.sum.sims <- simulateResiduals(new.sum)
plot(new.sum.sims)

summary(new.sum)

# percent infested

new.pinfest <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode + infest_sum_std_new + perc.infested + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.pinfest.sims <- simulateResiduals(new.pinfest)
plot(new.pinfest.sims)

summary(new.pinfest)

# number in neighborhood

new.n.nhinfest <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + beetle_recode + n + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.n.nhinfest.sims <- simulateResiduals(new.n.nhinfest)
plot(new.n.nhinfest.sims)

summary(new.n.nhinfest)

# add basal area host

new.ba_host <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode  + infest_sum_std_new + n + basal_area_host_std + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.ba_host.sims <- simulateResiduals(new.ba_host)
plot(new.ba_host.sims)

summary(new.ba_host)

# Final Model!

new.final <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + infest_sum_std_new + avg.dbh + (1|unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))
 
new.final.sims <- simulateResiduals(new.final)
plot(new.final.sims)
 
summary(new.final)

testDispersion(new.final.sims)
```

## New infestations with spatial weights

### Data prep
```{r}
proportions_list_new <- list()

for (site in unique(all_merged$site)) {
  
  # Filter focal and complete datasets for the site
  site_focal <- filter(new_three_visits, site == !!site)   # Focal points to buffer
  site_complete <- filter(all_merged, site == !!site)  # All points for bounding box
  
  # Convert focal points (already in UTM) to sf
  site_focal_sf <- st_as_sf(site_focal, coords = c("Longitude", "Latitude"), crs = 32612)
  
  # Convert complete dataset (in decimal degrees) to sf and transform to UTM
  site_complete_sf <- st_as_sf(site_complete, coords = c("Longitude", "Latitude"), crs = 4326)
  site_complete_utm <- st_transform(site_complete_sf, 32612)  # Transform to UTM
  
  # Get bounding box for the site in UTM
  site_bbox <- st_as_sfc(st_bbox(site_complete_utm))
  
  # Loop through each focal point and calculate proportions
  site_proportions <- sapply(1:nrow(site_focal_sf), function(i) {
    
    # Create a 10m buffer for the focal point
    buffer <- suppressWarnings(st_buffer(st_geometry(site_focal_sf)[i], dist = 10))
    
    # Compute the intersection between the buffer and the bounding box
    intersection <- suppressWarnings(st_intersection(buffer, site_bbox))
    
    # Calculate areas (buffer area and intersection area)
    circle_area <- suppressWarnings(st_area(buffer))
    intersection_area <- suppressWarnings(st_area(intersection))
    
    # Calculate proportion of buffer inside the bounding box
    proportion <- as.numeric(intersection_area / circle_area)
    
    return(proportion)
  })
  
  # Store the proportions for the current site
  proportions_list_new[[site]] <- site_proportions
  
  # Print progress for debugging
  cat("Processed", site, "with", length(site_proportions), "focal points.\n")
}


# Now append these proportions to the original focal dataset
new_with_proportions <- new_three_visits %>%
  mutate(proportion = unlist(proportions_list_new))

# Spatial weighting
model_data_new <- new_with_proportions
model_data_new$infest_sum_std_new <- model_data_new$infest_sum_std_new * model_data_new$proportion
model_data_new$avg.dbh <- model_data_new$avg.dbh * model_data_new$proportion
model_data_new$mean_percent_live_crown_std <- model_data_new$mean_percent_live_crown_std * model_data_new$proportion
model_data_new$brooms_nh <- model_data_new$brooms_nh * model_data_new$proportion
model_data_new$perc.infested <- model_data_new$perc.infested * model_data_new$proportion
model_data_new$basal_area_host_std <- model_data_new$basal_area_host_std * model_data_new$proportion
model_data_new$mean_dist_infested_recode <- model_data_new$mean_dist_infested_recode * model_data_new$proportion
model_data_new$n <- model_data_new$n * model_data_new$proportion

# Filter live
model_data_new$`Dead / Alive`[is.na(model_data_new$`Dead / Alive`)] <- 0
model_data_new <- filter(model_data_new, model_data_new$`Dead / Alive` == 1)
```


### Model Building - New infestation with Spatial weights

```{r}
cor(model_data_new[, c("logdbh", "logDISTANCE", "lag_logDISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])

## Building model below

# Start a-modelling

new.base.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.base.sp.sims <- simulateResiduals(new.base.sp)
plot(new.base.sp.sims)

summary(new.base.sp)

# Add gouting

new.gout.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.gout.sp.sims <- simulateResiduals(new.gout.sp)
plot(new.gout.sp.sims)

summary(new.gout.sp)

# Add brooms

new.broom.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.broom.sp.sims <- simulateResiduals(new.broom.sp)
plot(new.broom.sp.sims)

summary(new.broom.sp)

# Brooms - nh
new.broom.sp.nh <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + brooms_nh + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.broom.sp.sims <- simulateResiduals(new.broom.sp.nh)
plot(new.broom.sp.sims)

summary(new.broom.sp.nh)


# Add beetle

new.beetle.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + beetle_recode + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.beetle.sims.sp <- simulateResiduals(new.beetle.sp)
plot(new.beetle.sims.sp)

summary(new.beetle.sp)

# percent infested

new.pinfest.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.pinfest.sims.sp <- simulateResiduals(new.pinfest.sp)
plot(new.pinfest.sims.sp)

summary(new.pinfest.sp)

# number in neighborhood

new.n.nh.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.n.nh.sims.sp <- simulateResiduals(new.n.nh.sp)
plot(new.n.nh.sims.sp)

summary(new.n.nh.sp)

# add basal area host

new.ba_host.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + basal_area_host_std + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.ba_host.sims.sp <- simulateResiduals(new.ba_host.sp)
plot(new.ba_host.sims.sp)

summary(new.ba_host.sp)

# avg.gouting

new.gouting.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + avg.gouting + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.gouting.sims.sp <- simulateResiduals(new.gouting.sp)
plot(new.gouting.sims.sp)

summary(new.gouting.sp)

# avg.dbh

new.avgdbh.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + avg.dbh + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.avgdbh.sims.sp <- simulateResiduals(new.avgdbh.sp)
plot(new.avgdbh.sims.sp)

summary(new.avgdbh.sp)

# Mean percent live crown

new.crown.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.crown.sims.sp <- simulateResiduals(new.crown.sp)
plot(new.crown.sims.sp)

summary(new.crown.sp)

# distance to nearest tree

new.dist.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + mean_percent_live_crown_std + n + dist_nearest_tree + (1|unique_ID) + (1|site), data = model_data_new, family = binomial(link = "logit"))

new.dist.sims.sp <- simulateResiduals(new.dist.sp)
plot(new.dist.sims.sp)

summary(new.dist.sp)

# Final Model!

# lag_logDISTANCE doesn't come out, but lag percentage infested does

# Beetle has very large confidence interval due to small number of observations, which warrants exclusion. 

new.final.sp <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + lag_perc_inf + mean_percent_live_crown_std + n + DISTANCE_noinfest + (1|site), data = model_data_new, family = binomial(link = "logit"))
 
new.final.sims.sp <- simulateResiduals(new.final.sp)
plot(new.final.sims.sp)
 
summary(new.final.sp)
r2(new.final.sp)
check_singularity(new.final.sp)
```

### Coef plot - new infestation spatial weights

```{r}
new.final <- broom::tidy(new.final.sp) # Create a tidy dataframe for the model

# Add a column to identify the model
new.final$model <- "new.final.sp" 

dw_new_final <- dwplot(new.final) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "perc.infested" = "Percent Infested (neighborhood)",
               "lag_perc_inf" = "Lagged percent infested (neighborhood)",
               "DISTANCE_noinfest" = "Distance to nearest tree > 2.54cm",
                "gouting_recode" = "Gouting",
                "mean_percent_live_crown_std" = "Mean percent live crown (neighborhood)",
                "logdbh" = "DBH",
                "n" = "Number of trees in neighborhood"
      ),
    limits = c( "logDISTANCE",
                "mean_percent_live_crown_std",
                "lag_perc_inf",
                "n",
                "perc.infested",
                "DISTANCE_noinfest",
                "gouting_recode",
                "logdbh"
                ) # Specify the desired order
  )


print(dw_new_final+geom_vline(xintercept=0,lty=2))
```

### New infestation model removing observations after new infestation

```{r}
new_data_remove <- model_data_new %>% filter(!(infested == 1 & new_infested == 0))

new.base.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.base.rem.sims <- simulateResiduals(new.base.rem)
plot(new.base.rem.sims)

summary(new.base.rem)

# Add gouting

new.gout.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.gout.rem.sims <- simulateResiduals(new.gout.rem)
plot(new.gout.rem.sims)

summary(new.gout.rem)

# Add brooms

new.broom.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.broom.rem.sims <- simulateResiduals(new.broom.rem)
plot(new.broom.rem.sims)

summary(new.broom.rem)

# Brooms - nh
new.broom.rem.nh <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + brooms_nh + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.broom.rem.sims <- simulateResiduals(new.broom.rem.nh)
plot(new.broom.rem.sims)

summary(new.broom.rem.nh)


# Add beetle

new.beetle.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.beetle.sims.rem <- simulateResiduals(new.beetle.rem)
plot(new.beetle.sims.rem)

summary(new.beetle.rem)

# percent infested

new.pinfest.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.pinfest.sims.rem <- simulateResiduals(new.pinfest.rem)
plot(new.pinfest.sims.rem)

summary(new.pinfest.rem)

# number in neighborhood

new.n.nh.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.n.nh.sims.rem <- simulateResiduals(new.n.nh.rem)
plot(new.n.nh.sims.rem)

summary(new.n.nh.rem)

# add basal area host

new.ba_host.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + basal_area_host_std + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.ba_host.sims.rem <- simulateResiduals(new.ba_host.rem)
plot(new.ba_host.sims.rem)

summary(new.ba_host.rem)

# avg.gouting

new.gouting.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + avg.gouting + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.gouting.sims.rem <- simulateResiduals(new.gouting.rem)
plot(new.gouting.sims.rem)

summary(new.gouting.rem)

# avg.dbh

new.avgdbh.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + avg.dbh + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.avgdbh.sims.rem <- simulateResiduals(new.avgdbh.rem)
plot(new.avgdbh.sims.rem)

summary(new.avgdbh.rem)

# Mean percent live crown

new.crown.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + n + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.crown.sims.rem <- simulateResiduals(new.crown.rem)
plot(new.crown.sims.rem)

summary(new.crown.rem)

# distance to nearest tree

new.dist.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + perc.infested + mean_percent_live_crown_std + n + dist_nearest_tree + (1|unique_ID) + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.dist.sims.rem <- simulateResiduals(new.dist.rem)
plot(new.dist.sims.rem)

summary(new.dist.rem)

# Lagged vars

new.lag.rem <- glmmTMB(cbind(new_infested, 1 - new_infested) ~ logdbh + logDISTANCE + gouting_recode 
  #+ beetle_recode 
  + lag_perc_inf + perc.infested + mean_percent_live_crown_std + n + dist_nearest_tree + (1|site), data = new_data_remove, family = binomial(link = "logit"))

new.lag.sims.rem <- simulateResiduals(new.lag.rem)
plot(new.dist.sims.rem)

summary(new.lag.rem)
r2(new.lag.rem)

```

### Coef plot - new infestation spatial weights removing trees after infestation

```{r}
new.rem <- broom::tidy(new.lag.rem) # Create a tidy dataframe for the model

# Add a column to identify the model
new.rem$model <- "new.lag.rem" 

dw_new_rem <- dwplot(new.rem) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "perc.infested" = "Percent Infested (neighborhood)",
               "lag_perc_inf" = "Lagged percent infested (neighborhood)",
               "dist_nearest_tree" = "Distance to nearest tree",
                "gouting_recode" = "Gouting",
                #"beetle_recode" = "Beetle - presence/absence",
                #"mean_percent_live_crown_std" = "Mean percent live crown (neighborhood)",
                "logdbh" = "DBH",
                "n" = "Number of trees in neighborhood"
      ),
    limits = c( "logDISTANCE",
                #"beetle_recode",
                #"mean_percent_live_crown_std",
                "lag_perc_inf",
                "n",
                "perc.infested",
                "dist_nearest_tree",
                "gouting_recode",
                "logdbh"
                ) # Specify the desired order
  )


print(dw_new_rem+geom_vline(xintercept=0,lty=2))
```


# Question 4: What factors influence changes in infestation level?

```{r}
cor(data[, c("logdbh", "logDISTANCE_v1", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new_v1", "dist_nearest_tree", "basal_area_host_std", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])

# Start a-modelling

sum.base <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.base.sims <- simulateResiduals(sum.base)
plot(sum.base.sims)

summary(sum.base)

# Add gouting

sum.gout <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + logheight + Gouting_v1 + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.gout.sims <- simulateResiduals(sum.gout)
plot(sum.gout.sims)

summary(sum.gout)

# Add brooms

sum.broom <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + logheight + Gouting_v1 + brooms_recode + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.broom.sims <- simulateResiduals(sum.broom)
plot(sum.broom.sims)

summary(sum.broom)

# Brooms - nh

sum.broom.nh <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + logheight + Gouting_v1 + brooms_nh + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.broom.sims <- simulateResiduals(sum.broom.nh)
plot(sum.broom.sims)

summary(sum.broom.nh)

# Add beetle

sum.beetle <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + Gouting_v1 + beetle_recode + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.beetle.sims <- simulateResiduals(sum.beetle)
plot(sum.beetle.sims)

summary(sum.beetle)

# Add infest sum

sum.sum <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.sum.sims <- simulateResiduals(sum.sum)
plot(sum.sum.sims)

summary(sum.sum)

# number in neighborhood

sum.n.nhinfest <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + n + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.n.nhinfest.sims <- simulateResiduals(sum.n.nhinfest)
plot(sum.n.nhinfest.sims)

summary(sum.n.nhinfest)

# add basal area host

sum.ba_host <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + n + basal_area_host_std + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.ba_host.sims <- simulateResiduals(sum.ba_host)
plot(sum.ba_host.sims)

summary(sum.ba_host)

# Average DBH

sum.avg.dbh <- glmmTMB(infestation_sum_v1 ~ logdbh + logDISTANCE_v1 + Gouting_v1 + infest_sum_std_new_v1 + n + avg.dbh + (1|unique_ID) + (1|site), data = data, family = nbinom2)

sum.avg.dbh.sims <- simulateResiduals(sum.avg.dbh)
plot(sum.avg.dbh.sims)

summary(sum.avg.dbh)

# distance to nearest tree
sum.dist <- glmmTMB(infestation_sum_v1 ~ infest_sum_std_new_v1 + dist_nearest_tree + Gouting_v1 + logDISTANCE_v1 + logdbh + n + avg.dbh + (1|site) + (1|unique_ID), data = data, family = nbinom2)

sum.dist.sims <- simulateResiduals(sum.dist)
plot(sum.dist.sims)

summary(sum.dist)

# mean percent live crown
sum.mean.perc <- glmmTMB(infestation_sum_v1 ~ infest_sum_std_new_v1 + dist_nearest_tree + Gouting_v1 + logDISTANCE_v1 + logdbh + n + avg.dbh + mean_percent_live_crown_std + (1|site) + (1|unique_ID), data = data, family = nbinom2)

sum.mperc.sims <- simulateResiduals(sum.mean.perc)
plot(sum.mperc.sims)

summary(sum.mean.perc)

# Final

sum.final <- glmmTMB(infestation_sum_v1 ~ infest_sum_std_new_v1 + dist_nearest_tree + Gouting_v1 + logDISTANCE_v1 + logdbh + avg.dbh + (1|site) + (1|unique_ID), data = data, family = nbinom2)

sum.final.sims <- simulateResiduals(sum.final)
plot(sum.final.sims)

summary(sum.final)

# Final model with lagged variables

cor(sm[, c("logdbh", "logDISTANCE", "lag_logDISTANCE", "lag_infest_sum_std_new", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new", "dist_nearest_tree", "basal_area_host_std", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n")])


sum.time <- glmmTMB(infestation_sum ~  logdbh + logDISTANCE + lag_logDISTANCE + dist_nearest_tree + gouting_recode + avg.dbh + infest_sum_std_new + lag_infest_sum_std_new + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = sm, family = nbinom2)

sum.time.sims <- simulateResiduals(sum.time)
plot(sum.time.sims)
testDispersion(sum.time.sims)

summary(sum.time)
summary(time.glmm.lag.interact) # lagged model for infestation status

coef_plot_sum <- ggcoef_model(sum.time)
coef_plot_status <- ggcoef_model(time.glmm.lag.interact)

tidy_sum <- broom::tidy(sum.time) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_sum$model <- "sum" 

#new.final <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE_v1 + Gouting_v1 + beetle_recode + infest_sum_std_new_v1 + avg.dbh + (1|unique_ID) + (1|site), data = new, family = binomial(link = "logit"))


dw_sum <- dwplot(tidy_sum) +
  scale_y_discrete(
    labels = c( "logDISTANCE" = "Distance to nearest infested",
                "gouting_recode" = "Gouting",
                "lag_logDISTANCE" = "Lagged distance to nearest infested",
                "mean_percent_live_crown_std" = "Mean percent live crown (neighborhood)",
                "dist_nearest_tree" = "Distance to nearest tree",
                "logdbh" = "DBH",
                "lag_infest_sum_std_new" = "Lagged infestation sum (neighborhood)",
                "infest_sum_std_new" = "Infestation sum (neighborhood)",
                "avg.dbh" = "Average DBH (neighborhood)"),
    limits = c( "logDISTANCE",
                "lag_logDISTANCE",
                "mean_percent_live_crown_std",
                "avg.dbh",
                #"lag_infest_sum_std_new",
                "gouting_recode",
                "infest_sum_std_new",
                "dist_nearest_tree",
                "logdbh") # Specify the desired order
  )
print(dw_sum+geom_vline(xintercept=0,lty=2))

```

## Building sum with long data

```{r}

# Building sum with long data

sum.base_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.base_long.sims <- simulateResiduals(sum.base_long)
plot(sum.base_long.sims)

summary(sum.base_long)

# Add gouting

sum.gout_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.gout_long.sims <- simulateResiduals(sum.gout_long)
plot(sum.gout_long.sims)

summary(sum.gout_long)

# Add brooms

sum.broom_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + brooms_recode + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.broom_long.sims <- simulateResiduals(sum.broom_long)
plot(sum.broom_long.sims)

summary(sum.broom_long)

# Brooms - nh

sum.broom.nh_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + brooms_nh + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.broom_long.sims <- simulateResiduals(sum.broom.nh_long)
plot(sum.broom_long.sims)

summary(sum.broom.nh_long)

# Add beetle

sum.beetle_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + beetle_recode + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.beetle_long.sims <- simulateResiduals(sum.beetle_long)
plot(sum.beetle_long.sims)

summary(sum.beetle_long)

# Add infest sum

sum.sum_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.sum_long.sims <- simulateResiduals(sum.sum_long)
plot(sum.sum_long.sims)

summary(sum.sum_long)

# number in neighborhood

sum.n.nhinfest_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + n + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.n.nhinfest_long.sims <- simulateResiduals(sum.n.nhinfest_long)
plot(sum.n.nhinfest_long.sims)

summary(sum.n.nhinfest_long)

# add basal area host

sum.ba_host_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + n + basal_area_host_std + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.ba_host_long.sims <- simulateResiduals(sum.ba_host_long)
plot(sum.ba_host_long.sims)

summary(sum.ba_host_long)

# Average DBH

sum.avg.dbh_long <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + n + avg.dbh + (1|unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.avg.dbh_long.sims <- simulateResiduals(sum.avg.dbh_long)
plot(sum.avg.dbh_long.sims)

summary(sum.avg.dbh_long)

# distance to nearest tree
sum.dist_long <- glmmTMB(infestation_sum ~ infest_sum_std_new + dist_nearest_tree + Gouting + logDISTANCE + logdbh + n + avg.dbh + (1|site) + (1|unique_ID), data = data_long, family = nbinom2)

sum.dist_long.sims <- simulateResiduals(sum.dist_long)
plot(sum.dist_long.sims)

summary(sum.dist_long)

# Final

sum.final_long <- glmmTMB(log(infestation_sum + 1) ~ infest_sum_std_new + dist_nearest_tree + Gouting + logDISTANCE + logdbh + avg.dbh + (1|site) + (1|unique_ID), data = data_long, family = nbinom2)

sum.final_long.sims <- simulateResiduals(sum.final_long)
plot(sum.final_long.sims)

summary(sum.final_long)

coef_plot_sum <- ggcoef_model(sum.time)
coef_plot_sum_long <- ggcoef_model(sum.final_long)

testDispersion(sum.final_long.sims)

print(coef_plot_sum)
print(coef_plot_sum_long)

```


## Sum model with spatial weights

### Data prep

```{r}
all_merged$site <- as.factor(all_merged$site)

proportions_list <- list()

for (site in unique(all_merged$site)) {
  
  # Filter focal and complete datasets for the site
  site_focal <- filter(sm, site == !!site)   # Focal points to buffer
  site_complete <- filter(all_merged, site == !!site)  # All points for bounding box
  
  # Convert focal points (already in UTM) to sf
  site_focal_sf <- st_as_sf(site_focal, coords = c("Longitude", "Latitude"), crs = 32612)
  
  # Convert complete dataset (in decimal degrees) to sf and transform to UTM
  site_complete_sf <- st_as_sf(site_complete, coords = c("Longitude", "Latitude"), crs = 4326)
  site_complete_utm <- st_transform(site_complete_sf, 32612)  # Transform to UTM
  
  # Get bounding box for the site in UTM
  site_bbox <- st_as_sfc(st_bbox(site_complete_utm))
  
  # Loop through each focal point and calculate proportions
  site_proportions <- sapply(1:nrow(site_focal_sf), function(i) {
    
    # Create a 10m buffer for the focal point
    buffer <- suppressWarnings(st_buffer(st_geometry(site_focal_sf)[i], dist = 10))
    
    # Compute the intersection between the buffer and the bounding box
    intersection <- suppressWarnings(st_intersection(buffer, site_bbox))
    
    # Calculate areas (buffer area and intersection area)
    circle_area <- suppressWarnings(st_area(buffer))
    intersection_area <- suppressWarnings(st_area(intersection))
    
    # Calculate proportion of buffer inside the bounding box
    proportion <- as.numeric(intersection_area / circle_area)
    
    return(proportion)
  })
  
  # Store the proportions for the current site
  proportions_list[[site]] <- site_proportions
  
  # Print progress for debugging
  cat("Processed", site, "with", length(site_proportions), "focal points.\n")
}


# Now append these proportions to the original focal dataset
sm_with_proportions <- sm %>%
  mutate(proportion = unlist(proportions_list))

# View the updated focal dataset with proportions
View(sm_with_proportions)

# update neighborhood metrics with spatial weights
model_data <- sm_with_proportions
model_data$infest_sum_std_new <- model_data$infest_sum_std_new * model_data$proportion
model_data$avg.dbh <- model_data$avg.dbh * model_data$proportion
model_data$mean_percent_live_crown_std <- model_data$mean_percent_live_crown_std * model_data$proportion
model_data$brooms_nh <- model_data$brooms_nh * model_data$proportion
model_data$perc.infested <- model_data$perc.infested * model_data$proportion
model_data$basal_area_host_std <- model_data$basal_area_host_std * model_data$proportion
model_data$mean_dist_infested_recode <- model_data$mean_dist_infested_recode * model_data$proportion
model_data$n <- model_data$n * model_data$proportion
model_data$basal_area_nh_std <- model_data$basal_area_nh_std * model_data$proportion
model_data$comp_index <- model_data$comp_index * model_data$proportion

model_data$log_DISTANCE_noinfest <- log(model_data$DISTANCE_noinfest)


# Filter live
model_data$`Dead / Alive`[is.na(model_data$`Dead / Alive`)] <- 0
model_data <- filter(model_data, model_data$`Dead / Alive` == 1)
```

### Building sum model
```{r}

sum.base.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.base.sp.sims <- simulateResiduals(sum.base.sp)
plot(sum.base.sp.sims)

summary(sum.base.sp)

# Add gouting

sum.gout.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + gouting_recode + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.gout.sp.sims <- simulateResiduals(sum.gout.sp)
plot(sum.gout.sp.sims)

summary(sum.gout.sp)

# Add brooms

sum.broom.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + gouting_recode + brooms_recode + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.broom.sp.sims <- simulateResiduals(sum.broom.sp)
plot(sum.broom.sp.sims)

summary(sum.broom.sp)

# Brooms - nh

sum.broom.nh.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + gouting_recode + brooms_nh + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.broom.sp.sims <- simulateResiduals(sum.broom.nh.sp)
plot(sum.broom.sp.sims)

summary(sum.broom.nh.sp)

# Add beetle

sum.beetle.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.beetle.sp.sims <- simulateResiduals(sum.beetle.sp)
plot(sum.beetle.sp.sims)

summary(sum.beetle.sp)

# percent infested neighborhood

sum.perc.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + gouting_recode + perc.infested + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.perc.sp.sims <- simulateResiduals(sum.perc.sp)
plot(sum.perc.sp.sims)

summary(sum.perc.sp)

# number in neighborhood

sum.n.nhinfest.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + gouting_recode + perc.infested + n + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.n.nhinfest.sp.sims <- simulateResiduals(sum.n.nhinfest.sp)
plot(sum.n.nhinfest.sp.sims)

summary(sum.n.nhinfest.sp)

# add basal area host

sum.ba_host.sp <- glmmTMB(log(infestation_sum + 1) ~ logdbh + logDISTANCE + gouting_recode + perc.infested  + basal_area_host_std + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.ba_host.sp.sims <- simulateResiduals(sum.ba_host.sp)
plot(sum.ba_host.sp.sims)

summary(sum.ba_host.sp)

# Average DBH

sum.avg.dbh.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + gouting_recode + perc.infested + avg.dbh + (1|unique_ID) + (1|site),
  data = model_data, 
  family = nbinom2)


sum.avg.dbh.sp.sims <- simulateResiduals(sum.avg.dbh.sp)
plot(sum.avg.dbh.sp.sims)

summary(sum.avg.dbh.sp)

# distance to nearest tree
sum.dist.sp <- glmmTMB(infestation_sum ~ perc.infested + dist_nearest_tree + gouting_recode + logDISTANCE + logdbh + (1|site) + (1|unique_ID), data = model_data, family = nbinom2)

sum.dist.sp.sims <- simulateResiduals(sum.dist.sp)
plot(sum.dist.sp.sims)

summary(sum.dist.sp)

# avg. gouting

sum.avg.dbh.sp <- glmmTMB(log(infestation_sum + 1) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + avg.dbh + dist_nearest_tree + avg.gouting + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.avg.dbh.sp.sims <- simulateResiduals(sum.avg.dbh.sp)
plot(sum.avg.dbh.sp.sims)

summary(sum.avg.dbh.sp)

# mean percent live crown

sum.crown.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + avg.gouting + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = model_data, family = nbinom2)

sum.crown.sp.sims <- simulateResiduals(sum.crown.sp)
plot(sum.crown.sp.sims)

summary(sum.crown.sp)

# Outlier problem
# Define outlier threshold
upper_bound <- quantile(model_data$infestation_sum, 0.95)

# Find outliers
outliers <- which(model_data$infestation_sum > upper_bound)

# View outliers
print(model_data[outliers, ])

model_data_clean <- model_data[-outliers, ]

# Final

# re-scale sum
model_data$infestation_sum_scaled <- scale(model_data$infestation_sum)

# standardize infestation sum
model_data$infestation_sum_std <- (model_data$infestation_sum - mean(model_data$infestation_sum))/sd(model_data$infestation_sum)

# These two do the same thing ^^

sum.final.sp <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + avg.gouting + mean_percent_live_crown_std + exp(Longitude, Latitude) + (1|unique_ID) + (1|site), 
  data = model_data, 
  #ziformula = ~ 1,
  family = nbinom2)

sum.final.sp.sims <- simulateResiduals(sum.final.sp.poly)
plot(sum.final.sp.sims)

# Attempting a GAM - takes way too long to fit
# sum.final.sp <- gam(
#   infestation_sum ~ s(Longitude, Latitude, bs = "gp") + logdbh + logDISTANCE + 
#     gouting_recode + perc.infested + dist_nearest_tree + avg.gouting + 
#     mean_percent_live_crown_std + s(unique_ID, bs = "re") + s(site, bs = "re"),
#   data = model_data,
#   family = nb()
# )

# Attempting spatial process

# Standardize coordinates within each site
model_data <- model_data %>%
  group_by(site) %>%
  mutate(X_scaled = scale(Longitude), Y_scaled = scale(Latitude)) %>%
  ungroup()

# Scaling attempt

model_data_scaled <- model_data %>%
  mutate(
    logdbh = scale(logdbh),
    logDISTANCE = scale(logDISTANCE),
    gouting_recode = scale(gouting_recode),
    perc.infested = scale(perc.infested),
    dist_nearest_tree = scale(dist_nearest_tree),
    avg.gouting = scale(avg.gouting)
  )

spatial_model <- glmmTMB(
  infestation_sum ~ logdbh + logDISTANCE + gouting_recode + perc.infested + 
    dist_nearest_tree + avg.gouting + mean_percent_live_crown_std +
    (1 | unique_ID) + (1 | site),  
  data = model_data_scaled,
  family = nbinom1
)

spatial.sims <- simulateResiduals(spatial_model)
plot(spatial.sims)

summary(spatial_model)

library(ggstats)
coef_plot_sum <- ggcoef_model(sum.time)
coef_plot_sum.sp <- ggcoef_model(sum.final.sp)

testDispersion(sum.final.sp.sims)
testZeroInflation(sum.final.sp.sims)

print(coef_plot_sum)
print(coef_plot_sum.sp)

coords <- model_data[, c("Longitude", "Latitude")]
nb <- dnearneigh(as.matrix(coords), 0, 10)  # Adjust range as needed
moran.test(residuals(sum.final.sp), nb2listw(nb))

# None of these are working, so we'll try to change dependent variable to change in infestation at each time step

# Get spatial weights for all of our points, because we need to calculate change in infesatio

all_merged$site <- as.factor(all_merged$site)

proportions_list <- list()

for (site in unique(all_merged$site)) {
  
  # Filter focal and complete datasets for the site
  site_focal <- filter(data_long, site == !!site)   # Focal points to buffer
  site_complete <- filter(all_merged, site == !!site)  # All points for bounding box
  
  # Convert focal points (already in UTM) to sf
  site_focal_sf <- st_as_sf(site_focal, coords = c("Longitude", "Latitude"), crs = 32612)
  
  # Convert complete dataset (in decimal degrees) to sf and transform to UTM
  site_complete_sf <- st_as_sf(site_complete, coords = c("Longitude", "Latitude"), crs = 4326)
  site_complete_utm <- st_transform(site_complete_sf, 32612)  # Transform to UTM
  
  # Get bounding box for the site in UTM
  site_bbox <- st_as_sfc(st_bbox(site_complete_utm))
  
  # Loop through each focal point and calculate proportions
  site_proportions <- sapply(1:nrow(site_focal_sf), function(i) {
    
    # Create a 10m buffer for the focal point
    buffer <- suppressWarnings(st_buffer(st_geometry(site_focal_sf)[i], dist = 10))
    
    # Compute the intersection between the buffer and the bounding box
    intersection <- suppressWarnings(st_intersection(buffer, site_bbox))
    
    # Calculate areas (buffer area and intersection area)
    circle_area <- suppressWarnings(st_area(buffer))
    intersection_area <- suppressWarnings(st_area(intersection))
    
    # Calculate proportion of buffer inside the bounding box
    proportion <- as.numeric(intersection_area / circle_area)
    
    return(proportion)
  })
  
  # Store the proportions for the current site
  proportions_list[[site]] <- site_proportions
  
  # Print progress for debugging
  cat("Processed", site, "with", length(site_proportions), "focal points.\n")
}


# Now append these proportions to the original focal dataset
long_with_proportions <- data_long %>%
  mutate(proportion = unlist(proportions_list))

# View the updated focal dataset with proportions
View(long_with_proportions)

# update neighborhood metrics with spatial weights
sum_model_data <- long_with_proportions
sum_model_data$infest_sum_std_new <- sum_model_data$infest_sum_std_new * sum_model_data$proportion
sum_model_data$avg.dbh <- sum_model_data$avg.dbh * sum_model_data$proportion
sum_model_data$mean_percent_live_crown_std <- sum_model_data$mean_percent_live_crown_std * sum_model_data$proportion
sum_model_data$brooms_nh <- sum_model_data$brooms_nh * sum_model_data$proportion
sum_model_data$perc.infested <- sum_model_data$perc.infested * sum_model_data$proportion
sum_model_data$basal_area_host_std <- sum_model_data$basal_area_host_std * sum_model_data$proportion
sum_model_data$mean_dist_infested_recode <- sum_model_data$mean_dist_infested_recode * sum_model_data$proportion
sum_model_data$n <- sum_model_data$n * sum_model_data$proportion

# Filter live
sum_model_data$`Dead / Alive`[is.na(sum_model_data$`Dead / Alive`)] <- 0
sum_model_data <- filter(sum_model_data, sum_model_data$`Dead / Alive` == 1)

# make change in infestation metric
# re-scale sum
sum_model_data$infestation_sum_scaled <- scale(sum_model_data$infestation_sum)

sum_model_data <- sum_model_data %>%
  arrange(unique_ID, t) %>%
  group_by(unique_ID) %>%
  mutate(change_in_infestation = infestation_sum_scaled - lag(infestation_sum_scaled)) %>%
  ungroup()

sum_model_data <- subset(sum_model_data, t %in% c(2, 3, 4))

# remove NA observations in change

sum_model_data <- filter(sum_model_data, !is.na(sum_model_data$change_in_infestation))
sum_model_data$perc.infested_std <- scale(sum_model_data$perc.infested)
#sum_model_data$change_in_infestation_positive <- sqrt(sum_model_data$change_in_infestation + abs(min(sum_model_data$change_in_infestation)) + 1)
sum_model_data$lag_infestation_sum[is.na(sum_model_data$lag_infestation_sum)] <- 0
sum_model_data$lag_infestation_sum_std <- scale(sum_model_data$lag_infestation_sum)

sum_model_data$logdbh_std <- scale(sum_model_data$logdbh)
sum_model_data$logDISTANCE_std <- scale(sum_model_data$logDISTANCE)

sum_model <- glmmTMB(
  change_in_infestation_positive ~ logdbh + logDISTANCE + gouting_recode + perc.infested_std + dist_nearest_tree + avg.gouting + mean_percent_live_crown_std +
    (1 | unique_ID) + (1 | site),  
  data = sum_model_data,
  family = gamma
)

sum_model_std <- glmmTMB(change_in_infestation ~ logdbh + logDISTANCE + lag_infestation_sum, 
  #+ gouting_recode + perc.infested_std + dist_nearest_tree + avg.gouting + mean_percent_live_crown_std + (1|site),
  ziformula = ~ logdbh + logDISTANCE + lag_infestation_sum,
  #+ gouting_recode + perc.infested_std + dist_nearest_tree + avg.gouting + mean_percent_live_crown_std,
  data = sum_model_data, 
  family = tweedie(link = "log"))

summary(sum_model_std)
r2(sum_model_std)

sum.sims <- simulateResiduals(sum_model_std)
plot(sum.sims)

pairs(sum_model_data[, c("change_in_infestation", "logdbh", "logDISTANCE", "perc.infested_std", "dist_nearest_tree", "avg.gouting", "mean_percent_live_crown_std")])

cor(sum_model_data[, c("logdbh", "logDISTANCE", "lag_logDISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "dist_nearest_tree", "basal_area_host_std","infest_sum_std", "perc.infested", "perc.infested_std", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std", "n", "lag_infested", "lag_infestation_sum")])

# zip model
zero_inflated_model <- glmmTMB(
  infestation_sum ~ logdbh + lag_infestation_sum + logDISTANCE + (1 | unique_ID) + (1 | site),  
  data = sum_model_data,  
  family = nbinom2,  
  zi = ~ logdbh + lag_infestation_sum + logDISTANCE + (1 | unique_ID) + (1 | site)
)

summary(zero_inflated_model)

zip.sims <- simulateResiduals(zero_inflated_model)
plot(zip.sims)
r2(zero_inflated_model)

sum_model_data$log_DISTANCE_noinfest <- log(sum_model_data$DISTANCE_noinfest)






# Data tweaks after meetings (2/12/24)

sum_model_data$season <- ifelse(sum_model_data$t == 3, 0,1)

# Standardize all variables
sum_model_data$dbh_std <- scale(sum_model_data$logdbh)
sum_model_data$DISTANCE_std <- scale(sum_model_data$logDISTANCE)
sum_model_data$DISTANCE_noinfest_std <- scale(sum_model_data$log_DISTANCE_noinfest)
sum_model_data$n_std <- scale(sum_model_data$n)
sum_model_data$perc.infested_std <- scale(sum_model_data$perc.infested)
sum_model_data$brooms_nh_std <- scale(sum_model_data$brooms_nh)
sum_model_data$lag_perc.infested_std <- scale(sum_model_data$lag_perc.infested)

# Filter out non-infested trees
sum_model_data <- sum_model_data %>% filter(infestation_sum != 0)

# Get max infestation
sections <- sum_model_data %>% select(c("NL", "NM", "NH", "SL", "SM", "SH"))
sections$max_infestation <- apply(sections, 1, max, na.rm = TRUE)
sum_model_data$max_infestation <- sections$max_infestation

# Only one working
sum_model_std <- glmmTMB(max_infestation ~ dbh_std + DISTANCE_std + perc.infested_std + mean_percent_live_crown_std + lag_perc.infested_std + gouting_recode + DISTANCE_noinfest_std + brooms_nh + season + (1|site),
  #ziformula = ~ dbh_std + lag_perc.infested_std + DISTANCE_noinfest_std + (1|site),
  data = sum_model_data, 
  family = poisson(link = "log"))

summary(sum_model_std)
r2(sum_model_std)

sum.sims <- simulateResiduals(sum_model_std)
plot(sum.sims)

# Add brooms - not sig
# Brooms - nh - sig
# Add beetle - not sig
# avg. gouting - sig
# gouting - sig
# number in neighborhood - killed it
# add basal area host - slight (.)
# basal area all trees - not sig
# distance to nearest non-infested above 2.54 - sig

# Ordinal Regression
library(ordinal)

sum_model_data$max_infestation <- as.factor(sum_model_data$max_infestation)
sum_ord <- clm(max_infestation ~ dbh_std + DISTANCE_std + mean_percent_live_crown_std + lag_infestation_sum_std + lag_perc.infested_std + gouting_recode + DISTANCE_noinfest_std + brooms_nh_std, data = sum_model_data)

summary(sum_ord)

# maximum likelihood plots
slice.fm1 <- slice(sum_ord, lambda = 5)
par(mfrow = c(2, 5))
plot(slice.fm1)

# same but zoomed in - looking for any deviation of dotted from solid line
slice2.fm1 <- slice(sum_ord, lambda = 1e-5)
plot(slice2.fm1)

# check convergence
convergence(sum_ord)
```

### Coef plot - infestation sum model

```{r}
sum.rem <- broom::tidy(sum_model_std) # Create a tidy dataframe for the model
sum.rem <- sum.rem %>% filter(component != "zi")

# Add a column to identify the model
sum.rem$model <- "sum_model_std" 

dw_sum_rem <- dwplot(sum.rem) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "perc.infested_std" = "Percent Infested (neighborhood)",
               "lag_perc.infested" = "Lagged percent infested (neighborhood)",
               "dist_nearest_tree" = "Distance to nearest tree",
                "gouting_recode" = "Gouting",
                "brooms_nh" = "Broom count (neighborhood)",
                "mean_percent_live_crown_std" = "Mean percent live crown (neighborhood)",
                "logdbh" = "DBH",
                "log_DISTANCE_noinfest" = "Distance to nearest uninfested tree > 2.54cm"
      ),
    limits = c( "logDISTANCE",
                "mean_percent_live_crown_std",
                "lag_perc.infested",
                "brooms_nh",
                "log_DISTANCE_noinfest",
                "logdbh",
                "gouting_recode",
                "perc.infested_std"
                ) # Specify the desired order
  )


print(dw_sum_rem+geom_vline(xintercept=0,lty=2))
```

# Main infestation status workflow again!!

```{r}
data$mean_percent_live_crown_std <- (data$mean_percent_live_crown - mean(data$mean_percent_live_crown, na.rm = TRUE))/sd(data$mean_percent_live_crown, na.rm = TRUE)

cor(data_long[, c("logdbh", "DISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new", "mean_dist_infested", "dist_nearest_tree", "basal_area_host_std", "avg.dbh", "avg.gouting", "mean_percent_live_crown", "n")])

# Infestation sum and number of infested come out as 0.88 which may be close enough to just use one if considered in final model

```

# Building with long dataset

## Infestation final - data_long

- I've been building these models based on first visit and then putting in visits, why haven't I built them with the long dataset and use all observations at all times, idk! Let's do it

```{r}

data_long <- data_long %>% filter(`Dead / Alive` == 1)

tree.re.glmm <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + beetle_recode + (1|site) + (1|unique_ID), data = data_long, family = binomial(link = "logit"))

tree.re.sims <- simulateResiduals(tree.re.glmm)
plot(tree.re.sims)

summary(tree.re.glmm)
summary(infest.final)

residrecal <- recalculateResiduals(tree.re.sims, group = data_long$t)
testTemporalAutocorrelation(residrecal, time = unique(data_long$t))
acf(residuals(tree.re.glmm))
```

# Autocorrelation structures on long dataset

```{r}
# temporal autocorrelated structure
data_long$t <- as.factor(data_long$t)
data_long_no_beetle <- data_long %>% filter(beetle_recode == 0)


tree.re.glmm.ar <- glmmTMB(cbind(infested, 1 - infested) ~ -1 + logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + 
    ar1(0 + t | unique_ID) # autocorrelation term
  + (1|site), data = data_long_no_beetle, family = binomial(link = "logit"))

tree.re.sims.ar <- simulateResiduals(tree.re.glmm.ar)
plot(tree.re.sims.ar)
residrecal.ar <- recalculateResiduals(tree.re.sims.ar, group = data_long$t)
testTemporalAutocorrelation(residrecal.ar, time = unique(data_long$t))
acf(residuals(tree.re.glmm.ar))

summary(tree.re.glmm.ar)

tidy_main_long <- broom::tidy(tree.re.glmm.ar) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_main_long$model <- "Infestation - AR" 

dw_main <- dwplot(tidy_main_long) +
  scale_y_discrete(
    labels = c("logDISTANCE" = "Distance to nearest infested",
               "infest_sum_std_new" = "Infestation sum (neighborhood)",
                "Gouting1" = "Gouting - 1",
                "Gouting2" = "Gouting - 2",
                "Gouting3" = "Gouting - 3",
                "beetle_recode" = "Beetle - presence/absence",
                "avg.dbh" = "Average DBH in neighborhood",
                "avg.gouting" = "Average Gouting",
                "logdbh" = "DBH"),
    limits = c( "avg.dbh",
                "avg.gouting",
                "beetle_recode",
                "Gouting3",
                "Gouting2",
                "Gouting1",
                "infest_sum_std_new",
                "logDISTANCE", 
                "logdbh") # Specify the desired order
  )
print(dw_main+geom_vline(xintercept=0,lty=2))

```

# Site variograms

```{r}

# spatial autocorrelation structure

emig_long <- data_long %>% filter(site == "EMIG")

tree.re.glmm.emig <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + gouting_recode + infest_sum_std_new + avg.dbh + dist_nearest_tree + (1|unique_ID) + (1|site),
    #beetle_recode + - rank-deficient in this model
   data = emig_long, family = binomial(link = "logit"))

residuals.emig <- residuals(tree.re.glmm.emig, type = "pearson")

residual_data <- data.frame(
  residuals = residuals.emig,
  x = emig_long$Longitude,
  y = emig_long$Latitude
)

gstat_obj <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data
)

# Calculate the empirical semivariogram
semivariogram <- variogram(gstat_obj)

plot(semivariogram, main = "Semivariogram of GLMM Residuals - Emigration Canyon - all visits")

# Trying one visit to see difference

emig_v1 <- data_long %>% filter(site == "EMIG" & t == 1)

tree.re.glmm.emig.v1 <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + mean_percent_live_crown_std + n + perc.infested + (1|unique_ID),
    #beetle_recode + - rank-deficient in this model
   data = emig_v1, family = binomial(link = "logit"))

residuals.emig.v1 <- residuals(tree.re.glmm.emig.v1, type = "pearson")

residual_data.v1 <- data.frame(
  residuals = residuals.emig.v1,
  x = emig_v1$Longitude,
  y = emig_v1$Latitude
)

gstat_obj.v1 <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.v1
)

# Calculate the empirical semivariogram
semivariogram.v1 <- variogram(gstat_obj.v1)

plot(semivariogram.v1, main = "Semivariogram of GLMM Residuals - Emigration Canyon - visit 1")

# Testing all sites at one visit

all_v1 <- data_long %>% filter(t == 1)

tree.re.glmm.all.v1 <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + mean_percent_live_crown_std + n + perc.infested + (1|unique_ID),
    #beetle_recode + - rank-deficient in this model
   data = all_v1, family = binomial(link = "logit"))

residuals.all.v1 <- residuals(tree.re.glmm.all.v1,  type = "pearson")

residual_data.v1 <- data.frame(
  residuals = residuals.all.v1,
  x = all_v1$Longitude,
  y = all_v1$Latitude
)

gstat_obj.v1 <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.v1
)

# Calculate the empirical semivariogram
semivariogram.v1.all <- variogram(gstat_obj.v1)

plot(semivariogram.v1.all, main = "Semivariogram of GLMM Residuals - all sites - visit 1")

# Testing all sites at all visits

tree.re.glmm.all <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + mean_percent_live_crown_std + n + perc.infested + (1|unique_ID),
    #beetle_recode + - rank-deficient in this model
   data = data_long, family = binomial(link = "logit"))

residuals.all <- residuals(tree.re.glmm.all,  type = "pearson")

residual_data <- data.frame(
  residuals = residuals.all,
  x = data_long$Longitude,
  y = data_long$Latitude
)

gstat_obj.all <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data
)

# Calculate the empirical semivariogram
semivariogram.all <- variogram(gstat_obj.all)

plot(semivariogram.all, main = "Semivariogram of GLMM Residuals - all sites - all visits")

# Franklin Basin

frnk_long <- data_long %>% filter(site == "FRNK")

tree.re.glmm.ar.frnk <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown_std + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = frnk_long, family = binomial(link = "logit"))

residuals.frnk <- residuals(tree.re.glmm.ar.frnk)

residual_data.frnk <- data.frame(
  residuals = residuals.frnk,
  x = frnk_long$Latitude,
  y = frnk_long$Longitude
)

gstat_obj.frnk <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.frnk
)

# Calculate the empirical semivariogram
semivariogram.frnk <- variogram(gstat_obj.frnk)

plot(semivariogram.frnk, main = "Semivariogram of GLMM Residuals - Franklin Basin")

# Monte Cristo

mocr_long <- data_long %>% filter(site == "MOCR")

tree.re.glmm.ar.mocr <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = mocr_long, family = binomial(link = "logit"))

residuals.mocr <- residuals(tree.re.glmm.ar.mocr)

residual_data.mocr <- data.frame(
  residuals = residuals.mocr,
  x = mocr_long$Latitude,
  y = mocr_long$Longitude
)

gstat_obj.mocr <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.mocr
)

# Calculate the empirical semivariogram
semivariogram.mocr <- variogram(gstat_obj.mocr)

plot(semivariogram.mocr, main = "Semivariogram of GLMM Residuals - Monte Cristo")

# Mount Timp

mttp_long <- data_long %>% filter(site == "MTTP")

tree.re.glmm.ar.mttp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = mttp_long, family = binomial(link = "logit"))

residuals.mttp <- residuals(tree.re.glmm.ar.mttp)

residual_data.mttp <- data.frame(
  residuals = residuals.mttp,
  x = mttp_long$Latitude,
  y = mttp_long$Longitude
)

gstat_obj.mttp <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.mttp
)

# Calculate the empirical semivariogram
semivariogram.mttp <- variogram(gstat_obj.mttp)

plot(semivariogram.mttp, main = "Semivariogram of GLMM Residuals - Mount Timpanogos")

# Smith and Morehouse

smmh_long <- data_long %>% filter(site == "SMMH")

tree.re.glmm.ar.smmh <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = smmh_long, family = binomial(link = "logit"))

residuals.smmh <- residuals(tree.re.glmm.ar.smmh)

residual_data.smmh <- data.frame(
  residuals = residuals.smmh,
  x = smmh_long$Latitude,
  y = smmh_long$Longitude
)

gstat_obj.smmh <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.smmh
)

# Calculate the empirical semivariogram
semivariogram.smmh <- variogram(gstat_obj.smmh)

plot(semivariogram.smmh, main = "Semivariogram of GLMM Residuals - Smith and Morehouse")

# Soapstone

soap_long <- data_long %>% filter(site == "SOAP")

tree.re.glmm.ar.soap <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = soap_long, family = binomial(link = "logit"))

residuals.soap <- residuals(tree.re.glmm.ar.soap)

residual_data.soap <- data.frame(
  residuals = residuals.soap,
  x = soap_long$Latitude,
  y = soap_long$Longitude
)

gstat_obj.soap <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.soap
)

# Calculate the empirical semivariogram
semivariogram.soap <- variogram(gstat_obj.soap)

# Templr Fork

tmpf_long <- data_long %>% filter(site == "TMPF")

tree.re.glmm.ar.tmpf <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = tmpf_long, family = binomial(link = "logit"))

residuals.tmpf <- residuals(tree.re.glmm.ar.tmpf)

residual_data.tmpf <- data.frame(
  residuals = residuals.tmpf,
  x = tmpf_long$Latitude,
  y = tmpf_long$Longitude
)

gstat_obj.tmpf <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.tmpf
)

# Calculate the empirical semivariogram
semivariogram.tmpf <- variogram(gstat_obj.tmpf)

# Cox Springs

cxsp_long <- data_long %>% filter(site == "CXSP")

tree.re.glmm.ar.cxsp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + avg.dbh + Gouting + infest_sum_std_new + avg.gouting + mean_percent_live_crown + 
    beetle_recode +
    ar1(0 + t | unique_ID), # autocorrelation term
   data = cxsp_long, family = binomial(link = "logit"))

residuals.cxsp <- residuals(tree.re.glmm.ar.cxsp)

residual_data.cxsp <- data.frame(
  residuals = residuals.cxsp,
  x = cxsp_long$Latitude,
  y = cxsp_long$Longitude
)

gstat_obj.cxsp <- gstat(
  id = "residuals",
  formula = residuals ~ 1,
  locations = ~x + y,
  data = residual_data.cxsp
)

# Calculate the empirical semivariogram
semivariogram.cxsp <- variogram(gstat_obj.cxsp)


plot(semivariogram.cxsp, main = "Semivariogram of GLMM Residuals - Cox Springs")
plot(semivariogram.tmpf, main = "Semivariogram of GLMM Residuals - Temple Fork")
plot(semivariogram.smmh, main = "Semivariogram of GLMM Residuals - Smith and Morehouse")
plot(semivariogram.soap, main = "Semivariogram of GLMM Residuals - Soapstone Basin")
plot(semivariogram.mttp, main = "Semivariogram of GLMM Residuals - Mount Timpanogos")
plot(semivariogram.mocr, main = "Semivariogram of GLMM Residuals - Monte Cristo")
plot(semivariogram.frnk, main = "Semivariogram of GLMM Residuals - Franklin Basin")
plot(semivariogram, main = "Semivariogram of GLMM Residuals - Emigration Canyon")

summary(tree.re.glmm.ar.cxsp)
summary(tree.re.glmm.ar.emig)
summary(tree.re.glmm.ar.frnk)
summary(tree.re.glmm.ar.mttp)
summary(tree.re.glmm.ar.mocr)
summary(tree.re.glmm.ar.smmh)
summary(tree.re.glmm.ar.soap)
summary(tree.re.glmm.ar.tmpf)

# FRNK and MTTPare showinf NANs, let's investigate
# appears something wrong with avg.dbh

cor(frnk_long[, c("logDISTANCE", "logdbh", "brooms_nh", "infest_sum_std_new", "avg.dbh", "avg.gouting", "mean_percent_live_crown_std")])

step_model.cxsp <- step(tree.re.glmm.ar.cxsp, direction = "both", trace = TRUE)
step_model.emig <- step(tree.re.glmm.ar.emig, direction = "both", trace = TRUE)
step_model.frnk <- step(tree.re.glmm.ar.frnk, direction = "both", trace = TRUE)
step_model.mocr <- step(tree.re.glmm.ar.mocr, direction = "both", trace = TRUE)
step_model.mttp <- step(tree.re.glmm.ar.mttp, direction = "both", trace = TRUE)
step_model.smmh <- step(tree.re.glmm.ar.smmh, direction = "both", trace = TRUE)
step_model.soap <- step(tree.re.glmm.ar.soap, direction = "both", trace = TRUE)
step_model.tmpf <- step(tree.re.glmm.ar.tmpf, direction = "both", trace = TRUE)

summary(step_model.cxsp)
summary(step_model.tmpf)
summary(step_model.smmh)
summary(step_model.soap)
summary(step_model.mttp)
summary(step_model.mocr)
summary(step_model.emig)
summary(step_model.frnk)
```

## Infestation - clump

```{r}
clump_long <- filter(data_long, dbh == 1.27)

cor(clump_long[, c("logheight", "DISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "avg.dbh", "avg.gouting", "mean_percent_live_crown", "n")])

## Building model below

 clump.base <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.base.sims <- simulateResiduals(clump.base)
plot(clump.base.sims)

summary(clump.base)

# Add brooms

clump.broom <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.broom.sims <- simulateResiduals(clump.broom)
plot(clump.broom.sims)

summary(clump.broom)

clump.broom.nh <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + brooms_nh + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.broom.sims <- simulateResiduals(clump.broom.nh)
plot(clump.broom.sims)

summary(clump.broom.nh)

# Add beetle

clump.beetle <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + beetle_recode + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.beetle.sims <- simulateResiduals(clump.beetle)
plot(clump.beetle.sims)

summary(clump.beetle)

# Add infest sum

clump.sum <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + infest_sum_std_new + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.sum.sims <- simulateResiduals(clump.sum)
plot(clump.sum.sims)

summary(clump.sum)

# percent infested

clump.pinfest <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + infest_sum_std_new + perc.infested + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.pinfest.sims <- simulateResiduals(clump.pinfest)
plot(clump.pinfest.sims)

summary(clump.pinfest)

# distance to nearest tree

clump.dist.nearinfest <- glmmTMB(cbind(infested, 1 - infested) ~ Gouting + infest_sum_std_new + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.dist.nearinfest.sims <- simulateResiduals(clump.dist.nearinfest)
plot(clump.dist.nearinfest.sims)

summary(clump.dist.nearinfest)

# avg. dbh in neighborhood

clump.dbh.infest <- glmmTMB(cbind(infested, 1 - infested) ~ Gouting + infest_sum_std_new + avg.dbh + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.dbh.infest.sims <- simulateResiduals(clump.dbh.infest)
plot(clump.dbh.infest.sims)

summary(clump.dbh.infest)

# Building model above
clump.final.long <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + brooms_recode + (1|unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.final.sims <- simulateResiduals(clump.final.long)
plot(clump.final.sims)

summary(clump.final.long)

testZeroInflation(clump.final.sims)

```

## Infestation - large

```{r}
large_long <- filter(data_long, dbh != 1.27)

cor(large_long[, c("logheight", "DISTANCE", "brooms_recode", "brooms_nh", "beetle_recode", "infest_sum_std_new", "dist_nearest_tree", "basal_area_host_std", "avg.dbh", "avg.gouting", "mean_percent_live_crown", "n")])

## Building model below

 large.base <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.base.sims <- simulateResiduals(large.base)
plot(large.base.sims)

summary(large.base)

# Add brooms

large.broom <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.broom.sims <- simulateResiduals(large.broom)
plot(large.broom.sims)

summary(large.broom)

large.broom.nh <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + brooms_nh + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.broom.sims <- simulateResiduals(large.broom.nh)
plot(large.broom.sims)

summary(large.broom.nh)

# Add beetle

large.beetle <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + beetle_recode + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.beetle.sims <- simulateResiduals(large.beetle)
plot(large.beetle.sims)

summary(large.beetle)

# Add infest sum

large.sum <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + infest_sum_std_new + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.sum.sims <- simulateResiduals(large.sum)
plot(large.sum.sims)

summary(large.sum)

# distance to nearest tree

large.dist.nearinfest <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.dist.nearinfest.sims <- simulateResiduals(large.dist.nearinfest)
plot(large.dist.nearinfest.sims)

summary(large.dist.nearinfest)

# avg. dbh in neighborhood

large.dbh.infest <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.dbh.infest.sims <- simulateResiduals(large.dbh.infest)
plot(large.dbh.infest.sims)

summary(large.dbh.infest)

# Building model above
large.final.long <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + logdbh + Gouting + infest_sum_std_new + avg.dbh + dist_nearest_tree + (1|unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.final.sims <- simulateResiduals(large.final.long)
plot(large.final.sims)

summary(large.final.long)

testZeroInflation(large.final.sims)

```

# Long data infestation status beta coefficient plot

```{r}
tidy_large_long <- broom::tidy(large.final.long) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_large_long$model <- "large.long" 

dw_large_long <- dwplot(tidy_large_long)

tidy_small_long <- broom::tidy(clump.final.long)

tidy_small_long$model <- "clump.long"

dw_small_long <- dwplot(tidy_small_long)

tidy_all_long <- rbind(tidy_large_long, tidy_small_long)

dw_all_long <- dwplot(tidy_all_long) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "infest_sum_std_new" = "Infestation sum (neighborhood)",
               "dist_nearest_tree" = "Distance to nearest tree",
                "Gouting1" = "Gouting - 1",
                "Gouting2" = "Gouting - 2",
                "Gouting3" = "Gouting - 3",
                "brooms_recode" = "Brooms - presence/absence",
                "avg.dbh" = "Average dbh in neighborhood" ,
                "logdbh" = "DBH"
      ),
    limits = c("dist_nearest_tree",
               "avg.dbh",
                "Gouting3",
                "Gouting2",
                "Gouting1",
                "logdbh",
                "infest_sum_std_new",
                "brooms_recode",
                "logDISTANCE"
                ) # Specify the desired order
  )


print(dw_all_long+geom_vline(xintercept=0,lty=2))
   
```

# Large infestation - AR Structure

```{r}
## Building model below

large_long$t <- as.factor(large_long$t)

 large.base <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.base.sims <- simulateResiduals(large.base)
plot(large.base.sims)

summary(large.base)

# Add brooms

large.broom <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + brooms_recode + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.broom.sims <- simulateResiduals(large.broom)
plot(large.broom.sims)

summary(large.broom)

large.broom.nh <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + brooms_nh + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.broom.sims <- simulateResiduals(large.broom.nh)
plot(large.broom.sims)

summary(large.broom.nh)

# Add beetle

large.beetle.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + brooms_recode + beetle_recode + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.beetle.sims.ar <- simulateResiduals(large.beetle.ar)
plot(large.beetle.sims.ar)

summary(large.beetle.ar)

# Add infest sum

large.sum.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.sum.sims.ar <- simulateResiduals(large.sum.ar)
plot(large.sum.sims.ar)

summary(large.sum.ar)

# distance to nearest tree

large.dist.near.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + dist_nearest_tree + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.dist.near.sims.ar <- simulateResiduals(large.dist.near.ar)
plot(large.dist.near.sims.ar)

summary(large.dist.near.ar)

# avg. dbh in neighborhood

large.dbh.infest.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + dist_nearest_tree + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.dbh.infest.sims.ar <- simulateResiduals(large.dbh.infest.ar)
plot(large.dbh.infest.sims.ar)

summary(large.dbh.infest.ar)

# Building model above
large.final.long.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + logdbh + Gouting + infest_sum_std_new + avg.dbh + dist_nearest_tree + ar1(0 + t | unique_ID) + (1|site), data = large_long, family = binomial(link = "logit"))

large.final.sims.ar <- simulateResiduals(large.final.long.ar)
plot(large.final.sims.ar)

summary(large.final.long.ar)

tidy_large_long.ar <- broom::tidy(large.final.long.ar) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_large_long.ar$model <- "large.long.ar" 

dw_large_ar <- dwplot(tidy_large_long.ar) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "infest_sum_std_new" = "Infestation sum (neighborhood)",
               "dist_nearest_tree" = "Distance to nearest tree",
                "Gouting1" = "Gouting - 1",
                "Gouting2" = "Gouting - 2",
                "Gouting3" = "Gouting - 3",
                "avg.dbh" = "Average dbh in neighborhood" ,
                "logdbh" = "DBH"
      ),
    limits = c("dist_nearest_tree",
               "avg.dbh",
                "Gouting3",
                "Gouting2",
                "Gouting1",
                "logdbh",
                "infest_sum_std_new",
                "logDISTANCE"
                ) # Specify the desired order
  )


print(dw_large_ar+geom_vline(xintercept=0,lty=2))
```

# Clump infestation status - AR structure

```{r}
## Building model below

clump_long$t <- as.factor(clump_long$t)

 clump.base <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.base.sims <- simulateResiduals(clump.base)
plot(clump.base.sims)

summary(clump.base)

# Add brooms

clump.broom <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.broom.sims <- simulateResiduals(clump.broom)
plot(clump.broom.sims)

summary(clump.broom)

clump.broom.nh <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + brooms_nh + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.broom.sims <- simulateResiduals(clump.broom.nh)
plot(clump.broom.sims)

summary(clump.broom.nh)

# Add beetle

clump.beetle.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + brooms_recode + beetle_recode + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.beetle.sims.ar <- simulateResiduals(clump.beetle.ar)
plot(clump.beetle.sims.ar)

summary(clump.beetle.ar)

# Add infest sum

clump.sum.ar <- glmmTMB(cbind(infested, 1 - infested) ~logdbh + logDISTANCE + Gouting + infest_sum_std_new + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.sum.sims.ar <- simulateResiduals(clump.sum.ar)
plot(clump.sum.sims.ar)

summary(clump.sum.ar)

# distance to nearest tree

clump.dist.near.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + dist_nearest_tree + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.dist.near.sims.ar <- simulateResiduals(clump.dist.near.ar)
plot(clump.dist.near.sims.ar)

summary(clump.dist.near.ar)

# avg. dbh in neighborhood

clump.dbh.infest.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + dist_nearest_tree + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.dbh.infest.sims.ar <- simulateResiduals(clump.dbh.infest.ar)
plot(clump.dbh.infest.sims.ar)

summary(clump.dbh.infest.ar)

# Building model above
clump.final.long.ar <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + Gouting + infest_sum_std_new + avg.dbh + dist_nearest_tree + ar1(0 + t | unique_ID) + (1|site), data = clump_long, family = binomial(link = "logit"))

clump.final.sims.ar <- simulateResiduals(clump.final.long.ar)
plot(clump.final.sims.ar)

summary(clump.final.long.ar)
```


# Infestation Sum - AR structure

```{r}
sum.base.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.base.sims.ar <- simulateResiduals(sum.base.ar)
plot(sum.base.sims.ar)

summary(sum.base.ar)

# Add gouting

sum.gout.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.gout.sims.ar <- simulateResiduals(sum.gout.ar)
plot(sum.gout.sims.ar)

summary(sum.gout.ar)

# Add brooms

sum.broom.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + brooms_recode + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2)

sum.broom.sims.ar <- simulateResiduals(sum.broom.ar)
plot(sum.broom.sims.ar)

summary(sum.broom.ar)

# Brooms - nh

sum.broom.nh.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + brooms_nh + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2(link = "log"))

sum.broom.sims.ar <- simulateResiduals(sum.broom.nh.ar)
plot(sum.broom.sims.ar)

summary(sum.broom.nh.ar)

# gout - nbinom1 - still performing bad, maybe worse

sum.gout.ar.1 <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + logheight + Gouting + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom1(link = "log"))

sum.gout.sims.ar.1 <- simulateResiduals(sum.gout.ar.1)
plot(sum.gout.sims.ar.1)

summary(sum.gout.ar.1)

# Add beetle

sum.beetle.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + beetle_recode + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2(link = "log"))

sum.beetle.sims.ar <- simulateResiduals(sum.beetle.ar)
plot(sum.beetle.sims.ar)

summary(sum.beetle.ar)

# Add infest sum

sum.sum.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2(link = "log"))

sum.sum.sims.ar <- simulateResiduals(sum.sum.ar)
plot(sum.sum.sims.ar)

summary(sum.sum.ar)

# number in neighborhood

sum.n.nhinfest.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + n + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2(link = "log"))

sum.n.nhinfest.sims.ar <- simulateResiduals(sum.n.nhinfest.ar)
plot(sum.n.nhinfest.sims.ar)

summary(sum.n.nhinfest.ar)

# add basal area host

sum.ba_host.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + n + basal_area_host_std + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2(link = "log"))

sum.ba_host.sims.ar <- simulateResiduals(sum.ba_host.ar)
plot(sum.ba_host.sims.ar)

summary(sum.ba_host.ar)

# Average DBH

sum.avg.dbh.ar <- glmmTMB(infestation_sum ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + n + avg.dbh + ar1(0 + t | unique_ID) + (1|site), data = data_long, family = nbinom2(link = "log"))

sum.avg.dbh.sims.ar <- simulateResiduals(sum.avg.dbh.ar)
plot(sum.avg.dbh.sims.ar)

summary(sum.avg.dbh.ar)

# distance to nearest tree
sum.dist.ar <- glmmTMB(infestation_sum ~ infest_sum_std_new + dist_nearest_tree + Gouting + logDISTANCE + logdbh + n + avg.dbh + (1|site) + ar1(0 + t | unique_ID), data = data_long, family = nbinom2(link = "log"))

sum.dist.sims.ar <- simulateResiduals(sum.dist.ar)
plot(sum.dist.sims.ar)

summary(sum.dist.ar)

# Final

sum.final.ar <- glmmTMB(infestation_sum ~ infest_sum_std_new + dist_nearest_tree + Gouting + logDISTANCE + logdbh + avg.dbh + (1|site) + ar1(0 + t | unique_ID), data = data_long, family = nbinom2(link = "log"))

sum.final.sims.ar <- simulateResiduals(sum.final.ar)
plot(sum.final.sims.ar)

summary(sum.final.ar)
```

# New infestation - AR structure

```{r}

## Building model below

new_long$t <- as.factor(new_long$t)

# Start a-modelling

new.base.ar <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.base.sims.ar <- simulateResiduals(new.base.ar)
plot(new.base.sims.ar)

summary(new.base.ar)

# Add gouting

new.gout <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.gout.sims <- simulateResiduals(new.gout)
plot(new.gout.sims)

summary(new.gout)

# Add brooms

new.broom <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + brooms_recode + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.broom.sims <- simulateResiduals(new.broom)
plot(new.broom.sims)

summary(new.broom)

# Brooms - nh
new.broom.nh <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + brooms_nh + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.broom.sims <- simulateResiduals(new.broom.nh)
plot(new.broom.sims)

summary(new.broom.nh)


# Add beetle

new.beetle <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + brooms_nh + beetle_recode + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.beetle.sims <- simulateResiduals(new.beetle)
plot(new.beetle.sims)

summary(new.beetle)

# Add infest sum

new.sum.ar <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode + infest_sum_std_new + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.sum.sims.ar <- simulateResiduals(new.sum.ar)
plot(new.sum.sims.ar)

summary(new.sum.ar)

# percent infested

new.pinfest.ar <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + perc.infested + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.pinfest.sims.ar <- simulateResiduals(new.pinfest.ar)
plot(new.pinfest.sims.ar)

summary(new.pinfest.ar)

# number in neighborhood

new.n.nhinfest.ar <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + infest_sum_std_new + perc.infested + n + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.n.nhinfest.sims.ar <- simulateResiduals(new.n.nhinfest.ar)
plot(new.n.nhinfest.sims.ar)

summary(new.n.nhinfest.ar)

# add basal area host

new.ba_host.ar <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting  + infest_sum_std_new + perc.infested + n + basal_area_host_std + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))

new.ba_host.sims.ar <- simulateResiduals(new.ba_host.ar)
plot(new.ba_host.sims.ar)

summary(new.ba_host.ar)

# Final Model!

new.final.ar <- glmmTMB(cbind(new_infestation, 1 - new_infestation) ~ logdbh + logDISTANCE + Gouting + beetle_recode + infest_sum_std_new + avg.dbh + ar1(0 + t | unique_ID) + (1|site), data = new_long, family = binomial(link = "logit"))
 
new.final.sims.ar <- simulateResiduals(new.final.ar)
plot(new.final.sims.ar)
 
summary(new.final.ar)
```


# Full-factorial for neighborhood effect

So in our final infestation model, we have average dbh and infestation sum that come out as significantly impacting probability of infestation. What we want to do is see how different distances affect this probability. We'll make a dataframe with all the combinations of average dbh and infest sum over a multidude of distances and then run those combinations in the model we've built. What we'll get is an idea of how these neighborhoods interact with each other and try to get the best model out of it.

```{r}

# Create a list of all data frames
local_list <- list(
  local_1, local_2, local_4, local_6, local_8, local_10,
  local_12, local_14, local_16, local_18, local_20
)

# Make template
template <- Reduce(function(x, y) merge(x, y, all = TRUE), lapply(local_list, function(df) df["unique_ID"]))

# align data
aligned_local_list <- lapply(local_list, function(df) merge(template, df, by = "unique_ID", all.x = TRUE))

# Keep as separate data frames
names(aligned_local_list) <- c("local_1", "local_2", "local_4", "local_6", "local_8", 
                               "local_10", "local_12", "local_14", "local_16", "local_18", "local_20")

# New local data - same lengths

local_data_20_new <- data %>% inner_join(aligned_local_list$local_20, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")

 local_data_18_new <- data %>% inner_join(aligned_local_list$local_18, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_16_new <- data %>% inner_join(aligned_local_list$local_16, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_14_new <- data %>% inner_join(aligned_local_list$local_14, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_12_new <- data %>% inner_join(aligned_local_list$local_12, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_10_new <- data %>% inner_join(aligned_local_list$local_10, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
local_data_8_new <- data %>% inner_join(aligned_local_list$local_8,by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_6_new <- data %>% inner_join(aligned_local_list$local_6, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_4_new <- data %>% inner_join(aligned_local_list$local_4, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_2_new <- data %>% inner_join(aligned_local_list$local_2, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")
 local_data_1_new <- data %>% inner_join(aligned_local_list$local_1, by = "unique_ID") %>% filter(species == "ABLA" | species == "ABCO")

# List of avg.dbh vectors
avg.dbh <- data.frame(
  avg.dbh_1 = local_data_1_new$avg.dbh.y,
  avg.dbh_2 = local_data_2_new$avg.dbh.y,
  avg.dbh_4 = local_data_4_new$avg.dbh.y,
  avg.dbh_6 = local_data_6_new$avg.dbh.y,
  avg.dbh_8 = local_data_8_new$avg.dbh.y,
  avg.dbh_10 = local_data_10_new$avg.dbh.y,
  avg.dbh_12 = local_data_12_new$avg.dbh.y,
  avg.dbh_14 = local_data_14_new$avg.dbh.y,
  avg.dbh_16 = local_data_16_new$avg.dbh.y,
  avg.dbh_18 = local_data_18_new$avg.dbh.y,
  avg.dbh_20 = local_data_20_new$avg.dbh.y
)

# same for infest sum
infest_sum_std_new_v1 <- data.frame(
  infest_sum_std_new_v1_1 = local_data_1_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_2 = local_data_2_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_4 = local_data_4_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_6 = local_data_6_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_8 = local_data_8_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_10 = local_data_10_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_12 = local_data_12_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_14 = local_data_14_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_16 = local_data_16_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_18 = local_data_18_new$infest_sum_std_new_v1,
  infest_sum_std_new_v1_20 = local_data_20_new$infest_sum_std_new_v1
)

  
combinations <- expand.grid(var1 = infest_sum_std_new_v1, var2 = avg.dbh)

```



## Building infestation status model with weights

```{r}
# # update neighborhood metrics with spatial weights
# model_data <- sm_with_proportions
# model_data$infest_sum_std_new <- model_data$infest_sum_std_new * model_data$proportion
# model_data$avg.dbh <- model_data$avg.dbh * model_data$proportion
# model_data$mean_percent_live_crown_std <- model_data$mean_percent_live_crown_std * model_data$proportion
# model_data$brooms_nh <- model_data$brooms_nh * model_data$proportion
# model_data$perc.infested <- model_data$perc.infested * model_data$proportion
# model_data$basal_area_host_std <- model_data$basal_area_host_std * model_data$proportion
# model_data$mean_dist_infested_recode <- model_data$mean_dist_infested_recode * model_data$proportion
# model_data$n <- model_data$n * model_data$proportion
# 
# # Filter live
# model_data$`Dead / Alive`[is.na(model_data$`Dead / Alive`)] <- 0
# model_data <- filter(model_data, model_data$`Dead / Alive` == 1)

cor(model_data[, c("logDISTANCE", "lag_logDISTANCE", "infest_sum_std_new","lag_infested", "lag_perc.infested",  "lag_infest_sum_std_new", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "avg.dbh", "gouting_recode", "avg.gouting", "mean_percent_live_crown", "n", "comp_index")])

# Scale coordinates
model_data$x_scaled <- scale(model_data$Longitude)  
model_data$y_scaled <- scale(model_data$Latitude)  
model_data$spatial <- paste0(model_data$x_scaled, "_", model_data$y_scaled)

# Build model

# Standardize variables
model_data$logdbh_std <- scale(model_data$logdbh)
model_data$logDISTANCE_std <- scale(model_data$logDISTANCE)
model_data$log_DISTANCE_noinfest_std <- scale(model_data$log_DISTANCE_noinfest)
model_data$lag_infested_std <- scale(model_data$lag_infested)
model_data$n_std <- scale(model_data$n)
model_data$perc.infested_std <- scale(model_data$perc.infested)

prop.base <- glmmTMB(
  cbind(infested, 1 - infested) ~ 1 + exp(logdbh + logDISTANCE + gouting_recode + 1|unique_ID + 0|site),
  data = model_data,
  family = binomial(link = "logit")
)

prop.base.sims <- simulateResiduals(prop.base)
plot(prop.base.sims)

summary(prop.base)

# Add brooms

prop.broom.nh <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_nh + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.broom.sims <- simulateResiduals(prop.broom.nh)
plot(prop.broom.sims)

summary(prop.broom.nh)

# Add beetle

prop.beetle <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.beetle.sims <- simulateResiduals(prop.beetle)
plot(prop.beetle.sims)

summary(prop.beetle)

# Add infest sum

prop.sum <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + infest_sum_std_new + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.sum.sims <- simulateResiduals(prop.sum)
plot(prop.sum.sims)

summary(prop.sum)

# distance to nearest tree

prop.dist.nearinfest <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + infest_sum_std_new + dist_nearest_tree + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.dist.nearinfest.sims <- simulateResiduals(prop.dist.nearinfest)
plot(prop.dist.nearinfest.sims)

summary(prop.dist.nearinfest)

# avg. dbh in neighborhood

prop.dbh.infest <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + infest_sum_std_new + avg.dbh + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.dbh.infest.sims <- simulateResiduals(prop.dbh.infest)
plot(prop.dbh.infest.sims)

summary(prop.dbh.infest)

# lag infestation status

prop.lag.infest <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + infest_sum_std_new + + dist_nearest_tree + lag_infested + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.lag.infest.sims <- simulateResiduals(prop.lag.infest)
plot(prop.lag.infest.sims)

summary(prop.lag.infest)

# mean percent live crown

prop.crown.infest <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + infest_sum_std_new + + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.crown.infest.sims <- simulateResiduals(prop.crown.infest)
plot(prop.crown.infest.sims)

summary(prop.crown.infest)

# N

prop.n <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + infest_sum_std_new + + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.n.sims <- simulateResiduals(prop.n)
plot(prop.n.sims)

summary(prop.n)

# percent infested

prop.perc <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + perc.infested + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.perc.sims <- simulateResiduals(prop.perc)
plot(prop.perc.sims)

summary(prop.perc)

testSpatialAutocorrelation(prop.perc, x = model_data$Longitude, y = model_data$Latitude)

# avg gouting

prop.gouting <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + perc.infested + avg.gouting + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.gouting.sims <- simulateResiduals(prop.gouting)
plot(prop.gouting.sims)

summary(prop.gouting)

# Basal area host

prop.ba <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + perc.infested + basal_area_host_std + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.ba.sims <- simulateResiduals(prop.ba)
plot(prop.ba.sims)

summary(prop.ba)

# basal area all trees in neighborhood

prop.ba.all <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + perc.infested + basal_area_nh_std + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.ba.all.sims <- simulateResiduals(prop.ba.all)
plot(prop.ba.all.sims)

summary(prop.ba.all)

# distance to nearest non-infested and non-clump (replacing nearest tree variable)

prop.dist <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + log_DISTANCE_noinfest + lag_infested + mean_percent_live_crown_std + n + perc.infested + (1|unique_ID) + (1|site), data = model_data, family = binomial(link = "logit"))

prop.dist.sims <- simulateResiduals(prop.dist)
plot(prop.dist.sims)

summary(prop.dist)

# Building model above
prop.final.long <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + log_DISTANCE_noinfest + lag_infested + mean_percent_live_crown_std + n + perc.infested + comp_index + (1|site), data = model_data, family = binomial(link = "logit"))

prop.final.sims <- simulateResiduals(prop.final.long)
plot(prop.final.sims)

summary(prop.final.long)
r2(prop.final.long)
check_singularity(prop.final.long)

testZeroInflation(prop.final.sims)

r2(prop.final.long)

prop.final.infest <- broom::tidy(prop.final.long) # Create a tidy dataframe for the model

# Add a column to identify the model
prop.final.infest$model <- "prop.final.infest" 

dw_prop_final <- dwplot(prop.final.infest) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "perc.infested" = "Percent Infested (neighborhood)",
               "log_DISTANCE_noinfest" = "Distance to nearest non-infested tree > 2.54cm",
                "gouting_recode" = "Gouting",
                "lag_infested" = "Lagged infestation status",
                "mean_percent_live_crown_std" = "Mean percent live crown (neighborhood)",
                "n" = "Number of trees (neighborhood)" ,
                "logdbh" = "DBH"
      ),
    limits = c( "logDISTANCE",
                "mean_percent_live_crown_std",
                "n",
                "perc.infested",
                "log_DISTANCE_noinfest",
                "gouting_recode",
                "logdbh",
                "lag_infested"
                ) # Specify the desired order
  )


print(dw_prop_final+geom_vline(xintercept=0,lty=2))

```
### Bootstrapping on final status model

```{r}
detach("package:boot", unload = TRUE)
library(boot)

model_data <- as.data.frame(model_data)

boot_fn <- function(data, indices) {
  model_data <- data[indices, ]
  model <- glmmTMB(
   cbind(infested, 1 - infested) ~ logdbh_std + logDISTANCE_std + gouting_recode + log_DISTANCE_noinfest_std + lag_infested + mean_percent_live_crown_std + n_std + perc.infested_std + (1|site), data = model_data, family = binomial(link = "logit")
  )
  return(fixef(model)$cond)  # Extract fixed effects
}

# bootstrap

boot_results <- boot(model_data, boot_fn, R = 1000)

# Get coefficient variability
boot.ci(boot_results, type = "perc")


# Alternatively, manually define:
 coef_names <- c("Intercept", "logdbh_std", "logDISTANCE_std", "gouting_recode", "log_DISTANCE_noinfest_std", "lag_infested", "mean_percent_live_crown_std", "n_std", "perc.infested_std")

boot_ci_df <- data.frame(
  term = coef_names,
  estimate = colMeans(boot_results$t),  # Mean estimate
  lower = apply(boot_results$t, 2, quantile, probs = 0.025),  # 2.5% CI
  upper = apply(boot_results$t, 2, quantile, probs = 0.975)   # 97.5% CI
)

print(boot_ci_df)

ggplot(boot_ci_df, aes(x = reorder(term, estimate), y = estimate)) +
  geom_point(color = "red", size = 3) +  # Mean estimate
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "blue") +  # CI bars
  coord_flip() +  # Flip for better readability
  labs(title = "Bootstrap Confidence Intervals for Model Coefficients",
       x = "Predictor Variables", y = "Coefficient Estimate") +
  theme_minimal()
```


## Building Large and Small with spatial weights
###Small

```{r}
clump_sp <- filter(model_data, dbh == 1.27)

cor(clump_sp[, c("logDISTANCE", "lag_infested", "dist_nearest_tree", "basal_area_host_std", "perc.infested", "avg.dbh", "avg.gouting", "mean_percent_live_crown", "n")])

## Building model below

 clump.base.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.base.sims.sp <- simulateResiduals(clump.base.sp)
plot(clump.base.sims.sp)

summary(clump.base.sp)

# Add brooms

clump.broom.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + brooms_recode + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.broom.sims.sp <- simulateResiduals(clump.broom.sp)
plot(clump.broom.sims.sp)

summary(clump.broom.sp)

clump.broom.nh.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + brooms_recode + brooms_nh + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.broom.sims.sp <- simulateResiduals(clump.broom.nh.sp)
plot(clump.broom.sims.sp)

summary(clump.broom.nh.sp)

# Add beetle

clump.beetle.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + brooms_recode + beetle_recode + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.beetle.sims.sp <- simulateResiduals(clump.beetle.sp)
plot(clump.beetle.sims.sp)

summary(clump.beetle.sp)

# Add perc. infested

clump.sum.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + brooms_recode + perc.infested + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.sum.sims.sp <- simulateResiduals(clump.sum.sp)
plot(clump.sum.sims.sp)

summary(clump.sum.sp)

# distance to nearest tree

clump.dist.nearinfest.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + perc.infested + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.dist.nearinfest.sims.sp <- simulateResiduals(clump.dist.nearinfest.sp)
plot(clump.dist.nearinfest.sims.sp)

summary(clump.dist.nearinfest.sp)

# avg. dbh in neighborhood

clump.dbh.infest.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + perc.infested + avg.dbh + brooms_recode + dist_nearest_tree + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.dbh.infest.sims.sp <- simulateResiduals(clump.dbh.infest.sp)
plot(clump.dbh.infest.sims.sp)

summary(clump.dbh.infest.sp)

# mean percent live crown

clump.crown.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + perc.infested + mean_percent_live_crown + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.crown.sims.sp <- simulateResiduals(clump.crown.sp)
plot(clump.crown.sims.sp)

summary(clump.crown.sp)

# lag infested status

clump.lag.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + perc.infested + lag_infested + mean_percent_live_crown + avg.dbh + dist_nearest_tree + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.lag.sims.sp <- simulateResiduals(clump.lag.sp)
plot(clump.lag.sims.sp)

summary(clump.lag.sp)

# n

clump.n.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + perc.infested + lag_infested + dist_nearest_tree + n + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.n.sims.sp <- simulateResiduals(clump.n.sp)
plot(clump.n.sims.sp)

summary(clump.n.sp)

# Building model above
clump.final.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logDISTANCE + gouting_recode + perc.infested + lag_infested + dist_nearest_tree + (1|unique_ID) + (1|site), data = clump_sp, family = binomial(link = "logit"))

clump.final.sp.sims <- simulateResiduals(clump.final.sp)
plot(clump.final.sp.sims)

summary(clump.final.sp)

testZeroInflation(clump.final.sp.sims)

r2(clump.final.sp)
check_singularity(clump.final.sp)

```

###Large
```{r}

large_sp <- filter(model_data, dbh != 1.27)
# Build model

base.large.sp <- glmmTMB(
  cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + (1|unique_ID) + (1|site),
  data = large_sp,
  family = binomial(link = "logit")
)

base.large.sp.sims <- simulateResiduals(base.large.sp)
plot(base.large.sp.sims)

summary(base.large.sp)

# gouting

gout.large.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

gout.large.sp.sims <- simulateResiduals(gout.large.sp)
plot(gout.large.sp.sims)

summary(gout.large.sp)

# Add brooms (focal and neighborhood)

broom.large.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_recode + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

broom.large.sp.sims <- simulateResiduals(broom.large.sp)
plot(broom.large.sp.sims)

summary(broom.large.sp)

broom.nh.large.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + brooms_nh + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

broom.nh.large.sp.sims <- simulateResiduals(broom.nh.large.sp)
plot(broom.nh.large.sp.sims)

summary(broom.nh.large.sp)

# Add beetle
large.beetle.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + beetle_recode + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.beetle.sp.sims <- simulateResiduals(large.beetle.sp)
plot(large.beetle.sp.sims)

summary(large.beetle.sp)

# Add perc. infested

large.perc.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.perc.sp.sims <- simulateResiduals(large.perc.sp)
plot(large.perc.sp.sims)

summary(large.perc.sp)

# distance to nearest tree

large.dist.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.dist.sp.sims <- simulateResiduals(large.dist.sp)
plot(large.dist.sp.sims)

summary(large.dist.sp)

# avg. dbh in neighborhood

large.dbh.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + avg.dbh + dist_nearest_tree + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.dbh.sp.sims <- simulateResiduals(large.dbh.sp)
plot(large.dbh.sp.sims)

summary(large.dbh.sp)

# lag infestation status

large.lag.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + + dist_nearest_tree + lag_infested + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.lag.sp.sims <- simulateResiduals(large.lag.sp)
plot(large.lag.sp.sims)

summary(large.lag.sp)

# mean percent live crown

large.crown.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.crown.sp.sims <- simulateResiduals(large.crown.sp)
plot(large.crown.sp.sims)

summary(large.crown.sp)

# N

large.n.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + (1|unique_ID) + (1|site), data = large_sp, family = binomial(link = "logit"))

large.n.sp.sims <- simulateResiduals(large.n.sp)
plot(large.n.sp.sims)

summary(large.n.sp)

# 

# Building model above
large.final.sp <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + (1|site), data = large_sp, family = binomial(link = "logit"))

large.final.sp.sims <- simulateResiduals(large.final.sp)
plot(large.final.sp.sims)

summary(large.final.sp)

testZeroInflation(large.final.sp.sims)

# test spatial autocorrelation

# Get all unique coordinates
large_sp_spatial <- large_sp %>% dplyr::filter(!duplicated(large_sp$Latitude))

# build model with these
large.final.spatial <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + perc.infested + dist_nearest_tree + lag_infested + mean_percent_live_crown_std + n + (1|unique_ID) + (1|site), data = large_sp_spatial, family = binomial(link = "logit"))

summary(large.final.spatial)

testSpatialAutocorrelation(large.final.spatial, x = large_sp_spatial$Longitude, y = large_sp_spatial$Latitude)

r2(large.final.sp)
check_singularity(large.final.sp)
```

### Coef plot - spatial weights

```{r}
tidy_large_sp <- broom::tidy(large.final.sp) # Create a tidy dataframe for the model

# Add a column to identify the model
tidy_large_sp$model <- "large.sp" 

dw_large_sp <- dwplot(tidy_large_sp)

tidy_small_sp <- broom::tidy(clump.final.sp)

tidy_small_sp$model <- "clump.sp"

dw_small_sp <- dwplot(tidy_small_sp)

tidy_all_sp <- rbind(tidy_large_sp, tidy_small_sp)

dw_all_sp <- dwplot(tidy_all_sp) +
  scale_y_discrete(
    labels = c(
              "logDISTANCE" = "Distance to nearest infested",
               "infest_sum_std_new" = "Infestation sum (neighborhood)",
               "dist_nearest_tree" = "Distance to nearest tree",
                "lag_infested" = "Lagged Infestation Status",
                "mean_percent_live_crown" = "Mean Percent Live Crown (neighborhood)",
                "n" = "Number of infested (neighborhood)",
                "gouting_recode" = "Gouting",
                "logdbh" = "DBH"
      ),
    limits = c("logDISTANCE",
                "mean_percent_live_crown_std",
                "n",
                "perc.infested",
                "dist_nearest_tree",
                "gouting_recode",
                "logdbh",
                "lag_infested"
                ) # Specify the desired order
  )


print(dw_all_sp+geom_vline(xintercept=0,lty=2))
```

# Fir Apparency

```{r}
app_data <- merge_data

#fix NA hts
app_data$height[app_data$site=="EMIG" & is.na(app_data$height)] <-   app_data$dbh[is.na(app_data$height) & app_data$site == "EMIG"] * emig_ratio
app_data$height <- as.numeric(app_data$height)
app_data$height[app_data$site=="FRNK" & is.na(app_data$height)] <-   app_data$dbh[is.na(app_data$height) & app_data$site == "FRNK"] * frnk_ratio
app_data$height <- as.numeric(app_data$height)
app_data$height[app_data$site=="MOCR" & is.na(app_data$height)] <-   app_data$dbh[is.na(app_data$height) & app_data$site == "MOCR"] * mocr_ratio
app_data$height <- as.numeric(app_data$height)
app_data$height[app_data$site=="MTTP" & is.na(app_data$height)] <-   app_data$dbh[is.na(app_data$height) & app_data$site == "MTTP"] * mttp_ratio
app_data$height <- as.numeric(app_data$height)
app_data$height[app_data$site=="SOAP" & is.na(app_data$height)] <-   app_data$dbh[is.na(app_data$height) & app_data$site == "SOAP"] * soap_ratio
app_data$height <- as.numeric(app_data$height)

# Filter for comparable heights
app_data <- app_data %>% filter(!is.na(dbh)) # take out clumps
app_data <- app_data %>% filter(dbh >= 5.08) # take out trees less than 2 inches (our cutoff for non-hosts)


  hts <- app_data %>% group_by(site, species) %>%
    summarize(mean_height = mean(height, na.rm = TRUE))
  
# Extract ABLA mean height per site
abla_heights <- hts %>%
  filter(species == "ABLA") %>%
  rename(abla_mean_height = mean_height) %>%
  select(site, abla_mean_height)

# Compute competition index for each site
competition_index <- hts %>%
  filter(species != "ABLA") %>%  # Exclude ABLA itself
  left_join(abla_heights, by = "site") %>%  # Merge ABLA heights
  mutate(difference = (abla_mean_height - mean_height) / abla_mean_height) %>%
  group_by(site) %>%
  summarize(
    competition_index = sum(difference, na.rm = TRUE) * (1 / n()) * 100
  )

# Build model 


```

# Cox Survival Model

## Data prep

We need to redo spatial weights for all visits, not just 2,3,4

```{r}
proportions_list_all <- list()

for (site in unique(all_merged$site)) {
  
  # Filter focal and complete datasets for the site
  site_focal <- filter(local_data_10, site == !!site)   # Focal points to buffer
  site_complete <- filter(all_merged, site == !!site)  # All points for bounding box
  
  # Convert focal points (already in UTM) to sf
  site_focal_sf <- st_as_sf(site_focal, coords = c("Longitude", "Latitude"), crs = 32612)
  
  # Convert complete dataset (in decimal degrees) to sf and transform to UTM
  site_complete_sf <- st_as_sf(site_complete, coords = c("Longitude", "Latitude"), crs = 4326)
  site_complete_utm <- st_transform(site_complete_sf, 32612)  # Transform to UTM
  
  # Get bounding box for the site in UTM
  site_bbox <- st_as_sfc(st_bbox(site_complete_utm))
  
  # Loop through each focal point and calculate proportions
  site_proportions <- sapply(1:nrow(site_focal_sf), function(i) {
    
    # Create a 10m buffer for the focal point
    buffer <- suppressWarnings(st_buffer(st_geometry(site_focal_sf)[i], dist = 10))
    
    # Compute the intersection between the buffer and the bounding box
    intersection <- suppressWarnings(st_intersection(buffer, site_bbox))
    
    # Calculate areas (buffer area and intersection area)
    circle_area <- suppressWarnings(st_area(buffer))
    intersection_area <- suppressWarnings(st_area(intersection))
    
    # Calculate proportion of buffer inside the bounding box
    proportion <- as.numeric(intersection_area / circle_area)
    
    return(proportion)
  })
  
  # Store the proportions for the current site
  proportions_list_all[[site]] <- site_proportions
  
  # Print progress for debugging
  cat("Processed", site, "with", length(site_proportions), "focal points.\n")
}


# Now append these proportions to the original focal dataset
all_visits_with_weights <- local_data_10 %>%
  mutate(proportion = unlist(proportions_list_all))


# change to long
model_data_all <- all_visits_with_weights %>%
  pivot_longer(
    cols = ends_with("_v1") | ends_with("_v2") | ends_with("_v3") | ends_with("_v4"),  # columns with visit data
    names_to = c(".value", "visit"),       # use ".value" to keep the original variable name
    names_pattern = "(.*)_(v[0-9])"        # regex to separate variable name and visit
  ) %>%
  mutate(visit = as.numeric(gsub("v", "", visit)))  # Convert "v1", "v2", etc. to 1, 2, etc.

names(model_data_all)[50] <- "t"

# make logged and lagged variables

model_data_all <- model_data_all %>%
  arrange(unique_ID, t) %>%
  group_by(unique_ID) %>%
  mutate(
    logDISTANCE = log(DISTANCE),
    logDISTANCE_noclump = log(DISTANCE_noclump),
    logDISTANCE_noinfest = log(DISTANCE_noinfest),
    logdbh = log(dbh),
    lag_logDISTANCE = lag(logDISTANCE),
    lag_infest_sum_std_new = lag(infest_sum_std_new),
    lag_infest_sum = lag(infestation_sum),
    lag_Gouting = lag(Gouting),
    lag_infested = lag(infested),
    lag_perc.infested = lag(perc.infested),
    lag_infestation_sum = lag(infestation_sum)
  ) %>%
  ungroup()

model_data_all$qmd_under <- sqrt(((model_data_all$basal_area_non_host + model_data_all$basal_area_host)/(model_data_all$n*31.83))/0.00007854)

#spatial weights
model_data_all$infest_sum_std <- model_data_all$infest_sum_std * model_data_all$proportion
model_data_all$avg.dbh <- model_data_all$avg.dbh * model_data_all$proportion
model_data_all$mean_percent_live_crown_std <- model_data_all$mean_percent_live_crown_std * model_data_all$proportion
model_data_all$brooms_nh <- model_data_all$brooms_nh * model_data_all$proportion
model_data_all$perc.infested <- model_data_all$perc.infested * model_data_all$proportion
model_data_all$basal_area_host_std <- model_data_all$basal_area_host_std * model_data_all$proportion
model_data_all$mean_dist_infested_recode <- model_data_all$mean_dist_infested_recode * model_data_all$proportion
model_data_all$n <- model_data_all$n * model_data_all$proportion

# Filter live
model_data_all$`Dead / Alive`[is.na(model_data_all$`Dead / Alive`)] <- 0
model_data_all <- filter(model_data_all, model_data_all$`Dead / Alive` == 1)
```


```{r}
#prop.final.long <- glmmTMB(cbind(infested, 1 - infested) ~ logdbh + logDISTANCE + gouting_recode + log_DISTANCE_noinfest + lag_infested + mean_percent_live_crown_std + n + perc.infested + comp_index + (1|site), data = model_data, family = binomial(link = "logit"))

model_data_all$season <- ifelse(model_data_all$t == 1 | model_data_all$t == 3, 0,1)

# Standardize all variables
model_data_all$dbh_std <- scale(model_data_all$logdbh)
model_data_all$DISTANCE_std <- scale(model_data_all$logDISTANCE)
model_data_all$DISTANCE_noinfest_std <- scale(model_data_all$logDISTANCE_noinfest)
model_data_all$n_std <- scale(model_data_all$n)
model_data_all$perc.infested_std <- scale(model_data_all$perc.infested)


cfit <- coxph(Surv(t, infested) ~ logdbh + logDISTANCE + gouting_recode + logDISTANCE_noinfest + lag_infested + mean_percent_live_crown_std + n + perc.infested + season, data = model_data_all )

cfit_site <- coxph(Surv(t, infested) ~ logdbh + logDISTANCE + gouting_recode + logDISTANCE_noinfest + lag_infested + mean_percent_live_crown_std + n + perc.infested + season + strata(site), data = model_data_all )

print(cfit)
summary(cfit)

round(cbind(simple= coef(cfit), stratified=coef(cfit_site)), 4)

```



# GLMMfields for each site

```{r}
data_long <- as.data.frame(data_long)

infest_glmm_rf <- glmmfields(infested ~ logDISTANCE + avg.dbh + gouting_recode + logdbh + 
    #lag_infested + 
    perc.infested + dist_nearest_tree + mean_percent_live_crown_std, 
  data = data_long, 
  lon = "Longitude", 
  lat = "Latitude",
  nknots = 6,
  chains = 2,
  iter = 500,
  time = "t", 
  covariance = "matern",
  cluster = "kmeans", 
  binomial_N = "infested",
  family = binomial(link = "logit"))

rf_tidy <- tidy(infest_glmm_rf_emig)

dw_rf <- dwplot(rf_tidy) +
  scale_y_discrete(
    labels = c("B[1]" = "Distance to nearest infested",
               #"lag_infested" = "Lagged infestation status",
               "B[6]" = "Distance to nearest tree",
               "B[5]" = "Percent infested (neighborhood)",
                "B[3]" = "Gouting",
                "B[2]" = "Average dbh in neighborhood",
                "B[4]" = "DBH",
                "B[7]" = "Mean percent live crown (Neighborhood)",
                "gp_sigma" = "Sigma",
                "gp_theta" = "Theta"
      ),
    limits = c( "gp_sigma",
                "gp_theta",
                "B[1]",
                "B[7]",
                "B[2]",
                "B[5]",
                "B[6]",
                "B[4]",
                "B[3]" 
                #"lag_infested"
      ) # Specify the desired order
  )


print(dw_rf+geom_vline(xintercept=0,lty=2))

```

# Stem maps

```{r}
plot_data <- merge_data

plot_data <- plot_data %>%
  mutate(
    status_v1 = case_when(
      `Dead / Alive_v1` == 0 ~ "Dead",                             
      host_non == 0 ~ "NonHost",                           
      host_non == 1 & infested_v1 == 0 ~ "NonInfested", 
      host_non == 1 & infested_v1 == 1 ~ "Infested",     
      TRUE ~ NA_character_                            
    ),
    status_v2 = case_when(
      `Dead / Alive_v2` == 0 ~ "Dead",                             
      host_non == 0 ~ "NonHost",                           
      host_non == 1 & infested_v2 == 0 ~ "NonInfested", 
      host_non == 1 & infested_v2 == 1 ~ "Infested", 
      host_non == 1 & infested_v1 == 0 & infested_v2 == 1 ~ "New Infested",
      TRUE ~ NA_character_                            
    ),
    status_v3 = case_when(
      `Dead / Alive_v3` == 0 ~ "Dead",                             
      host_non == 0 ~ "NonHost",                           
      host_non == 1 & infested_v3 == 0 ~ "NonInfested", 
      host_non == 1 & infested_v3 == 1 ~ "Infested",
      host_non == 1 & infested_v2 == 0 & infested_v3 == 1 ~ "New Infested",
      TRUE ~ NA_character_                            
    ),
    status_v4 = case_when(
      `Dead / Alive_v4` == 0 ~ "Dead",                             
      host_non == 0 ~ "NonHost",                           
      host_non == 1 & infested_v4 == 0 ~ "NonInfested", 
      host_non == 1 & infested_v4 == 1 ~ "Infested", 
      host_non == 1 & infested_v3 == 0 & infested_v4 == 1 ~ "New Infested",
      TRUE ~ NA_character_                            
    )
  )


```

# New plots

```{r}
data_new <- local_data_10

data_new <- data_new %>%
  mutate(size_class = cut(dbh, breaks = seq(0, max(dbh, na.rm = TRUE) + 5, by = 5), right = FALSE))

# Summarize new infestations by size class
size_summary <- data_new %>%
  group_by(size_class) %>%
  summarize(
    new_infestations_v2 = sum(new_infested_v2, na.rm = TRUE),
    new_infestations_v3 = sum(new_infested_v3, na.rm = TRUE),
    new_infestations_v4 = sum(new_infested_v4, na.rm = TRUE),
    new_infestations_all = sum(new_infested_v2 + new_infested_v3 + new_infested_v4), na.rm = TRUE)
    .groups = "drop"
  )

# Plot the data
ggplot(size_summary, aes(x = size_class, y = new_infestations_all)) +
  geom_bar(stat = "identity", fill = "skyblue", alpha = 0.8) +
  labs(
    title = "New Infestations by DBH Size Class",
    x = "Size Class (cm)",
    y = "Number of New Infestations"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# checking totals for models and data
total_model_new <- sum(model_data_new$new_infested)
total_data_new <- sum(size_summary$new_infestations_all)
```

