---
title: "Random things"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

This scripts purpose is to be somewhere where I can do a bunch of random things. The first data object I'm bringing in is our data with local statistics at 8 meters, which is what I should be working with from now on, I'm hoping that instead of one distinct purpose, this is where I'll mix and match files from other markdowns to get what I want.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
knitr::opts_knit$set(root.dir = "~/Project Files/Analysis/Cleaning and Compilation/Cleaning")
```

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

# Clean out environment

```{r}
rm(list = ls())
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(lme4) # to use glmer
library(data.table)
library(glmmTMB)
library(DHARMa)
library(MuMIn)
```

# Load data

```{r}

data <- read_xlsx("local_data_8.xlsx")

```

# New infestations vs infestation load

I want to plot out what the overal infestation loads are at each site and plot that against new infestations at that plot. We expect that higher infestation loads will lead to more new infestations so let's see if that's true.


```{r}

data <- data %>% mutate(new_infestation = ifelse(infested_v1 == 0 & (infested_v2 == 1 | infested_v3 == 1 | infested_v4 == 1), 1, 0))

data <- data %>% filter(`Dead / Alive_v1` == 1)

count <- data %>% group_by(site) %>% filter(infested_v1 == 0) %>% count()

infest_stats <- data %>% group_by(site) %>% dplyr::summarise(
  perc_infested_v1 = (sum(infested_v1) / n()) * 100,
  sum_newly_infested_v4 = sum(new_infestation)) %>% full_join(count, by = "site")

infest_stats$perc_new_infested <- (infest_stats$sum_newly_infested_v4 / infest_stats$n) *100

# plot 
infest_plot <- ggplot(infest_stats, aes(x = perc_infested_v1, y = perc_new_infested, label = site)) +
  geom_point(size = 3, color = "blue") +
  geom_text_repel(size = 3) +
  labs(
    y = "Percentage of Newly Infested Trees (Whole Study)",
    x = "Percentage of Infested Trees (Start of Study)",
    title = "Infested vs Newly Infested Percentage"
  ) +
  theme_minimal()

```

# Summary statistic table for infested trees at each visit

We'll create a summary table here looking at a couple things.

First: We'll make a table that's infested trees vs non-infested trees at each visit that includes pretty much all the factors we're inculding in the models:

- Distance to nearest infested
- Distance to nearest infested over 1 in dbh
- Average dbh of focal tree
- local metrics
- focal tree metrics

```{r}
names(data)[103] <- "TARGET_v1"
names(data)[104] <- "DISTANCE_v1"
names(data)[105] <- "TARGET_noclump_v1"
names(data)[106] <- "DISTANCE_noclump_v1"
names(data)[107] <- "TARGET_v2"
names(data)[108] <- "DISTANCE_v2"
names(data)[109] <- "TARGET_noclump_v2"
names(data)[110] <- "DISTANCE_noclump_v2"
names(data)[111] <- "TARGET_v3"
names(data)[112] <- "DISTANCE_v3"
names(data)[113] <- "TARGET_noclump_v3"
names(data)[114] <- "DISTANCE_noclump_v3"
names(data)[115] <- "TARGET_v4"
names(data)[116] <- "DISTANCE_v4"
names(data)[117] <- "TARGET_noclump_v4"
names(data)[118] <- "DISTANCE_noclump_v4"
names(data)[122] <- "n.infested_v1"
names(data)[123] <- "n.infested_v2"
names(data)[124] <- "n.infested_v3"
names(data)[125] <- "n.infested_v4"
names(data)[126] <- "perc.infested_v1"
names(data)[127] <- "perc.infested_v2"
names(data)[128] <- "perc.infested_v3"
names(data)[129] <- "perc.infested_v4"
names(data)[130] <- "infested_sum_v1"
names(data)[131] <- "infested_sum_v2"
names(data)[132] <- "infested_sum_v3"
names(data)[133] <- "infested_sum_v4"
names(data)[139] <- "mean_dist_infested_v1"
names(data)[140] <- "mean_dist_infested_v2"
names(data)[141] <- "mean_dist_infested_v3"
names(data)[142] <- "mean_dist_infested_v4"

data$height <- data$height * 0.3048

data_long <- data %>%
  pivot_longer(
    cols = ends_with("_v1") | ends_with("_v2") | ends_with("_v3") | ends_with("_v4"),  # columns with visit data
    names_to = c(".value", "visit"),       # use ".value" to keep the original variable name
    names_pattern = "(.*)_(v[0-9])"        # regex to separate variable name and visit
  ) %>%
  mutate(visit = as.numeric(gsub("v", "", visit)))  # Convert "v1", "v2", to 1, 2, etc.

vars_to_summarize <- c("dbh", "height", "canopy_ht", "brooms", "Gouting", "infestation_sum", "DISTANCE", "DISTANCE_noclump", "n", "n.infested", "perc.infested", "infested_sum",  "average.dbh_under",          
"brooms_under", "avg.gouting_under" , "gouting.sum_under", "avg_crown_ht_under",  "mean_dist_infested", "basal_area_non_host", "basal_area_host", "percent_basal_area_non_host", "mean_percent_live_crown", "dist_nearest_tree")

# Function to calculate summary statistics
summary_stats <- function(data, vars) {
  data %>%
    group_by(visit, infested) %>%  
    summarise(
      across(
        all_of(vars),
        list(
          mean = ~ mean(.x, na.rm = TRUE),
          sd = ~ sd(.x, na.rm = TRUE),
          n = ~ n(),
          se = ~ sd(.x, na.rm = TRUE) / sqrt(n())
        ),
        .names = "{.col}_{.fn}"
      ),
      .groups = "drop"
    )
}

summary_table <- summary_stats(data_long, vars_to_summarize)

names(summary_table)[11] <- "canopyht_mean"
names(summary_table)[12] <- "canopyht_sd"
names(summary_table)[13] <- "canopyht_n"
names(summary_table)[14] <- "canopyht_se"
names(summary_table)[23] <- "infestsum_mean"
names(summary_table)[24] <- "infestsum_sd"
names(summary_table)[25] <- "infestsum_n"
names(summary_table)[26] <- "infestsum_se"
names(summary_table)[31] <- "DISTnoclump_mean"
names(summary_table)[32] <- "DISTnoclum_sd"
names(summary_table)[33] <- "DISTnoclum_n"
names(summary_table)[34] <- "DISTnoclum_se"
names(summary_table)[47] <- "infestedsum_mean"
names(summary_table)[48] <- "infestedsum_sd"
names(summary_table)[49] <- "infestedsum_n"
names(summary_table)[50] <- "infestedsum_se"
names(summary_table)[51] <- "avg.dbh_mean"
names(summary_table)[52] <- "avg.dbh_sd"
names(summary_table)[53] <- "avg.dbh_n"
names(summary_table)[54] <- "avg.dbh_se"
names(summary_table)[55] <- "brooms.under_mean"
names(summary_table)[56] <- "brooms.under_sd"
names(summary_table)[57] <- "brooms.under_n"
names(summary_table)[58] <- "brooms.under_se"
names(summary_table)[59] <- "avg.gouting.under_mean"
names(summary_table)[60] <- "avg.gouting.under_sd"
names(summary_table)[61] <- "avg.gouting.under_n"
names(summary_table)[62] <- "avg.gouting.under_se"
names(summary_table)[63] <- "gouting.sum.under_mean"
names(summary_table)[64] <- "gouting.sum.under_sd"
names(summary_table)[65] <- "gouting.sum.under_n"
names(summary_table)[66] <- "gouting.sum.under_se"
names(summary_table)[67] <- "avg.crown.ht.under_mean"
names(summary_table)[68] <- "avg.crown.ht.under_sd"
names(summary_table)[69] <- "avg.crown.ht.under_n"
names(summary_table)[70] <- "avg.crown.ht.under_se"
names(summary_table)[71] <- "mean.dist.infested.under_mean"
names(summary_table)[72] <- "mean.dist.infested.under_sd"
names(summary_table)[73] <- "mean.dist.infested.under_n"
names(summary_table)[74] <- "mean.dist.infested.under_se"
names(summary_table)[75] <- "ba.nonhost.under_mean"
names(summary_table)[76] <- "ba.nonhost.under_sd"
names(summary_table)[77] <- "ba.nonhost.under_n"
names(summary_table)[78] <- "ba.nonhost.under_se"
names(summary_table)[79] <- "ba.host.under_mean"
names(summary_table)[80] <- "ba.host.under_sd"
names(summary_table)[81] <- "ba.host.under_n"
names(summary_table)[82] <- "ba.host.under_se"
names(summary_table)[83] <- "perc.ba.nonhost.under_mean"
names(summary_table)[84] <- "perc.ba.nonhost.under_sd"
names(summary_table)[85] <- "perc.ba.nonhost.under_n"
names(summary_table)[86] <- "perc.ba.nonhost.under_se"
names(summary_table)[87] <- "mean.perc.live.crown.under_mean"
names(summary_table)[88] <- "mean.perc.live.crown.under_sd"
names(summary_table)[89] <- "mean.perc.live.crown.under_n"
names(summary_table)[90] <- "mean.perc.live.crown.under_se"
names(summary_table)[91] <- "dist.nearest.tree_mean"
names(summary_table)[92] <- "dist.nearest.tree_sd"
names(summary_table)[93] <- "dist.nearest.tree_n"
names(summary_table)[94] <- "dist.nearest.tree_se"


# Reshape the data to prepare for plotting

plot_data <-  summary_table %>%
    # Pivot to long format
    pivot_longer(cols = -c(visit, infested), 
                 names_to = "variable_stat", 
                 values_to = "value") %>%
    # Separate variable name and statistic type
    separate(variable_stat, into = c("variable", "stat_type"), sep = "_") %>%
    # Pivot wider to create columns for each stat type (mean, se, etc.)
    pivot_wider(names_from = stat_type, values_from = value) %>%
    # Remove rows with NA in `mean` or `se` (optional, depending on data structure)
    filter(!is.na(mean) & !is.na(se))

# Plotting function
plot_group <- function(data, variables) {
  ggplot(data %>% filter(variable %in% variables), 
         aes(x = factor(visit), y = mean, fill = as.factor(infested))) +
    geom_bar(stat = "identity", position = position_dodge(), alpha = 0.8) +
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
                  position = position_dodge(0.9), width = 0.25) +
    facet_wrap(~ variable, scales = "free_y", ncol = 2) +
    scale_fill_manual(
      values = c("0" = "lightblue", "1" = "red"),  
      name = "Infestation Status",                
      labels = c("Non-Infested", "Infested")      
    ) +
    labs(
      title = "Mean with Standard Error by Visit",
      x = "Visit",
      y = "Mean Value",
      fill = "Infestation Status"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "top"
    )
}

# Create subsets of variables for grouping
variables_group1 <- c("dbh", "canopyht", "brooms", "height", "dist.nearest.tree")
variables_group2 <- c("Gouting", "infestedsum", "DISTANCE", "DISTnoclump")
variables_group3 <- c("n", "n.infested", "perc.infested", "infestedsum", "avg.dbh")
variables_group4 <- c("avg.crown.ht.under",  "mean.dist.infested.under", "ba.nonhost.under", "ba.host.under")
variables_group5 <- c("brooms.under", "avg.gouting.under", "gouting.sum.under")
variables_group6 <- c("ba.host.under", "perc.ba.nonhost.under", "mean.perc.live.crown.under")


# Generate plots for each group
plot1 <- plot_group(plot_data, variables_group1)
plot2 <- plot_group(plot_data, variables_group2)
plot3 <- plot_group(plot_data, variables_group3)
plot4 <- plot_group(plot_data, variables_group4)
plot5 <- plot_group(plot_data, variables_group5)
plot6 <- plot_group(plot_data, variables_group6)

# Display the plots
print(plot1)
print(plot2)
print(plot3)
print(plot4)
print(plot5)
print(plot6)
```

# New infestations by size class

```{r}
data$dbh[is.na(data$dbh)] <- 2.54/2

data <- data %>%
  mutate(size_class = cut(dbh, breaks = seq(0, max(dbh, na.rm = TRUE) + 5, by = 5), right = FALSE))

# Summarize new infestations by size class
size_summary <- data %>%
  group_by(size_class) %>%
  summarize(
    new_infestations = sum(new_infestation, na.rm = TRUE),
    .groups = "drop"
  )

# Plot the data
ggplot(size_summary, aes(x = size_class, y = new_infestations)) +
  geom_bar(stat = "identity", fill = "skyblue", alpha = 0.8) +
  labs(
    title = "New Infestations by DBH Size Class",
    x = "Size Class (cm)",
    y = "Number of New Infestations"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

# Change in infestation sum from start vs infestation percentage at start of study

```{r}

infest_stats_sum <- data %>% group_by(site) %>% dplyr::summarise(
  perc_infested_v1 = (sum(infested_v1) / n()) * 100, # percent infested at time 1
  infest_sum_change = (sum(infestation_sum_v4, na.rm = TRUE) - sum(infestation_sum_v1, na.rm = TRUE))) # change in infestation sum over study

infest_sum_plot <- ggplot(infest_stats_sum, aes(x = perc_infested_v1, y = infest_sum_change, label = site)) +
  geom_point(size = 3, color = "blue") +
  geom_text_repel(size = 3) +
  labs(
    y = "Change in Infestation Sum (Whole Study)",
    x = "Percentage of Infested Trees (Start of Study)",
    title = "Infested at start vs Change in Infestation Sum"
  ) +
  theme_minimal()

infest_sum_plot
infest_plot

```

# Look at Gouting change at each visit

```{r}
data$Gouting_v2[is.na(data$Gouting_v2)] <- 0
data$Gouting_v3[is.na(data$Gouting_v3)] <- 0
data$Gouting_v4[is.na(data$Gouting_v4)] <- 0

gout_stats <- data %>% group_by(site) %>% dplyr::summarise(
  "Mean - Visit 1" = mean(Gouting_v1),
  "Mean - Visit 2" = mean(Gouting_v2),
  "Mean - Visit 3" = mean(Gouting_v3),
  "Mean - Visit 4" = mean(Gouting_v4))

# reshape to long
gout_long <- gout_stats %>%
  pivot_longer(
    cols = starts_with("Mean"), # Columns to reshape
    names_to = "Visit", # New column to identify visits
    values_to = "Mean"  # Values for the means
  )

gout_long$Visit <- gsub("Mean -", "", gout_long$Visit)
# The palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

gout_plot <- ggplot(gout_long, aes(x = site, y = Mean, fill = Visit)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Gouting Mean Across Sites and Visits",
    x = "Site",
    y = "Mean Value"
  ) +
  theme_minimal() +
  scale_color_manual(name="Visit", values = cbPalette) 
```

# Combine all points with cleaned data

```{r}
all_points <- read_xlsx("C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/all_sites_with_points.xlsx") # Visit 4 for all sites




```