---
title: "Full Site Model"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Full Site Model
3/27/24
Michael Wayman

This will be probably my most ambitious Markdown so far (possibly too ambitious). What I want ot do is look at one site where we have all of our data, which in this case will be Emigration Canyon. We have all of our host and non-host data and two visits. We should hopefully be able to build a model that incorporates spatial and temporal dependency as well as all local statistics around a focal tree. 

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(lme4) # to use glmer
library(data.table)
library(R2jags)

```

# Clean out environment

```{r}
rm(list = ls())
```

# Load data

At time of creation, I want to load in original data, combine visits, then add point data, just to make this as clean as possible(ie. not having duplicate columns with multiple distances and targets and stuff)

```{r, warning=FALSE, echo=FALSE}

combined <- read_xlsx("combined_cleaned2.xlsx")
points <- read_xlsx("emig_points.xlsx")

```

# Collapse visits and combine points

```{r}
# Binary infestation metric
combined[combined == "-"] <- "0" # change all "-" to a zero

combined <- combined %>%
  mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH")), 
            funs(if_else(startsWith(., "1-"), "1", .))) # Change all "1-" ratings to just a singular 1 value

combined <- combined %>%
  mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH")), as.numeric)

combined$NL[is.na(combined$NL)] <- 0
combined$NM[is.na(combined$NM)] <- 0
combined$NH[is.na(combined$NH)] <- 0
combined$SL[is.na(combined$SL)] <- 0
combined$SM[is.na(combined$SM)] <- 0
combined$SH[is.na(combined$SH)] <- 0

combined <- combined %>%
  mutate(infested = ifelse(NL + NM + NH + SL + SM + SH != 0, 1, 0))

names(combined)[1] <- "site"
emig <- combined %>% dplyr::filter(site == "EMIG")

# First we'll subset by Visit
visit1 <- emig %>% filter(Visit == 1)
visit1$tag_ID <- paste(visit1$Species, visit1$`Tag #`, sep = "")
visit2 <- emig %>% filter(Visit == 2)
visit2$tag_ID <- paste(visit2$Species, visit2$`Tag #`, sep = "")
visit3 <- emig %>% filter(Visit == 3)
visit3$tag_ID <- paste(visit3$Species, visit3$`Tag #`, sep = "")
visit4 <- emig %>% filter(Visit == 4)
visit4$tag_ID <- paste(visit4$Species, visit4$`Tag #`, sep = "")

# We'll merge all 4
visit3 <- visit3 %>%
  rename_with(~ paste0(., "_v3"), -tag_ID) # add suffixes to visit 3 and 4
visit4 <- visit4 %>%
  rename_with(~ paste0(., "_v4"), -tag_ID) 

emig_merged <- visit1 %>%
  left_join(visit2, by = "tag_ID", suffix = c("_v1", "_v2")) %>%
  left_join(visit3, by = "tag_ID", suffix = c("", "_v3")) %>%
  left_join(visit4, by = "tag_ID", suffix = c("", "_v4")) 
  
#emig_merged <- inner_join(v1, visit2, by = "tag_ID", suffix = c("_v1", "_Visit2")) # we have duplicates
emig_merged <- emig_merged %>% distinct(tag_ID, .keep_all = TRUE) # Get rid of duplicates

# Make a host/non-host variable
emig_merged$host <- if_else(emig_merged$Species_v1 == "ABLA" | emig_merged$Species_v1 == "ABCO", 1, 0)

# Merge with points
points$Name <- sub("^(\\d+)$", "abla\\1", points$Name)
points$Name <- toupper(points$Name)
emig_merged$Name <- paste(emig_merged$Species_v1, emig_merged$`Tag #_v1`, sep = "")


emig_merged <- full_join(points, emig_merged, by = "Name")

# Omit points with no data associated or no point associated
emig_merged <- emig_merged %>% filter(!is.na(site_v1))
#emig_merged <- emig_merged %>% filter(!is.na(Latitude))
```

# Distance to all trees

We need to bring in our distance calculator in order to calculate all the local statistics that we want to put into our analysis.

```{r}

# So we need to convert our decimel degrees to UTM for calculation
all_merged <- emig_merged

# Convert to sf
all_merged_sf <- st_as_sf(all_merged, coords = c("Longitude", "Latitude"), crs = 4326)

# Define the target UTM zone
target_crs <- 32612  # EPSG code for UTM zone 12

# Transform the coordinates
all_merged_utm <- st_transform(all_merged_sf, crs = target_crs)
all_coords <- st_coordinates(all_merged_utm)
all_coords <- as.data.frame(all_coords)

# Put back into original datframe

all_merged_utm <- all_merged
all_merged_utm$Longitude <- all_coords$X
all_merged_utm$Latitude <- all_coords$Y

# Use Name as unique ID
# all_merged_utm$unique_ID <- paste(all_merged_utm$site_v1, all_merged_utm$`Tag #`, rownames(all_merged_utm), sep = "_")

all_infested <- all_merged_utm[all_merged_utm$infested_v1 == 1, ] # Infested trees with UTM coordinates


# all_merged <- all_merged[!duplicated(all_merged), ] #- there shouldn't be duplicates now
rownames(all_merged_utm) <- all_merged_utm$Name

# Remove non-host for these calculations
all_merged_utm_host <- all_merged_utm %>% filter(all_merged_utm$Species_v1 == "ABLA" | all_merged_utm$Species_v1 == "ABCO")

rm(Holder)
Holder = as.data.frame(matrix(nrow=0,ncol=3)) 
colnames(Holder) =c("FROM", "TARGET", "DISTANCE")

#Assign rownames as Tree-ID, dataset also needs UTM_X, UTM_Y, and PLOT
  subset_frame = all_merged_utm_host
TEMP_DIST = vegdist(subset_frame[,c("Longitude","Latitude")], method = "euclid")
TEMP_DIST.df = as.data.frame(as.matrix(TEMP_DIST))
rownames(TEMP_DIST.df) = (subset_frame$Name)
colnames(TEMP_DIST.df) = (subset_frame$Name)
TEMP_DIST.df$`Tag #` = rownames(TEMP_DIST.df)
TEMP_DIST_MELT = reshape2::melt(TEMP_DIST.df, id.vars = c("Tag #"))
colnames(TEMP_DIST_MELT) = c("FROM", "TARGET", "DISTANCE")

Holder <- TEMP_DIST_MELT

Holder_new <- merge(Holder, all_merged_utm, by.x = "TARGET", by.y = "Name")
holder_new.dt <- as.data.table(Holder_new)
length(unique(holder_new.dt$tag_ID))
nrow(holder_new.dt)
# Change infestation to binary variable so we can index it
# holder_new.dt$infested_1[holder_new.dt$infested_1 == "-"] <- 0
# unique(holder_new.dt$infested_1)
# holder_new.dt$infested_1[holder_new.dt$infested_1 == "BO" | holder_new.dt$infested_1 == "BR, BO" | holder_new.dt$infested_1 == "BR"] <- 1
# holder_new.dt$infested_1 <- as.numeric(holder_new.dt$infested_1)
# 
# holder_new.dt$infested_2[holder_new.dt$infested_2 == "-"] <- 0
# unique(holder_new.dt$infested_2)
# holder_new.dt$infested_2[holder_new.dt$infested_2 == "BO" | holder_new.dt$infested_2 == "BR, BO" | holder_new.dt$infested_2 == "BR"] <- 1
# holder_new.dt$infested_2 <- as.numeric(holder_new.dt$infested_2)

#min_dist_holder <- holder_new.dt[`infested` == 1 & (DISTANCE == min(DISTANCE[(DISTANCE != 0)])) ,.(DISTANCE, TARGET),by=.("unique_ID")]

# distance to nearest infested at time 1

# This isn't working since we have non-host with NA's in the infested variable

min_dist_holder <- holder_new.dt[infested_v1 == 1 ,.(TARGET, DISTANCE),by=.(FROM)][, .SD[which.min(DISTANCE[DISTANCE != 0])],by=.(FROM)]

# distance to nearest infested at time 2
min_dist_holder2 <- holder_new.dt[infested_v2 == 1 ,.(TARGET, DISTANCE),by=.(FROM)][, .SD[which.min(DISTANCE[DISTANCE != 0])],by=.(FROM)]

# Time 3
min_dist_holder3 <- holder_new.dt[infested_v3 == 1 ,.(TARGET, DISTANCE),by=.(FROM)][, .SD[which.min(DISTANCE[DISTANCE != 0])],by=.(FROM)]
# Time 4
min_dist_holder4 <- holder_new.dt[infested_v4 == 1 ,.(TARGET, DISTANCE),by=.(FROM)][, .SD[which.min(DISTANCE[DISTANCE != 0])],by=.(FROM)]

names(min_dist_holder2)[2] <- "TARGET2"
names(min_dist_holder2)[3] <- "DISTANCE2"
names(min_dist_holder3)[2] <- "TARGET3"
names(min_dist_holder3)[3] <- "DISTANCE3"
names(min_dist_holder4)[2] <- "TARGET4"
names(min_dist_holder4)[3] <- "DISTANCE4"

# add distances to data
# visit 1 distances
names(min_dist_holder)[1] <- "Name"
data_with_nearest_infested_dis <- all_merged_utm_host %>% inner_join(min_dist_holder, by = "Name")

# visit 2 distances
names(min_dist_holder2)[1] <- "Name"
data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder2, by = "Name")

# visit 3
names(min_dist_holder3)[1] <- "Name"
data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder2, by = "Name")

# visit 4
names(min_dist_holder4)[1] <- "Name"
data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder2, by = "Name")

# Now we have distances to nearest infested neighbors for host trees on the site


#write_xlsx(data_with_nearest_infested_dis,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/data_with_nearest_infested.xlsx")

```

# Infestation Sum

```{r}

# Visit 1
# data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")), 
#             funs(if_else(startsWith(., "1-"), "1", .))) %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")), as.numeric)
# Change all "1-" ratings to just a singular 1 value and make numeric

# data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")),~coalesce(., 0)) # change all NA's to 0

data_with_nearest_infested_dis$infestation_sum_v1_focal <- data_with_nearest_infested_dis$NL_v1 + data_with_nearest_infested_dis$NM_v1 + data_with_nearest_infested_dis$NH_v1 + data_with_nearest_infested_dis$SL_v1 + data_with_nearest_infested_dis$SM_v1 + data_with_nearest_infested_dis$SH_v1

# Visit 2
# data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")), 
#             funs(if_else(startsWith(., "1-"), "1", .))) %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")), as.numeric)
# # Change all "1-" ratings to just a singular 1 valueand make numeric
# 
# data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")),~coalesce(., 0)) # change all NA's to 0

data_with_nearest_infested_dis$infestation_sum_v2_focal <- data_with_nearest_infested_dis$NL_v2 + data_with_nearest_infested_dis$NM_v2 + data_with_nearest_infested_dis$NH_v2 + data_with_nearest_infested_dis$SL_v2 + data_with_nearest_infested_dis$SM_v2 + data_with_nearest_infested_dis$SH_v2

data_with_nearest_infested_dis$infestation_sum_v3_focal <- data_with_nearest_infested_dis$NL_v3 + data_with_nearest_infested_dis$NM_v3 + data_with_nearest_infested_dis$NH_v3 + data_with_nearest_infested_dis$SL_v3 + data_with_nearest_infested_dis$SM_v3 + data_with_nearest_infested_dis$SH_v3

data_with_nearest_infested_dis$infestation_sum_v4_focal <- data_with_nearest_infested_dis$NL_v4 + data_with_nearest_infested_dis$NM_v4 + data_with_nearest_infested_dis$NH_v4 + data_with_nearest_infested_dis$SL_v4 + data_with_nearest_infested_dis$SM_v4 + data_with_nearest_infested_dis$SH_v4

# Subset target trees to get infesation sums
targets <- as.character(data_with_nearest_infested_dis$TARGET)
infested_subset_inf_sum <- data_with_nearest_infested_dis %>% filter(Name %in% targets) %>% dplyr::select(c("Name", "infestation_sum_v1_focal", "infestation_sum_v2_focal", "infestation_sum_v3_focal", "infestation_sum_v4_focal"))
names(infested_subset_inf_sum)[1] <- "TARGET"
names(infested_subset_inf_sum)[2] <- "inf_sum_v1_target"
names(infested_subset_inf_sum)[3] <- "inf_sum_v2_target"
names(infested_subset_inf_sum)[4] <- "inf_sum_v3_target"
names(infested_subset_inf_sum)[5] <- "inf_sum_v4_target"


# append sums to our main dataframe
# Append target coordinates to data frame
data_with_nearest_infested_dis <- inner_join(data_with_nearest_infested_dis, infested_subset_inf_sum, by = "TARGET")
str(data_with_nearest_infested_dis)



# Again for all data so we can extract values for full model

# Visit 1
# all_merged_utm <- all_merged_utm %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")), 
#             funs(if_else(startsWith(., "1-"), "1", .))) %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")), as.numeric)
# # Change all "1-" ratings to just a singular 1 value and make numeric
# 
# all_merged_utm <- all_merged_utm %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1")),~coalesce(., 0)) # change all NA's to 0

all_merged_utm$infestation_sum_v1_focal <- all_merged_utm$NL_v1 + all_merged_utm$NM_v1 + all_merged_utm$NH_v1 + all_merged_utm$SL_v1 + all_merged_utm$SM_v1 + all_merged_utm$SH_v1

# Visit 2
# all_merged_utm <- all_merged_utm %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")), 
#             funs(if_else(startsWith(., "1-"), "1", .))) %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")), as.numeric)
# # Change all "1-" ratings to just a singular 1 valueand make numeric
# 
# all_merged_utm <- all_merged_utm %>%
#   mutate_at(vars(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2")),~coalesce(., 0)) # change all NA's to 0

all_merged_utm$infestation_sum_v2_focal <- all_merged_utm$NL_v2 + all_merged_utm$NM_v2 + all_merged_utm$NH_v2 + all_merged_utm$SL_v2 + all_merged_utm$SM_v2 + all_merged_utm$SH_v2

all_merged_utm$infestation_sum_v3_focal <- all_merged_utm$NL_v3 + all_merged_utm$NM_v3 + all_merged_utm$NH_v3 + all_merged_utm$SL_v3 + all_merged_utm$SM_v3 + all_merged_utm$SH_v3

all_merged_utm$infestation_sum_v4_focal <- all_merged_utm$NL_v4 + all_merged_utm$NM_v4 + all_merged_utm$NH_v4 + all_merged_utm$SL_v4 + all_merged_utm$SM_v4 + all_merged_utm$SH_v4

```

# Infestation as a binary

```{r}
# Time 1
# all_merged_utm$`Inf Location_v1`[ all_merged_utm$`Inf Location_v1`== "-"] <- 0
# all_merged_utm$`Inf Location_v1`[all_merged_utm$`Inf Location_v1` == "BO" | all_merged_utm$`Inf Location_v1` == "BR, BO" | all_merged_utm$`Inf Location_v1` == "BR"] <- 1
# all_merged_utm$`Inf Location_v1`[is.na(all_merged_utm$`Inf Location_v1`)] <- 0 # change NA's to 0
# all_merged_utm$`Inf Location_v1` <- as.numeric(all_merged_utm$`Inf Location_v1`)
# unique(all_merged_utm$`Inf Location_v1`)
# # Time 2
# all_merged_utm$`Inf Location_v2`[ all_merged_utm$`Inf Location_v2`== "-"] <- 0
# all_merged_utm$`Inf Location_v2`[all_merged_utm$`Inf Location_v2` == "BO" | all_merged_utm$`Inf Location_v2` == "BR, BO" | all_merged_utm$`Inf Location_v2` == "BR"] <- 1
# all_merged_utm$`Inf Location_v2`[is.na(all_merged_utm$`Inf Location_v2`)] <- 0 # change NA's to 0
# all_merged_utm$`Inf Location_v2` <- as.numeric(all_merged_utm$`Inf Location_v2`)
# unique(all_merged_utm$`Inf Location_v2`)

```

# Function to extract info from target trees

```{r}

# Change names to make easier
names(all_merged_utm)[12] <- "dead_alive"
names(all_merged_utm)[15] <- "brooms"

# Recode some things for extraction
all_merged_utm$Species_v1 <- as.factor(all_merged_utm$Species_v1)

# Make things!

# BA in cubed meters or something
all_merged_utm$DBH_v1 <- all_merged_utm$DBH_v1 * 2.54 # convert DBH to cm
all_merged_utm <- mutate(all_merged_utm, BA = (0.00007854 * (DBH_v1 ^ 2)))

# Crown Height in meters
all_merged_utm$crown_height <- (all_merged_utm$Height_v1 - all_merged_utm$`Canopy Height_v1`) * 0.3048

# Function to get distances for every tree 10m and under
TEMP_DIST_10 = vegdist(all_merged_utm[,c("Longitude","Latitude")], method = "euclid")
TEMP_DIST_10.df = as.data.frame(as.matrix(TEMP_DIST_10))
rownames(TEMP_DIST_10.df) = (all_merged_utm$Name)
colnames(TEMP_DIST_10.df) = (all_merged_utm$Name)
TEMP_DIST_10.df$`Tag #` = rownames(TEMP_DIST_10.df)
TEMP_DIST_MELT_10 = reshape2::melt(TEMP_DIST_10.df, id.vars = c("Tag #"))
colnames(TEMP_DIST_MELT_10) = c("FROM", "TARGET", "DISTANCE")

Holder_10 <- TEMP_DIST_MELT_10

Holder_new_10 <- merge(Holder_10, all_merged_utm, by.x = "TARGET", by.y = "Name")
all_merged_utm.dt <- as.data.table(Holder_new_10)

# Get all trees under 10m from focal tree
under10_holder <- all_merged_utm.dt[,.(TARGET, DISTANCE), by=.(FROM)][,.SD[which(DISTANCE <= 10 & DISTANCE != 0)], by=.(FROM)]

# Distance to nearest infested neighbor
min_dist_holder <- all_merged_utm.dt[infested_v1 == 1 ,.(TARGET, DISTANCE),by=.(FROM)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(FROM)]

min_dist_holder <- min_dist_holder[, -2]

names(min_dist_holder)[1] <- "Name"


df <- under10_holder
# Create a function to extract the variable based on the linked tree ID
extractVariable <- function(TARGET, variable) {
  linkedTreeRow <- all_merged_utm[all_merged_utm$Name == TARGET, ]
  if (nrow(linkedTreeRow) > 0) {
    return(linkedTreeRow[[variable]])
  } else {
    return(NA)  # or any other value to indicate missing data
  }
}

str(all_merged_utm)

# Apply the function to create a new column with the extracted variable - need to change function for each extraction !!
df$infestation_under10 <- sapply(df$TARGET, extractVariable, variable = "infested_v1")
df$infestation.sum.1_under10 <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v1_focal")
df$dbh_under10 <- sapply(df$TARGET, extractVariable, variable = "DBH_v1")
df$brooms_under10 <- sapply(df$TARGET, extractVariable, variable = "brooms")
df$gouting_under10 <- sapply(df$TARGET, extractVariable, variable = "Gouting_v1")
df$species_under10 <- sapply(df$TARGET, extractVariable, variable = "Species_v1")
df$BA_under10 <- sapply(df$TARGET, extractVariable, variable = "BA")
df$crown_ht__under10 <- sapply(df$TARGET, extractVariable, variable = "crown_height")
df$dead_alive <- sapply(df$TARGET, extractVariable, variable = "dead_alive")

# Host and non-host BA
df$host_non <- if_else(df$species_under10 == "ABLA", 1, 0)

# create a summary table of local stats for each tree
local.stats <- df %>% 
  dplyr::group_by(FROM) %>% 
  dplyr::summarise(n = n(),
                   number.infested_under10 = sum(infestation_under10), # number of infested
                   percent.infested_under10 = (sum(infestation_under10)/n) * 100, # percent infested
                   infestation.sum_under10 = sum(infestation.sum.1_under10), # infestation load or propagule pressure
                   average.dbh_under10 = mean(dbh_under10, na.rm = TRUE), # average dbh
                   brooms_under10 = sum(brooms_under10), # number of brooms (this may not be anything)
                   # avg.gouting_under10 = mean(gouting_under10, na.rm = TRUE), # Since we only have 1's at EMIG, average isn't useful
                   gouting.sum_under = sum(gouting_under10, na.rm = TRUE), # gouting sum
                   basal_area_non_host = sum(BA_under10[!is.na(host_non) & host_non != 1], na.rm = TRUE), # non-host BA
                   basal_area_host = sum(BA_under10[host_non == 1], na.rm = TRUE), # host BA
                   percent_basal_area_non_host = basal_area_non_host / (basal_area_non_host + basal_area_host) * 100, # percent BA of non-host
                   crown_ht_sum = sum(crown_ht__under10, na.rm = TRUE),
                   mean_distance_infested = ifelse(sum(infestation_under10) == 0, NA, mean(DISTANCE[infestation_under10 == 1], na.rm = TRUE))) # If there's no infestation within 10m, it'll come out as NA 

names(local.stats)[1] <- "Name"

# combine with data
data_with_local_stats <- all_merged_utm %>% inner_join(local.stats, by = "Name")

# append nearest infested neighbor distance
data_with_local_stats <- inner_join(data_with_local_stats, min_dist_holder, by = "Name")

#write_xlsx(data_with_local_stats, "under10m_data.xlsx")
```

# Now we can set up a Bayesian Model

We'll set up the probability of being infested by local and focal tree stats and see what ends up coming out

```{r}
data_with_local_stats <- read_xlsx("under10m_data.xlsx")
# Filter out hosts to not skew infested metric
host_local <- data_with_local_stats %>% filter(Species_v1 == "ABLA")

pow <- function(base, exponent) {
  return(base^exponent)
}

# Change NA DBH to 0.5
host_local$DBH_v1[is.na(host_local$DBH_v1)] <- 2.54/2
host_local$Gouting_v1[is.na(host_local$Gouting_v1)] <- 0

# Change NA's in mean distance infested to 0 - this is just to get it to run!! - NEED TO CHANGE FOR FINAL MODEL - 0 might be right since I think we're getting NA's because there's no infested within 10 meters
host_local$mean_distance_infested[is.na(host_local$mean_distance_infested)] <- 0

# Check for NA's

na_subset <- host_local[, c("infested_v1", "DISTANCE", "DBH_v1", "Gouting_v1", "brooms", "infestation.sum_under10", "mean_distance_infested", "percent_basal_area_non_host", "basal_area_host", "average.dbh_under10", "number.infested_under10", "percent.infested_under10", "crown_ht_sum", "Name"), drop = FALSE]

# Check for NA values in each specified column
na_count <- colSums(is.na(na_subset))

# General linear setup

# newinflog <- glm(cbind(new_infested, 1 - new_infested) ~
#                    DISTANCE + logDBH + Gouting_v1 + `Broom (#)_v1` + broom_binary + bearings1 + logInfestsum,
#                    data = live_host,
#                    family = binomial(link = "logit"))

# Trying without RE of tree to try to get this to run

host_local$crown_ht_sum_std <- (host_local$crown_ht_sum - mean(host_local$crown_ht_sum))/sd(host_local$crown_ht_sum)
host_local$infest_sum_10_std <- (host_local$infestation.sum_under10 - mean(host_local$infestation.sum_under10))/sd(host_local$infestation.sum_under10)
host_local$basal_area_host_std <- (host_local$basal_area_host - mean(host_local$basal_area_host))/sd(host_local$basal_area_host)

infested_jags_data_re <- list(infested = host_local$infested_v1, 
                            logDISTANCE = log(host_local$DISTANCE),
                            # note that we have to convert the FoodTreatment factor to
                            # a number to work appropriately in JAGS. Also, numeric coding needs to start at 0, not 1
                            DBH_focal = log(host_local$DBH_v1), 
                            Gouting_focal = host_local$Gouting_v1,
                            Brooms_focal = host_local$brooms,
                            infest_sum_10 = host_local$infest_sum_10_std,
                            mean_dist_infest_10 = host_local$mean_distance_infested,
                            # non_host_ba_perc = host_local$percent_basal_area_non_host,
                            host_ba = host_local$basal_area_host_std,
                            # avg_dbh_10 = host_local$average.dbh_under10,
                            n_infest_10 = host_local$number.infested_under10,
                            percent_infested_10 = host_local$percent.infested_under10,
                            # crown_ht_sum = host_local$crown_ht_sum_std,
                            tree = as.numeric(as.factor(host_local$Name)),
                            N = nrow(host_local)
                            , N_tree = length(levels(factor(host_local$Name)))
                                                              )
# Model
infested_jags_naive_re <- function(){
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)
  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
  # Likelihood:
  for (i in 1:N){
    lin_comb[i] <- beta_0 + (beta_distance * logDISTANCE[i]) + (beta_dbh * DBH_focal[i]) +
      (beta_gouting * Gouting_focal[i]) + (beta_brooms * Brooms_focal[i]) + (beta_infestsum * infest_sum_10[i]) + (beta_mean_dist_infest * mean_dist_infest_10[i]) + (beta_host_ba * host_ba[i])+ beta_n_infest * n_infest_10[i] + beta_percent_infest * percent_infested_10[i]
    #+ beta_non_host_ba_perc * non_host_ba_perc[i] + beta_host_ba * host_ba[i] + beta_avg_dbh * avg_dbh_10[i]  + beta_crown_ht * crown_ht_sum[i] + (beta_non_host_ba_perc * non_host_ba_perc[i])
    + beta_tree[tree[i]]
    # here's our linear combination of predictors, just like in a standard regression
    
     p[i] <- exp(lin_comb[i])/(1 + exp(lin_comb[i]))
    infested[i] ~ dbinom(p[i], 1) # Here's the Binomial likelihood,
  }
}

infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, 1), 
       beta_distance = rnorm(1, 0, 1),
       beta_dbh = rnorm(1, 0, 1),
       beta_gouting = rnorm(1, 0, 1),
       beta_brooms = rnorm(1, 0, 1),
       beta_infestsum = rnorm(1, 0, 1),
       beta_mean_dist_infest = rnorm(1, 0, 1),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, 1),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, 1),
       beta_percent_infest = rnorm(1, 0, 1),
       # beta_crown_ht = rnorm(1, 0, 1))
    #,
       beta_tree = rep(0, length(levels(factor(host_local$Name)))))
      # p = runif(1, 0.1, 0.9))
}

 infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest",# "beta_non_host_ba_perc", "beta_avg_dbh",  "beta_crown_ht",
   "sigma_sq_tree",
   "p")

infested_fit_naive_re <- jags(data = infested_jags_data_re, 
                            inits = infested_inits_re, 
                            parameters.to.save = infested_params_re, 
                            model.file = infested_jags_naive_re,
                            n.chains = 3,
                            n.iter = 12000, 
                            n.burnin = 2000, 
                            n.thin = 10, 
                            DIC = F)

mcmcplots::mcmcplot(infested_fit_naive_re)

infested_fit_naive_re

intensity.glm <- glmmTMB(infestation_sum_v1_focal ~ 
                      logDISTANCE + DBH_v1 + Gouting_v1 + `Broom (#)_v1` + (1|site),
                      family = nbinom1,
                      data = live_host)

confint(intensity.glm) 
# These intervals are pretty close to the 2.5th and 97.5th quantiles of beta_food 
# and beta_parent in our Bayesian model. 

# check for convergence
traceplot(intensity_fit_naive_re, mfrow = c(2, 2), ask = F)

dev.off()

intensity_mcmc_re <- as.mcmc(intensity_fit_naive_re)
plot(intensity_mcmc_re)

```

# Regular GLM model

Since we're getting local stats and are doing a Bayesian implementation, we should do a regular GLM and see what comes out

```{r}

emig.inf <- glm(cbind(infested_v1, 1 - infested_v1) ~ DISTANCE + DBH_v1 + Gouting_v1 + brooms + number.infested_under10 + percent.infested_under10 + infestation.sum_under10 + average.dbh_under10 + basal_area_non_host + percent_basal_area_non_host + mean_distance_infested, data = host_local, family = binomial(link = "logit"))

inf.sims <- simulateResiduals(emig.inf)
plot(inf.sims)

summary(emig.inf)

```

# Bayesian Model for all sites

```{r}
all_sites <- read_xlsx("under10m_data.xlsx")

# Host and non-host BA
df$host_non <- if_else(df$species_under10 == "ABLA", 1, 0)

# create a summary table of local stats for each tree
local.stats <- df %>% 
  dplyr::group_by(FROM) %>% 
  dplyr::summarise(n = n(),
                   number.infested_under10 = sum(infestation_under10), # number of infested
                   percent.infested_under10 = (sum(infestation_under10)/n) * 100, # percent infested
                   infestation.sum_under10 = sum(infestation.sum.1_under10), # infestation load or propagule pressure
                   average.dbh_under10 = mean(dbh_under10, na.rm = TRUE), # average dbh
                   brooms_under10 = sum(brooms_under10), # number of brooms (this may not be anything)
                   # avg.gouting_under10 = mean(gouting_under10, na.rm = TRUE), # Since we only have 1's at EMIG, average isn't useful
                   gouting.sum_under = sum(gouting_under10, na.rm = TRUE), # gouting sum
                   basal_area_non_host = sum(BA_under10[!is.na(host_non) & host_non != 1], na.rm = TRUE), # non-host BA
                   basal_area_host = sum(BA_under10[host_non == 1], na.rm = TRUE), # host BA
                   percent_basal_area_non_host = basal_area_non_host / (basal_area_non_host + basal_area_host) * 100, # percent BA of non-host
                   crown_ht_sum = sum(crown_ht__under10, na.rm = TRUE),
                   mean_distance_infested = ifelse(sum(infestation_under10) == 0, NA, mean(DISTANCE[infestation_under10 == 1], na.rm = TRUE))) # If there's no infestation within 10m, it'll come out as NA 

names(local.stats)[1] <- "Name"

# combine with data
data_with_local_stats <- all_merged_utm %>% inner_join(local.stats, by = "Name")

# append nearest infested neighbor distance
data_with_local_stats <- inner_join(data_with_local_stats, min_dist_holder, by = "Name")



# Change NA DBH to 0.5
all_sites$dbh[is.na(all_sites$dbh)] <- 2.54/2
all_sites$Gouting_v1[is.na(all_sites$Gouting_v1)] <- 0

# Change NA's in mean distance infested to 0 - this is just to get it to run!! - NEED TO CHANGE FOR FINAL MODEL - 0 might be right since I think we're getting NA's because there's no infested within 10 meters
all_sites$mean_distance_infested[is.na(all_sites$mean_distance_infested)] <- 0

# Check for NA's

na_subset <- all_sites[, c("infested_v1", "DISTANCE", "dbh", "Gouting_v1", "brooms", "infestation.sum_under10", "mean_distance_infested", "percent_basal_area_non_host", "basal_area_host", "average.dbh_under10", "number.infested_under10", "percent.infested_under10", "crown_ht_sum", "Name"), drop = FALSE]

# Check for NA values in each specified column
na_count <- colSums(is.na(na_subset))

# General linear setup

# newinflog <- glm(cbind(new_infested, 1 - new_infested) ~
#                    DISTANCE + logDBH + Gouting_v1 + `Broom (#)_v1` + broom_binary + bearings1 + logInfestsum,
#                    data = live_host,
#                    family = binomial(link = "logit"))

# Trying without RE of tree to try to get this to run

all_sites$crown_ht_sum_std <- (all_sites$crown_ht_sum - mean(all_sites$crown_ht_sum))/sd(all_sites$crown_ht_sum)
all_sites$infest_sum_10_std <- (all_sites$infestation.sum_under10 - mean(all_sites$infestation.sum_under10))/sd(all_sites$infestation.sum_under10)
all_sites$basal_area_host_std <- (all_sites$basal_area_host - mean(all_sites$basal_area_host))/sd(all_sites$basal_area_host)

infested_jags_data_re <- list(infested = all_sites$infested_v1, 
                            logDISTANCE = log(all_sites$DISTANCE),
                            # note that we have to convert the FoodTreatment factor to
                            # a number to work appropriately in JAGS. Also, numeric coding needs to start at 0, not 1
                            DBH_focal = log(all_sites$dbh), 
                            Gouting_focal = all_sites$Gouting_v1,
                            Brooms_focal = all_sites$brooms,
                            infest_sum_10 = all_sites$infest_sum_10_std,
                            mean_dist_infest_10 = all_sites$mean_distance_infested,
                            # non_host_ba_perc = all_sites$percent_basal_area_non_host,
                            host_ba = all_sites$basal_area_host_std,
                            # avg_dbh_10 = all_sites$average.dbh_under10,
                            n_infest_10 = all_sites$number.infested_under10,
                            percent_infested_10 = all_sites$percent.infested_under10,
                            # crown_ht_sum = all_sites$crown_ht_sum_std,
                            tree = as.numeric(as.factor(all_sites$Name)),
                            N = nrow(all_sites)
                            , N_tree = length(levels(factor(all_sites$Name)))
                                                              )
# Model
infested_jags_naive_re <- function(){
  # Priors:
  beta_0 ~ dnorm(0, 0.5) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.5) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.5)
  beta_gouting ~ dnorm(0, 0.5)
  beta_brooms ~ dnorm(0, 0.5)
  beta_infestsum ~ dnorm(0, 0.5)
  beta_mean_dist_infest ~ dnorm(0, 0.5)
  #beta_non_host_ba_perc ~ dnorm(0, 0.5)
  beta_host_ba ~ dnorm(0, 0.5)
  # beta_avg_dbh ~ dnorm(0, 0.5)
  beta_n_infest ~ dnorm(0, 0.5)
  beta_percent_infest ~ dnorm(0, 0.5)
  # beta_crown_ht ~ dnorm(0, 0.5)
  
  for(t in 1:N_tree){
    beta_tree[t] ~ dnorm(0, tau_tree)
  }
  
  sigma_sq_tree ~ dunif(0, 100) # here's my prior for tree
  tau_tree <- pow(sigma_sq_tree,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
  # Likelihood:
  for (i in 1:N){
    lin_comb[i] <- beta_0 + (beta_distance * logDISTANCE[i]) + (beta_dbh * DBH_focal[i]) +
      (beta_gouting * Gouting_focal[i]) + (beta_brooms * Brooms_focal[i]) + (beta_infestsum * infest_sum_10[i]) + (beta_mean_dist_infest * mean_dist_infest_10[i]) + (beta_host_ba * host_ba[i])+ beta_n_infest * n_infest_10[i] + beta_percent_infest * percent_infested_10[i]
    #+ beta_non_host_ba_perc * non_host_ba_perc[i] + beta_host_ba * host_ba[i] + beta_avg_dbh * avg_dbh_10[i]  + beta_crown_ht * crown_ht_sum[i] + (beta_non_host_ba_perc * non_host_ba_perc[i])
    + beta_tree[tree[i]]
    # here's our linear combination of predictors, just like in a standard regression
    
     p[i] <- exp(lin_comb[i])/(1 + exp(lin_comb[i]))
    infested[i] ~ dbinom(p[i], 1) # Here's the Binomial likelihood,
  }
}

infested_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, 1), 
       beta_distance = rnorm(1, 0, 1),
       beta_dbh = rnorm(1, 0, 1),
       beta_gouting = rnorm(1, 0, 1),
       beta_brooms = rnorm(1, 0, 1),
       beta_infestsum = rnorm(1, 0, 1),
       beta_mean_dist_infest = rnorm(1, 0, 1),
       #beta_non_host_ba_perc = rnorm(1, 0, 1),
       beta_host_ba = rnorm(1, 0, 1),
       # beta_avg_dbh = rnorm(1, 0, 1),
       beta_n_infest = rnorm(1, 0, 1),
       beta_percent_infest = rnorm(1, 0, 1),
       # beta_crown_ht = rnorm(1, 0, 1))
    #,
       beta_tree = rep(0, length(levels(factor(all_sites$Name)))))
      # p = runif(1, 0.1, 0.9))
}

 infested_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_infestsum",
   "beta_mean_dist_infest", "beta_host_ba", "beta_n_infest", "beta_percent_infest",# "beta_non_host_ba_perc", "beta_avg_dbh",  "beta_crown_ht",
   "sigma_sq_tree",
   "p")

infested_fit_naive_re <- jags(data = infested_jags_data_re, 
                            inits = infested_inits_re, 
                            parameters.to.save = infested_params_re, 
                            model.file = infested_jags_naive_re,
                            n.chains = 3,
                            n.iter = 12000, 
                            n.burnin = 2000, 
                            n.thin = 10, 
                            DIC = F)

mcmcplots::mcmcplot(infested_fit_naive_re)
```
