---
title: "GLM - Infestation in Space"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Spatial Indices
2/22/24
Michael Wayman

The purpose of this script is to attempt to run through the glm workflow with the intention of creating a glm model that addresses the probability of an individual tree just being infested at all. We'll set up a model for all sites and then proceed to doing each site individually.

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(Matrix) # to use glmer

```

## Clean out global environment

```{r}
rm(list = ls())
```

## Load Data

```{r}
all_data <- read_xlsx("combined_cleaned.xlsx")
site_stats <- read_xlsx("site_stats.xlsx")
data_with_dis <- read_xlsx("data_with_nearest_info.xlsx")
names(data_with_dis)[8] <- "infested"

```

# Model set up for all sites

```{r}


```

