---
title: "Spatial_data"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Spatial Indices
1/27/23 * approx date
Michael Wayman

The purpose of this script is to attempt to run through the glm workfloe with the intention of creating a glm model that addresses the probability of an individual tree getting infested based on individual attributes. Let's do it!

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(Matrix) # to use glmer
library(lme4) # to use glmer

```

## Clean out global environment

```{r}
rm(list = ls())
```

## Load Data

```{r, echo = F, warning=FALSE}
all_data <- read_xlsx("combined_cleaned2.xlsx")
all_data <- all_data[!duplicated(all_data), ] # - there shouldn't be duplicates now
site_stats <- read_xlsx("site_stats.xlsx")
data_with_dis <- read_xlsx("data_with_nearest_info.xlsx")

```

# First, we'll plot histograms for all of our covariates and do any transformations that we may need to do (eg. Site as a factor)

# Histograms
## DBH

```{r}

hist(all_data$DBH)
max(all_data$DBH, na.rm = TRUE)
hist(log(all_data$DBH))
all_data$logDBH <- log(all_data$DBH)

```

## Height

```{r}

hist(all_data$Height)
max(all_data$Height, na.rm = TRUE)
all_data[which.max(all_data$Height),] # Need to fix max height, incorrect
hist(log(all_data$Height))
all_data$logHeight <- log(all_data$Height)

```

## Canopy Height

```{r}

hist(all_data$`Canopy Height`)
hist(log(all_data$`Canopy Height`))
all_data$logCanopyHeight <- log(all_data$`Canopy Height`)

```

## Gouting

```{r}

hist(all_data$Gouting)
# can be input as categorical

```

## Brooms

```{r}

hist(all_data$`Broom (#)`)
hist(log(all_data$`Broom (#)`))
# could look at binary (has some/don't) AND
# another covariate with how many

all_data$broom_binary <- ifelse(all_data$`Broom (#)` > 0, 1, 0)

```


## Sections and infestation is all ordinal

# Pair-wise correlations

```{r}

pairs(all_data[, c("DBH", "Height", "Canopy Height", "Gouting", "Broom (#)")], pch = 16, col = "blue", main = "Pair-wise Correlations")

cor(all_data[, c("DBH", "Height", "Canopy Height", "Gouting", "Broom (#)")], use = "complete.obs")

cor(all_data[, c("logDBH", "logHeight", "Canopy Height", "Gouting", "Broom (#)")], use = "complete.obs")

# add columns with log tranformed values

```
We see from this correlation matrix that we're not going to want to include both DBH and height in a model.

# Create a summed infestation statistic for model

```{R}
str(all_data)
# We need to change all of our 1-somethings into just 1
working <- all_data

working <- working %>%
  mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH")), 
            funs(if_else(startsWith(., "1-"), "1", .))) # Change all "1-" ratings to just a singular 1 value

working <- working %>%
  mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH")), as.numeric)

working <- working %>%
  mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH")),~coalesce(., 0)) # change all NA's to 0

str(working)
head(working)


working$infestation_sum <- working$NL + working$NM + working$NH + working$SL + working$SM + working$SH

```

## Plotting summed statistic

```{r}

hist(working$infestation_sum) # A lot of zeros
hist(log(working$infestation_sum))

```
We have a lot of zeros, this may require ZIP (zero-inflated proabability model) later on

# Response variable - infestation as a binary response

```{r}

# first we need to make a binary infested vs not variable
working$infested <- ifelse(working$`Inf Location` == "-", 0, 1)

# that worked but there are mistakes in the data, let's find them all
# Create a vector indicating whether the conditions are met
conditions_met <- with(working, (infestation_sum != 0 & infested == 1) | (infestation_sum == 0 & infested == 0))

# Print rows where conditions are not met
print(working[!conditions_met, ])

```

# First model

Let's look at the probability of a tree getting infested for the first time based on a group of predictors. In order to do this setup, we'll need to get a response variable that is 1 if infested for the first time on visit 2. So we'll colapse our data into visits, code a new infestation variable and use this as our response. We'll then add in our predictors and work through our modelling process.


## Re-coding
```{r}
# data_with_dis$site_tag <- paste(data_with_dis$site, data_with_dis$tag, sep = "")

host <- data_with_dis %>% filter(species == "ABLA" | species == "ABCO") # Filter out hosts

# live_host <- host %>% filter(`Dead / Alive_Visit1` == 1) # Filter out live hosts


# new_infestation column will be our binary new infestation variable
#host$new_infestation <- host$`Inf Location_Visit2` - host$`Inf Location_Visit1`
#host$new_infestation[host$new_infestation == -1] <- 0

# this change in infestation may be used later (-1 corresponds to trees that lost infestation, 1 gained it, 0 had no change)
#host$changeininfestation <- host$`Inf Location_Visit2` - host$`Inf Location_Visit1`

# tack on distances and bearings to nearest infested
#data_with_dis$site_tag <- paste(data_with_dis$`Site ID_Visit1`, data_with_dis$`Tag #`, sep = "")
#data_with_dis <- data_with_dis %>% distinct(`Site ID_Visit1`, `site_tag`, .keep_all = TRUE) # Get rid of duplicates

```

## Beetles

I want to make one column that's just binary beetles or not and then code a variable that represents our groups

```{r}
unique(host$Beetle_v1)
host$beetle.binary_v1 <- if_else(host$Beetle_v1 == "-" | is.na(host$Beetle_v1), 0, 1)
host$beetle.binary_v2 <- if_else(host$Beetle_v2 == "-" | is.na(host$Beetle_v2), 0, 1)
host$beetle.binary_v3 <- if_else(host$Beetle_v3 == "-" | is.na(host$Beetle_v3), 0, 1)
host$beetle.binary_v4 <- if_else(host$Beetle_v4 == "-" | is.na(host$Beetle_v4), 0, 1)

```

## Model set-up

```{r}
# Think
table(host$new_infestation)

# Looking at bearings
hist(host$bearings1)

# convert to metric and then log variables
host$dbh <- host$dbh * 2.54
host$logDBH <- log(host$dbh)

host$height <- host$height * 0.3048
host$logHeight <- log(host$height)

host$canopy_ht <- host$canopy_ht * 0.3048

host$broom_binary <- ifelse(host$brooms > 0, 1, 0)

host$logInfestsum <- log(host$inf_sum_Visit1_target)
host$logInfestsum2 <- log(host$inf_sum_Visit2_target)
host$logInfestsum3 <- log(host$inf_sum_Visit3_target)

host$Gouting_v1 <- as.ordered(host$Gouting_v1)
host$Gouting_v1 <- ifelse(is.na(host$Gouting_v1), 0, host$Gouting_v1)

newinf <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings1 + inf_sum_Visit1_target,
                    data = host,
                    family = binomial(link = "logit"))
library(DHARMa)
library(lme4)

newinfSims <- simulateResiduals(newinf)
plot(newinfSims)

summary(newinf)

# model with logged covariates and Gouting as an ordered factor

newinflog <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings1 + logInfestsum,
                    data = host,
                    family = binomial(link = "logit"))


newinflogSims <- simulateResiduals(newinflog)
plot(newinflogSims)

summary(newinflog)

# Base model with interaction

newinf_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings1,
                    data = host,
                    family = binomial(link = "logit"))

newinf_interactSims <- simulateResiduals(newinf_interact)
plot(newinf_interactSims)

summary(newinf_interact)

# Logged model with distance and infestation level as an interaction

newinflog_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE * logInfestsum + logDBH + Gouting_v1 + brooms + broom_binary + bearings1,
                    data = host,
                    family = binomial(link = "logit"))

newinflog_interact_sims <- simulateResiduals(newinflog_interact)
plot(newinflog_interact_sims)

summary(newinflog_interact)

anova(newinf, newinflog, newinf_interact, newinflog_interact)

```

# Add in Site as RE

```{r}
library(lme4)

newinf_RE <- glmer(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings1 + inf_sum_Visit1_target + (1|site),
                    data = host,
                    family = binomial(link = "logit"))

# Random effects with logged things

newinf_RE <- glmer(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings1 + logInfestsum + (1|site),
                    data = host,
                    family = binomial(link = "logit"))
#library(DHARMa)

newinfSims <- simulateResiduals(newinf_RE)
plot(newinfSims)

summary(newinf_RE)

anova(newinf, newinflog, newinf_interact, newinflog_interact, newinf_RE)
```

# Just infested GLM for all sites

```{r}

inf_RE <- glmer(cbind(infested_v1, 1 - infested_v1) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings1 + inf_sum_Visit1_target + (1|site),
                    data = host,
                    family = binomial(link = "logit"))

infSims <- simulateResiduals(inf_RE)
plot(infSims)

summary(inf_RE)

```

# Models by site

I'd like to run these models on each site themselves. Maybe by just looking at each site, we may be able to glean different effects of variables. 

```{r}

unique(host$site)
names(host)[1] <- "site"
host$Gouting_v1 <- as.numeric(host$Gouting_v1)
host$dbh[is.na(host$dbh)] <- 1.27 # change clumps to 0.5 in or 1.27cm
CXSP <- host %>% dplyr::filter(site == "CXSP")
EMIG <- host %>% dplyr::filter(site == "EMIG")
SMMH <- host %>% dplyr::filter(site == "SMMH")
FRNK <- host %>% dplyr::filter(site == "FRNK")
MOCR <- host %>% dplyr::filter(site == "MOCR")
MTTP <- host %>% dplyr::filter(site == "MTTP")
SOAP <- host %>% dplyr::filter(site == "SOAP")
TMPF <- host %>% dplyr::filter(site == "TMPF")

# Rescale bearings for each site to reflect general wind patterns taken from RAWS data

# Emig - 127 degrees prevailing wind
original_min <- -180
original_max <- 180
new_center <- 127

# Rescale function
rescale_value <- function(value, original_min, original_max, new_center) {
  scaled_value <- (value - (original_max + original_min) / 2) + new_center
  return(scaled_value)
}

EMIG$bearings_scaled1 <- rescale_value(EMIG$bearings1, original_min, original_max, new_center)

# CXSP - 215 degrees
new_center <- 215
CXSP$bearings_scaled1 <- rescale_value(CXSP$bearings1, original_min, original_max, new_center)

# FRNK - 274 degrees
new_center <- 274
FRNK$bearings_scaled1 <- rescale_value(FRNK$bearings1, original_min, original_max, new_center)

# MOCR - 282 degrees
new_center <- 282
MOCR$bearings_scaled1 <- rescale_value(MOCR$bearings1, original_min, original_max, new_center)

# MTTP - 129 degrees
new_center <- 129
MTTP$bearings_scaled1 <- rescale_value(MTTP$bearings1, original_min, original_max, new_center)

# SMMH - 279 degrees
new_center <- 279
SMMH$bearings_scaled1 <- rescale_value(SMMH$bearings1, original_min, original_max, new_center)

# SOAP - 282 degrees
new_center <- 282
SOAP$bearings_scaled1 <- rescale_value(SOAP$bearings1, original_min, original_max, new_center)

# TMPF - 274 degrees
new_center <- 274
TMPF$bearings_scaled1 <- rescale_value(TMPF$bearings1, original_min, original_max, new_center)

```

## CXSP

```{r}
# base model - NOTHING

newinf_cxsp <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target + beetle.binary_v1,
                    data = CXSP,
                    family = binomial(link = "logit"))

newinf_cxspSims <- simulateResiduals(newinf_cxsp)
plot(newinf_cxspSims)

summary(newinf_cxsp)

# logged variables and binary broom - NOTHING

cxsp_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = CXSP,
                    family = binomial(link = "logit"))

cxsp_logSims <- simulateResiduals(cxsp_log)
plot(cxsp_logSims)

summary(cxsp_log)

# distance and infest sum as interaction - distnace and interaction appear significant

newinf_cxsp_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = CXSP,
                    family = binomial(link = "logit"))

newinf_cxsp_interactSims <- simulateResiduals(newinf_cxsp_interact)
plot(newinf_cxsp_interactSims)

summary(newinf_cxsp_interact)

anova(newinf_cxsp, newinf_cxsp_interact, newinf_cxsp_interactSims, newinf_cxsp_interactSims)

```

## EMIG

```{r}
# base model - Not working for some reason?? coming out as "large glm" - it was gouting with NA's that was the problem!!

#unique(EMIG$dbh) # NA's - how to deal with this??
#unique(EMIG$Gouting_v1)

# We want to treat NA's as a seperate category, but not disregard them. Here, we'll switch out NA's with 0.5 to give it a numerical value
#EMIG$dbh <- ifelse(is.na(EMIG$dbh), 0.5, EMIG$dbh)

# Deal with Gouting NA's
EMIG$Gouting_v1 <- ifelse(is.na(EMIG$Gouting_v1), 0, EMIG$Gouting_v1)
unique(EMIG$Gouting_v1)

# Base model - infestation sum of nearest infested significant
newinf_emig <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = EMIG,
                    family = binomial(link = "logit"))

newinf_emigSims <- simulateResiduals(newinf_emig)
plot(newinf_emigSims)

summary(newinf_emig)

# logged variables and binary broom - same as base

emig_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = EMIG,
                    family = binomial(link = "logit"))

emig_logSims <- simulateResiduals(emig_log)
plot(emig_logSims)

summary(emig_log)

# distance and infest sum as interaction - distance appears significant as well as sum of nearest infested

newinf_emig_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = EMIG,
                    family = binomial(link = "logit"))

newinf_emig_interactSims <- simulateResiduals(newinf_emig_interact)
plot(newinf_emig_interactSims)

summary(newinf_emig_interact)

# Logged model with interaction - distance, interanction, and brooms come out

emig_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * logInfestsum) + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1,
                    data = EMIG,
                    family = binomial(link = "logit"))

emig_loginteractSims <- simulateResiduals(emig_log_interact)
plot(emig_loginteractSims)

summary(emig_log_interact)

```

## SMMH

```{r}
# base model - NOTHING

newinf_smmh <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = SMMH,
                    family = binomial(link = "logit"))

newinf_smmhSims <- simulateResiduals(newinf_smmh)
plot(newinf_smmhSims)

summary(newinf_smmh)

# logged variables and binary broom - NOTHING

smmh_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = SMMH,
                    family = binomial(link = "logit"))

smmh_logSims <- simulateResiduals(smmh_log)
plot(smmh_logSims)

summary(smmh_log)

# distance and infest sum as interaction - nothing

newinf_smmh_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = SMMH,
                    family = binomial(link = "logit"))

newinf_smmh_interactSims <- simulateResiduals(newinf_smmh_interact)
plot(newinf_smmh_interactSims)

summary(newinf_smmh_interact)

# logged variables, binary broom, and interaction - Nothing

smmh_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * logInfestsum) + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1,
                    data = SMMH,
                    family = binomial(link = "logit"))

smmh_loginteractSims <- simulateResiduals(smmh_log_interact)
plot(smmh_loginteractSims)

summary(smmh_log_interact)
```

## FRNK

```{r}
# base model - DBH coming out relatively strong

unique(FRNK$Gouting_v1)
FRNK$Gouting_v1 <- ifelse(is.na(FRNK$Gouting_v1), 0, FRNK$Gouting_v1)

newinf_frnk <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = FRNK,
                    family = binomial(link = "logit"))

newinf_frnkSims <- simulateResiduals(newinf_frnk)
plot(newinf_frnkSims)

summary(newinf_frnk)

# logged variables and binary broom - logDBH coming out strong

frnk_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = FRNK,
                    family = binomial(link = "logit"))

frnk_logSims <- simulateResiduals(frnk_log)
plot(frnk_logSims)

summary(frnk_log)

# distance and infest sum as interaction - just DBH still

newinf_frnk_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = FRNK,
                    family = binomial(link = "logit"))

newinf_frnk_interactSims <- simulateResiduals(newinf_frnk_interact)
plot(newinf_frnk_interactSims)

summary(newinf_frnk_interact)

# logged variables, binary broom, and interaction - logDBH again

frnk_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * logInfestsum) + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1,
                    data = FRNK,
                    family = binomial(link = "logit"))

frnk_loginteractSims <- simulateResiduals(frnk_log_interact)
plot(frnk_loginteractSims)

summary(frnk_log_interact)
```

## MOCR

```{r}
# base model - NOTHING

newinf_mocr <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = MOCR,
                    family = binomial(link = "logit"))

newinf_mocrSims <- simulateResiduals(newinf_mocr)
plot(newinf_mocrSims)

summary(newinf_mocr)

# logged variables and binary broom - NOTHING

mocr_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = MOCR,
                    family = binomial(link = "logit"))

mocr_logSims <- simulateResiduals(mocr_log)
plot(mocr_logSims)

summary(mocr_log)

# distance and infest sum as interaction

newinf_mocr_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = MOCR,
                    family = binomial(link = "logit"))

newinf_mocr_interactSims <- simulateResiduals(newinf_mocr_interact)
plot(newinf_mocr_interactSims)

summary(newinf_mocr_interact)

# logged variables, binary broom and interaction - NOTHING

mocr_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * logInfestsum) + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1,
                    data = MOCR,
                    family = binomial(link = "logit"))

mocr_loginteractSims <- simulateResiduals(mocr_log_interact)
plot(mocr_loginteractSims)

summary(mocr_log_interact)
```

## MTTP

```{r}
# base model - NOTHING

newinf_mttp <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = MTTP,
                    family = binomial(link = "logit"))

newinf_mttpSims <- simulateResiduals(newinf_mttp)
plot(newinf_mttpSims)

summary(newinf_mttp)

# logged variables and no binary broom (small n) - NOTHING

mttp_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + bearings_scaled1 + logInfestsum,
                    data = MTTP,
                    family = binomial(link = "logit"))

mttp_logSims <- simulateResiduals(mttp_log)
plot(mttp_logSims)

summary(mttp_log)

# distance and infest sum as interaction

newinf_mttp_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = MTTP,
                    family = binomial(link = "logit"))

newinf_mttp_interactSims <- simulateResiduals(newinf_mttp_interact)
plot(newinf_mttp_interactSims)

summary(newinf_mttp_interact)

# logged variables, no binary broom (small sample size I think) and interaction - NOTHING

mttp_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * logInfestsum) + logDBH + Gouting_v1 + brooms + bearings_scaled1,
                    data = MTTP,
                    family = binomial(link = "logit"))

mttp_loginteractSims <- simulateResiduals(mttp_log_interact)
plot(mttp_loginteractSims)

summary(mttp_log_interact)
```

## SOAP

```{r}
# base model
unique(SOAP$Gouting_v1)
SOAP$Gouting_v1 <- ifelse(is.na(SOAP$Gouting_v1), 0, SOAP$Gouting_v1)

# Gouting won't go into this site because there is no gouting observed atm

newinf_soap <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = SOAP,
                    family = binomial(link = "logit"))

newinf_soapSims <- simulateResiduals(newinf_soap)
plot(newinf_soapSims)

summary(newinf_soap)

# logged variables and binary broom

soap_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = SOAP,
                    family = binomial(link = "logit"))

soap_logSims <- simulateResiduals(soap_log)
plot(soap_logSims)

summary(soap_log)

# distance and infest sum as interaction

newinf_soap_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + brooms + bearings_scaled1,
                    data = SOAP,
                    family = binomial(link = "logit"))

newinf_soap_interactSims <- simulateResiduals(newinf_soap_interact)
plot(newinf_soap_interactSims)

summary(newinf_soap_interact)

# logged variables, binary broom and interaction - logDBH again

soap_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * logInfestsum) + logDBH + brooms + broom_binary + bearings_scaled1,
                    data = SOAP,
                    family = binomial(link = "logit"))

soap_loginteractSims <- simulateResiduals(soap_log_interact)
plot(soap_loginteractSims)

summary(soap_log_interact)
```
LogDBH coming out slightly significant

## TMPF

```{r}
# base model
unique(TMPF$Gouting_v1)
TMPF$Gouting_v1 <- ifelse(is.na(TMPF$Gouting_v1), 0, TMPF$Gouting_v1)

newinf_tmpf <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + dbh + Gouting_v1 + brooms + bearings_scaled1 + inf_sum_Visit1_target,
                    data = TMPF,
                    family = binomial(link = "logit"))

newinf_tmpfSims <- simulateResiduals(newinf_tmpf)
plot(newinf_tmpfSims)

summary(newinf_tmpf)

# logged variables and binary broom

tmpf_log <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    DISTANCE + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1 + logInfestsum,
                    data = TMPF,
                    family = binomial(link = "logit"))

tmpf_logSims <- simulateResiduals(tmpf_log)
plot(tmpf_logSims)

summary(tmpf_log)

# distance and infest sum as interaction

newinf_tmpf_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE * inf_sum_Visit1_target) + dbh + Gouting_v1 + brooms + bearings_scaled1,
                    data = TMPF,
                    family = binomial(link = "logit"))

newinf_tmpf_interactSims <- simulateResiduals(newinf_tmpf_interact)
plot(newinf_tmpf_interactSims)

summary(newinf_tmpf_interact)

# logged variables, binary broom and interaction

tmpf_log_interact <- glm(cbind(new_infestation, 1 - new_infestation) ~
                    (DISTANCE *  logInfestsum) + logDBH + Gouting_v1 + brooms + broom_binary + bearings_scaled1,
                    data = TMPF,
                    family = binomial(link = "logit"))

tmpf_loginteractSims <- simulateResiduals(tmpf_log_interact)
plot(tmpf_loginteractSims)

summary(tmpf_log_interact)

```
# Semivariograms for each site intercept only or naive models

## CXSP

```{r}
names(CXSP)[8] <- "infested"

coordinates.cxsp <- CXSP[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.cxsp <- CXSP$infested

# Create a SpatialPointsDataFrame
cxsp_sp <- SpatialPointsDataFrame(coords = coordinates.cxsp, data = data.frame(infestation.cxsp), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_cxsp <- variogram(infestation.cxsp ~ 1, data = cxsp_sp)
class(variogram_cxsp)

# Plot the variogram
plot(variogram_cxsp, main = "cxsp_variogram")

# refit with autocorrelation fxns
# get rid of any duplicate locations for spatial autocorrelation
CXSP_nodupe <- CXSP[!duplicated(CXSP[, c("Latitude.focal", "Longitude.focal")]), ]
## spherical
cxsp_sph <- gls(infested ~ 1, data = CXSP_nodupe, correlation = corSpher(value = c(0.02,
                                                                            0.14),
                                                                          nugget = T))

## Linear
cxsp_lin <- gls(infested ~ 1, data = CXSP_nodupe, correlation = corLin(value = c(0.02,
                                                                          0.14),
                                                          form = ~ Latitude.focal + Longitude.focal, nugget = T))

## Gaussian
cxsp_gaus <- gls(infested ~ 1, data = CXSP_nodupe, correlation = corGaus(value = c(0.02,
                                                                          0.14),
                                                          form = ~ Latitude.focal + Longitude.focal, nugget = T))

## Exponential 
cxsp_exp <- gls(infested ~ 1, data = CXSP_nodupe, correlation = corExp(value = c(0.02,
                                                                          0.14),
                                                          form = ~ Latitude.focal + Longitude.focal, nugget = T))

## Ratio
cxsp_ratio <- gls(infested ~ 1, data = CXSP_nodupe, correlation = corRatio(value = c(0.02,
                                                                          0.14),
                                                          form = ~ Latitude.focal + Longitude.focal, nugget = T))

## Compare
summary(cxsp_sph)
anova(cxsp_lin, cxsp_gaus, cxsp_exp, cxsp_ratio)


```
## EMIG

```{r}
names(EMIG)[8] <- "infested"

coordinates.emig <- EMIG[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.emig <- EMIG$infested

# Create a SpatialPointsDataFrame
emig_sp <- SpatialPointsDataFrame(coords = coordinates.emig, data = data.frame(infestation.emig), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_emig <- variogram(infestation.emig ~ 1, data = emig_sp)

# Plot the variogram
plot(variogram_emig)

# refit with autocorrelation fxns


```

## FRNK

```{r}
names(FRNK)[8] <- "infested"

coordinates.frnk <- FRNK[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.frnk <- FRNK$infested

# Create a SpatialPointsDataFrame
frnk_sp <- SpatialPointsDataFrame(coords = coordinates.frnk, data = data.frame(infestation.frnk), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_frnk <- variogram(infestation.frnk ~ 1, data = frnk_sp)

# Plot the variogram
plot(variogram_frnk)

# refit with autocorrelation fxns


```
## MOCR

```{r}
names(MOCR)[8] <- "infested"

coordinates.mocr <- MOCR[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.mocr <- MOCR$infested

# Create a SpatialPointsDataFrame
mocr_sp <- SpatialPointsDataFrame(coords = coordinates.mocr, data = data.frame(infestation.mocr), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_mocr <- variogram(infestation.mocr ~ 1, data = mocr_sp)

# Plot the variogram
plot(variogram_mocr)

# refit with autocorrelation fxns


```

## MTTP

```{r}
names(MTTP)[8] <- "infested"

coordinates.mttp <- MTTP[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.mttp <- MTTP$infested

# Create a SpatialPointsDataFrame
mttp_sp <- SpatialPointsDataFrame(coords = coordinates.mttp, data = data.frame(infestation.mttp), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_mttp <- variogram(infestation.mttp ~ 1, data = mttp_sp)

# Plot the variogram
plot(variogram_mttp)

# refit with autocorrelation fxns


```

## SMMH

```{r}
names(SMMH)[8] <- "infested"

coordinates.smmh <- SMMH[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.smmh <- SMMH$infested

# Create a SpatialPointsDataFrame
smmh_sp <- SpatialPointsDataFrame(coords = coordinates.smmh, data = data.frame(infestation.smmh), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_smmh <- variogram(infestation.smmh ~ 1, data = smmh_sp)

# Plot the variogram
plot(variogram_smmh)

# refit with autocorrelation fxns


```

## SOAP

```{r}
names(SOAP)[8] <- "infested"

coordinates.soap <- SOAP[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.soap <- SOAP$infested

# Create a SpatialPointsDataFrame
soap_sp <- SpatialPointsDataFrame(coords = coordinates.soap, data = data.frame(infestation.soap), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_soap <- variogram(infestation.soap ~ 1, data = soap_sp)

# Plot the variogram
plot(variogram_soap)

# refit with autocorrelation fxns


```
## TMPF

```{r}
names(TMPF)[8] <- "infested"

coordinates.tmpf <- TMPF[, c("Latitude.focal", "Longitude.focal")] # create coordinate dataframe
infestation.tmpf <- TMPF$infested

# Create a SpatialPointsDataFrame
tmpf_sp <- SpatialPointsDataFrame(coords = coordinates.tmpf, data = data.frame(infestation.tmpf), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_tmpf <- variogram(infestation.tmpf ~ 1, data = tmpf_sp)

# Plot the variogram
plot(variogram_tmpf)

# refit with autocorrelation fxns


```

# Spatial set-up

In this section we'll attempt to analyze the plots spatially in different ways. First, we'll try a good ol' Poisson point pattern. This involves laying a grid over the plot every 5 to ten meters, which creates a point pattern and then we can look at means across a site for each "section" and look at trends. I want to look at infestation sum vs prevalence FIRST, because that seems the most interesting and obvious thread to pull. We can then look at other variables that we've brought up before.

** I ended up doing this in the grid data markdown sheet, head over there to see results!!

# Intensity models

What I want to do now is look at intensity as a response variable. I think I want to try Justin's method of slowly adding in variables to see how our variance changes with the addition of variables. Basically: we start simple and get more complex. We'll start by looking at intensity as a funciton of DBH, then add in other tree characteristics like brooms, gouting, then add in other tree characteristics such as nearest neighbor variables.

This is on Posit browser since this version of R can't support the packages needed to run mixed models

## All sites

```{r}
library(glmmTMB)

hist(host$DISTANCE)
host$logDISTANCE <- log(host$DISTANCE)
host$absvalbearings <- abs(host$bearings1)

intensity.glm <- glmmTMB(infestation_sum_Visit1_focal ~ 
                      logDISTANCE + dbh + Gouting_v1 + brooms + absvalbearings + (1|site),
                      family = nbinom1,
                      data = host)

# simulate residuals and not plot
intensitySims <- simulateResiduals(intensity.glm)
plot(intensitySims)

testZeroInflation(intensitySims)

summary(intensity.glm)

# Let's use continuation ratio with logit link function and see how that fits 

library(VGAM) # let's us use continuation ratio function



```

# Intensity in a Bayesian Framework

```{r}
library(glmmTMB)

intensity.glm <- glmmTMB(infestation_sum_Visit1_focal ~ 
                      log(DISTANCE) + dbh + Gouting_v1 + brooms + abs(bearings1) + (1|site),
                      family = nbinom1,
                      data = host)

# Get rid of NA's for model
host_no_na <- host %>% dplyr::filter(!is.na(dbh))

intensity_jags_data <- list(infest_sum = host_no_na$infestation_sum_Visit1_focal, 
                            logDISTANCE = host_no_na$logDISTANCE,
                            # note that we have to convert the FoodTreatment factor to
                            # a number to work appropriately in JAGS. Also, numeric coding needs to start at 0, not 1
                            DBH_focal = host_no_na$dbh, 
                            Gouting_focal = host_no_na$Gouting_v1,
                            Brooms_focal = host_no_na$brooms,
                            # same with SexParent. Also, numeric coding needs to start at 0
                            N = nrow(host_no_na))

intensity_jags_naive <- function(){
  # Priors:
  beta_0 ~ dnorm(0, 0.01) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.01) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.01)
  beta_gouting ~ dnorm(0, 0.01)
  beta_brooms ~ dnorm(0, 0.01)
  # Note that we don't have any "tau" or "sigma" here, since we're taking the data to be Poisson, so
  # if we know lambda, we know both the mean and the variance. 
  
  # Likelihood:
  for (i in 1:N){
    lin_comb[i] <- beta_0 + beta_distance * logDISTANCE[i] + beta_dbh * DBH_focal[i] +
      beta_gouting * Gouting_focal[i] + beta_brooms * Brooms_focal[i]
    # here's our linear combination of predictors, just like in a standard regression
    lambda[i] <- exp(lin_comb[i]) # here we apply the link function to that 
    # linear combination to get us back to the scale of lambda.
    infest_sum[i] ~ dpois(lambda[i]) # Here's the Poisson likelihood,
    # with a sampling-unit-specific value of the rate parameter, lambda[i].
  }
}

intensity_inits <- function(){
  list(beta_0 = rnorm(1, 0, 1), 
       beta_distance = rnorm(1, 0, 1),
       beta_dbh = rnorm(1, 0, 1),
       beta_gouting = rnorm(1, 0, 1),
       beta_brooms = rnorm(1, 0, 1))
}

intensity_params <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms")

library(R2jags)

intensity_fit_naive <- jags(data = intensity_jags_data, 
                            inits = intensity_inits, 
                            parameters.to.save = intensity_params, 
                            model.file = intensity_jags_naive,
                            n.chains = 3,
                            n.iter = 12000, 
                            n.burnin = 2000, 
                            n.thin = 10, 
                            DIC = F)
intensity_fit_naive

intensity.glm <- glm(infestation_sum_Visit1_focal ~ 
                      logDISTANCE + dbh + Gouting_v1 + brooms,
                      family = poisson,
                      data = host_no_na)

confint(intensity.glm) 
# These intervals are pretty close to the 2.5th and 97.5th quantiles of beta_food 
# and beta_parent in our Bayesian model. 

# check for convergence
traceplot(intensity_fit_naive, mfrow = c(2, 2), ask = F)

intensity_mcmc <- as.mcmc(intensity_fit_naive)
plot(intensity_mcmc)

# Add in Random Effect

intensity.glm <- glmmTMB(infestation_sum_Visit1_focal ~ 
                      logDISTANCE + dbh + Gouting_v1 + brooms + (1|site),
                      family = nbinom1,
                      data = host)

# Get rid of NA's for model

pow <- function(base, exponent) {
  return(base^exponent)
}

intensity_jags_data_re <- list(infest_sum = host_no_na$infestation_sum_Visit1_focal, 
                            logDISTANCE = host_no_na$logDISTANCE,
                            # note that we have to convert the FoodTreatment factor to
                            # a number to work appropriately in JAGS. Also, numeric coding needs to start at 0, not 1
                            DBH_focal = host_no_na$dbh, 
                            Gouting_focal = host_no_na$Gouting_v1,
                            Brooms_focal = host_no_na$brooms,
                            site = as.numeric(as.factor(host_no_na$site)),
                            # same with SexParent. Also, numeric coding needs to start at 0
                            N = nrow(host_no_na),
                            N_site = length(levels(factor(host_no_na$site))))

intensity_jags_naive_re <- function(){
  # Priors:
  beta_0 ~ dnorm(0, 0.01) # intercept (use diffuse normal just as we would for standard regression)
  beta_distance ~ dnorm(0, 0.01) # slope (use diffuse normal just as we would for standard regression)
  beta_dbh ~ dnorm(0, 0.01)
  beta_gouting ~ dnorm(0, 0.01)
  beta_brooms ~ dnorm(0, 0.01)
  # Note that we don't have any "tau" or "sigma" here, since we're taking the data to be Poisson, so
  # if we know lambda, we know both the mean and the variance. 
  
  for(i in 1:N_site){
    beta_site[i] ~ dnorm(0, tau_site)
  }
  
  sigma_sq_site ~ dunif(0, 100) # here's my prior to nest variance
  tau_site <- pow(sigma_sq_site,-2) # here, I'm converting from that variance to the precision (tau = 1/sqrt(var))
  
  # prior for dispersion parameter
  r ~ dunif(0, 50) # r is the dispersion parameter. As r -> infinity, 
  # the negative binomial goes to a Poisson, so this prior is not particularly
  # restrictive (once r is 50, we're essentially working with something very like a Poisson)
  
  # Likelihood:
  for (i in 1:N){
    lin_comb[i] <- beta_0 + beta_site[site[i]] + beta_distance * logDISTANCE[i] + beta_dbh * DBH_focal[i] +
      beta_gouting * Gouting_focal[i] + beta_brooms * Brooms_focal[i]
    # here's our linear combination of predictors, just like in a standard regression
    lambda[i] <- exp(lin_comb[i]) # here we apply the link function to that 
    # linear combination to get us back to the scale of lambda.
     p[i] <- r / (r + lambda[i]) # here's where we incorporate the dispersion parameter.
    infest_sum[i] ~ dnegbin(p[i], r) # Here's the Neg Bin likelihood,
    # with a sampling-unit-specific value of the rate parameter, lambda[i].
  }
}

intensity_inits_re <- function(){
  list(beta_0 = rnorm(1, 0, 1), 
       beta_distance = rnorm(1, 0, 1),
       beta_dbh = rnorm(1, 0, 1),
       beta_gouting = rnorm(1, 0, 1),
       beta_brooms = rnorm(1, 0, 1),
       beta_site = rep(0, length(levels(factor(host_no_na$site)))),
       r = runif(1, 0, 50))
}

intensity_params_re <- c("beta_0", "beta_distance", "beta_dbh", "beta_gouting", "beta_brooms", "beta_site", "sigma_sq_site", "r")

intensity_fit_naive_re <- jags(data = intensity_jags_data_re, 
                            inits = intensity_inits_re, 
                            parameters.to.save = intensity_params_re, 
                            model.file = intensity_jags_naive_re,
                            n.chains = 3,
                            n.iter = 12000, 
                            n.burnin = 2000, 
                            n.thin = 10, 
                            DIC = F)
intensity_fit_naive_re

intensity.glm <- glmmTMB(infestation_sum_Visit1_focal ~ 
                      logDISTANCE + dbh + Gouting_v1 + brooms + (1|site),
                      family = nbinom1,
                      data = host)

confint(intensity.glm) 
# These intervals are pretty close to the 2.5th and 97.5th quantiles of beta_food 
# and beta_parent in our Bayesian model. 

# check for convergence
traceplot(intensity_fit_naive_re, mfrow = c(2, 2), ask = F)

dev.off()

intensity_mcmc_re <- as.mcmc(intensity_fit_naive_re)
plot(intensity_mcmc_re)
```

## CXSP

```{r}

intensity.cxsp <- glm(infestation_sum_Visit1_focal ~ 
                      DISTANCE + dbh + Gouting_v1 + brooms + bearings1 + beetle.bianry_v1,
                      family = poisson,
                      data = CXSP)
plot(intensity.cxsp)
summary(intensity.cxsp)

```

## EMIG

NO beetle included because there were no beetle presence in live trees at this site

```{r}

intensity.emig <- glm(infestation_sum_Visit1_focal ~ 
                      DISTANCE + dbh + Gouting_v1 + brooms + bearings1,
                      family = poisson,
                      data = EMIG)
plot(intensity.emig)
summary(intensity.emig)

```


