---
title: "Local Statistics"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

New Infested Trees
3/1/23 * approx date
Michael Wayman

The goal of this script is to look at our newly infested trees at each plot and what characteristics may be playing a part in their transition. We'll look at all of our variables that we've collected and see if any of them correlate with being infested for the first time. For most of the script, we'll just be using our measured variables excluding spatial aspects. At the end, we may try to incorporate the spatial element and see how that changes things. Let's do it!

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(rfishnet2)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(plyr)
library(reshape2)
library(geosphere)
library(data.table)
library(writexl)
```

## Clean out global environment

```{r}
rm(list = ls())
```

## Load Data

```{r}
combined <- read_xlsx("data_with_bearings.xlsx")
```

## Final Solution to distances


```{r}

# So we need to convert our decimel degrees to UTM for calculation
all_merged <- combined

# Convert to sf
all_merged_sf <- st_as_sf(all_merged, coords = c("Longitude.focal", "Latitude.focal"), crs = 4326)

# Define the target UTM zone
target_crs <- 32612  # EPSG code for UTM zone 12

# Transform the coordinates
all_merged_utm <- st_transform(all_merged_sf, crs = target_crs)
all_coords <- st_coordinates(all_merged_utm)
all_coords <- as.data.frame(all_coords)

# Put back into original datframe

all_merged_utm <- all_merged
all_merged_utm$Longitude.focal <- all_coords$X
all_merged_utm$Latitude.focal <- all_coords$Y

#all_merged_utm$unique_ID <- paste(all_merged_utm$site, all_merged_utm$`Tag #`, rownames(all_merged_utm), sep = "_")

all_infested <- all_merged_utm[all_merged_utm$infestation_visit_1 == 1, ] # Infested trees with UTM coordinates


#write_clip(all_merged_utm) # copy dataframe without making file
all_merged <- all_merged_utm[!duplicated(all_merged_utm$unique_ID), ] #- there shouldn't be duplicates now
rownames(all_merged) <- all_merged$unique_ID

# change crown height NA's to zero
all_merged$canopy_ht[is.na(all_merged$canopy_ht)] <- 0

rm(Holder)
Holder = as.data.frame(matrix(nrow=0,ncol=4)) 
colnames(Holder) =c("PLOT", "FROM", "TARGET", "DISTANCE")

#Assign rownames as Tree-ID, dataset also needs UTM_X, UTM_Y, and PLOT
for( i in levels(as.factor(all_merged$site))){
  subset_frame = subset(all_merged, site == i)
TEMP_DIST = vegdist(subset_frame[,c("Longitude.focal","Latitude.focal")], method = "euclid")
TEMP_DIST.df = as.data.frame(as.matrix(TEMP_DIST))
rownames(TEMP_DIST.df) = (subset_frame$unique_ID)
colnames(TEMP_DIST.df) = (subset_frame$unique_ID)
TEMP_DIST.df$`Tag #` = rownames(TEMP_DIST.df)
TEMP_DIST_MELT = reshape2::melt(TEMP_DIST.df, id.vars = c("Tag #"))
colnames(TEMP_DIST_MELT) = c("FROM", "TARGET", "DISTANCE")
#Make a new column with plot ID
TEMP_DIST_MELT$PLOT = i
Holder = rbind(Holder, TEMP_DIST_MELT) #This is your exit
}

#Merge with infestation data
names(Holder)[1] <- "unique_ID"

Holder_new <- merge(Holder, all_merged_utm, by.x = "TARGET", by.y = "unique_ID")
holder_new.dt <- as.data.table(Holder_new)
length(unique(holder_new.dt$unique_ID))
nrow(holder_new.dt)
names(holder_new.dt)[12] <- "infested_1"
names(holder_new.dt)[39] <- "infested_2"
library(data.table)
#min_dist_holder <- holder_new.dt[`infested` == 1 & (DISTANCE == min(DISTANCE[(DISTANCE != 0)])) ,.(DISTANCE, TARGET),by=.("unique_ID")]

# Get all trees under 10m from focal tree
under10_holder <- holder_new.dt[,.(TARGET, DISTANCE.x), by=.(unique_ID)][,.SD[which(DISTANCE.x <= 10 & DISTANCE.x != 0)], by=.(unique_ID)]

# add distances to data
# visit 1 distances
#data_with_nearest_infested_dis <- all_merged_utm %>% inner_join(min_dist_holder, by = "unique_ID")

#write_xlsx(data_with_nearest_infested_dis,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/data_with_nearest_infested.xlsx")

```

# Function to extract info from target trees

```{r}
# Infestation sum because its not here for some reason
all_merged_utm$infestation_sum_v1 <- all_merged_utm$NL_v1 + all_merged_utm$NM_v1 + all_merged_utm$NH_v1 + all_merged_utm$SL_v1 + all_merged_utm$SM_v1 + all_merged_utm$SH_v1

all_merged_utm$infestation_sum_v2 <- all_merged_utm$NL_v2 + all_merged_utm$NM_v2 + all_merged_utm$NH_v2 + all_merged_utm$SL_v2 + all_merged_utm$SM_v2 + all_merged_utm$SH_v2

all_merged_utm$infestation_sum_v3 <- all_merged_utm$NL_v3 + all_merged_utm$NM_v3 + all_merged_utm$NH_v3 + all_merged_utm$SL_v3 + all_merged_utm$SM_v3 + all_merged_utm$SH_v3

all_merged_utm$infestation_sum_v4 <- all_merged_utm$NL_v4 + all_merged_utm$NM_v4 + all_merged_utm$NH_v4 + all_merged_utm$SL_v4 + all_merged_utm$SM_v4 + all_merged_utm$SH_v4

# New way to do this
all_merged_utm$dbh <- all_merged_utm$dbh * 2.54 # convert DBH to cm
all_merged_utm <- mutate(all_merged_utm, BA = (0.00007854 * (dbh ^ 2)))

df <- under10_holder
# Create a function to extract the variable based on the linked tree ID
extractVariable <- function(TARGET, variable) {
  linkedTreeRow <- all_merged_utm[all_merged_utm$unique_ID == TARGET, ]
  if (nrow(linkedTreeRow) > 0) {
    return(linkedTreeRow[[variable]])
  } else {
    return(NA)  # or any other value to indicate missing data
  }
}

str(all_merged_utm)

# Apply the function to create a new column with the extracted variable - need to change function for each extraction !!
df$infestation_under10 <- sapply(df$TARGET, extractVariable, variable = "infestation_visit_1")
df$infestation.sum.1_under10 <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_v1")
df$dbh_under10 <- sapply(df$TARGET, extractVariable, variable = "dbh")
df$brooms_under10 <- sapply(df$TARGET, extractVariable, variable = "brooms")
df$gouting_under10 <- sapply(df$TARGET, extractVariable, variable = "Gouting_v1")
df$species_under10 <- sapply(df$TARGET, extractVariable, variable = "species")
df$BA_under10 <- sapply(df$TARGET, extractVariable, variable = "BA")
df$crown_ht__under10 <- sapply(df$TARGET, extractVariable, variable = "canopy_ht")
df$dead_alive <- sapply(df$TARGET, extractVariable, variable = "Dead / Alive_v1")

#combined$
# Need to run this to get rid of duplicates before extracting DBH's
#df <- df %>%
      #distinct(unique_ID, .keep_all = TRUE)

# Apply the function to create a new column with the extracted variable - need to change function for each extraction !!
# df$infestation_under10 <- sapply(df$TARGET, extractVariable)
# df$infestation.sum.1_under10 <- sapply(df$TARGET, extractVariable)
# df$dbh_under10 <- sapply(df$TARGET, extractVariable)
# df$brooms_under10 <- sapply(df$TARGET, extractVariable)
# df$gouting_under10 <- sapply(df$TARGET, extractVariable)

# create a summary table of local stats for each tree
local.stats <- df %>% 
  dplyr::group_by(unique_ID) %>% 
  dplyr::summarise(n = n(),
                   number.infested_under10 = sum(infestation_under10),
                   percent.infested_under10 = (sum(infestation_under10)/n) * 100,
                   infestation.sum_under10 = sum(infestation.sum.1_under10),
                   average.dbh_under10 = mean(dbh_under10, na.rm = TRUE),
                   brooms_under10 = sum(brooms_under10),
                   avg.gouting_under10 = mean(gouting_under10, na.rm = TRUE),
                   gouting.sum_under = sum(gouting_under10, na.rm = TRUE),
                   avg_crown_ht_under_10 = mean(crown_ht__under10), na.rm = TRUE)



# combine with data
data_with_local_stats <- combined %>% inner_join(local.stats, by = "unique_ID")
local <- data_with_local_stats

library(writexl)

#write_xlsx(data_with_local_stats, "under10m_data.xlsx")
```

# Modelling with local data

## Histograms of local stats

```{r}

hist(local$n)
hist(local$number.infested_under10)
hist(local$percent.infested_under10) # use log percent infested
hist(local$infestation.sum_under10) # use log infest sum
hist(local$average.dbh_under10) # use log dbh
hist(local$brooms_under10) # use log brooms
hist(local$avg.gouting_under10)
hist(local$gouting.sum_under) # use log gouting sum

# Logs look better, but give us infinate values for all these categories, how do you deal with this??


# Correlations
cor(local[, c("n", "number.infested_under10", "percent.infested_under10", "infestation.sum_under10", "average.dbh_under10", "brooms_under10", "avg.gouting_under10", "gouting.sum_under")], use = "complete.obs")

# Seems like the main one to avoid is number infested/infestation sum - let's use percent infested and infestation sum. By using percent, we get number infested when we input n.

```

## All sites newly infested

```{r}
# change infestation variable

local.newinf <- glm(cbind(new_infestation, 1 - new_infestation) ~ n + percent.infested_under10 + infestation.sum_under10 + average.dbh_under10 + brooms_under10 + avg.gouting_under10 + gouting.sum_under, data = local, family = binomial(link = "logit"))

local.new.sims <- simulateResiduals(local.newinf)
plot(local.new.sims)

summary(local.newinf)

```

## All sites - just infested vs not

```{r}

local.inf <- glm(cbind(infested_v1, 1 - infested_v1) ~ n + percent.infested_under10 + infestation.sum_under10 + average.dbh_under10 + brooms_under10 + avg.gouting_under10 + gouting.sum_under, data = local, family = binomial(link = "logit"))

local.inf.sims <- simulateResiduals(local.inf)
plot(local.inf.sims)

summary(local.inf)
```
