---
title: "More GLMs"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Full Site Model
4/11/24
Michael Wayman

This script is going to be an attempt to coalesce all the sites into GLM's that incorporate local stats. This will be a reflection of the Bayesian implementation, but for all sites and each individual site. I'm doing a new markdown so that I can get all the distances and local things calculated in one script, which will hopefully make things easier down the line.

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

# Clean out environment

```{r}
rm(list = ls())
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(lme4) # to use glmer
library(data.table)
library(R2jags)
library(data.table)
library(rfishnet2)
library(glmmTMB)
library(DHARMa)
```

# Load data

```{r}

combined <- read_xlsx("combined_cleaned.xlsx")
points <- read.csv("all_points.csv", header = T)
rest_of_points <- read.csv("rest_of_points.csv", header = T)

points <- points %>% bind_rows(rest_of_points)

```

# Collapse visits and combine points

```{r}
names(combined)[1] <- "site"

# Change Infestation to binary variable
combined$`Inf Location`[combined$`Inf Location`== "-"] <- 0
combined$`Inf Location`[combined$`Inf Location` == "BO" | combined$`Inf Location` == "BR, BO" | combined$`Inf Location` == "BR"] <- 1
unique(combined$`Inf Location`)
combined$`Inf Location` <-  if_else(combined$`Inf Location` == 1, 1, 0)
combined$`Inf Location`[is.na(combined$`Inf Location`)] <- 0

# First we'll subset by Visit
visit1 <- combined %>% filter(Visit == 1)

visit1$unique_ID <- paste(visit1$site, visit1$Species, visit1$`Tag #`, sep = "")

visit2 <- combined %>% filter(Visit == 2)

visit2$unique_ID <- paste(visit2$site, visit2$Species, visit2$`Tag #`, sep = "")

# We'll merge the two by Tag #, we have a few more in the second visit due to adding trees
combined_merged <- inner_join(visit1, visit2, by = "unique_ID", suffix = c("_Visit1", "_Visit2")) # we have duplicates
combined_merged <- combined_merged %>% distinct(unique_ID, .keep_all = TRUE) # Get rid of duplicates

# Make a host/non-host variable
combined_merged$host <- if_else(combined_merged$Species_Visit1 == "ABLA", 1, 0)

# Merge with points
points$Name <- sub("^(\\d+)$", "abla\\1", points$Name)
points$Name <- toupper(points$Name)

fishnet_points <- points[, c("Name", "Latitude", "Longitude", "Elevation")]

# To make sure we're combining the right points, we need to filter by site, add site name to Name and then combine back into our data

cxsp_points <- spatial_search(fishnet_points, 41.641310, -111.491606, 1) # search 1km around point found
ggplot(cxsp_points, aes(x = Latitude, y = Longitude)) + 
  geom_point() # check that it worked - hell yes spatial search

cxsp_points$Site <- "CXSP"
cxsp_points$unique_ID <- paste(cxsp_points$Site, cxsp_points$Name, sep = "")


emig_points <- spatial_search(fishnet_points, 42.375017, -111.555346, 1)
ggplot(emig_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

emig_points$Site <- "EMIG"
emig_points$unique_ID <- paste(emig_points$Site, emig_points$Name, sep = "")


smmh_points <- spatial_search(fishnet_points, 40.769878, -111.107718, 1)
ggplot(smmh_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

smmh_points$Site <- "SMMH"
smmh_points$unique_ID <- paste(smmh_points$Site, smmh_points$Name, sep = "")


frnk_points <- spatial_search(fishnet_points, 41.933229, -111.570967, 1) # search 1km around point found
ggplot(frnk_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

frnk_points$Site <- "FRNK"
frnk_points$unique_ID <- paste(frnk_points$Site, frnk_points$Name, sep = "")


mocr_points <- spatial_search(fishnet_points, 41.409149, -111.532901, 1) # search 1km around point found
ggplot(mocr_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

mocr_points$Site <- "MOCR"
mocr_points$unique_ID <- paste(mocr_points$Site, mocr_points$Name, sep = "")


mttp_points <- spatial_search(fishnet_points, 40.435067, -111.638473, 1) # search 1km around point found
ggplot(mttp_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

mttp_points$Site <- "MTTP"
mttp_points$unique_ID <- paste(mttp_points$Site, mttp_points$Name, sep = "")

#write_xlsx(mttp_points,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/mttp_points.xlsx")

soap_points <- spatial_search(fishnet_points, 40.546185, -111.029622, 1) # search 1km around point found
ggplot(soap_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

soap_points$Site <- "SOAP"
soap_points$unique_ID <- paste(soap_points$Site, soap_points$Name, sep = "")

#write_xlsx(soap_points,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/soap_points.xlsx")

tmpf_points <- spatial_search(fishnet_points, 41.798592, -111.518826, 1) # search 1km around point found
ggplot(tmpf_points, aes(x = Latitude, y = Longitude)) + 
  geom_point()

tmpf_points$Site <- "TMPF"
tmpf_points$unique_ID <- paste(tmpf_points$Site, tmpf_points$Name, sep = "")

all_points <- rbind(cxsp_points, emig_points, frnk_points, mocr_points, mttp_points, smmh_points, soap_points, tmpf_points)

combined_merged$Name <- paste(combined_merged$Species_Visit1, combined_merged$`Tag #_Visit1`, sep = "")


all_merged <- full_join(all_points, combined_merged, by = "unique_ID")

# Omit points with no data associated or no point associated
all_merged <- all_merged %>% filter(!is.na(site_Visit1))
all_merged <- all_merged %>% filter(!is.na(Latitude))

# writexl::write_xlsx(all_merged, "all_data_with_points.xlsx")
```

# Distances to all trees

```{r}

# Convert to sf
all_merged_sf <- st_as_sf(all_merged, coords = c("Longitude", "Latitude"), crs = 4326)

# Define the target UTM zone
target_crs <- 32612  # EPSG code for UTM zone 12

# Transform the coordinates
all_merged_utm <- st_transform(all_merged_sf, crs = target_crs)
all_coords <- st_coordinates(all_merged_utm)
all_coords <- as.data.frame(all_coords)

# Put back into original datframe

all_merged_utm$Longitude <- as.numeric(all_coords$X)
all_merged_utm$Latitude <- as.numeric(all_coords$Y)

all_infested <- all_merged_utm[all_merged_utm$`Inf Location_Visit1` == 1, ] # Infested trees with UTM coordinates


# write_clip(all_merged_utm) # copy dataframe without making file
all_merged_utm <- all_merged_utm[!duplicated(all_merged_utm$unique_ID), ] #- there shouldn't be duplicates now
rownames(all_merged_utm) <- all_merged_utm$unique_ID


euclidean_distance <- function(x1, y1, x2, y2) {
    distance <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
    return(distance)
}


rm(Holder)
Holder = as.data.frame(matrix(nrow=0,ncol=4)) 
colnames(Holder) =c("PLOT", "FROM", "TARGET", "DISTANCE")

# New loop 

for (i in levels(as.factor(all_merged_utm$site_Visit1))) {
  # Subset by plot
  subset_frame <- subset(all_merged_utm, site_Visit1 == i)
  
  # Create matrix to store distances
  n <- nrow(subset_frame)
  distance_matrix <- matrix(0, nrow = n, ncol = n)
  rownames(distance_matrix) <- subset_frame$unique_ID
  colnames(distance_matrix) <- subset_frame$unique_ID
  
  # Populate distance matrix
  for (row in 1:n) {
    for (col in 1:n) {
      distance_matrix[row, col] <- euclidean_distance(subset_frame$Longitude[row], subset_frame$Latitude[row],
                                                      subset_frame$Longitude[col], subset_frame$Latitude[col])
    }
  }
  
  # Convert matrix to data frame
  TEMP_DIST.df <- as.data.frame(distance_matrix)
  
  # Add tag_ID column
  TEMP_DIST.df$tag_ID <- rownames(TEMP_DIST.df)
  
  # Reshape data
  TEMP_DIST_MELT <- reshape2::melt(TEMP_DIST.df, id.vars = "tag_ID")
  colnames(TEMP_DIST_MELT) <- c("FROM", "TARGET", "DISTANCE")
  
  # Add plot ID column
  TEMP_DIST_MELT$PLOT <- i
  
  # Append to Holder
  Holder <- rbind(Holder, TEMP_DIST_MELT)
}


#Merge with infestation data
names(Holder)[1] <- "unique_ID"

Holder_new <- merge(Holder, all_merged_utm, by.x = "TARGET", by.y = "unique_ID")
holder_new.dt <- as.data.table(Holder_new)
length(unique(holder_new.dt$unique_ID))
nrow(holder_new.dt)
names(holder_new.dt)[16] <- "infested_1"
names(holder_new.dt)[42] <- "infested_2"
#min_dist_holder <- holder_new.dt[`infested` == 1 & (DISTANCE == min(DISTANCE[(DISTANCE != 0)])) ,.(DISTANCE, TARGET),by=.("unique_ID")]

# distance to nearest infested at time 1
min_dist_holder <- holder_new.dt[infested_1 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]

# distance to nearest infested at time 2
min_dist_holder2 <- holder_new.dt[infested_2 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]

names(min_dist_holder2)[2] <- "TARGET2"
names(min_dist_holder2)[3] <- "DISTANCE2"

# add distances to data
# visit 1 distances
data_with_nearest_infested_dis <- all_merged_utm %>% inner_join(min_dist_holder, by = "unique_ID")

# visit 2 distances
data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder2, by = "unique_ID")

data_with_nearest_infested_dis <- as.data.frame(data_with_nearest_infested_dis)


#write_xlsx(data_with_nearest_infested_dis,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/data_with_nearest_infested.xlsx")

```

# Infestation Sum

```{r}

# Visit 1
data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
  mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")), 
            funs(if_else(startsWith(., "1-"), "1", .))) %>%
  mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")), as.numeric)
# Change all "1-" ratings to just a singular 1 value and make numeric

data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
  mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")),~coalesce(., 0)) # change all NA's to 0

data_with_nearest_infested_dis$infestation_sum_Visit1_focal <- data_with_nearest_infested_dis$NL_Visit1 + data_with_nearest_infested_dis$NM_Visit1 + data_with_nearest_infested_dis$NH_Visit1 + data_with_nearest_infested_dis$SL_Visit1 + data_with_nearest_infested_dis$SM_Visit1 + data_with_nearest_infested_dis$SH_Visit1

# Visit 2
data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
  mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")), 
            funs(if_else(startsWith(., "1-"), "1", .))) %>%
  mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")), as.numeric)
# Change all "1-" ratings to just a singular 1 valueand make numeric

data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>%
  mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")),~coalesce(., 0)) # change all NA's to 0

data_with_nearest_infested_dis$infestation_sum_Visit2_focal <- data_with_nearest_infested_dis$NL_Visit2 + data_with_nearest_infested_dis$NM_Visit2 + data_with_nearest_infested_dis$NH_Visit2 + data_with_nearest_infested_dis$SL_Visit2 + data_with_nearest_infested_dis$SM_Visit2 + data_with_nearest_infested_dis$SH_Visit2

# Subset target trees to get infesation sums
targets <- as.character(data_with_nearest_infested_dis$TARGET)
infested_subset_inf_sum <- data_with_nearest_infested_dis %>% filter(unique_ID %in% targets) %>% dplyr::select(c("unique_ID", "infestation_sum_Visit1_focal", "infestation_sum_Visit2_focal"))
names(infested_subset_inf_sum)[1] <- "TARGET"
names(infested_subset_inf_sum)[2] <- "inf_sum_Visit1_target"
names(infested_subset_inf_sum)[3] <- "inf_sum_Visit2_target"



# append sums to our main dataframe
# Append target coordinates to data frame
data_with_nearest_infested_dis <- inner_join(data_with_nearest_infested_dis, infested_subset_inf_sum, by = "TARGET")
str(data_with_nearest_infested_dis)



# Again for all data so we can extract values for full model

# Visit 1
all_merged_utm <- all_merged_utm %>%
  mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")), 
            funs(if_else(startsWith(., "1-"), "1", .))) %>%
  mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")), as.numeric)
# Change all "1-" ratings to just a singular 1 value and make numeric

all_merged_utm <- all_merged_utm %>%
  mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")),~coalesce(., 0)) # change all NA's to 0

all_merged_utm$infestation_sum_Visit1_focal <- all_merged_utm$NL_Visit1 + all_merged_utm$NM_Visit1 + all_merged_utm$NH_Visit1 + all_merged_utm$SL_Visit1 + all_merged_utm$SM_Visit1 + all_merged_utm$SH_Visit1

# Visit 2
all_merged_utm <- all_merged_utm %>%
  mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")), 
            funs(if_else(startsWith(., "1-"), "1", .))) %>%
  mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")), as.numeric)
# Change all "1-" ratings to just a singular 1 valueand make numeric

all_merged_utm <- all_merged_utm %>%
  mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")),~coalesce(., 0)) # change all NA's to 0

all_merged_utm$infestation_sum_Visit2_focal <- all_merged_utm$NL_Visit2 + all_merged_utm$NM_Visit2 + all_merged_utm$NH_Visit2 + all_merged_utm$SL_Visit2 + all_merged_utm$SM_Visit2 + all_merged_utm$SH_Visit2

```

# Infestation as a binary

```{r}
# Time 1
all_merged_utm$`Inf Location_Visit1`[ all_merged_utm$`Inf Location_Visit1`== "-"] <- 0
all_merged_utm$`Inf Location_Visit1`[all_merged_utm$`Inf Location_Visit1` == "BO" | all_merged_utm$`Inf Location_Visit1` == "BR, BO" | all_merged_utm$`Inf Location_Visit1` == "BR"] <- 1
all_merged_utm$`Inf Location_Visit1`[is.na(all_merged_utm$`Inf Location_Visit1`)] <- 0 # change NA's to 0
all_merged_utm$`Inf Location_Visit1` <- as.numeric(all_merged_utm$`Inf Location_Visit1`)
unique(all_merged_utm$`Inf Location_Visit1`)
# Time 2
all_merged_utm$`Inf Location_Visit2`[ all_merged_utm$`Inf Location_Visit2`== "-"] <- 0
all_merged_utm$`Inf Location_Visit2`[all_merged_utm$`Inf Location_Visit2` == "BO" | all_merged_utm$`Inf Location_Visit2` == "BR, BO" | all_merged_utm$`Inf Location_Visit2` == "BR"] <- 1
all_merged_utm$`Inf Location_Visit2`[is.na(all_merged_utm$`Inf Location_Visit2`)] <- 0 # change NA's to 0
all_merged_utm$`Inf Location_Visit2` <- as.numeric(all_merged_utm$`Inf Location_Visit2`)
unique(all_merged_utm$`Inf Location_Visit2`)

```

# Function to extract info from target trees

```{r}

# # Change names to make easier
# names(all_merged_utm)[12] <- "dead_alive"
# names(all_merged_utm)[13] <- "infested_1"
# names(all_merged_utm)[15] <- "brooms"
# names(all_merged_utm)[39] <- "infested_2"
# 
# # Recode some things for extraction
# all_merged_utm$Species_Visit1 <- as.factor(all_merged_utm$Species_Visit1)
# 
# # Make things!
# 
# # BA in cubed meters or something
# all_merged_utm$DBH_Visit1 <- all_merged_utm$DBH_Visit1 * 2.54 # convert DBH to cm
# all_merged_utm <- mutate(all_merged_utm, BA = (0.00007854 * (DBH_Visit1 ^ 2)))
# 
# # Crown Height in meters
# all_merged_utm$crown_height <- (all_merged_utm$Height_Visit1 - all_merged_utm$`Canopy Height_Visit1`) * 0.3048
# 
# # Get all trees under 10m from focal tree
# under10_holder <- holder_new.dt[,.(TARGET, DISTANCE), by=.(unique_ID)][,.SD[which(DISTANCE <= 10 & DISTANCE != 0)], by=.(unique_ID)]
# 
# df <- under10_holder
# # Create a function to extract the variable based on the linked tree ID
# extractVariable <- function(TARGET, variable) {
#   linkedTreeRow <- all_merged_utm[all_merged_utm$unique_ID == TARGET, ]
#   if (nrow(linkedTreeRow) > 0) {
#     return(linkedTreeRow[[variable]])
#   } else {
#     return(NA)  # or any other value to indicate missing data
#   }
# }
# 
# str(all_merged_utm)
# 
# # Apply the function to create a new column with the extracted variable - need to change function for each extraction !!
# df$infestation_under10 <- sapply(df$TARGET, extractVariable, variable = "infested_1")
# df$infestation.sum.1_under10 <- sapply(df$TARGET, extractVariable, variable = "infestation_sum_Visit1_focal")
# df$dbh_under10 <- sapply(df$TARGET, extractVariable, variable = "DBH_Visit1")
# df$brooms_under10 <- sapply(df$TARGET, extractVariable, variable = "brooms")
# df$gouting_under10 <- sapply(df$TARGET, extractVariable, variable = "Gouting_Visit1")
# df$species_under10 <- sapply(df$TARGET, extractVariable, variable = "Species_Visit1")
# df$BA_under10 <- sapply(df$TARGET, extractVariable, variable = "BA")
# df$crown_ht__under10 <- sapply(df$TARGET, extractVariable, variable = "crown_height")
# df$dead_alive <- sapply(df$TARGET, extractVariable, variable = "dead_alive")
# 
# # Host and non-host BA
# df$host_non <- if_else(df$species_under10 == "ABLA", 1, 0)
# 
# # create a summary table of local stats for each tree
# local.stats <- df %>% 
#   dplyr::group_by(unique_ID) %>% 
#   dplyr::summarise(n = n(),
#                    number.infested_under10 = sum(infestation_under10), # number of infested
#                    percent.infested_under10 = (sum(infestation_under10)/n) * 100, # percent infested
#                    infestation.sum_under10 = sum(infestation.sum.1_under10), # infestation load or propagule pressure
#                    average.dbh_under10 = mean(dbh_under10, na.rm = TRUE), # average dbh
#                    brooms_under10 = sum(brooms_under10), # number of brooms (this may not be anything)
#                    # avg.gouting_under10 = mean(gouting_under10, na.rm = TRUE), # Since we only have 1's at EMIG, average isn't useful
#                    gouting.sum_under = sum(gouting_under10, na.rm = TRUE), # gouting sum
#                    basal_area_non_host = sum(BA_under10[!is.na(host_non) & host_non != 1], na.rm = TRUE), # non-host BA
#                    basal_area_host = sum(BA_under10[host_non == 1], na.rm = TRUE), # host BA
#                    percent_basal_area_non_host = basal_area_non_host / (basal_area_non_host + basal_area_host) * 100, # percent BA of non-host
#                    crown_ht_sum = sum(crown_ht__under10, na.rm = TRUE),
#                    mean_distance_infested = ifelse(sum(infestation_under10) == 0, NA, mean(DISTANCE[infestation_under10 == 1], na.rm = TRUE))) # If there's no infestation within 10m, it'll come out as NA 
# 
# names(local.stats)[1] <- "unique_ID"
# 
# # combine with data
# data_with_local_stats <- all_merged_utm %>% inner_join(local.stats, by = "unique_ID")
# 
# # append nearest infested neighbor distance
# data_with_local_stats <- inner_join(data_with_local_stats, min_dist_holder, by = "unique_ID")
# 
# write_xlsx(data_with_local_stats, "all_trees_data_under10m.xlsx")

data_with_local_stats <- read_xlsx("all_trees_data_under10m.xlsx")
```

# GLMM with site as RE

```{r}

# just load in data from above to skip allocation of df from previous chunk (takes a long time)
# glm_data <- read_xlsx("/Big Glm/all_trees_data_under10m.xlsx")
# check covariates and transform

host_data <- data_with_local_stats %>% filter(Species_Visit1 == "ABLA")

host_data$infest_sum_10_std <- (host_data$infestation.sum_under10 - mean(host_data$infestation.sum_under10))/sd(host_data$infestation.sum_under10)
host_data$basal_area_host_std <- (host_data$basal_area_host - mean(host_data$basal_area_host))/sd(host_data$basal_area_host)

hist(data_with_local_stats$mean_distance_infested)

# Look for NA's

sum(is.na(host_data$DISTANCE))
sum(is.na(host_data$DBH_Visit1))
sum(is.na(host_data$Gouting_Visit1))
sum(is.na(host_data$brooms))
sum(is.na(host_data$number.infested_under10))
sum(is.na(host_data$percent.infested_under10))
sum(is.na(host_data$infestation.sum_under10))
sum(is.na(host_data$average.dbh_under10))
sum(is.na(host_data$basal_area_non_host))
sum(is.na(host_data$percent_basal_area_non_host))
sum(is.na(host_data$mean_distance_infested))

# Deal with NA's
host_data$DBH_Visit1[is.na(host_data$DBH_Visit1)] <- 0.5 # Change DBH to 0.5 to include clumps
host_data$Gouting_Visit1[is.na(host_data$Gouting_Visit1)] <- 0 # change NA's to 0
host_data <- host_data[!is.na(host_data$mean_distance_infested), ]

# Take out TMPF becasue of lack of points

host_data <- host_data %>% filter(host_data$Site != "TMPF")


inf.glmm <- glmmTMB(cbind(infested_1, 1 - infested_1) ~ log(DISTANCE) + log(DBH_Visit1) + Gouting_Visit1 + brooms + log(number.infested_under10) + log(percent.infested_under10) + log(infestation.sum_under10) + log(average.dbh_under10) + mean_distance_infested + (1|Site), data = host_data, family = binomial(link = "logit"))

inf.sims <- simulateResiduals(inf.glmm)
plot(inf.sims)

summary(inf.glmm)

r2 <- r.squaredGLMM(inf.glmm)

stepAIC(inf.glmm)

inf.glmm.nore <- glmmTMB(cbind(infested_1, 1 - infested_1) ~ log(DISTANCE) + log(DBH_Visit1) + Gouting_Visit1 + brooms + log(number.infested_under10) + log(percent.infested_under10) + log(infestation.sum_under10) + log(average.dbh_under10) + basal_area_host + mean_distance_infested, data = host_data, family = binomial(link = "logit"))

inf.sims.nore <- simulateResiduals(inf.glmm.nore)
plot(inf.sims.nore)

summary(inf.glmm.nore)

# Take out non-significant variables

inf.glmm.clean <- glmmTMB(cbind(infested_1, 1 - infested_1) ~ log(DISTANCE) + log(DBH_Visit1) + Gouting_Visit1 + brooms + log(infestation.sum_under10) + mean_distance_infested + (1|Site), data = host_data, family = binomial(link = "logit"))

inf.sims.clean <- simulateResiduals(inf.glmm.clean)
plot(inf.sims.clean)

summary(inf.glmm)
summary(inf.glmm.clean)

r2.clean <- r.squaredGLMM(inf.glmm)
r2.clean
```

# GLMM with newly infested

```{r}
# Make newly infested column
host_data$newly_infested <- if_else(host_data$infested_1 == 0 & host_data$infested_2 == 1, 1, 0)

inf.glmm.newly.infested <- glmmTMB(cbind(newly_infested, 1 - newly_infested) ~ log(DISTANCE) + log(DBH_Visit1) + Gouting_Visit1 + brooms + log(number.infested_under10) + log(percent.infested_under10) + log(infestation.sum_under10) + log(average.dbh_under10) + basal_area_host + mean_distance_infested + (1|Site), data = host_data, family = binomial(link = "logit"))

inf.sims.newly <- simulateResiduals(inf.glmm.newly.infested)
plot(inf.sims.newly)

summary(inf.glmm.newly.infested)

```