---
title: "New infested"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

New Infested Trees
11/27/23
Michael Wayman

The goal of this script is to look at our newly infested trees at each plot and what characteristics may be playing a part in their transition. We'll look at all of our variables that we've collected and see if any of them correlate with being infested for the first time. For most of the script, we'll just be using our measured variables excluding spatial aspects. At the end, we may try to incorporate the spatial element and see how that changes things. Let's do it!

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(rfishnet2)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(plyr)
library(reshape2)
library(geosphere)
library(data.table)
library(writexl)
```

## Clean out global environment

```{r}
rm(list = ls())
```

## Load Data

```{r, echo = F, warning=F}
combined <- read_xlsx("combined_cleaned2.xlsx")
site_stats <- read_xlsx("site_stats.xlsx")
```

We'll start by filtering out all of our newly infested trees

```{r}
# host <- combined %>% filter(Species == "ABLA" | Species == "ABCO") # Filter out hosts
# live_host <- host %>% filter(`Dead / Alive` == 1) # Filter out live hosts

live_host <- combined

# # Change Infestation to binary variable
# live_host$`Inf Location`[ live_host$`Inf Location`== "-"] <- 0
# unique(live_host$`Inf Location`)
# live_host$`Inf Location`[live_host$`Inf Location` == "BO" | live_host$`Inf Location` == "BR, BO" | live_host$`Inf Location` == "BR"] <- 1
# live_host[which(is.na(live_host$`Inf Location`)), ]
# live_host <- live_host[!is.na(live_host$`Inf Location`),] # Drop NA's
# live_host$`Inf Location` <- as.numeric(live_host$`Inf Location`)
# str(live_host)

# We want to make sure our binary variable is accurate, so instead of counting on ratings like BO, BR, we'll change all of our ratings to numbers and have a count for each visit.

unique(live_host$NL)
unique(live_host$NM)
unique(live_host$NH)
unique(live_host$SL)
unique(live_host$SM)
unique(live_host$SH)
live_host[live_host == "-"] <- "0" # change all "-" to a zero

live_host <- live_host %>%
  dplyr::mutate(across(c("NL", "NM", "NH", "SL", "SM", "SH"), 
                ~ if_else(startsWith(., "1"), "1", .)))# Change all "1-" ratings to just a singular 1 value

unique(live_host$NL)

live_host <- live_host %>%
  mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH")), as.numeric)

live_host$NL[is.na(live_host$NL)] <- 0
live_host$NM[is.na(live_host$NM)] <- 0
live_host$NH[is.na(live_host$NH)] <- 0
live_host$SL[is.na(live_host$SL)] <- 0
live_host$SM[is.na(live_host$SM)] <- 0
live_host$SH[is.na(live_host$SH)] <- 0

live_host <- live_host %>%
  mutate(infested = ifelse(NL + NM + NH + SL + SM + SH != 0, 1, 0))

# First we'll subset by Visit
visit1 <- live_host %>% filter(Visit == 1)
visit1$tag_ID <- paste(visit1$Species, visit1$`Tag #`, sep = "")
visit1$unique_ID <- paste(visit1$`Site ID`, visit1$tag_ID, sep = "")
visit2 <- live_host %>% filter(Visit == 2)
visit2$tag_ID <- paste(visit2$Species, visit2$`Tag #`, sep = "")
visit2$unique_ID <- paste(visit2$`Site ID`, visit2$tag_ID, sep = "")
visit3 <- live_host %>% filter(Visit == 3)
visit3$tag_ID <- paste(visit3$Species, visit3$`Tag #`, sep = "")
visit3$unique_ID <- paste(visit3$`Site ID`, visit3$tag_ID, sep = "")
visit4 <- live_host %>% filter(Visit == 4)
visit4$tag_ID <- paste(visit4$Species, visit4$`Tag #`, sep = "")
visit4$unique_ID <- paste(visit4$`Site ID`, visit4$tag_ID, sep = "")

visit3 <- visit3 %>%
  rename_with(~ paste0(., "_v3"), -unique_ID) # add suffixes to visit 3 and 4
visit4 <- visit4 %>%
  rename_with(~ paste0(., "_v4"), -unique_ID) 

# merged <- visit1 %>%
#   full_join(visit2, by = "unique_ID", suffix = c("_v1", "_v2")) %>%
#   full_join(visit3, by = "unique_ID", suffix = c("", "_v3")) %>%
#   full_join(visit4, by = "unique_ID", suffix = c("", "_v4"))

# Troubleshooting merge

# Create a complete list of unique_IDs
all_ids <- unique(c(visit1$unique_ID, visit2$unique_ID, visit3$unique_ID, visit4$unique_ID))

# Add missing IDs to each dataset
visit1 <- visit1 %>% right_join(data.frame(unique_ID = all_ids), by = "unique_ID")
visit2 <- visit2 %>% right_join(data.frame(unique_ID = all_ids), by = "unique_ID")
visit2 <- filter(visit2, !duplicated(visit2$unique_ID))
visit3 <- visit3 %>% right_join(data.frame(unique_ID = all_ids), by = "unique_ID")
visit3 <- filter(visit3, !duplicated(visit3$unique_ID))
visit4 <- visit4 %>% right_join(data.frame(unique_ID = all_ids), by = "unique_ID")
visit4 <- filter(visit4, !duplicated(visit4$unique_ID))

# Merge the datasets
merged <- visit1 %>%
  right_join(visit2, by = "unique_ID", suffix = c("_v1", "_v2")) %>%
  right_join(visit3, by = "unique_ID", suffix = c("", "_v3")) %>%
  right_join(visit4, by = "unique_ID", suffix = c("", "_v4"))

# Coalesce takes the first non-missing value from the list of variables and takes that
  
merged <- merged %>% mutate(
    site = coalesce(`Site ID_v1`, `Site ID_v2`, `Site ID_v3`, `Site ID_v4`),
    tag = coalesce(`Tag #_v1`, `Tag #_v2`, `Tag #_v3`, `Tag #_v4`),
    species = coalesce(Species_v1, Species_v2, Species_v3, Species_v4),
    dbh = coalesce(DBH_v1, DBH_v2, DBH_v3, DBH_v4),
    height = coalesce(Height_v1, Height_v2, Height_v3, Height_v4),
    canopy_ht = coalesce(`Canopy Height_v1`, `Canopy Height_v2`, `Canopy Height_v3`, `Canopy Height_v4`),
    brooms = coalesce(`Broom (#)_v1`, `Broom (#)_v2`, `Broom (#)_v3`, `Broom (#)_v4`),
    rot = coalesce(Rot_v1, Rot_v2, Rot_v3, Rot_v4),
  ) 
  
  # Select combined columns and add suffixes to non-identical columns
  merged <- merged %>% select(
    site, 
    tag, 
    unique_ID,
    species, 
    dbh,
    height,
    canopy_ht,
    brooms,
    rot,
    `Dead / Alive_v1`, `Dead / Alive_v2`, `Dead / Alive_v3`, `Dead / Alive_v4`, 
    `Inf Location_v1`, `Inf Location_v2`, `Inf Location_v3`, `Inf Location_v4`,
    Gouting_v1, Gouting_v2, Gouting_v3, Gouting_v4,
    Beetle_v1, Beetle_v2, Beetle_v3, Beetle_v4,
    NL_v1, NL_v2, NL_v3, NL_v4,
    NLW_v1, NLW_v2, NLW_v3, NLW_v4,
    NM_v1, NM_v2, NM_v3, NM_v4,
    NMW_v1, NMW_v2, NMW_v3, NMW_v4,
    NH_v1, NH_v2, NH_v3, NH_v4,
    NHW_v1, NHW_v2, NHW_v3, NHW_v4,
    SL_v1, SL_v2, SL_v3, SL_v4,
    SLW_v1, SLW_v2, SLW_v3, SLW_v4,
    SM_v1, SM_v2, SM_v3, SM_v4,
    SMW_v1, SMW_v2, SMW_v3, SMW_v4,
    SH_v1, SH_v2, SH_v3, SH_v4,
    SHW_v1, SHW_v2, SHW_v3, SHW_v4,
    Notes_v1, Notes_v2, Notes_v3, Notes_v4, 
    WSB_v1, WSB_v2, WSB_v3, WSB_v4,
    infested_v1, infested_v2, infested_v3, infested_v4
  )

colnames(merged)
setdiff(visit1$tag_ID, visit2$tag_ID)
str(merged)

# Check the count of unique Tag Numbers in each dataset
count_visit1 <- n_distinct(visit1$unique_ID)
count_visit2 <- n_distinct(visit2$unique_ID)
count_visit3 <- n_distinct(visit3$unique_ID)
count_visit4 <- n_distinct(visit4$unique_ID)

# Now we'll look at the change from Visit 1 to Visit 2 - anything that's not zero should be newly infested
merged <- merged %>%
  mutate(
    new_infestation = if_else(infested_v1 == 0 & (infested_v2 == 1 | infested_v3 == 1 | infested_v4 == 1), 1, 0)
  )

# Subset data to identify trees with changes 
new_infest_trees <- merged[merged$new_infestation == 1, ]
other_trees <- merged[merged$new_infestation == 0, ]
#unchanged_trees <- merged[merged$ChangeInVariable != 1, ]
#non_infested_trees <- merged %>% filter(`Inf Location_Visit1` == 0 & `Inf Location_Visit2` == 0) # Filter trees that weren't infested during either visit
infested_t1_trees <- merged[merged$infested_v1 == 1, ]
uninfested_t1_trees <- merged[merged$infested_v1 == 0, ]

```

# Now that we have our changed trees, we can look at individual variables to see if anything stands out between trees that changed and those that did not - we'll start by vizualizing te differences in DBH, Height, Gouting and Brooms

** We're now looking at newly infested vs everything else

```{r}
new_infested_trees_above1 <- new_infest_trees[!is.na(new_infest_trees$dbh),] # drop NA's for stats
other_trees_above1 <- other_trees[!is.na(other_trees$dbh),] 
# unchanged_trees_above1 <- unchanged_trees[!is.na(unchanged_trees$DBH_Visit1),] # drop NA's for stats
# non_infested_trees <- non_infested_trees[!is.na(non_infested_trees$DBH_Visit1), ] # drop NA's for stats

ggplot() +
  geom_histogram(data = new_infested_trees_above1, aes(x = dbh, fill = "Newly infested trees"), position = "stack", bins = 10, alpha = 0.7) +
  geom_histogram(data = other_trees_above1, aes(x = dbh, fill = "No change in infestation Status"), position = "stack", bins = 10, alpha = 0.7) +
  labs(title = "DBH of newly infested trees vs others", x = "DBH", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "All other trees" = "red")) +
  theme_minimal()

# Make a model to compare by combining data first and coding changed to 1 and unchanged to 0
new_infested_trees_above1$Dataset <- 1
other_trees_above1$Dataset <- 0
combined_data <- rbind(new_infested_trees_above1, other_trees_above1)

combined_data$Dataset <- as.factor(combined_data$Dataset)
ggplot(combined_data, aes(x = site, y = dbh, fill = Dataset)) + 
  labs(title="DBH (in.) of trees that got infested vs all others",x="Site", y = "DBH") +
  scale_fill_discrete(name = "", labels = c("Others", "Newly infested")) +
  theme(legend.position = c(0.8, 0.8), legend.title = element_blank()) +
    geom_boxplot()


model <- lm(dbh ~ Dataset, data = combined_data)
summary(model)

# Infested vs not - Time 1
infested_t1_trees$Dataset <- 1
uninfested_t1_trees$Dataset <- 0

merged_t1 <- rbind(infested_t1_trees, uninfested_t1_trees)

merged_t1$Dataset <- as.factor(merged_t1$infested_v1)

ggplot(merged_t1, aes(x = site, y = dbh, fill = Dataset)) + 
  labs(title="DBH (in.) of infested vs non-infested trees - Time 1",x="Site", y = "DBH", fill = "Infested Status") +
  scale_fill_discrete(name = "Infested Status", labels = c("Uninfested", "Infested")) +
  theme(legend.position = c(0.8, 0.8)) +
    geom_boxplot()
```
### Un-infested vs newly infested Histogram, Box plot and model

```{r}

non_infested_trees <- merged %>% dplyr::filter(infested_v1 == 0 & infested_v2 == 0 & infested_v3 == 0 & infested_v4 == 0)
non_infested_trees <- non_infested_trees[!is.na(non_infested_trees$dbh),]


# Histogram of all trees that were infested vs not
ggplot() +
  geom_histogram(data = new_infested_trees_above1, aes(x = dbh, fill = "Newly infested trees"), position = "stack", bins = 10, alpha = 0.7) +
  geom_histogram(data = non_infested_trees, aes(x = dbh, fill = "Trees that haven't gotten infested"), position = "stack", bins = 10, alpha = 0.7) +
  labs(title = "DBH of trees that were infested vs not infested", x = "DBH", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "Not infested" = "red")) +
  theme_minimal()



# Make a model to compare by combining data first and coding changed to 1 and unchanged to 0
new_infested_trees_above1$Dataset <- 1
non_infested_trees$Dataset <- 0
combined_data_1 <- rbind(new_infested_trees_above1, non_infested_trees)

combined_data_1$Dataset <- as.factor(combined_data_1$Dataset)
ggplot(combined_data_1, aes(x = site, y = dbh, fill = Dataset)) + 
  labs(title="DBH of trees that got infested vs those that did not",x="Site", y = "DBH") +
    geom_boxplot()


model <- lm(dbh ~ Dataset, data = combined_data_1)
summary(model)

```

## We'll do the same with Height

** Why do we have NA heights??

```{r}

merged[is.na(merged$height), ]

```

```{r}
new_infested_trees_above1 <- new_infested_trees_above1[!is.na(new_infested_trees_above1$height),] # drop NA's for stats
non_infested_trees <- non_infested_trees[!is.na(non_infested_trees$height),] # drop NA's for stats



ggplot() +
  geom_histogram(data = new_infested_trees_above1, aes(x = dbh, fill = "Newly infested trees"), position = "stack", bins = 10, alpha = 0.7) +
  geom_histogram(data = non_infested_trees, aes(x = dbh, fill = "Never infested"), position = "stack", bins = 10, alpha = 0.7) +
  labs(title = "Height of trees that got infested vs those that did not", x = "Height", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "Never infested" = "red")) +
  theme_minimal()

model <- lm(height ~ Dataset, data = combined_data)
summary(model)

# Make a model to compare by combining data first and coding changed to 1 and unchanged to 0
changed_trees_above1$Dataset <- 1
non_infested_trees$Dataset <- 0
combined_data_1 <- rbind(changed_trees_above1, non_infested_trees)

combined_data_1$Dataset <- as.factor(combined_data_1$Dataset)
ggplot(combined_data_1, aes(x = `Site ID_Visit1`, y = Height_Visit1, fill = Dataset)) + 
  labs(title="Height of trees that got infested vs those that did not",x="Site", y = "Height") +
    geom_boxplot()


model <- lm(Height_Visit1 ~ Dataset, data = combined_data_1)
summary(model)
```

## And the same with Gouting

```{r}

changed_trees_above1 <- changed_trees %>% mutate(Gouting_Visit1 = ifelse(is.na(Gouting_Visit1), 0, Gouting_Visit1)) # We want to change NA's this time around to specify "no observable gouting" instead of dropping all of our NA's

unchanged_trees_above1 <- unchanged_trees %>% mutate(Gouting_Visit1 = ifelse(is.na(Gouting_Visit1), 0, Gouting_Visit1))

ggplot() +
  geom_histogram(data = changed_trees_above1, aes(x = Gouting_Visit1, fill = "Newly infested trees"), position = "stack", bins = 4, alpha = 0.7) +
  geom_histogram(data = unchanged_trees_above1, aes(x = Gouting_Visit1, fill = "No change in infestation Status"), position = "stack", bins = 4, alpha = 0.7) +
  labs(title = "Gouting of trees that changed status vs no status change - Visit 1", x = "Gouting", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "No change in infestation status" = "red")) +
  theme_minimal()

# Let's do the same, but for Gouting at time 2
ggplot() +
  geom_histogram(data = changed_trees_above1, aes(x = Gouting_Visit2, fill = "Newly infested trees"), position = "stack", bins = 4, alpha = 0.7) +
  geom_histogram(data = unchanged_trees_above1, aes(x = Gouting_Visit2, fill = "No change in infestation Status"), position = "stack", bins = 4, alpha = 0.7) +
  labs(title = "Gouting of trees that changed status vs no status change - Visit 2", x = "Gouting", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "No change in infestation status" = "red")) +
  theme_minimal()


model <- lm(Gouting_Visit1 ~ Dataset, data = combined_data)
summary(model)

# Make a model to compare by combining data first and coding changed to 1 and unchanged to 0

ggplot() +
  geom_histogram(data = changed_trees_above1, aes(x = Gouting_Visit1, fill = "Newly infested trees"), position = "stack", bins = 4, alpha = 0.7) +
  geom_histogram(data = non_infested_trees, aes(x = Gouting_Visit1, fill = "Not infested"), position = "stack", bins = 4, alpha = 0.7) +
  labs(title = "Gouting of trees that changed status vs not infested - Visit 2", x = "Gouting", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "No change in infestation status" = "red")) +
  theme_minimal()


model <- lm(Gouting_Visit1 ~ Dataset, data = combined_data_1)
summary(model)
```

## And with brooms

```{r}

broom_plot <- ggplot() +
  geom_histogram(data = changed_trees_above1, aes(x = `Broom (#)_Visit1`, fill = "Newly infested trees"), position = "stack", bins = 15, alpha = 0.7) +
  geom_histogram(data = unchanged_trees_above1, aes(x = `Broom (#)_Visit1`, fill = "No change in infestation Status"), position = "stack", bins = 15, alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 15, by = 1)) +
  labs(title = "# of Brooms of trees that changed status vs no status change", x = "Height", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "No change in infestation status" = "red")) +
  theme_minimal()

print(broom_plot)

model <- lm(`Broom (#)_Visit1` ~ Dataset, data = combined_data)
summary(model)

# newly vs not infested
ggplot(combined_data_1, aes(x = `Site ID_Visit1`, y = `Broom (#)_Visit1`, fill = Dataset)) + 
  labs(title="Broom number on trees that got infested vs those that did not",x="Site", y = "Height") +
    geom_boxplot()
```
So it looks like we're seeing a relationship in newly infested trees that suggests larger trees are being infested more so than other trees.


# Nearest neighbor of newly infested trees

What I want to do is calculate the nearest neighbor distances for every newly infested tree to the closest infested tree. I will then compare these distances with trees that remain uninfested (were uninfested at time 1 and time 2). I'm sure I'll run into a lot of issues as this is the first time I'm really incorporating the spatial aspect, but let's try it.

## To do this the easiest, I want to try and get all of our host tree points attached to the data we're working with in here

## Revision to above - I'm going to combine all points and then we can just filter for live hosts out of that

```{r}
  most_points = read.csv("all_points.csv", header = T)
  rest_of_points <- read.csv("rest_of_points.csv", header = T)
  point24 <- read.csv("waypoints2024.csv", header = T)
  points <- most_points %>% bind_rows(rest_of_points) %>% bind_rows(point24)
  names(points)[2] <- "Tag #" # rename Tag # for combination

  # remove non-host trees
  # points <- points %>% filter(!grepl("psmeg", `Tag #`))
  # points <- points %>% filter(!grepl("potr", `Tag #`))
  # points <- points %>% filter(!grepl("pico", `Tag #`))
  # points$`Tag #` <- as.numeric(points$`Tag #`)

  # the main problem is that we have duplicates across sites, so we'll first separate our points by site, then separate our data by site, combine our points with our data and then combine our sites. It's kind of a slog, but I think it's the best way to avoid confusion

  # Get points by site
# cxsp_points <- read_xlsx("cxsp_points.xlsx")
# emig_points <- read_xlsx("emig_points.xlsx")
# smmh_points <- read_xlsx("smmh_points.xlsx")
# frnk_points <- read_xlsx("frnk_points.xlsx")
# mocr_points <- read_xlsx("mocr_points.xlsx")
# mttp_points <- read_xlsx("mttp_points.xlsx")
# soap_points <- read_xlsx("soap_points.xlsx")
# tmpf_points <- read_xlsx("tmpf_points.xlsx")
  
  
   cxsp_points <- spatial_search(points, 41.641310, -111.491606, 1) # search 1km around point found
  # Get rid of duplicates - these were moved, but points were taken, so we're taking the most recent points with these tags
   cxsp_points <- cxsp_points[-which(cxsp_points$Number == 3057), ]
   cxsp_points <- cxsp_points[-which(cxsp_points$Number == 3085), ]
  
   emig_points <- spatial_search(points, 42.375017, -111.555346, 1)
   smmh_points <- spatial_search(points, 40.769878, -111.107718, 1)
   # Get rid of dupes
   smmh_points <- smmh_points[-which(smmh_points$Number == 327), ]
   smmh_points <- smmh_points[-which(smmh_points$Number == 518), ]
   # Looks like a group was mislabelled
  smmh_points$`Tag #`[smmh_points$Number == 580 & smmh_points$`Tag #` == 5661] <- 5561
  smmh_points$`Tag #`[smmh_points$Number == 581 & smmh_points$`Tag #` == 5662] <- 5562
  smmh_points$`Tag #`[smmh_points$Number == 582 & smmh_points$`Tag #` == 5663] <- 5563
  smmh_points$`Tag #`[smmh_points$Number == 583 & smmh_points$`Tag #` == 5664] <- 5564
  smmh_points$`Tag #`[smmh_points$Number == 584 & smmh_points$`Tag #` == 5665] <- 5565
  smmh_points$`Tag #`[smmh_points$Number == 608 & smmh_points$`Tag #` == 5789] <- 5589
  
   frnk_points <- spatial_search(points, 41.933229, -111.570967, 1) # search 1km around point found
   # Get rid of dupes and mislabelling
  frnk_points <- frnk_points[-which(frnk_points$Number == 1211), ] # tag moved, two points
  frnk_points <- frnk_points[-which(frnk_points$Number == 1331), ] # tag moved, two points
  frnk_points <- frnk_points[-which(frnk_points$Number == 1427), ] # tag moved, two points
  frnk_points <- frnk_points[-which(frnk_points$Number == 1263), ] # tag moved, two points
  frnk_points <- frnk_points[-which(frnk_points$Number == 1022), ] # tag moved, two points
  frnk_points$`Tag #`[frnk_points$Number == 968 & frnk_points$`Tag #` == 5819] <- 5919 # mislabel
  frnk_points$`Tag #`[frnk_points$Number == 889 & frnk_points$`Tag #` == 5949] <- 5849 # mislabel
  frnk_points$`Tag #`[frnk_points$Number == 890 & frnk_points$`Tag #` == 5950] <- 5850 # mislabel
  frnk_points$`Tag #`[frnk_points$Number == 891 & frnk_points$`Tag #` == 5951] <- 5851 # mislabel
  frnk_points$`Tag #`[frnk_points$Number == 1136 & frnk_points$`Tag #` == 7071] <- 7171 # mislabel
  frnk_points$`Tag #`[frnk_points$Number == 1137 & frnk_points$`Tag #` == 7072] <- 7172 # mislabel
  frnk_points <- frnk_points[-which(frnk_points$Number == 1258), ] # done twice, taking most recent
  frnk_points$`Tag #`[frnk_points$Number == 987 & frnk_points$`Tag #` == 5838] <- 5938
  frnk_points$`Tag #`[frnk_points$Number == 988 & frnk_points$`Tag #` == 5839] <- 5939
  
   mocr_points <- spatial_search(points, 41.409149, -111.532901, 1) # search 1km around point found
   # Get rid of dupes and mislabelling
   mocr_points <- mocr_points[-which(mocr_points$Number == 91), ] # done twice, taking most recent
   mttp_points <- spatial_search(points, 40.435067, -111.638473, 1) # search 1km around point found
  soap_points <- spatial_search(points, 40.546185, -111.029622, 1) # search 1km around point found
  tmpf_points <- spatial_search(points, 41.798592, -111.518826, 1) # search 1km around point found

  # Get site data and make unique IDs including species
  cxsp <- merged %>% filter(site == "CXSP") 
   cxsp$tag_ID <- paste(cxsp$species, cxsp$tag, sep = "")
  # cxsp <- cxsp %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
 # relocate(unique_ID, .after = "tag_ID")
   
  emig <- merged %>% filter(site == "EMIG") 
   emig$tag_ID <- paste(emig$species, emig$tag, sep = "")
 #   emig <- emig %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
  
  smmh <- merged %>% filter(site == "SMMH") 
   smmh$tag_ID <- paste(smmh$species, smmh$tag, sep = "")
   # smmh <- smmh %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
    
  frnk <- merged %>% filter(site == "FRNK") 
   frnk$tag_ID <- paste(frnk$species, frnk$tag, sep = "")
   # frnk <- frnk %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
    
  mocr <- merged %>% filter(site == "MOCR")
   mocr$tag_ID <- paste(mocr$species, mocr$tag, sep = "")
   # mocr <- mocr %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
    
  mttp <- merged %>% filter(site == "MTTP")
   mttp$tag_ID <- paste(mttp$species, mttp$tag, sep = "")
   # mttp <- mttp %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
    
  soap <- merged %>% filter(site == "SOAP")
   soap$tag_ID <- paste(soap$species, soap$tag, sep = "")
   # soap <- soap %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
    
  tmpf <- merged %>% filter(site == "TMPF")
   tmpf$tag_ID <- paste(tmpf$species, tmpf$tag, sep = "")
   # tmpf <- tmpf %>%
  #mutate(unique_ID = paste(site, tag_ID, sep = "")) %>%
  #relocate(unique_ID, .after = "tag_ID")
    

  # Merge points with data (I'm making a tag_ID column to get a unique ID for every point and then linking that with the unique ID in the data)
    names(cxsp_points)[2] <- "tag_ID"
  cxsp_points <- cxsp_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  cxsp_points$tag_ID <- str_to_upper(cxsp_points$tag_ID)
  
  sum(duplicated(cxsp$tag_ID)) # For the main data
  sum(duplicated(cxsp_points$tag_ID)) # For the GPS data
  
  #cxsp <- filter(cxsp, !duplicated(cxsp$tag_ID))
  cxsp_points <- filter(cxsp_points, !duplicated(cxsp_points$tag_ID))
  
  cxsp_merged <- left_join(cxsp, cxsp_points, by = "tag_ID")
  
  cxsp_merged <- cxsp_merged[!duplicated(cxsp_merged$tag_ID),]
  
  #EMIG
  
  names(emig_points)[2] <- "tag_ID"
  emig_points <- emig_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  emig_points$tag_ID <- str_to_upper(emig_points$tag_ID)
  
  emig <- filter(emig, !duplicated(emig$tag_ID))
  emig_points <- filter(emig_points, !duplicated(emig_points$tag_ID))
  
  emig_merged <- left_join(emig, emig_points, by = "tag_ID")
  
  emig_merged <- emig_merged[!duplicated(emig_merged$tag_ID),]
  
  #SMMH
  
  names(smmh_points)[2] <- "tag_ID"
  smmh_points <- smmh_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  smmh_points$tag_ID <- str_to_upper(smmh_points$tag_ID)
  
  smmh <- filter(smmh, !duplicated(smmh$tag_ID))
  
  # smmh$tag_ID <- sub("^(ABCO|ABLA)", "", smmh$tag_ID) # remove species from tag ID for combo
  # smmh_points$tag_ID <- sub("^(ABCO|ABLA)", "", smmh_points$tag_ID)
  
  smmh_merged <- left_join(smmh, smmh_points, by = "tag_ID") # left join to ensure all data, but we will have missing points
  
  smmh_merged <- smmh_merged[!duplicated(smmh_merged$tag_ID),]
  
  #FRNK
  
 names(frnk_points)[2] <- "tag_ID"
  frnk_points <- frnk_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  frnk_points$tag_ID <- str_to_upper(frnk_points$tag_ID)
  
  frnk <- filter(frnk, !duplicated(frnk$tag_ID))
  
  frnk_merged <- left_join(frnk, frnk_points, by = "tag_ID")
  
  frnk_merged <- frnk_merged[!duplicated(frnk_merged$tag_ID),]
  
  #MOCR
  
  names(mocr_points)[2] <- "tag_ID"
  mocr_points <- mocr_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  mocr_points$tag_ID <- str_to_upper(mocr_points$tag_ID)
  
  mocr <- filter(mocr, !duplicated(mocr$tag_ID))
  
  mocr_merged <- left_join(mocr, mocr_points, by = "tag_ID")
  
  mocr_merged <- mocr_merged[!duplicated(mocr_merged$tag_ID),]
  
  #MTTP
  
  names(mttp_points)[2] <- "tag_ID"
  mttp_points <- mttp_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  mttp_points$tag_ID <- str_to_upper(mttp_points$tag_ID)
  
  mttp <- filter(mttp, !duplicated(mttp$tag_ID))
  
  mttp_merged <- left_join(mttp, mttp_points, by = "tag_ID")
  
  #SOAP
  
 names(soap_points)[2] <- "tag_ID"
  soap_points <- soap_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  soap_points$tag_ID <- str_to_upper(soap_points$tag_ID)
  
  soap <- filter(soap, !duplicated(soap$tag_ID))
  
  soap_merged <- left_join(soap, soap_points, by = "tag_ID")
  
  soap_merged <- soap_merged[!duplicated(soap_merged$tag_ID),]
  
  #TMPF
  
 names(tmpf_points)[2] <- "tag_ID"
  tmpf_points <- tmpf_points %>%
  mutate(tag_ID = if_else(grepl("^[0-9]+$", tag_ID), 
                            paste0("abla", tag_ID), 
                            tag_ID))
  tmpf_points$tag_ID <- str_to_upper(tmpf_points$tag_ID)
  
  tmpf <- filter(tmpf, !duplicated(tmpf$tag_ID))
  
  tmpf_merged <- left_join(tmpf, tmpf_points, by = "tag_ID")

  all_merged <- rbind(cxsp_merged, emig_merged, smmh_merged, frnk_merged, mocr_merged, mttp_merged, soap_merged, tmpf_merged)
  
  all_merged$tagID <- paste(all_merged$site, all_merged$tag_ID, sep = "")
  
  
  # Check for duplicates
  duplicates <- all_merged %>% filter(duplicated(all_merged$unique_ID))
  
  all_merged <- all_merged %>% filter(!duplicated(all_merged$unique_ID))

  write_clip(all_merged)

  # We'll filter again for infested and non-infested trees
  #all_merged$ChangeInVariable <- all_merged$`Inf Location_Visit2` - all_merged$`Inf Location_Visit1`

 #write_xlsx(all_merged,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/all_merged_all_spp.xlsx")
  
```

# All-merged complete

```{r}
  
# all previous lines to create all_merged below
all_merged <- read_xlsx("all_merged_all_spp.xlsx")


all_merged <- all_merged %>%
  dplyr::mutate(
    # Create binary infestation variable for visit_1
     infestation_v1 = if_else(
      rowSums(across(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1"))) > 0,
      1, 0
    ),

    # Repeat for other visits
    infestation_v2 = if_else(
      rowSums(across(c("NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2"))) > 0,
      1, 0
    ),

    infestation_v3 = if_else(
      rowSums(across(c("NL_v3", "NM_v3", "NH_v3", "SL_v3", "SM_v3", "SH_v3"))) > 0,
      1, 0
    ),

    infestation_v4 = if_else(
      rowSums(across(c("NL_v4", "NM_v4", "NH_v4", "SL_v4", "SM_v4", "SH_v4"))) > 0,
      1, 0
    )
  )

# Make a new infestation variable to check whether a tree gained an infestation at any point during study

all_merged <- all_merged %>%
  dplyr::mutate(
    new_infestation = if_else(infested_v1 == 0 & (infested_v2 == 1 | infested_v3 == 1 | infested_v4 == 1), 1, 0)
  )
# Subset data to identify trees with changes 
newly_inf_trees <- all_merged[all_merged$new_infestation == 1, ]
#no_change_trees <- all_merged[all_merged$ChangeInVariable != 1, ]
all_non_infested_trees <- all_merged %>% dplyr::filter(infested_v1 == 0 & infested_v2 == 0 & infested_v3 == 0 & infested_v4 == 0) 
# Filter trees that weren't infested during either visit

# checking number of points for live hosts
all_points <- rbind(cxsp_points, emig_points, frnk_points, mocr_points, mttp_points, smmh_points, soap_points, tmpf_points)
all_points <- all_points[!is.na(all_points$tag_ID),] # dropping all of our non-host
  

```

## Final Solution to distances

```{r}

# # So we need to convert our decimel degrees to UTM for calculation
 all_merged <- all_merged %>% filter(species == "ABLA" | species == "ABCO")

 # Convert to sf
 all_merged_sf <- st_as_sf(all_merged, coords = c("Longitude", "Latitude"), crs = 4326)

 # Define the target UTM zone
 target_crs <- 32612  # EPSG code for UTM zone 12

 # Transform the coordinates
 all_merged_utm <- st_transform(all_merged_sf, crs = target_crs)
 all_coords <- st_coordinates(all_merged_utm)
 all_coords <- as.data.frame(all_coords)

 # Put back into original datframe

 all_merged_utm <- all_merged
 all_merged_utm$Longitude <- all_coords$X
 all_merged_utm$Latitude <- all_coords$Y

 all_infested <- all_merged_utm[all_merged_utm$infested_v1 == 1, ] # Infested trees with UTM coordinates


 #write_clip(all_merged_utm) # copy dataframe without making file
  all_merged <- all_merged[!duplicated(all_merged), ] #- there shouldn't be duplicates now
 rownames(all_merged_utm) <- all_merged_utm$unique_ID

 rm(Holder)
 Holder = as.data.frame(matrix(nrow=0,ncol=4))
 colnames(Holder) =c("PLOT", "FROM", "TARGET", "DISTANCE")

 #Assign rownames as Tree-ID, dataset also needs UTM_X, UTM_Y, and PLOT
 for( i in levels(as.factor(all_merged_utm$site))){
   subset_frame = subset(all_merged_utm, site == i)
 TEMP_DIST = vegdist(subset_frame[,c("Longitude","Latitude")], method = "euclid")
 TEMP_DIST.df = as.data.frame(as.matrix(TEMP_DIST))
 rownames(TEMP_DIST.df) = (subset_frame$unique_ID)
 colnames(TEMP_DIST.df) = (subset_frame$unique_ID)
 TEMP_DIST.df$`Tag #` = rownames(TEMP_DIST.df)
 TEMP_DIST_MELT = reshape2::melt(TEMP_DIST.df, id.vars = c("Tag #"))
 colnames(TEMP_DIST_MELT) = c("FROM", "TARGET", "DISTANCE")
 #Make a new column with plot ID
 TEMP_DIST_MELT$PLOT = i
 Holder = rbind(Holder, TEMP_DIST_MELT) #This is your exit
 }

 #Merge with infestation data
 names(Holder)[1] <- "unique_ID"

 Holder_new <- merge(Holder, all_merged_utm, by.x = "TARGET", by.y = "unique_ID")
 holder_new.dt <- as.data.table(Holder_new)
 length(unique(holder_new.dt$unique_ID))
 nrow(holder_new.dt)
 
 library(data.table)
 
 # Change NA's in infested columns
 holder_new.dt$infested_v1[is.na(holder_new.dt$infested_v1)] <- 0
  holder_new.dt$infested_v2[is.na(holder_new.dt$infested_v2)] <- 0
   holder_new.dt$infested_v3[is.na(holder_new.dt$infested_v3)] <- 0
    holder_new.dt$infested_v4[is.na(holder_new.dt$infested_v4)] <- 0

 # distance to nearest infested at time 1
 min_dist_holder <- holder_new.dt[infested_v1 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]
  length(unique(min_dist_holder$unique_ID))
  min_dist_holder <- min_dist_holder[!duplicated(min_dist_holder$unique_ID)]

 # distance to nearest infested at time 2
 min_dist_holder2 <- holder_new.dt[infested_v2 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]
 min_dist_holder2 <- min_dist_holder2[!duplicated(min_dist_holder2$unique_ID)]

 # distance to nearest infested at time 3
 min_dist_holder3 <- holder_new.dt[infested_v3 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]
 min_dist_holder3 <- min_dist_holder3[!duplicated(min_dist_holder3$unique_ID)]

 # distance to nearest infested at time 4
 min_dist_holder4 <- holder_new.dt[infested_v4 == 1 ,.(TARGET, DISTANCE),by=.(unique_ID)][, .SD[which(DISTANCE == min(DISTANCE[which(DISTANCE!=0)]))],by=.(unique_ID)]
 min_dist_holder4 <- min_dist_holder4[!duplicated(min_dist_holder4$unique_ID)]


 names(min_dist_holder2)[2] <- "TARGET2"
 names(min_dist_holder2)[3] <- "DISTANCE2"
  names(min_dist_holder3)[2] <- "TARGET3"
 names(min_dist_holder3)[3] <- "DISTANCE3"
  names(min_dist_holder4)[2] <- "TARGET4"
 names(min_dist_holder4)[3] <- "DISTANCE4"

 # add distances to data
 # visit 1 distances
 data_with_nearest_infested_dis <- all_merged_utm %>% inner_join(min_dist_holder, by = "unique_ID")

 # visit 2 distances
 data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder2, by = "unique_ID")
 
 # visit 3 distances
 data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder3, by = "unique_ID")
 
 # visit 4 distances
 data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% inner_join(min_dist_holder4, by = "unique_ID")
 
 data_with_nearest_infested_dis <- data_with_nearest_infested_dis %>% filter(!duplicated(unique_ID)) # get rid of duplicates

#write_xlsx(data_with_nearest_infested_dis,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/data_with_nearest_infested.xlsx")

# Don't need to run code above anymore for the time being

#data_with_nearest_infested_dis <- read_xlsx("data_with_nearest_infested.xlsx")

```

Now we can compare distance to nearest infested in trees that were newly infested and trees that never got infested and see if it's different.

```{r}
newly_inf_trees <- data_with_nearest_infested_dis[data_with_nearest_infested_dis$new_infestation == 1, ]
never_infested_trees <- data_with_nearest_infested_dis %>% filter(infestation_visit_1 == 0 & infestation_visit_2 == 0 & infestation_visit_3 == 0 & infestation_visit_4 == 0) # Filter trees that weren't infested during either visit

ggplot() +
  geom_histogram(data = newly_inf_trees, aes(x = DISTANCE, fill = "Newly infested trees"), position = "stack", bins = 10, alpha = 0.7) +
  geom_histogram(data = never_infested_trees, aes(x = DISTANCE, fill = "Trees that never got infested"), position = "stack", bins = 10, alpha = 0.7) +
  labs(title = "Distance to nearest infested of trees that got infested vs non-infested trees", x = "Distance to nearest infested", y = "Frequency") +
  scale_fill_manual(values = c("Newly infested trees" = "blue", "No change in infestation status" = "red")) +
  theme_minimal()

# Make a model to compare by combining data first and coding changed to 1 and unchanged to 0
newly_inf_trees$Dataset <- 1
never_infested_trees$Dataset <- 0
combined_data_dis <- rbind(newly_inf_trees, never_infested_trees)

combined_data_dis$Dataset <- as.factor(combined_data_dis$Dataset)
ggplot(combined_data_dis, aes(x = site, y = DISTANCE, fill = Dataset)) + 
  labs(title="Distance to nearest infested of trees that got infested vs those that did not - Visit 1",x="Site", y = "Distance to nearest infested tree") +
    geom_boxplot()

# Time 2 
ggplot(combined_data_dis, aes(x = site, y = DISTANCE2, fill = Dataset)) + 
  labs(title="Distance to nearest infested of trees that got infested vs those that did not - Visit 2",x="Site", y = "Distance to nearest infested tree") +
    geom_boxplot()

# Time 3 
ggplot(combined_data_dis, aes(x = site, y = DISTANCE3, fill = Dataset)) + 
  labs(title="Distance to nearest infested of trees that got infested vs those that did not - Visit 2",x="Site", y = "Distance to nearest infested tree") +
    geom_boxplot()

# Time 4 
ggplot(combined_data_dis, aes(x = site, y = DISTANCE4, fill = Dataset)) + 
  labs(title="Distance to nearest infested of trees that got infested vs those that did not - Visit 2",x="Site", y = "Distance to nearest infested tree") +
    geom_boxplot()

# Infested vs Non-infested
combined_data_dis$infestation_visit_1 <- as.factor(combined_data_dis$infestation_visit_1)

# Removing TMPF because of lack of GPS points

# ggplot(combined_data_dis, aes(x = site, y = DISTANCE, fill = as.factor(infestation_visit_1))) + 
#   labs(title="Distance to nearest infested of infested trees vs non-infested - Visit 1",x="Site", y = "Distance to nearest infested tree") +
#   scale_fill_discrete(name = "", labels = c("Uninfested", "Infested")) +
#   theme(legend.position = c(0.8, 0.8), legend.title = element_blank()) +
#     geom_boxplot()


# Model
model <- lm(DISTANCE2 ~ Dataset, data = combined_data_dis)
summary(model)

# T-tests

# Newly infested vs never infested
t.test(newly_inf_trees$DISTANCE, never_infested_trees$DISTANCE)

# infested vs not - Time 1

infested_t1 <- data_with_nearest_infested_dis[data_with_nearest_infested_dis$`Inf Location_Visit1` == 1, ]
non_infested_t1 <- data_with_nearest_infested_dis[data_with_nearest_infested_dis$`Inf Location_Visit1` == 0, ]

t.test(infested_t1$DISTANCE, non_infested_t1$DISTANCE)

# infested vs not - Time 2
infested_t2 <- data_with_nearest_infested_dis[data_with_nearest_infested_dis$`Inf Location_Visit2` == 1, ]
non_infested_t2 <- data_with_nearest_infested_dis[data_with_nearest_infested_dis$`Inf Location_Visit2` == 0, ]

t.test(infested_t2$DISTANCE, non_infested_t2$DISTANCE)
```

## Bearings

I'd like to incorporate bearing from points to each other as a stand-in for wind direction. If there's distinct differences in the bearings from infested trees to non-infested, this may indicate wind directionality is playing an important role in dispersal.

```{r}

# we need to convert our points back from UTM for bearings
str(data_with_nearest_infested_dis)
data_with_nearest_infested_dis <- st_as_sf(data_with_nearest_infested_dis, coords = c("Longitude", "Latitude"), crs = 32612)

# Transform the coordinates
data_with_nearest_infested_dis_4326 <- st_transform(data_with_nearest_infested_dis, crs = 4326)

# put back into dataframe

coords <- st_coordinates(data_with_nearest_infested_dis_4326)
coords <- as.data.frame(coords)
data_with_nearest_infested_dis_4326$Latitude <- coords$Y
data_with_nearest_infested_dis_4326$Longitude <- coords$X

# let's take all of our coordinates
all_coords <- data_with_nearest_infested_dis_4326[, c("unique_ID", "Latitude", "Longitude")]
all_coords.df <- as.data.frame(all_coords)

#####

# Identify our targets and match class with coordinate matrix
targets <- as.character(data_with_nearest_infested_dis_4326$TARGET)

# subset our coordinates by the ID's of our infested trees from Visit 1
infested_subset <- all_coords.df %>% filter(unique_ID %in% targets)
names(infested_subset)[1] = "TARGET"


# Append target coordinates to data frame
data_with_target_lat_long <- inner_join(data_with_nearest_infested_dis_4326, infested_subset, by = "TARGET")

# bearings
# make a data frame to extract coordinate columns
data_with_target_lat_long <- as.data.frame(data_with_target_lat_long)
# calculate bearings
data_with_target_lat_long$bearings1 <- bearing(as.matrix(data_with_target_lat_long[, c("Longitude.x", "Latitude.x")]), as.matrix(data_with_target_lat_long[, c("Longitude.y", "Latitude.y")]))

# change names to make this easier in future
names(data_with_target_lat_long)[104] = "Latitude.focal"
names(data_with_target_lat_long)[105] = "Longitude.focal"
names(data_with_target_lat_long)[106] = "Latitude.target"
names(data_with_target_lat_long)[107] = "Longitude.target"

# Do same with visit 2
# Identify our targets and match class with coordinate matrix
targets2 <- as.character(data_with_nearest_infested_dis_4326$TARGET2)

# subset our coordinates by the ID's of our infested trees from Visit 2
infested_subset2 <- all_coords.df %>% filter(unique_ID %in% targets2)
names(infested_subset2)[1] = "TARGET2"


# Append target coordinates to data frame
data_with_target_lat_long <- inner_join(data_with_target_lat_long, infested_subset2, by = "TARGET2")

# change names
names(data_with_target_lat_long)[110] = "Latitude.target2"
names(data_with_target_lat_long)[111] = "Longitude.target2"


# bearings - visit2
# calculate bearings
data_with_target_lat_long$bearings2 <- bearing(as.matrix(data_with_target_lat_long[, c("Longitude.focal", "Latitude.focal")]), as.matrix(data_with_target_lat_long[, c("Longitude.target2", "Latitude.target2")]))

# Visit 3
# Identify our targets and match class with coordinate matrix
targets3 <- as.character(data_with_nearest_infested_dis_4326$TARGET3)

# subset our coordinates by the ID's of our infested trees from Visit 2
infested_subset3 <- all_coords.df %>% filter(unique_ID %in% targets3)
names(infested_subset3)[1] = "TARGET3"


# Append target coordinates to data frame
data_with_target_lat_long <- inner_join(data_with_target_lat_long, infested_subset3, by = "TARGET3")

# change names
names(data_with_target_lat_long)[114] = "Latitude.target3"
names(data_with_target_lat_long)[115] = "Longitude.target3"


# bearings - visit3
# calculate bearings
data_with_target_lat_long$bearings3 <- bearing(as.matrix(data_with_target_lat_long[, c("Longitude.focal", "Latitude.focal")]), as.matrix(data_with_target_lat_long[, c("Longitude.target3", "Latitude.target3")]))

# Identify our targets and match class with coordinate matrix
targets4 <- as.character(data_with_nearest_infested_dis_4326$TARGET4)

# subset our coordinates by the ID's of our infested trees from Visit 2
infested_subset4 <- all_coords.df %>% filter(unique_ID %in% targets4)
names(infested_subset4)[1] = "TARGET4"


# Append target coordinates to data frame
data_with_target_lat_long <- inner_join(data_with_target_lat_long, infested_subset4, by = "TARGET4")

# change names
names(data_with_target_lat_long)[118] = "Latitude.target4"
names(data_with_target_lat_long)[119] = "Longitude.target4"


# bearings - visit4
# calculate bearings
data_with_target_lat_long$bearings4 <- bearing(as.matrix(data_with_target_lat_long[, c("Longitude.focal", "Latitude.focal")]), as.matrix(data_with_target_lat_long[, c("Longitude.target4", "Latitude.target4")]))

# Now we have a dataframe with all of our trees, the closest infested neighbor as well as its bearing - Fuck yeah

duplicates <- data_with_target_lat_long[duplicated(data_with_target_lat_long), ]
data_with_target_lat_long <- unique(data_with_target_lat_long)

#write_xlsx(data_with_target_lat_long,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/data_with_bearings.xlsx")

data_with_target_lat_long <- read_xlsx("data_with_bearings.xlsx")

```

# Infestation Sum

Having an infestation sum variable will allow us to look at how infested each tree is at every time step and incorporate it into our model

```{r}
# Visit 1
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #   mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")), 
 #             funs(if_else(startsWith(., "1-"), "1", .))) # Change all "1-" ratings to just a singular 1 value
 
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #   mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")), as.numeric)
 # 
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #   mutate_at(vars(c("NL_Visit1", "NM_Visit1", "NH_Visit1", "SL_Visit1", "SM_Visit1", "SH_Visit1")),~coalesce(., 0)) # change all NA's to 0
 
# ^ already did this

data_with_target_lat_long$infestation_sum_Visit1_focal <- data_with_target_lat_long$NL_v1 + data_with_target_lat_long$NM_v1 + data_with_target_lat_long$NH_v1 + data_with_target_lat_long$SL_v1 + data_with_target_lat_long$SM_v1 + data_with_target_lat_long$SH_v1
 
 # Visit 2
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #   mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")), 
 #             funs(if_else(startsWith(., "1-"), "1", .))) # Change all "1-" ratings to just a singular 1 value
 # 
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #   mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")), as.numeric)
 # 
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #   mutate_at(vars(c("NL_Visit2", "NM_Visit2", "NH_Visit2", "SL_Visit2", "SM_Visit2", "SH_Visit2")),~coalesce(., 0)) # change all NA's to 0

# ^ did this in earlier script
 
# Visit 2 infestation sum
data_with_target_lat_long$infestation_sum_Visit2_focal <- data_with_target_lat_long$NL_v2 + data_with_target_lat_long$NM_v2 + data_with_target_lat_long$NH_v2 + data_with_target_lat_long$SL_v2 + data_with_target_lat_long$SM_v2 + data_with_target_lat_long$SH_v2

# Visit 3 infestation sum
data_with_target_lat_long$infestation_sum_Visit3_focal <- data_with_target_lat_long$NL_v3 + data_with_target_lat_long$NM_v3 + data_with_target_lat_long$NH_v3 + data_with_target_lat_long$SL_v3 + data_with_target_lat_long$SM_v3 + data_with_target_lat_long$SH_v3

# Visit 4 infestation sum
data_with_target_lat_long$infestation_sum_Visit4_focal <- data_with_target_lat_long$NL_v4 + data_with_target_lat_long$NM_v4 + data_with_target_lat_long$NH_v4 + data_with_target_lat_long$SL_v4 + data_with_target_lat_long$SM_v4 + data_with_target_lat_long$SH_v4
 
# Subset target trees to get infestation sums
 infested_subset_inf_sum <- data_with_target_lat_long %>% filter(unique_ID %in% targets) %>% dplyr::select(c("unique_ID", "infestation_sum_Visit1_focal", "infestation_sum_Visit2_focal", "infestation_sum_Visit3_focal", "infestation_sum_Visit4_focal"))
 names(infested_subset_inf_sum)[1] <- "TARGET"
 names(infested_subset_inf_sum)[2] <- "inf_sum_Visit1_target"
 names(infested_subset_inf_sum)[3] <- "inf_sum_Visit2_target"
 names(infested_subset_inf_sum)[4] <- "inf_sum_Visit3_target"
 names(infested_subset_inf_sum)[5] <- "inf_sum_Visit4_target"
 
 
 
  # append sums to our main dataframe
# Append target coordinates to data frame
data_with_target_lat_long <- inner_join(data_with_target_lat_long, infested_subset_inf_sum, by = "TARGET")
 str(data_with_target_lat_long)
 
 # Add a split tree variable before we get rid of duplicates
 
 # data_with_target_lat_long <- unique(data_with_target_lat_long)
 # 
 # data_with_target_lat_long <- data_with_target_lat_long %>%
 #       distinct(unique_ID, .keep_all = TRUE)

#write_xlsx(data_with_target_lat_long,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/data_with_nearest_info.xlsx")

data_with_target_lat_long <- read_xlsx("data_with_nearest_info.xlsx")
```

# Function to extract info from target trees

```{r}
df <- data_with_target_lat_long
df <- df[!duplicated(df), ]
# Create a function to extract the variable based on the linked tree ID
extractVariable <- function(TARGET) {
  linkedTreeRow <- df[df$unique_ID == TARGET, ]
  if (nrow(linkedTreeRow) > 0) {
    return(linkedTreeRow$dbh)
  } else {
    return(NA)  # or any other value to indicate missing data
  }
}

# Need to run this to get rid of duplicates before extracting DBH's
df <- df %>%
      distinct(unique_ID, .keep_all = TRUE)

# Apply the function to create a new column with the extracted variable
df$DBH_TARGET <- sapply(df$TARGET, extractVariable)
df$DBH_TARGET2 <- sapply(df$TARGET2, extractVariable)
df$DBH_TARGET3 <- sapply(df$TARGET3, extractVariable)
df$DBH_TARGET4 <- sapply(df$TARGET4, extractVariable)

```

# Try to plot all sites for visual

```{r}

# Change small trees to 0.5 so we can visualize them

df_viz <- data_with_target_lat_long
df_viz$dbh[is.na(df_viz$dbh)] <- 0.5

for (i in unique(df_viz$site)) {
p <- ggplot(subset(df_viz, site == i), aes(x = Longitude.focal, y = Latitude.focal, size = dbh, color = factor(infestation_sum_Visit1_focal))) +
  geom_point() +
  
  scale_color_viridis_d(option = "plasma") +
  facet_wrap(~site, scales = "free") +
  labs(title = paste(i, "Infestation Sum and DBH"),
       x = "",
       y = "",
       color = "Infestation Sum",
       size = "DBH(in)") +
  theme_bw() +
   theme(axis.text.x = element_blank(),  # Remove x-axis values
          axis.text.y = element_blank(),
     legend.key.size = unit(0.5, "cm"))

print(p)
}

# NA's in this are beacause we have some trees that show up in later visits that aren't here because we're working with visit 1


```



