---
title: "Non-hosts"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Non-host stuff
12/11/23
Michael Wayman

The goal of this script is to look at our plots in which we have all of our host and non-host trees available (Cox Springs,     Emigration Canyon, and Smith and Morehouse) and calculate various measures of infestation spread as a function of non-host characeristics. We'll look at metrics such as non-host stocking, QMD and SDI and how these may affect our infestation metrics that we've used a few times up to this point such as change in infestation prevalence by basal area and tree count.

** We now have all of our data available

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```
## Clean out global environment

```{r}
rm(list = ls())
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(rfishnet2)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(plyr)
library(reshape2)
library(geosphere)
```

## Load Data

```{r}
combined <- read_xlsx("combined_cleaned2.xlsx")
```

## Basal area calculation

To get our stand basal area(BA), we'll calculate and sum every host tree BA. The formula for a tree's BA is: 

Tree BA ($Ft^{2}$) = $\left(\frac{\pi}{576}\right)$ * $DBH^{2}$

Let's first do this for all of our trees and add a column.

```{r}
combined$DBH <- combined$DBH * 2.54 # convert DBH to cm
combined <- mutate(combined, BA = (0.0000785 * (DBH ^ 2)))
combined$Species <- as.factor(combined$Species) # make Species a factor for easier summation later
combined_na_drop <- combined[!is.na(combined$BA),]
str(combined) # lots of NA's that we'll deal with later


```

## Then we'll just grab our non-host sites

```{r}

cxsp_data_1 <- combined_na_drop %>% filter(`Site ID` == "CXSP" & Visit == 1) # Data includes both visits
emig_data_1 <- combined_na_drop %>% filter(`Site ID` == "EMIG" & Visit == 1)
smmh_data_1 <- combined_na_drop %>% filter(`Site ID` == "SMMH" & Visit == 2) # non-host done on 2nd visit

```

## Then we can sum up BA by species for each site

```{r}

cxsp_ba_sum <- cxsp_data_1 %>% 
               dplyr::group_by(Species) %>%
               dplyr::summarise(ba_sum = (sum(BA)/0.25), # This gives us sq meter per hectare
                                count = gdata::nobs(Species, na.rm = T),
                                 TPHA = (gdata::nobs(Species)/0.25),
                                 QMD = (sqrt(ba_sum/TPHA))*13.54,
                                 SDI = TPHA*((QMD/25.4)^(1.605)))

emig_ba_sum <- emig_data_1 %>%
                dplyr::group_by(Species) %>%
                dplyr::summarise(ba_sum = (sum(BA)/0.25),
                                 count = gdata::nobs(Species, na.rm = T),
                                 TPHA = (gdata::nobs(Species)/0.25),
                                 QMD = (sqrt(ba_sum/TPHA))*13.54,
                                 SDI = TPHA*((QMD/25.4)^(1.605)))

smmh_ba_sum <- smmh_data_1 %>%
                dplyr::group_by(Species) %>%
                dplyr::summarise(ba_sum = (sum(BA)/0.25),
                                 count = gdata::nobs(Species, na.rm = T),
                                 TPHA = (gdata::nobs(Species)/0.25),
                                 QMD = (sqrt(ba_sum/TPHA))*13.54,
                                 SDI = TPHA*((QMD/25.4)^(1.605)))
cxsp_ba_sum
emig_ba_sum
smmh_ba_sum

all_non_host <- bind_rows(cxsp_ba_sum, emig_ba_sum, smmh_ba_sum)

# Group by the shared character column and summarize to sum the numeric columns
summarized_df <- all_non_host %>%
  group_by(Species) %>%
  summarize_all(sum)

summarized_df <- summarized_df[-2,]

write_clip(summarized_df)

```
