---
title: "Plot-level stats"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Preliminary Analysis
10/31/23
Michael Wayman

Plot-level statistics of Abies lasiocarpa (due to lack of non-host data) in association with Dispersal of Balsam Woolly Adelgid in the Intermountain Region. The goal is to attain plot-level statistics including:

1. Prevalence by tree count, basal area, and by section (check, check, and check) 
2. BA (check)
3. QMD (check)
4. Plot-level density (SDI) (check)
5. Average Heights (if available) - look at canopy heights, there's a mistake of 736 somewhere in there and get rid of 0's - this may not be hepful atm honestly, I think calculations using this will help in the future, but what does average canopy height tell us? not much
*
6. Degree days - I think I should do temp stuff on a different file
7. Mean temps
*

8. Change of prevalence from Time 1 to Time 2 (check)
9. Species composition (Unable to complete as of 10/31/23 due to lack of non-host data)

# Set up

## Directory

## Clean out global environment

```{r}
rm(list = ls())
```

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(Rcpp)
library(readr)
library(dplyr)
library(tinytex)
library(reshape2)
library(fastmap)
library(pandoc)
library(writexl)
```

## Data

Look at cleaning script for info on initial cleaning

```{r import}
combined <- read_xlsx("combined_cleaned.xlsx")
```

# Prevalence by tree count

For most calculations, we'll need to filter out all alive abla.

For prevalence, we're going to change infestation location to a binary variable. This will make it easier to calculate prevalence. For prevalence itself, we'll do it two different ways; percent prevalence by tree count and percent prevalence by stand basal area.

First, we'll filter all live ABLA.

```{r get live}
live_abla <- combined %>% dplyr::filter(Species == "ABLA" & combined$`Dead / Alive` == 1)
unique(live_abla$`Dead / Alive`) # check to make sure filter worked
unique(live_abla$Species)
```

Then, we'll make our Location a binary variable.

```{r code location}
live_abla$`Inf Location`[ live_abla$`Inf Location`== "-"] <- 0
unique(live_abla$`Inf Location`)
live_abla$`Inf Location`[live_abla$`Inf Location` == "BO" | live_abla$`Inf Location` == "BR, BO" | live_abla$`Inf Location` == "BR"] <- 1
```

Now we can get the prevalence by total trees for each visit.

```{r get percentages}
unique(live_abla$`Inf Location`) # found some NA's
live_abla <- live_abla[!is.na(live_abla$`Inf Location`),] # drop NA's

visit_1 <- live_abla %>% filter(live_abla$Visit == 1)
visit_2 <- live_abla %>% filter(live_abla$Visit == 2)

prev_table_1 <- table(visit_1$`Site ID`, visit_1$`Inf Location`) # make a table of status by site
prev_table_2 <- table(visit_2$`Site ID`, visit_2$`Inf Location`)

perc_table_1 <- prop.table(prev_table_1, margin = 1) %>% `*`(100) %>% round(2) # look at percentages by site

perc_table_2 <- prop.table(prev_table_2, margin = 1) %>% `*`(100) %>% round(2)


final_percentages_1 <- perc_table_1[,"1"]
final_percentages_2 <- perc_table_2[,"1"]
final_percentages_1
final_percentages_2
percent_difference <- final_percentages_2 - final_percentages_1
```

We're seeing percentages rise at every site by tree numbers alone, let's see how this changes when we look at infestation by basal area.

# Prevalence by basal area

## Basal area calculation by site

To get our stand basal area(BA), we'll calculate and sum every host tree BA. The formula for a tree's BA is: 

Tree BA ($Ft^{2}$) = $\left(\frac{\pi}{576}\right)$ * $DBH^{2}$

Let's first do this for all of our trees and add a column.

```{r calculate BA}
combined <- mutate(combined, BA = (0.005454154 * (DBH ^ 2)))
head(combined) # lots of NA's that we'll deal with later
```

Now we'll sum up our host tree BA for each infested/non-infested as well as site totals on every site

```{r BA calc and perc}
live_abla <- combined %>% dplyr::filter(Species == "ABLA" & combined$`Dead / Alive` == 1)# we only need infested and site total, so lets drop our non-infested trees

## we need to run this code again to get binary infestation again
live_abla$`Inf Location`[ live_abla$`Inf Location`== "-"] <- 0
unique(live_abla$`Inf Location`)
live_abla$`Inf Location`[live_abla$`Inf Location` == "BO" | live_abla$`Inf Location` == "BR, BO" | live_abla$`Inf Location` == "BR"] <- 1
##

infested_ba_1 <- live_abla %>% filter(`Inf Location` == 1 & Visit == 1)
infested_ba_2 <- live_abla %>% filter(`Inf Location` == 1 & Visit == 2) # we re-make our live host tree, now with BA

visit_1_ba <- live_abla %>% filter(live_abla$Visit == 1) # we'll remake our visits with BA
visit_2_ba <- live_abla %>% filter(live_abla$Visit == 2) # the stand BA may differ slightly between visits because of trees added/lost

visit_1_ba <- visit_1_ba[!is.na(visit_1_ba$BA),] # drop NA's for sum
site_ba_1 <- visit_1_ba %>% dplyr::group_by(`Site ID`) %>% summarise("Stand BA" = sum(BA/0.25) * 0.404686) # sum BA and convert to square feet per acre

visit_2_ba <- visit_2_ba[!is.na(visit_2_ba$BA),]
site_ba_2 <- visit_2_ba %>% group_by(`Site ID`) %>% summarise("Stand BA" = sum(BA/0.25) * 0.404686) # sum BA

# we had some NA's in the Inf location, so lets drop those before we group by status
visit_1_ba <- visit_1_ba[!is.na(visit_1_ba$`Inf Location`),]

# we only need infested and site total, so lets drop our non-infested trees
infested_ba_1 <- visit_1_ba %>% filter(`Inf Location` == 1)
infested_ba_2 <- visit_1_ba %>% filter(`Inf Location` == 1)

infested_ba_1 <- infested_ba_1 %>% group_by(`Site ID`) %>% summarise(BA_by_status = sum(BA/0.25) * 0.404686)
infested_ba_2 <- infested_ba_2 %>% group_by(`Site ID`) %>% summarise(BA_by_status = sum(BA/0.25) * 0.404686)


site_ba_1 # Remember, this is JUST Live ABLA, not all trees, so not all BA on the site
site_ba_2 # Remember, this is JUST Live ABLA, not all trees, so not all BA on the site

# Now we can combine our two summed variables and create a percentage based on basal area

ba_infested_1 <- inner_join(infested_ba_1,site_ba_1,by="Site ID") %>% mutate(Percentage_infested_by_BA = (BA_by_status/`Stand BA`)*100)
ba_infested_2 <- inner_join(infested_ba_2,site_ba_2,by="Site ID") %>% mutate(Percentage_infested_by_BA = (BA_by_status/`Stand BA`)*100)

# Let's take that percentage column and compare it with our percentages we got by pure tree count

perc_infested_ba_1 <- ba_infested_1 %>% select(`Site ID`, Percentage_infested_by_BA) # make a new object with just percentages
perc_infested_ba_2 <- ba_infested_2 %>% select(`Site ID`, Percentage_infested_by_BA)

# Try to combine these percentages with our tree count percentages above
# What we'll do is extract our percentages above and tack them on to our precentages that we just made
perc_by_tree_1 <- as.vector(final_percentages_1)
perc_by_tree_2 <- as.vector(final_percentages_2)

perc_infested_ba_1$perc_by_tree <- perc_by_tree_1
perc_infested_ba_1$visit <- 1 # gonna add a visit here so I can stack these two tables
perc_infested_ba_2$perc_by_tree <- perc_by_tree_2
perc_infested_ba_2$visit <- 2
perc_infested_ba <- perc_infested_ba_1 %>% bind_rows(perc_infested_ba_2) # Smash these two puppies together so I can graph it out

perc_infested_ba <- perc_infested_ba %>% group_by(visit)

## Now we have a table showing infestation percentages by BA as well as by tree count. They vary quite a bit
# I want to get a table with both BA percentages on there so I can easily subtract them, but I think I need to brute force it unfortunatley, 
perc_difference_by_ba <- perc_infested_ba %>% reshape2::dcast(`Site ID` ~ visit, value.var = "Percentage_infested_by_BA")
names(perc_difference_by_ba)[2:3] = c("visit1", "visit2") # cant mutate with numbers as variable names
perc_difference_by_ba <- perc_difference_by_ba %>% mutate(diff = visit2 - visit1)

```

I'm thinking now that I'd like a couple decent graphs just to show this prevalence disparity and to try and visualize what the prevalence changes are from visit 1 to visit 2. I've combined the two summary tables together up above so we can plot it out nicely.

```{r plot percentages}
ggplot(perc_infested_ba, aes(fill = factor(visit), y = Percentage_infested_by_BA, x = `Site ID`)) + 
    geom_bar(position="dodge", stat="identity") +
    labs(x="Sites", y="% Infestation by Basal Area") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + # gets rid of plot lines
    scale_fill_discrete(labels=c('Visit 1', 'Visit 2')) +
    scale_fill_discrete(name = "Visit")

ggplot(perc_infested_ba, aes(fill = factor(visit), y = perc_by_tree, x = `Site ID`)) + 
    geom_bar(position="dodge", stat="identity") +
    labs(x="Sites", y="% Infestation by Tree Count") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + # gets rid of plot lines
    scale_fill_discrete(labels=c('Visit 1', 'Visit 2')) +
    scale_fill_discrete(name = "Visit")
```

# Prevalence by section

So now I want to calculate prevalence by section. Meaning, I want to know what percentage of all the sections measured (6 per tree) were infested at the time of assessment. We don't care how infested they were, just if they were infested or not. So we'll convert our ratings for our live hosts into a binary (1 or 0) variable and then sum up the ratings for each site. This will give us percentages just like our previous prevalence calculations. I expect these percentages to be lower than the others because we have so many more sections with less of them being infested.

```{r prevalence by section, warning=FALSE}
abies <- combined %>% filter(Species == "ABLA" | Species == "ABCO") # Filter Hosts
live_hosts_1 <- abies %>% filter(`Dead / Alive` == 1 & Visit == 1) # Filter Live and first visit
section_prev_1 <- live_hosts_1 # make new object to change a bunch of shtuff
section_prev_1[section_prev_1 == "-"] <- "0" # change all "-" to a zero
section_prev_1$NL[section_prev_1$NL != "0"] <- 1 # replace all values in column that aren't 0
section_prev_1$NM[section_prev_1$NM != "0"] <- 1 # Do this for all sections
section_prev_1$NH[section_prev_1$NH != "0"] <- 1
section_prev_1$SL[section_prev_1$SL != "0"] <- 1
section_prev_1$SM[section_prev_1$SM != "0"] <- 1
section_prev_1$SH[section_prev_1$SH != "0"] <- 1
# to make tables easier later, we'll dchange all these sections to numeric
section_prev_1[,13:24] <- lapply(section_prev_1[,13:24], as.numeric)
section_prev_1[, 13:24][is.na(section_prev_1[, 13:24])] <- 0 # replace NA's with 0s

# Do the same for visit 2
live_hosts_2 <- abies %>% filter(`Dead / Alive` == 1 & Visit == 2)
section_prev_2 <- live_hosts_2
section_prev_2$NL[is.na(section_prev_2$NL)] <- 0
section_prev_2$NL <- as.numeric(section_prev_2$NL)
section_prev_2[section_prev_2 == "-"] <- "0" # change all "-" to a zero
section_prev_2$NL[section_prev_2$NL != "0"] <- 1 # replace all values in column that aren't 0
section_prev_2$NM[section_prev_2$NM != "0"] <- 1 # Do this for all sections
section_prev_2$NH[section_prev_2$NH != "0"] <- 1
section_prev_2$SL[section_prev_2$SL != "0"] <- 1
section_prev_2$SM[section_prev_2$SM != "0"] <- 1
section_prev_2$SH[section_prev_2$SH != "0"] <- 1
# to make tables easier later, we'll dchange all these sections to numeric
section_prev_2[,13:24] <- lapply(section_prev_2[,13:24], as.numeric)
section_prev_2[, 13:24][is.na(section_prev_2[, 13:24])] <- 0 # replace NA's with 0s

# Now that we have a binary variable in our sections, we can sum and get percentages 

section_prev_2 %>% 
  group_by(`Site ID`) %>% 
  summarise(
    NL_prop = mean(NL) * 100,
    NM_prop = mean(NM) * 100,
    NH_prop = mean(NH) * 100,
    SL_prop = mean(SL) * 100,
    SM_prop = mean(SM) * 100,
    SH_prop = mean(SH) * 100
  ) # This code gives us the percentage infested of each section per site

# Drop Window variables for now
section_prev_1 <- section_prev_1 %>% select(-c(NLW, NMW, NHW, SLW, SMW, SHW))
section_prev_2 <- section_prev_2 %>% select(-c(NLW, NMW, NHW, SLW, SMW, SHW))

# Calculate total count of 1's across sections
section_prev_1 <- section_prev_1 %>%
  mutate(total_infested = NL + NM + NH + SL + SM + SH)

section_prev_2 <- section_prev_2 %>%
  mutate(total_infested = NL + NM + NH + SL + SM + SH)
# Add a column for section total
section_prev_1 <- section_prev_1 %>%
  mutate(total_observations = 6)

section_prev_2 <- section_prev_2 %>%
  mutate(total_observations = 6)

# Calculate proportions
Percent_infested_by_section_1 <- section_prev_1 %>%
  group_by(`Site ID`) %>%
  summarise(
    prop_infested_1_by_section = sum(total_infested) / sum(total_observations) * 100
  )

Percent_infested_by_section_2 <- section_prev_2 %>%
  group_by(`Site ID`) %>%
  summarise(
    prop_infested_2_by_section = sum(total_infested) / sum(total_observations) * 100
  )

# View the result
print(Percent_infested_by_section_1)
print(Percent_infested_by_section_2)
Percent_infested_by_section_1$prop_infested_2_by_section <- Percent_infested_by_section_2$prop_infested_2_by_section
Percent_infested_by_section_1 # Final table to make plot with

# Plot Section prevalence
plot_section_perc <- melt(Percent_infested_by_section_1, id.vars='Site ID')
ggplot(plot_section_perc, aes(fill = variable, y = value, x = `Site ID`)) + 
    geom_bar(position="dodge", stat="identity") +
    labs(x="Sites", y="% Infestation by section") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + # gets rid of plot lines
    scale_fill_discrete(labels=c('Visit 1', 'Visit 2'))
```

# QMD

Now let's get QMD for all of our sites to get an idea of diameter distribution among our trees. The calculation for QMD of a stand is: QMD Equation: sqrt(Plot BA/Plot TPA) * 13.54 (I don't know why this isn't working in markdown, oh well)


```{r TPA calc}
abla <- combined %>% dplyr::filter(Species == "ABLA" & Visit == 1) # lets work with fir, but this time all fir including dead and lets just use trees from first visit to avoid confusion at the moment

counts <- abla %>% count(`Site ID`) # counts of fir for each site

counts <- counts %>% mutate(TPA = (n/0.25) * 0.404686) # This is where we are converting trees/(1/4) hectare to trees per acre
counts$TPA <- round(counts$TPA, digits = 0) # round result because there's no partial trees in real life

# We need to remake our site BA's to include dead fir too

all_abla <- combined %>% dplyr::filter(Species == "ABLA") # we re-make our host tree, now with BA
visit_1_ba_all_abla <- all_abla %>% filter(all_abla$Visit == 1) # we'll remake our visits with BA
visit_1_ba_all_abla <- visit_1_ba_all_abla[!is.na(visit_1_ba_all_abla$BA),] # drop NA's for sum
site_ba_1_all_abla <- visit_1_ba_all_abla %>% group_by(`Site ID`) %>% summarise("Stand BA" = sum(BA))


counts <- inner_join(counts, site_ba_1_all_abla, by = "Site ID")
counts <- counts %>% mutate(Average_Basal_Area_Per_Tree = `Stand BA` / TPA) %>% mutate(QMD = (sqrt(`Stand BA`/TPA))*13.54)
counts # This table gives us counts, TPA, Stand BA, Average basal area per tree, and QMD for every site *on visit 1*
```


For plot density, we'll be using Rieneke's SDI: SDI = N(Dq/10)^b where N = TPA, Dq = QMD and b = -1.605. We'll just add that onto our existing count table from above so we have all of our plot statistics in one nice table.

```{r final summary table}
counts <- counts %>% mutate(SDI = TPA*((QMD/10)^(-1.605)))
counts
counts <- counts %>% full_join(Percent_infested_by_section_1, by = "Site ID")
counts$Percent_by_tree_1 <- perc_by_tree_1
counts$Percent_by_tree_2 <- perc_by_tree_2
site_stats <- counts
write_xlsx(site_stats,"C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning/site_stats.xlsx")
```

# Heights

We only have heights for a few sites, but let's try to get the average heights for te sites where we do have them

After comparing heights after filtering with heights of all fir, it looks like we have all fir heights for:
- CXSP
- EMIG
- FRNK
- MTTP
- SMMH
```{r heights}
heights <- combined[!is.na(combined$Height),] # LEt's first drop all of our NA's in our height variable
heights <- heights %>% dplyr::filter(Species == "ABLA" & Visit == 1) # let's also jst focus on live ABLA from one visit


ht_count <- heights %>% count(`Site ID`) # let's see how many heights we have for each site
compare <- abla %>% count(`Site ID`)

heights <- heights %>% filter(`Site ID` == "CXSP" | `Site ID` == "EMIG" | `Site ID` == "FRNK" | `Site ID` == "MTTP" | `Site ID` == "SMMH") # filter for all the sites we have heights for

average_heights <- heights %>% group_by(`Site ID`) %>% summarise(mean_ht = mean(Height))

average_heights

# We'll do the same thing for canopy heights
canopy_hts <- combined[!is.na(combined$`Canopy Height`),]
canopy_hts <- canopy_hts %>% dplyr::filter(Species == "ABLA" & Visit == 1) # let's also jst focus on live ABLA from one visit


canopy_ht_count <- canopy_hts %>% count(`Site ID`) # let's see how many heights we have for each site
compare_canopy <- abla %>% count(`Site ID`)

canopy_hts <- canopy_hts %>% filter(`Site ID` == "CXSP" | `Site ID` == "EMIG" | `Site ID` == "FRNK" | `Site ID` == "MTTP" | `Site ID` == "SMMH") # filter for all the sites we have heights for

average_canopy_heights <- canopy_hts %>% group_by(`Site ID`) %>% summarise(mean_ht = mean(`Canopy Height`))
average_canopy_heights
```



```{r plot-2}

```

### 

```{r plot-3}

```



```{r plot-4}

```






## Vizualize

