---
title: "Point Processes Redux"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Point Processes again
2/8/24
Michael Wayman

The purpose of this script is to look at point processes using Rob Fletcher and Marie-Josie Fortin's Spatial Ecology and Conservation Modeling book. We want to get spatial statistics for our sites using marked models as well as including edge effects. Let's get to it!

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

# clean out environment

```{r}
rm(list = ls())

```

# Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(Matrix) # to use glmer
library(lme4) # to use glmer

```

# Load data

```{r}

trees <- read_xlsx("data_with_nearest_info.xlsx")

all_trees <- read_xlsx("trees_for_spatial_tests.xlsx")

```

# Coordinate conversion

We need to convert our coordinates to get results in an interpretable format

```{r}

# library(sp)
# xy = data.frame(trees$Longitude.focal, trees$Latitude.focal)
# coordinates(xy) <- c("trees.Longitude.focal", "trees.Latitude.focal")
# proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")
# NE <- spTransform(xy, CRS("+proj=utm +zone=12 ellps=WGS84"))
# NE <- as.data.frame(NE)
# 
# trees$Longitude.focal <- NE$coords.x1
# trees$Latitude.focal <- NE$coords.x2

```

# Combine non-host with host - leaving this out for first run !!

# ```{r}
# 
# # get rid of abla in points
# non_host_points_cxsp <- head(cxsp_points, 257)
# non_host_points_cxsp <- non_host_points_cxsp[-10,] # run twice to get rid of 10 and 11
# non_host_points_cxsp$site <- "CXSP" # add site for combination
# names(non_host_points_cxsp)[2] <- "Latitude.focal" # change names for combination
# names(non_host_points_cxsp)[3] <- "Longitude.focal"
# non_host_points_cxsp$Species <- str_sub(non_host_points_cxsp$Name, start = 1, end = 4) %>% stringr::str_to_upper() # Make species name for later
# 
# cxsp <- bind_rows(cxsp, non_host_points_cxsp) # add together
# 
# ```

# Create necessary objects

# IMPORTANT CAVEAT !!!!

- We'll need to get points for our corners to use any of the indices created from these point patterns to make it standardized. At the moment, we are just changing bounds to include all of our points, which will not reflect reality since we are considering edge effects.

```{r}



cxsp <- all_trees %>% dplyr::filter(all_trees$site == "CXSP")
# change names for combination with non-host
cxsp$Name <- as.character(cxsp$Name)

cxsp_points <- read_xlsx("cxsp_points.xlsx")

#ppp.window_cxsp <- owin(xrange = c(41.641, 41.6416), yrange = c(-111.4924, -111.4917))

#ppp.cxsp <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window_cxsp)

# minimun window

ppp.window_cxsp.min <- owin(xrange = c(min(cxsp$Latitude.focal), max(cxsp$Latitude.focal)), yrange = c(min(cxsp$Longitude.focal), max(cxsp$Longitude.focal)))

ppp.cxsp.min <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window_cxsp.min)

plot(ppp.cxsp.min)
summary(ppp.cxsp.min)

ggplot(cxsp, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```

## Quadrat counts and test

```{r}
q <- quadratcount(ppp.cxsp.min, nx = 5, ny = 5) # 10x10m quadrats
plot(ppp.cxsp.min)
plot(q)
quadrat.test(ppp.cxsp.min, nx = 5, ny = 5, method = "Chisq")
```

## Ripley's K and L functions

```{r}

knone <- Kest(ppp.cxsp.min, correction = "none") # no edge effect correction
plot(knone)

lnone <- Lest(ppp.cxsp.min, correction = "none")
plot(lnone) # standardized to 1:1 expectation
plot(lnone, .-r~r) # standardized to zero expectation
```

## With weights

```{r}
Liso <- Lest(ppp.cxsp.min, correction = "isotropic") # weights points at edge
plot(Liso, .-r~r)

Ltrans <- Lest(ppp.cxsp.min, correction = "translate") # pacman edge correction
plot(Ltrans, .-r~r)

```

## Monte Carlo simulations

```{r}
Lcsr <- envelope(ppp.cxsp.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)

```
# Estimate g

This uses a "ring" instead of a circle using r (radius)

```{r}
Ptrans <- pcf(ppp.cxsp.min, correction = "translate")
plot(Ptrans)

Penv <- envelope(ppp.cxsp.min, pcf, nsim = 999, rank = 1, correction = "translate", global = FALSE)
plot(Penv, shade = c("hi", "lo"), legend = FALSE)

```


```{r}
Gtrans = Gest(ppp.cxsp.min, correction = "rs") # reduced sample correction
plot(Gtrans, legend = F)

Genv <- envelope(ppp.cxsp.min, Gest, nsim = 999, rank = 1, correction = "rs", global = FALSE) # utilizes nearest neighbor distances so r tends to be a lot smaller in these plots
plot(Genv, shade = c("hi", "lo"), legend = FALSE)

```

# General conclusion (Second-order Univariate Point Pattern Analysis)

From these graphs, we see that overall for just the distribution of trees at Cox Spring, we are seeing distribution of trees that are more aggregated than random for all distances and that there is not a random distribution at any scale. This is to be expected. When we move to a marked design, we hope to see more interesting results

# Marked Point Pattern

```{r}
cxsp$infestation_visit_1[is.na(cxsp$infestation_visit_1)] <- 0 # change NA's to 0
cxsp$infestedPA <- as.factor(ifelse(cxsp$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window_cxsp.min, marks = cxsp$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA) # summary stats by mark
plot(split(ppp.PA)) # plots of presence/absence

```

# Infestation ignoring host distribution

```{r}

infest_data <- subset(cxsp, infestation_visit_1 > 0)
#ppp.bug <- ppp(infest_data$Latitude.focal, infest_data$Longitude.focal, window = ppp.window_cxsp)

#Lbug <- envelope(ppp.bug, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# minimum window
ppp.PA.cxsp.min <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window_cxsp.min, marks = cxsp$infestedPA)

Lmulti.cxsp.min <- envelope(ppp.PA.cxsp.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.cxsp.min)))

plot(Lmulti.cxsp.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

Kmulti.cxsp.min <- envelope(ppp.PA.cxsp.min, Kcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.cxsp.min)))

plot(Kmulti.cxsp.min)

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
Lmulti <- envelope(ppp.PA.cxsp.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.cxsp.min)))

plot(Lmulti,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

# Want to test out getting ride of duplicate points to see if anything changes with the end result - doesn't look too different
cxsp.nodupe <- cxsp %>% distinct(Latitude.focal, .keep_all = TRUE)


ppp.window_cxsp.min.nodupe <- owin(xrange = c(min(cxsp.nodupe$Latitude.focal), max(cxsp.nodupe$Latitude.focal)), yrange = c(min(cxsp.nodupe$Longitude.focal), max(cxsp.nodupe$Longitude.focal)))

ppp.cxsp.min.nodupe <- ppp(cxsp.nodupe$Latitude.focal, cxsp.nodupe$Longitude.focal, window = ppp.window_cxsp.min.nodupe)

cxsp.nodupe$infestedPA <- as.factor(ifelse(cxsp.nodupe$infestation_visit_1 > 0, "presence", "absence"))

ppp.PA.cxsp.min.nodupe <- ppp(cxsp.nodupe$Latitude.focal, cxsp.nodupe$Longitude.focal, window = ppp.window_cxsp.min.nodupe, marks = cxsp.nodupe$infestedPA)

Lmulti.cxsp.nodupe <- envelope(ppp.PA.cxsp.min.nodupe, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.cxsp.min.nodupe)))


plot(Lmulti.cxsp.nodupe,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF")

# Infested plotted over infested
Lmulti.cxsp.nodupe.inf <- envelope(ppp.PA.cxsp.min.nodupe, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.cxsp.min.nodupe)[rlabel(ppp.PA.cxsp.min.nodupe)$marks=='presence']))

plot(Lmulti.cxsp.nodupe.inf)


# Trying to compare Bivariate PCF of data 

# original ppp
cxsp$infestedPA <- as.character(cxsp$infestedPA)
# cxsp$infestedPA[cxsp$infestedPA == "presence"] <- "I"
# cxsp$infestedPA[cxsp$infestedPA == "absence"] <- "N"
cxsp$infestedPA <- as.factor(cxsp$infestedPA)
cxsp$infestedPA <- relevel(cxsp$infestedPA, "presence")
ppp.PA.cxsp.min <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window_cxsp.min, marks = cxsp$infestedPA)

Lmulti <- spatstat.explore::envelope(ppp.PA.cxsp.min, Lcross, nsim = 999, rank = 1, i = "absence", j = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.cxsp.min)))

plot(Lmulti,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

# new ppp
cxsp$marks_v1 <- as.factor(as.character(cxsp$marks_v1))
ppp.cxsp.min.i <- ppp(cxsp$Latitude.focal, cxsp$Longitude.focal, window = ppp.window_cxsp.min, marks = cxsp$marks_v1)

Lmulti.cxsp.v1 <- spatstat.explore::envelope(ppp.cxsp.min.i, Lcross, nsim = 999, rank = 1, i = "I", j = "N", global = FALSE, simulate = expression(rlabel(ppp.cxsp.min.i)))

plot(Lmulti,.-r~r,  shade = c("hi", "lo"), legend = FALSE)
plot(Lmulti.cxsp.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE)
```

# EMIG

## point pattern

```{r}
emig <- all_trees %>% dplyr::filter(all_trees$site == "EMIG")
# change names for combination with non-host
emig$Name <- as.character(emig$Name)

#ppp.window_emig <- owin(xrange = c(42.37375, 42.3743), yrange = c(-111.5538, -111.5531))

#ppp.emig <- ppp(emig$Latitude.focal, emig$Longitude.focal, window = ppp.window_emig)

# Minimum window
ppp.window_emig.min <- owin(xrange = c(min(emig$Latitude.focal), max(emig$Latitude.focal)), yrange = c(min(emig$Longitude.focal), max(emig$Longitude.focal)))

ppp.emig.min <- ppp(emig$Latitude.focal, emig$Longitude.focal, window = ppp.window_emig.min)

#plot(ppp.emig)
#summary(ppp.emig)

ggplot(emig, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```
# Marked Point Pattern

```{r}
emig$infestation_visit_1[is.na(emig$infestation_visit_1)] <- 0 # change NA's to 0
emig$infestedPA <- as.factor(ifelse(emig$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA.emig <- ppp(emig$Latitude.focal, emig$Longitude.focal, window = ppp.window_emig.min, marks = emig$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA.emig) # summary stats by mark
plot(split(ppp.PA.emig)) # plots of presence/absence

#infest_data.emig <- subset(emig, infested > 0)
#ppp.bug.emig <- ppp(infest_data.emig$Latitude.focal, infest_data.emig$Longitude.focal, window = ppp.window_emig)

#Lbug.emig <- envelope(ppp.bug.emig, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.emig, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# minimum window
ppp.PA.emig.min <- ppp(emig$Latitude.focal, emig$Longitude.focal, window = ppp.window_emig.min, marks = emig$infestedPA)

Lmulti.emig.min <- envelope(ppp.PA.emig.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.emig.min)))

plot(Lmulti.emig.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF")

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.emig <- envelope(ppp.PA.emig, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.emig)))

#plot(Lmulti.emig,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

```

# FRNK

## point pattern

```{r}

frnk <- all_trees %>% dplyr::filter(all_trees$site == "FRNK")
# change names for combination with non-host
frnk$Name <- as.character(frnk$Name)

#ppp.window_frnk <- owin(xrange = c(41.9334, 41.93386), yrange = c(-111.5718, -111.5711))

#ppp.frnk <- ppp(frnk$Latitude.focal, frnk$Longitude.focal, window = ppp.window_frnk)

# Testing minimun window (ie. different window boundaries)
ppp.window_frnk.min <- owin(xrange = c(min(frnk$Latitude.focal), max(frnk$Latitude.focal)), yrange = c(min(frnk$Longitude.focal), max(frnk$Longitude.focal)))

ppp.frnk.min <- ppp(frnk$Latitude.focal, frnk$Longitude.focal, window = ppp.window_frnk.min)

plot(ppp.frnk.min)
summary(ppp.frnk.min)

ggplot(frnk, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```
# Marked Point Pattern

```{r}
frnk$infestation_visit_1[is.na(frnk$infestation_visit_1)] <- 0
frnk$infestedPA <- as.factor(ifelse(frnk$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA.frnk <- ppp(frnk$Latitude.focal, frnk$Longitude.focal, window = ppp.window_frnk.min, marks = frnk$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA.frnk) # summary stats by mark
plot(split(ppp.PA.frnk)) # plots of presence/absence

#infest_data.frnk <- subset(frnk, infested > 0)
#ppp.bug.frnk <- ppp(infest_data.frnk$Latitude.focal, infest_data.frnk$Longitude.focal, window = ppp.window_frnk)

#Lbug.frnk <- envelope(ppp.bug.frnk, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.frnk, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# minimum window
ppp.PA.frnk.min <- ppp(frnk$Latitude.focal, frnk$Longitude.focal, window = ppp.window_frnk.min, marks = frnk$infestedPA)

Lmulti.frnk.min <- envelope(ppp.PA.frnk.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.frnk.min)))

plot(Lmulti.frnk.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF")

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.frnk <- envelope(ppp.PA.frnk, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.frnk)))

#plot(Lmulti.frnk,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

```

# MOCR

## point pattern

```{r}

mocr <- all_trees %>% dplyr::filter(all_trees$site == "MOCR")
# change names for combination with non-host
names(mocr)[1] <- "site"
names(mocr)[2] <- "Name"
names(mocr)[3] <- "Species"
names(mocr)[8] <- "infested"
mocr$Name <- as.character(mocr$Name)

#ppp.window_mocr <- owin(xrange = c(41.4090, 41.4095), yrange = c(-111.53415, -111.5335))

#ppp.mocr <- ppp(mocr$Latitude.focal, mocr$Longitude.focal, window = ppp.window_mocr)

# Testing minimun window (ie. different window boundaries)
ppp.window_mocr.min <- owin(xrange = c(min(mocr$Latitude.focal), max(mocr$Latitude.focal)), yrange = c(min(mocr$Longitude.focal), max(mocr$Longitude.focal)))

ppp.mocr.min <- ppp(mocr$Latitude.focal, mocr$Longitude.focal, window = ppp.window_mocr.min)

plot(ppp.mocr.min)
summary(ppp.mocr.min)

ggplot(mocr, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```
# Marked Point Pattern

```{r}
mocr$infestation_visit_1[is.na(mocr$infestation_visit_1)] <- 0
mocr$infestedPA <- as.factor(ifelse(mocr$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA.mocr <- ppp(mocr$Latitude.focal, mocr$Longitude.focal, window = ppp.window_mocr.min, marks = mocr$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA.mocr) # summary stats by mark
plot(split(ppp.PA.mocr)) # plots of presence/absence

plot(split(ppp.PA.mocr)[[2]])



#infest_data.mocr <- subset(mocr, infested > 0)
#ppp.bug.mocr <- ppp(infest_data.mocr$Latitude.focal, infest_data.mocr$Longitude.focal, window = ppp.window_mocr)

#Lbug.mocr <- envelope(ppp.bug.mocr, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.mocr, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# minimum window
ppp.PA.mocr.min <- ppp(mocr$Latitude.focal, mocr$Longitude.focal, window = ppp.window_mocr.min, marks = mocr$infestedPA)

Lmulti.mocr.min <- envelope(ppp.PA.mocr.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.mocr.min)))

plot(Lmulti.mocr.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF")

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.mocr <- envelope(ppp.PA.mocr, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.mocr)))

#plot(Lmulti.mocr,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

```

# MTTP

## point pattern

```{r}

mttp <- all_trees %>% dplyr::filter(all_trees$site == "MTTP")
# change names for combination with non-host
names(mttp)[1] <- "site"
names(mttp)[2] <- "Name"
names(mttp)[3] <- "Species"
names(mttp)[8] <- "infested"
mttp$Name <- as.character(mttp$Name)

#ppp.window_mttp <- owin(xrange = c(40.4347, 40.4352), yrange = c(-111.6393, -111.6386))

#ppp.mttp <- ppp(mttp$Latitude.focal, mttp$Longitude.focal, window = ppp.window_mttp)

# Testing minimun window (ie. different window boundaries)
ppp.window_mttp.min <- owin(xrange = c(min(mttp$Latitude.focal), max(mttp$Latitude.focal)), yrange = c(min(mttp$Longitude.focal), max(mttp$Longitude.focal)))

ppp.mttp.min <- ppp(mttp$Latitude.focal, mttp$Longitude.focal, window = ppp.window_mttp.min)

plot(ppp.mttp.min)
summary(ppp.mttp.min)

ggplot(mttp, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```
# Marked Point Pattern

```{r}
mttp$infestation_visit_1[is.na(mttp$infestation_visit_1)] <- 0
mttp$infestedPA <- as.factor(ifelse(mttp$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA.mttp <- ppp(mttp$Latitude.focal, mttp$Longitude.focal, window = ppp.window_mttp.min, marks = mttp$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA.mttp) # summary stats by mark
plot(split(ppp.PA.mttp)) # plots of presence/absence

#infest_data.mttp <- subset(mttp, infested > 0)
#ppp.bug.mttp <- ppp(infest_data.mttp$Latitude.focal, infest_data.mttp$Longitude.focal, window = ppp.window_mttp)

#Lbug.mttp <- envelope(ppp.bug.mttp, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.mttp, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# minimum window
ppp.PA.mttp.min <- ppp(mttp$Latitude.focal, mttp$Longitude.focal, window = ppp.window_mttp.min, marks = mttp$infestedPA)

Lmulti.mttp.min <- envelope(ppp.PA.mttp.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.mttp.min)))

plot(Lmulti.mttp.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF")

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.mttp <- envelope(ppp.PA.mttp, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.mttp)))

#plot(Lmulti.mttp,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

```

# SMMH

## point pattern

```{r}

smmh <- all_trees %>% dplyr::filter(all_trees$site == "SMMH")
# change names for combination with non-host
names(smmh)[1] <- "site"
names(smmh)[2] <- "Name"
names(smmh)[3] <- "Species"
names(smmh)[8] <- "infested"
smmh$Name <- as.character(smmh$Name)

#ppp.window_smmh <- owin(xrange = c(40.7705, 40.77105), yrange = c(-111.10755, -111.1069))

#ppp.smmh <- ppp(smmh$Latitude.focal, smmh$Longitude.focal, window = ppp.window_smmh)

# Testing minimun window (ie. different window boundaries)
ppp.window_smmh.min <- owin(xrange = c(min(smmh$Latitude.focal), max(smmh$Latitude.focal)), yrange = c(min(smmh$Longitude.focal), max(smmh$Longitude.focal)))

ppp.smmh.min <- ppp(smmh$Latitude.focal, smmh$Longitude.focal, window = ppp.window_smmh.min)

plot(ppp.smmh.min)
summary(ppp.smmh.min)

ggplot(smmh, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```
# Marked Point Pattern

```{r}
smmh$infestation_visit_1[is.na(smmh$infestation_visit_1)] <- 0
smmh$infestedPA <- as.factor(ifelse(smmh$infestation_visit_1 > 0, "presence", "absence"))
#ppp.PA.smmh <- ppp(smmh$Latitude.focal, smmh$Longitude.focal, window = ppp.window_smmh, marks = smmh$infestedPA) # create marked object using infestation presence/absence

ppp.PA.smmh.min <- ppp(smmh$Latitude.focal, smmh$Longitude.focal, window = ppp.window_smmh.min, marks = smmh$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA.smmh.min) # summary stats by mark
plot(split(ppp.PA.smmh.min)) # plots of presence/absence

#infest_data.smmh <- subset(smmh, infested > 0)
#ppp.bug.smmh <- ppp(infest_data.smmh$Latitude.focal, infest_data.smmh$Longitude.focal, window = ppp.window_smmh)

#Lbug.smmh <- envelope(ppp.bug.smmh, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.smmh, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.smmh <- envelope(ppp.PA.smmh, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.smmh)))

#plot(Lmulti.smmh,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

Lmulti.smmh.min <- envelope(ppp.PA.smmh.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.smmh.min)))

plot(Lmulti.smmh.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF")

```

# SOAP

## point pattern

```{r}

soap <- all_trees %>% dplyr::filter(all_trees$site == "SOAP")
# change names for combination with non-host
names(soap)[1] <- "site"
names(soap)[2] <- "Name"
names(soap)[3] <- "Species"
names(soap)[8] <- "infested"
soap$Name <- as.character(soap$Name)

#ppp.window_soap <- owin(xrange = c(40.5453, 40.5458), yrange = c(-111.03, -111.0293))

#ppp.soap <- ppp(soap$Latitude.focal, soap$Longitude.focal, window = ppp.window_soap)

# Testing minimun window (ie. different window boundaries)
ppp.window_soap.min <- owin(xrange = c(min(soap$Latitude.focal), max(soap$Latitude.focal)), yrange = c(min(soap$Longitude.focal), max(soap$Longitude.focal)))

ppp.soap.min <- ppp(soap$Latitude.focal, soap$Longitude.focal, window = ppp.window_soap.min)

plot(ppp.soap.min)
summary(ppp.soap.min)

ggplot(soap, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```
# Marked Point Pattern

```{r}
soap$infestation_visit_1[is.na(soap$infestation_visit_1)] <- 0
soap$infestedPA <- as.factor(ifelse(soap$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA.soap <- ppp(soap$Latitude.focal, soap$Longitude.focal, window = ppp.window_soap.min, marks = soap$infestedPA) # create marked object using infestation presence/absence

# Test minimun window
ppp.PA.soap.min <- ppp(soap$Latitude.focal, soap$Longitude.focal, window = ppp.window_soap.min, marks = soap$infestedPA)

split(ppp.PA.soap.min) # summary stats by mark
plot(split(ppp.PA.soap.min)) # plots of presence/absence

#infest_data.soap <- subset(soap, infested > 0)
#ppp.bug.soap <- ppp(infest_data.soap$Latitude.focal, infest_data.soap$Longitude.focal, window = ppp.window_soap)

#Lbug.soap <- envelope(ppp.bug.soap, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.soap, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.soap <- envelope(ppp.PA.soap, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.soap)))

Lmulti.soap.min <- envelope(ppp.PA.soap.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.soap.min)))


#par(mfrow = c(1, 2))
#plot(Lmulti.soap,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

# Looking for differences in window size
plot(Lmulti.soap.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF")


```

# TMPF

## point pattern

```{r}

tmpf <- all_trees %>% dplyr::filter(all_trees$site == "TMPF")
# change names for combination with non-host
names(tmpf)[1] <- "site"
names(tmpf)[2] <- "Name"
names(tmpf)[3] <- "Species"
names(tmpf)[8] <- "infested"
tmpf$Name <- as.character(tmpf$Name)

ppp.window_tmpf <- owin(xrange = c(40.4347, 40.4352), yrange = c(-111.6393, -111.6386))

ppp.tmpf <- ppp(tmpf$Latitude.focal, tmpf$Longitude.focal, window = ppp.window_tmpf)

ppp.window_tmpf.min <- owin(xrange = c(min(tmpf$Latitude.focal), max(tmpf$Latitude.focal)), yrange = c(min(tmpf$Longitude.focal), max(tmpf$Longitude.focal)))

ppp.tmpf.min <- ppp(tmpf$Latitude.focal, tmpf$Longitude.focal, window = ppp.window_tmpf.min)

plot(ppp.tmpf)
summary(ppp.tmpf)

ggplot(tmpf, aes(x = Latitude.focal, y = Longitude.focal)) + 
  geom_point() + coord_fixed()
```

# Marked Point Pattern

```{r}
tmpf$infestation_visit_1[is.na(tmpf$infestation_visit_1)] <- 0
tmpf$infestedPA <- as.factor(ifelse(tmpf$infestation_visit_1 > 0, "presence", "absence"))
ppp.PA.tmpf <- ppp(tmpf$Latitude.focal, tmpf$Longitude.focal, window = ppp.window_tmpf.min, marks = tmpf$infestedPA) # create marked object using infestation presence/absence

# Test minimun window
ppp.PA.tmpf.min <- ppp(tmpf$Latitude.focal, tmpf$Longitude.focal, window = ppp.window_tmpf.min, marks = tmpf$infestedPA)

split(ppp.PA.tmpf.min) # summary stats by mark
plot(split(ppp.PA.tmpf.min)) # plots of presence/absence

#infest_data.tmpf <- subset(tmpf, infested > 0)
#ppp.bug.tmpf <- ppp(infest_data.tmpf$Latitude.focal, infest_data.tmpf$Longitude.focal, window = ppp.window_tmpf)

#Lbug.tmpf <- envelope(ppp.bug.tmpf, Lest, nsim = 999, rank = 1, i = "presence", global = F)

#plot(Lbug.tmpf, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
#Lmulti.tmpf <- envelope(ppp.PA.tmpf, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.tmpf)))

Lmulti.tmpf.min <- envelope(ppp.PA.tmpf.min, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.tmpf.min)))


#par(mfrow = c(1, 2))
#plot(Lmulti.tmpf,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

# Looking for differences in window size
plot(Lmulti.tmpf.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF")


```

# Plot all presence plots

```{r}

par(mfrow = c(3, 3),
  pin = c(1, 1),
  mar = c(0.25, 0.25, 0.25, 0.25),
  oma = c(0, 0, 0.5, 0))
plot(split(ppp.PA.mocr.min)[[2]]) 
plot(split(ppp.PA.emig.min)[[2]]) 
plot(split(ppp.PA.mttp.min)[[2]]) 
plot(split(ppp.PA.frnk.min)[[2]]) 
plot(split(ppp.PA.smmh.min)[[2]]) 
plot(split(ppp.PA.soap.min)[[2]]) 
plot(split(ppp.PA.cxsp.min)[[2]])


```

# Marked Point Pattern

```{r}
tmpf$infestedPA <- as.factor(ifelse(tmpf$infested > 0, "presence", "absence"))
ppp.PA.tmpf <- ppp(tmpf$Latitude.focal, tmpf$Longitude.focal, window = ppp.window_tmpf, marks = tmpf$infestedPA) # create marked object using infestation presence/absence

split(ppp.PA.tmpf) # summary stats by mark
plot(split(ppp.PA.tmpf)) # plots of presence/absence

infest_data.tmpf <- subset(tmpf, infested > 0)
ppp.bug.tmpf <- ppp(infest_data.tmpf$Latitude.focal, infest_data.tmpf$Longitude.focal, window = ppp.window_tmpf)

Lbug.tmpf <- envelope(ppp.bug.tmpf, Lest, nsim = 999, rank = 1, i = "presence", global = F)

plot(Lbug.tmpf, .-r~r, shade = c("hi", "lo"), legend = FALSE)

# bivariate K function (bivariate pair correlation function) - accounting for host distribution
Lmulti.tmpf <- envelope(ppp.PA.tmpf, Lcross, nsim = 999, rank = 1, i = "presence", global = FALSE, simulate = expression(rlabel(ppp.PA.tmpf)))

plot(Lmulti.tmpf,.-r~r,  shade = c("hi", "lo"), legend = FALSE)

```

# 

# Continuous Marked point pattern

EMIG DBH

We'll do DBH for this example - we may not need this but it'll be nice to see

```{r}

ppp.dbh.emig <- ppp(emig$Latitude.focal, emig$Longitude.focal, window = ppp.window_emig.min, marks = !is.na(emig$DBH_Visit1)) # need to exclude NA's for this, if we want to unclude clumps, we need to give a value

mcf.dbh <- markcorr(ppp.dbh.emig) # mark correlation analysis - Values <1 and outside the envelope indicate negative relationships of patch area with patches r distance away, while values >1 indicate positive relationships

mcfEnv <- envelope(ppp.dbh.emig, markcorr, nsim = 999, correction = "iso", global = FALSE)

plot(mcfEnv, shade = c("hi", "lo"), legend = F)

```

# Alternative null models

The following null models will be Matern cluster process and Thomas cluster process. These processes use "parent" points and "offspring" points with differing distributions. This can simulate seed dispersal, but is this case, we'll look at it like propagule pressure coming from the host tree (or something like that)

- The following plot basically says that we do see an aggregated pattern as expected by one of the previous processes.

```{r}

Kthomas <- kppm(ppp.cxsp.min, ~1, "Thomas")
summary(Kthomas)

Kthomas.env <- envelope(Kthomas, Lest, nsim = 999, rank = 1, global = F)
plot(Kthomas.env, .-r~r, shade = c("hi", "lo"), legend = F)

```

# Spatial dependence and autocorrelation

## Correlogram

We're gonna run through this book's version of this and see if it differs from what we did in an earlier RMarkdown

```{r}

#calculate a distance matrix > 
coords <- cbind(cxsp$Latitude.focal, cxsp$Longitude.focal)  
colnames(coords) <- c("x", "y") 
distmat <- as.matrix(dist(coords)) 

#maximum distance to consider in correlogram/variogram
maxdist <- 2/3 * max(distmat)

# correlog from pgirmess
library(pgirmess)
corr <- correlog(coords, cxsp$infested, method = "Moran", nbclass = 14, alternative = "two.sided")

# correlogram plot - don't account for distances past our max dist calculation, which is 42.97
plot(corr) +
abline(h=0)

# same thing with ncf package and Monte Carlo sims
library(ncf)
corr_ncf <- ncf::correlog(x = cxsp$Latitude.focal, y = cxsp$Longitude.focal, z = cxsp$infested, increment = 5, resamp = 999)

plot(corr_ncf) +
  abline(h=0)

# Spline correlogram with 95% CI bootstrapping
spline.corr <- spline.correlog(x = cxsp$Latitude.focal, y = cxsp$Longitude.focal, z = cxsp$infested, xmax = maxdist, resamp = 100, type = "boot")

plot(spline.corr)

# spdep package - this is how we can do binary response or infestation response

# make neighborhood list using queen's rule - 8 neighbors
neigh <- dnearneigh(x = coords, d1 = 0, d2 = 3, longlat = F)

plot(neigh, coordinates(coords))

# make spatial weights list with row-standardized weights('W')
wts <- nb2listw(neighbours = neigh, style = 'W', zero.policy = T)

# Monte carlo permutations
mor.mc <- moran.mc(x = cxsp$infested, listw = wts, nsim = 999, zero.policy = T)
#moran test
mor.norm <- moran.test(x = cxsp$infested, listw = wts, randomisation = F, zero.policy = T)

mor.mc
mor.norm

# dataframe for storage
corr.sp <- data.frame(dist = seq(5, 0.5 * max(distmat), by = 5), MoransI = NA, Null.LCL = NA, Null.UCL = NA, Pvalue = NA)

# Calculate MOran's I fofr lag distances
for(i in 1:nrow(corr.sp)) {
  d.start <- corr.sp[i, "dist"] - 5
  d.end <- corr.sp[i, "dist"]
  neigh <- dnearneigh(x = coords, d1 = d.start, d2 = d.end, longlat = F)
  wts <- nb2listw(neighbours = neigh, style = 'W', zero.policy = T)
  mor.i <- moran.mc(x = cxsp$infested, listw = wts, nsim = 999, zero.policy = T)
  
  # summarize results
  corr.sp[i, "dist"] <- (d.end + d.start)/2 
  corr.sp[i, "MoransI"] <- mor.i$statistic 
  corr.sp[i, "Null.LCL"] <- quantile(mor.i$res, p = 0.025) 
  corr.sp[i, "Null.UCL"] <- quantile(mor.i$res, p = 0.975) 
  corr.sp[i, "Pvalue"] <- mor.i$p.value
}
plot(y = corr.sp$MoransI, x = corr.sp$dist) +
  abline(h=0) +
  lines(corr.sp$dist, corr.sp$Null.LCL, col = "red") +
  lines(corr.sp$dist, corr.sp$Null.UCL, col = "red")
```

## Semivariograms

```{r}

library(geoR)
library(gstat)

# need to feed site data with no duplicated coordinates - looking at DBH to have a continuous data col
geo.inf <- as.geodata(cxsp.nodupe, coords.col = 68:69, data.col = 8)
plot(geo.inf)

#Empirical semivariogram > 
emp.geoR <- variog(geo.inf, max.dist = maxdist)
plot(emp.geoR) 

#standardize break points to a minimum 3-m lag distance
emp.geoR <- variog(geo.inf, max.dist = maxdist, breaks = c(seq(0, maxdist, by = 3))) 
plot(emp.geoR)

#in geoR, variogram in each direction
emp4.geoR <- variog4(geo.inf, max.dist = maxdist) 
plot(emp4.geoR)
legend(4,2)

# Fitting model variograms to the data
## Exponential
mlexp <- likfit(geo.inf, cov.model = "exp", ini = c(700, 10))

## Spherical variogram
mlsph <- likfit(geo.inf, cov.model = "sph", ini = c(700, 10))
summary(mlexp)

AIC(mlexp, mlsph)

emp.env <- variog.mc.env(geo.inf, obj.var = emp.geoR) 
plot(emp.geoR, envelope = emp.env) 
lines(mlexp, col = "blue")

```

# Host Tree Distributions

Let's look at distributions for all sites to see if they are CSR or not

```{r}
Lcsr.cxsp <- envelope(ppp.cxsp.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F, main = "Cox Spring Host Distribution")


Lcsr.emig <- envelope(ppp.emig.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.emig, .-r~r, shade = c("hi", "lo"), legend = F, main = "Emigration Canyon Host Distribution")

Lcsr.frnk <- envelope(ppp.frnk.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.frnk, .-r~r, shade = c("hi", "lo"), legend = F, main = "Franklin Basin Host Distribution")

Lcsr.mocr <- envelope(ppp.mocr.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.mocr, .-r~r, shade = c("hi", "lo"), legend = F, main = "Monte Cristo Host Distribution")

Lcsr.mttp <- envelope(ppp.mttp.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.mttp, .-r~r, shade = c("hi", "lo"), legend = F, main = "Mount Timpanogos Host Distribution")

Lcsr.smmh <- envelope(ppp.smmh.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.smmh, .-r~r, shade = c("hi", "lo"), legend = F, main = "Smith and Morehouse Host Distribution")

Lcsr.soap <- envelope(ppp.soap.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.soap, .-r~r, shade = c("hi", "lo"), legend = F, main = "Soapstone Basin Host Distribution")

Lcsr.tmpf <- envelope(ppp.tmpf.min, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr.tmpf, .-r~r, shade = c("hi", "lo"), legend = F, main = "Temple Fork Host Distribution")

```