---
title: "Semivariance / Autocorrelation"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Semivariance and autocorrelation Indices
11/15/23
Michael Wayman



# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

## Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
```

# Cox Spring Semivariance

Getting semivariance allows us to select an appropriate spatial model for spatial interpolation by understanding how correlated our points are spatially.

```{r}
variogram_lowercase_fx <- function(x){
host <- x %>% filter(Species == "abla" | Species == "abco")
live_host <- host %>% filter(`Dead / Alive` == 1)

## we need to run this code again to get binary infestation again
live_host$`Inf Location`[live_host$`Inf Location`== "-"] <- 0
unique(live_host$`Inf Location`)
live_host$`Inf Location`[live_host$`Inf Location` == "BO" | live_host$`Inf Location` == "BR, BO"] <- 1
live_host$`Inf Location` <- as.numeric(live_host$`Inf Location`)

Lats <- live_host$Latitude # create latitude vector
Longs <- live_host$Longitude # create longitude vector

coordinates <- live_host[, c("Latitude", "Longitude")] # create coordinate dataframe
infestation <- live_host$`Inf Location`

# Create a SpatialPointsDataFrame
tree_sp <- SpatialPointsDataFrame(coords = coordinates, data = data.frame(infestation), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_model <- variogram(infestation ~ 1, data = tree_sp)

# Plot the variogram
plot(variogram_model)

variogram_model_fit <- fit.variogram(variogram_model, model = vgm(1, "Exp"))

plot<-plot(variogram_model, model = variogram_model_fit)
summary(variogram_model)

# Extract the range, sill, and nugget
range_value <- variogram_model_fit$range
sill_value <- variogram_model_fit$sill
nugget_value <- variogram_model_fit$nugget

# Print the values
cat("Range:", range_value, "\n")

# Calculate nugget and sill from the partial sill
nugget_value <- variogram_model_fit$psill - variogram_model_fit$range
sill_value <- variogram_model_fit$psill

# Print the values
cat("Nugget:", nugget_value, "\n")
cat("Sill:", sill_value, "\n")
return(plot)
}

variogram_lowercase_fx(cxsp_combined)
```

# Emigration Canyon Semivariance

```{r}
host <- emig_combined %>% filter(Species == "abla" | Species == "abco")
live_host <- host %>% filter(`Dead / Alive` == 1)

## we need to run this code again to get binary infestation again
live_host$`Inf Location`[live_host$`Inf Location`== "-"] <- 0
live_host$`Inf Location`[is.na(live_host$`Inf Location`)] <- 0
unique(live_host$`Inf Location`)
live_host$`Inf Location`[live_host$`Inf Location` == "BO" | live_host$`Inf Location` == "BR, BO" | live_host$`Inf Location` == "BR"] <- 1
live_host$`Inf Location` <- as.numeric(live_host$`Inf Location`)

Lats <- live_host$Latitude # create latitude vector
Longs <- live_host$Longitude # create longitude vector

coordinates <- live_host[, c("Latitude", "Longitude")] # create coordinate dataframe
infestation <- live_host$`Inf Location`

# Create a SpatialPointsDataFrame
tree_sp <- SpatialPointsDataFrame(coords = coordinates, data = data.frame(infestation), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_model <- variogram(infestation ~ 1, data = tree_sp)

# Plot the variogram
plot(variogram_model)

variogram_model_fit <- fit.variogram(variogram_model, model = vgm(1, "Exp"))

plot(variogram_model, model = variogram_model_fit)
summary(variogram_model)
summary(variogram_model_fit)

# Extract the range, sill, and nugget
range_value <- variogram_model_fit$range
sill_value <- variogram_model_fit$sill
nugget_value <- variogram_model_fit$nugget

# Print the values
cat("Range:", range_value, "\n")

# Calculate nugget and sill from the partial sill
nugget_value <- variogram_model_fit$psill - variogram_model_fit$range
sill_value <- variogram_model_fit$psill

# Print the values
cat("Nugget:", nugget_value, "\n")
cat("Sill:", sill_value, "\n")
```

# Franklin Basin Semivariance

```{r}
variogram_uppercase_fx <- function(x){
host <- x %>% filter(Species == "ABLA" | Species == "ABCO")
live_host <- host %>% filter(`Dead / Alive` == 1)

## we need to run this code again to get binary infestation again
live_host$`Inf Location`[live_host$`Inf Location`== "-"] <- 0
live_host$`Inf Location`[is.na(live_host$`Inf Location`)] <- 0
unique(live_host$`Inf Location`)
live_host$`Inf Location`[live_host$`Inf Location` == "BO" | live_host$`Inf Location` == "BR, BO" | live_host$`Inf Location` == "BR"] <- 1
live_host$`Inf Location` <- as.numeric(live_host$`Inf Location`)

Lats <- live_host$Latitude # create latitude vector
Longs <- live_host$Longitude # create longitude vector

coordinates <- live_host[, c("Latitude", "Longitude")] # create coordinate dataframe
infestation <- live_host$`Inf Location`

# Create a SpatialPointsDataFrame
tree_sp <- SpatialPointsDataFrame(coords = coordinates, data = data.frame(infestation), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_model <- variogram(infestation ~ 1, data = tree_sp)

# Plot the variogram
plot(variogram_model)

variogram_model_fit <- fit.variogram(variogram_model, model = vgm(1, "Exp"))

plot<-plot(variogram_model, model = variogram_model_fit)
summary(variogram_model)
summary(variogram_model_fit)

# Extract the range, sill, and nugget
range_value <- variogram_model_fit$range
sill_value <- variogram_model_fit$sill
nugget_value <- variogram_model_fit$nugget

# Print the values
cat("Range:", range_value, "\n")

# Calculate nugget and sill from the partial sill
nugget_value <- variogram_model_fit$psill - variogram_model_fit$range
sill_value <- variogram_model_fit$psill

# Print the values
cat("Nugget:", nugget_value, "\n")
cat("Sill:", sill_value, "\n")
return(plot)
}
```

# Monte Cristo Semivariance

```{r}
variogram_uppercase_fx(mocr_combined)
```

# Mount Timpanogos Semivariance

```{r}
variogram_uppercase_fx(mttp_combined)
```

# Smith and Morehouse Semivariance ( This on is coming out weird - wrong model? )

```{r}
host <- smmh_combined %>% filter(Species == "ABLA" | Species == "ABCO")
live_host <- host %>% filter(`Dead / Alive` == 1)

## we need to run this code again to get binary infestation again
live_host$`Inf Location`[live_host$`Inf Location`== "-"] <- 0
live_host$`Inf Location`[is.na(live_host$`Inf Location`)] <- 0
unique(live_host$`Inf Location`)
live_host$`Inf Location`[live_host$`Inf Location` == "BO" | live_host$`Inf Location` == "BR, BO" | live_host$`Inf Location` == "BR"] <- 1
live_host$`Inf Location` <- as.numeric(live_host$`Inf Location`)

Lats <- live_host$Latitude # create latitude vector
Longs <- live_host$Longitude # create longitude vector

coordinates <- live_host[, c("Latitude", "Longitude")] # create coordinate dataframe
infestation <- live_host$`Inf Location`

# Create a SpatialPointsDataFrame
tree_sp <- SpatialPointsDataFrame(coords = coordinates, data = data.frame(infestation), proj4string = CRS("+proj=longlat +datum=WGS84"))

# Calculate the variogram
variogram_model <- variogram(infestation ~ 1, data = tree_sp)

# Plot the variogram
plot(variogram_model)

variogram_model_fit <- fit.variogram(variogram_model, model = vgm(psill = 0.1486, "Exp", range = 0.003897))

plot(variogram_model, model = variogram_model_fit)
summary(variogram_model)
summary(variogram_model_fit)

# Extract the range, sill, and nugget
range_value <- variogram_model_fit$range
sill_value <- variogram_model_fit$sill
nugget_value <- variogram_model_fit$nugget

# Print the values
cat("Range:", range_value, "\n")

# Calculate nugget and sill from the partial sill
nugget_value <- variogram_model_fit$psill - variogram_model_fit$range
sill_value <- variogram_model_fit$psill

# Print the values
cat("Nugget:", nugget_value, "\n")
cat("Sill:", sill_value, "\n")
```

# 

