---
title: "Point Pattern Analysis 2"
author: "Michael Wayman"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# turn off annoying messages with this option
knitr::opts_chunk$set(warning = T, error = T, message = F)
```

# Description

Point Processes another time
3/25/24
Michael Wayman

The purpose of this script is to do point pattern analysis a slightly different way. 1) Get all of our points into the same window by using plot coordinates instead of real world coordinates (eg. 0 through 50 instead of UTM or degrees) 2) compare not only infested to host distribution, but infested to non-infested so we can compare various states.

# Set up

## Directory

```{r wd}
getwd()
# "C:/Users/bwap/Documents/Project Files/Analysis/Cleaning and Compilation/Cleaning"
```

# clean out environment

```{r}
rm(list = ls())

```

# Packages

```{r packages, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(tinytex)
library(fastmap)
library(pandoc)
library(sf)
library(spatstat)
library(sp)
library(gstat)
library(spdep)
library(geosphere)
library(nngeo)
library(vegan)
library(clipr)
library(reshape2)
library(Matrix) # to use glmer
library(lme4) # to use glmer

```

# Load data

```{r, warning=FALSE, echo=FALSE}

trees <- read_xlsx("data_with_nearest_info.xlsx")

# all_trees <- read_xlsx("all_sites_with_points.xlsx") # This is just visit 4

all_trees <- read_xlsx("all_merged_all_spp.xlsx")

```

# Coordinate conversion

We need to separate by site and then convert to plot coordinates. Then we can plot everything in the same window easily.

```{r}

# GEt infestation variable again...
all_trees[all_trees == "-"] <- "0" # change all "-" to a zero

# all_trees <- all_trees %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1","NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2","NL_v3", "NM_v3", "NH_v3", "SL_v3", "SM_v3", "SH_v3","NL_v4", "NM_v4", "NH_v4", "SL_v4", "SM_v4", "SH_v4" )), 
#             ~ if_else(startsWith(as.character(.), "1-"), "1", as.character(.))) # Change all "1-" ratings to just a singular 1 value
# 
# all_trees <- all_trees %>%
#   mutate_at(vars(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1","NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2","NL_v3", "NM_v3", "NH_v3", "SL_v3", "SM_v3", "SH_v3","NL_v4", "NM_v4", "NH_v4", "SL_v4", "SM_v4", "SH_v4" )), as.numeric)

# all_trees <- all_trees %>%
#    mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH" )), 
#              ~ if_else(startsWith(as.character(.), "1-"), "1", as.character(.))) # Change all "1-" ratings to just a singular 1 value
#  
#  all_trees <- all_trees %>%
#    mutate_at(vars(c("NL", "NM", "NH", "SL", "SM", "SH" )), as.numeric)

# all_trees$NL[is.na(all_trees$NL)] <- 0
# all_trees$NM[is.na(all_trees$NM)] <- 0
# all_trees$NH[is.na(all_trees$NH)] <- 0
# all_trees$SL[is.na(all_trees$SL)] <- 0
# all_trees$SM[is.na(all_trees$SM)] <- 0
# all_trees$SH[is.na(all_trees$SH)] <- 0

 #all_trees <- all_trees %>%
   #mutate(infested = ifelse(NL + NM + NH + SL + SM + SH != 0, 1, 0))
# 
# all_trees <- all_trees[!is.na(all_trees$Longitude),] # get rid of NAs


# convert to UTM to get nice "square" plots
xy = data.frame(all_trees$Longitude, all_trees$Latitude)
xy <- xy[!is.na(xy$all_trees.Longitude),] # Get rid of NAs
coordinates(xy) <- c("all_trees.Longitude", "all_trees.Latitude")
proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")
NE <- spTransform(xy, CRS("+proj=utm +zone=12 ellps=WGS84"))
NE <- as.data.frame(NE)

all_trees$Longitude <- NE$coords.x1
all_trees$Latitude <- NE$coords.x2

#Marks for status - we'll do marks for non-hosts and then hosts with and w/o infestation
# all_trees$infested = as.factor(all_trees$infested)
all_trees <- all_trees %>% mutate(host = ifelse(species == "ABLA" | species == "ABCO", 1, 0))
all_trees$host <- as.factor(all_trees$host)
# take care of NA's
all_trees$infested_v1[is.na(all_trees$infested_v1)] <- 0
all_trees$infested_v2[is.na(all_trees$infested_v2)] <- 0
all_trees$infested_v3[is.na(all_trees$infested_v3)] <- 0
all_trees$infested_v4[is.na(all_trees$infested_v4)] <- 0

all_trees <- all_trees %>%
  mutate(marks_v1 = case_when(
    host == 0 ~ factor("NH"),
    infested_v1 == 1 & host == 1 ~ factor("I"),
    infested_v1 == 0 & host == 1 ~ factor("N")),
    marks_v2 = case_when(
    host == 0 ~ factor("NH"),
    infested_v2 == 1 & host == 1 ~ factor("I"),
    infested_v2 == 0 & host == 1 ~ factor("N")),
    marks_v3 = case_when(
    host == 0 ~ factor("NH"),
    infested_v3 == 1 & host == 1 ~ factor("I"),
    infested_v3 == 0 & host == 1 ~ factor("N")),
    marks_v4 = case_when(
    host == 0 ~ factor("NH"),
    infested_v4 == 1 & host == 1 ~ factor("I"),
    infested_v4 == 0 & host == 1 ~ factor("N")),
    marks_new_v2 = case_when(
      host == 0 ~ factor("NH"),
      (infested_v1 == 0 & infested_v2) == 1 ~ factor("I"),
      (infested_v1 == 0 & infested_v2) == 0 ~ factor("N"),
      (infested_v1 == 1 & infested_v2) == 1 ~ factor("O"),
      (infested_v1 == 1 & infested_v2) == 0 ~ factor("O")
    ),
    marks_new_v3 = case_when(
      host == 0 ~ factor("NH"),
      (infested_v2 == 0 & infested_v3 == 1) & host == 1 ~ factor("I"),
      (infested_v2 == 0 & infested_v3 == 0) & host == 1 ~ factor("N"),
      (infested_v2 == 1 & infested_v3 == 1) & host == 1 ~ factor("O"),
      (infested_v2 == 1 & infested_v3 == 0) & host == 1 ~ factor("O")
    ),
     marks_new_v4 = case_when(
      host == 0 ~ factor("NH"),
      (infested_v3 == 0 & infested_v4 == 1) & host == 1 ~ factor("I"),
      (infested_v3 == 0 & infested_v4 == 0) & host == 1 ~ factor("N"),
      (infested_v3 == 1 & infested_v4 == 1) & host == 1 ~ factor("O"),
      (infested_v3 == 1 & infested_v4 == 0) & host == 1 ~ factor("O")))

#writexl::write_xlsx(all_trees, "trees_for_spatial_tests.xlsx")

#names(all_trees)[6] <- "site"
cxsp <- all_trees %>% dplyr::filter(site == "CXSP")
emig <- all_trees %>% dplyr::filter(site == "EMIG")
frnk <- all_trees %>% dplyr::filter(site == "FRNK")
mocr <- all_trees %>% dplyr::filter(site == "MOCR")
mttp <- all_trees %>% dplyr::filter(site == "MTTP")
smmh <- all_trees %>% dplyr::filter(site == "SMMH")
soap <- all_trees %>% dplyr::filter(site == "SOAP")
tmpf <- all_trees %>% dplyr::filter(site == "TMPF")

# Function to scale GPS coordinates
scale_coordinates <- function(coord, min_val, max_val) {
  scaled_coord <- (coord - min(coord)) / (max(coord) - min(coord)) * (max_val - min_val) + min_val
  return(scaled_coord)
}

# Scale longitude and latitude to plot coordinates (0 to 50)
cxsp$plot_x <- scale_coordinates(cxsp$Longitude, 0, 50)
cxsp$plot_y <- scale_coordinates(cxsp$Latitude, 0, 50)

# Plot the scaled coordinates to check
plot(cxsp$plot_x, cxsp$plot_y, type = 'p', pch = 16, col = 'blue', xlab = 'X', ylab = 'Y', main = 'GPS Data Scaled to Plot Coordinates')

# repeat for all plots
emig$plot_x <- scale_coordinates(emig$Longitude, 0, 50)
emig$plot_y <- scale_coordinates(emig$Latitude, 0, 50)

frnk$plot_x <- scale_coordinates(frnk$Longitude, 0, 50)
frnk$plot_y <- scale_coordinates(frnk$Latitude, 0, 50)

mocr$plot_x <- scale_coordinates(mocr$Longitude, 0, 50)
mocr$plot_y <- scale_coordinates(mocr$Latitude, 0, 50)

mttp$plot_x <- scale_coordinates(mttp$Longitude, 0, 50)
mttp$plot_y <- scale_coordinates(mttp$Latitude, 0, 50)

smmh$plot_x <- scale_coordinates(smmh$Longitude, 0, 50)
smmh$plot_y <- scale_coordinates(smmh$Latitude, 0, 50)

soap$plot_x <- scale_coordinates(soap$Longitude, 0, 50)
soap$plot_y <- scale_coordinates(soap$Latitude, 0, 50)

tmpf$plot_x <- scale_coordinates(tmpf$Longitude, 0, 50)
tmpf$plot_y <- scale_coordinates(tmpf$Latitude, 0, 50)
```

# Window and PPP objects

```{r}

#Window
window = owin(c(0,50),
              c(0,50))

# Loop all sites to get ppp and pcf objects

sites <- c("cxsp", "emig", "frnk", "mocr", "mttp", "smmh", "soap", "tmpf")

# this function takes a site, creates a ppp and pcf object, spits them out and plots the site's pcf
ppp_pcf <- function(i) {
obj_name <- deparse(substitute(i))
i.ppp <- ppp(i$plot_x, i$plot_y, window = window, marks = i$marks_v1)
i.pcf <- pcfinhom(i.ppp, r = 0:20, nsim =3)
assign(paste0(obj_name, ".ppp"), i.ppp, envir = .GlobalEnv)
assign(paste0(obj_name, ".pcf"), i.pcf, envir = .GlobalEnv)
plot(i.pcf)
}

ppp_pcf(cxsp)
ppp_pcf(emig)
ppp_pcf(frnk)
ppp_pcf(mocr)
ppp_pcf(mttp)
ppp_pcf(smmh)
ppp_pcf(soap)
ppp_pcf(tmpf)

```

# Envelopes

```{r}
cxsp.pcf.envelope = envelope(cxsp.ppp, fun = pcfinhom, r=seq(0,50,.5),nsim=99,simulate=expression(runifpoint(nrow(cxsp),win=window)))
plot(cxsp.pcf.envelope)

```

# Tests

## Cox Springs

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
cxsp_inf_all <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='I']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(cxsp_inf_all)


# Infested - Infested
cxsp_inf_inf <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(cxsp_inf_inf)

# Infested - Non-infested
cxsp_inf_non_1 <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(cxsp_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "CXSP - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))

legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)

# Infested - All host
cxsp_inf_host_1 <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(cxsp_inf_host_1)

# Non-infested - All
cxsp_noninf_all <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='I']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(cxsp_noninf_all)

# Hosts - All
cxsp_host_non <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I' | cxsp.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='I']),
                                          (rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
png("cxsp_inf_non_1.png", width = 800, height = 600)
plot(cxsp_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "CXSP - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 2 
cxsp.ppp <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v2)

cxsp_inf_non_2 <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("cxsp_inf_non_2.png", width = 800, height = 600)
plot(cxsp_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "CXSP - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
cxsp.ppp <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v3)

cxsp_inf_non_3 <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("cxsp_inf_non_3.png", width = 800, height = 600)
plot(cxsp_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "CXSP - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
cxsp.ppp <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v4)

cxsp_inf_non_4 <- envelope(Y=cxsp.ppp[cxsp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp)[rlabel(cxsp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("cxsp_inf_non_4.png", width = 800, height = 600)
plot(cxsp_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "CXSP - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(cxsp.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)


```

## Emigration Canyon

```{r, warning=FALSE}
#Tests - Time 1
#Infested - All
emig_inf_all <- envelope(Y=emig.ppp[emig.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N']),
                                          (rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='I']),
                                          (rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(emig_inf_all)

# Infested - Non-infested
emig_inf_non_1 <- envelope(Y=emig.ppp[emig.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("emig_inf_non_1.png", width = 800, height = 600)
plot(emig_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "EMIG - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Infested - All host
emig_inf_host_1 <- envelope(Y=emig.ppp[emig.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N']),
                                          (rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(emig_inf_host_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(3, 1, 2, 2),
  main = "EMIG - Infested vs Host Trees",
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))

# Non-infested - All
emig_noninf_all <- envelope(Y=emig.ppp[emig.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N']),
                                          (rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='I']),
                                          (rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(emig_noninf_all)

# Hosts - Non-hosts
emig_host_non <- envelope(Y=emig.ppp[emig.ppp$marks=='I' | emig.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(emig_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
emig.ppp <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v2)

emig_inf_non_2 <- envelope(Y=emig.ppp[emig.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("emig_inf_non_2.png", width = 800, height = 600)
plot(emig_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "EMIG - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
emig.ppp <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v3)

emig_inf_non_3 <- envelope(Y=emig.ppp[emig.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("emig_inf_non_3.png", width = 800, height = 600)
plot(emig_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "EMIG - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
emig.ppp <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v4)

emig_inf_non_4 <- envelope(Y=emig.ppp[emig.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp)[rlabel(emig.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("emig_inf_non_4.png", width = 800, height = 600)
plot(emig_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "EMIG - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr.emig <- envelope(emig.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)


```

## Smith and Morehouse

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
smmh_inf_all <- envelope(Y=smmh.ppp[smmh.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='N']),
                                          (rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(smmh_inf_all)

# Infested - Non-infested
smmh_inf_non_1 <- envelope(Y=smmh.ppp[smmh.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("smmh_inf_non_1.png", width = 800, height = 600)
plot(smmh_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SMMH - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Non-infested - All
smmh_noninf_all <- envelope(Y=smmh.ppp[smmh.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='N']),
                                          (rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(smmh_noninf_all)

# Hosts - Non-hosts
smmh_host_non <- envelope(Y=smmh.ppp[smmh.ppp$marks=='I' | smmh.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(smmh_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
smmh.ppp <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v2)

smmh_inf_non_2 <- envelope(Y=smmh.ppp[smmh.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("smmh_inf_non_2.png", width = 800, height = 600)
plot(smmh_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SMMH - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
smmh.ppp <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v3)

smmh_inf_non_3 <- envelope(Y=smmh.ppp[smmh.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("smmh_inf_non_3.png", width = 800, height = 600)
plot(smmh_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SMMH - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
smmh.ppp <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v4)

smmh_inf_non_4 <- envelope(Y=smmh.ppp[smmh.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp)[rlabel(smmh.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("smmh_inf_non_4.png", width = 800, height = 600)
plot(smmh_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SMMH - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(smmh.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)


```

## Franklin Basin

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
frnk_inf_all <- envelope(Y=frnk.ppp[frnk.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='N']),
                                          (rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(frnk_inf_all)

# Infested - Non-infested
frnk_inf_non_1 <- envelope(Y=frnk.ppp[frnk.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("frnk_inf_non_1.png", width = 800, height = 600)
plot(frnk_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "FRNK - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Non-infested - All
frnk_noninf_all <- envelope(Y=frnk.ppp[frnk.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='N']),
                                          (rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(frnk_noninf_all)

# Hosts - Non-hosts
frnk_host_non <- envelope(Y=frnk.ppp[frnk.ppp$marks=='I' | frnk.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(frnk_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
frnk.ppp <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v2)

frnk_inf_non_2 <- envelope(Y=frnk.ppp[frnk.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("frnk_inf_non_2.png", width = 800, height = 600)
plot(frnk_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "FRNK - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
frnk.ppp <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v3)

frnk_inf_non_3 <- envelope(Y=frnk.ppp[frnk.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("frnk_inf_non_3.png", width = 800, height = 600)
plot(frnk_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "FRNK - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
frnk.ppp <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v4)

frnk_inf_non_4 <- envelope(Y=frnk.ppp[frnk.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp)[rlabel(frnk.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("frnk_inf_non_4.png", width = 800, height = 600)
plot(frnk_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "FRNK - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(frnk.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)


```

## Monte Cristo

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
mocr_inf_all <- envelope(Y=mocr.ppp[mocr.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='N']),
                                          (rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mocr_inf_all)

# Infested - Non-infested
mocr_inf_non_1 <- envelope(Y=mocr.ppp[mocr.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mocr_inf_non_1.png", width = 800, height = 600)
plot(mocr_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MOCR - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Non-infested - All
mocr_noninf_all <- envelope(Y=mocr.ppp[mocr.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='N']),
                                          (rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mocr_noninf_all)

# Hosts - Non-hosts
mocr_host_non <- envelope(Y=mocr.ppp[mocr.ppp$marks=='I' | mocr.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mocr_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
mocr.ppp <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v2)

mocr_inf_non_2 <- envelope(Y=mocr.ppp[mocr.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mocr_inf_non_2.png", width = 800, height = 600)
plot(mocr_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MOCR - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
mocr.ppp <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v3)

mocr_inf_non_3 <- envelope(Y=mocr.ppp[mocr.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mocr_inf_non_3.png", width = 800, height = 600)
plot(mocr_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MOCR - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
mocr.ppp <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v4)

mocr_inf_non_4 <- envelope(Y=mocr.ppp[mocr.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp)[rlabel(mocr.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mocr_inf_non_4.png", width = 800, height = 600)
plot(mocr_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MOCR - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(mocr.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)


```

## Mount Timpanogos

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
mttp_inf_all <- envelope(Y=mttp.ppp[mttp.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='N']),
                                          (rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mttp_inf_all)

# Infested - Non-infested
mttp_inf_non_1 <- envelope(Y=mttp.ppp[mttp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mttp_inf_non_1.png", width = 800, height = 600)
plot(mttp_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MTTP - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Non-infested - All
mttp_noninf_all <- envelope(Y=mttp.ppp[mttp.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='N']),
                                          (rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mttp_noninf_all)

# Hosts - Non-hosts
mttp_host_non <- envelope(Y=mttp.ppp[mttp.ppp$marks=='I' | mttp.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mttp_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
mttp.ppp <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v2)

mttp_inf_non_2 <- envelope(Y=mttp.ppp[mttp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mttp_inf_non_2.png", width = 800, height = 600)
plot(mttp_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MTTP - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
mttp.ppp <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v3)

mttp_inf_non_3 <- envelope(Y=mttp.ppp[mttp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mttp_inf_non_3.png", width = 800, height = 600)
plot(mttp_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MTTP - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
mttp.ppp <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v4)

mttp_inf_non_4 <- envelope(Y=mttp.ppp[mttp.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp)[rlabel(mttp.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("mttp_inf_non_4.png", width = 800, height = 600)
plot(mttp_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "MTTP - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(mttp.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)

```

## Soapstone Basin

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
soap_inf_all <- envelope(Y=soap.ppp[soap.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='N']),
                                          (rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(soap_inf_all)

# Infested - Non-infested
soap_inf_non_1 <- envelope(Y=soap.ppp[soap.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("soap_inf_non_1.png", width = 800, height = 600)
plot(soap_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SOAP - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Non-infested - All
soap_noninf_all <- envelope(Y=soap.ppp[soap.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='N']),
                                          (rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(soap_noninf_all)

# Hosts - Non-hosts
soap_host_non <- envelope(Y=soap.ppp[soap.ppp$marks=='I' | soap.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(soap_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
soap.ppp <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v2)

soap_inf_non_2 <- envelope(Y=soap.ppp[soap.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("soap_inf_non_2.png", width = 800, height = 600)
plot(soap_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SOAP - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
soap.ppp <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v3)

soap_inf_non_3 <- envelope(Y=soap.ppp[soap.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("soap_inf_non_3.png", width = 800, height = 600)
plot(soap_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SOAP - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
soap.ppp <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v4)

soap_inf_non_4 <- envelope(Y=soap.ppp[soap.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp)[rlabel(soap.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("soap_inf_non_4.png", width = 800, height = 600)
plot(soap_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "SOAP - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(soap.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)


```

## Temple Fork

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
tmpf_inf_all <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='N']),
                                          (rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(tmpf_inf_all)

# Infested - Non-infested
tmpf_inf_non_1 <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("tmpf_inf_non_1.png", width = 800, height = 600)
plot(tmpf_inf_non_1,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "TMPF - Infested vs Non-infested Trees - Visit 1",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Non-infested - All
tmpf_noninf_all <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='N'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='N']),
                                          (rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(tmpf_noninf_all)

# Hosts - Non-hosts
tmpf_host_non <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='I' | tmpf.ppp$marks=="N"],fun=pcfinhom,r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='NH'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(tmpf_host_non)

# Infested vs non-infested - times 2,3,4 and then host distribution

# Time 2 
tmpf.ppp <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v2)

tmpf_inf_non_2 <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("tmpf_inf_non_2.png", width = 800, height = 600)
plot(tmpf_inf_non_2,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "TMPF - Infested vs Non-infested Trees - Visit 2",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 3 
tmpf.ppp <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v3)

tmpf_inf_non_3 <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("tmpf_inf_non_3.png", width = 800, height = 600)
plot(tmpf_inf_non_3,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "TMPF - Infested vs Non-infested Trees - Visit 3",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Time 4 
tmpf.ppp <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v4)

tmpf_inf_non_4 <- envelope(Y=tmpf.ppp[tmpf.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp)[rlabel(tmpf.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

png("tmpf_inf_non_4.png", width = 800, height = 600)
plot(tmpf_inf_non_4,
  col = c("red", "black", "blue", "grey"),
  lwd = c(1, 1, 2, 2),
  main = "TMPF - Infested vs Non-infested Trees - Visit 4",
  legend = FALSE,
  xlab = "Distance (r)",
  ylab = "pcf (inhomogeneous)", 
  ylim = c(0,10))
legend(
  "topright",
  legend = c("Simulation Envelope", "Observed", "Simulation Mean"),
  col = c("grey", "red", "black"),
  lwd = c(1, 1, 2),
  lty = c(1, 1, 2),
  cex = 0.8
)
dev.off()

# Host tree distribution
Lcsr <- envelope(tmpf.ppp, Lest, nsim = 999, rank = 1, correction = "translate", global = T) # 999 simulations, alpha = 0.01, "pointwise envelopes". Works better for L than K because of variance stabilizing properties
plot(Lcsr, .-r~r, shade = c("hi", "lo"), legend = F)

```

# Plot all host distributions

```{r}
# Extract values from sites
Lemig <- Lest(emig.ppp)   
r_emig <- Lemig$r         
Lr_emig <- Lemig$iso 
Lfrnk <- Lest(frnk.ppp)   
r_frnk <- Lfrnk$r         
Lr_frnk <- Lfrnk$iso 
Lmocr <- Lest(mocr.ppp)   
r_mocr <- Lmocr$r         
Lr_mocr <- Lmocr$iso 
Lmttp <- Lest(mttp.ppp)   
r_mttp <- Lmttp$r         
Lr_mttp <- Lmttp$iso 
Lsmmh <- Lest(smmh.ppp)   
r_smmh <- Lsmmh$r         
Lr_smmh <- Lsmmh$iso 
Lsoap <- Lest(soap.ppp)   
r_soap <- Lsoap$r         
Lr_soap <- Lsoap$iso 
Ltmpf <- Lest(tmpf.ppp)   
r_tmpf <- Ltmpf$r         
Lr_tmpf <- Ltmpf$iso 

png("host_distributions.png", width = 800, height = 600)
plot(
  Lcsr, 
  . - r ~ r,      
  shade = c("hi", "lo"),          
  legend = FALSE,                 
  main = "Host Distributions by site", 
  xlab = "Distance (r)", 
  ylab = "L(r) - r",
  ylim = c(-1,4)
)

# add lines
lines(r_emig, Lr_emig - r_emig, col = "blue", lwd = 2)  
lines(r_frnk, Lr_frnk - r_frnk, col = "green", lwd = 2) 
lines(r_mocr, Lr_mocr - r_mocr, col = "yellow", lwd = 2)
lines(r_mttp, Lr_mttp - r_mttp, col = "orange", lwd = 2)  
lines(r_smmh, Lr_smmh - r_smmh, col = "violet", lwd = 2)
lines(r_soap, Lr_soap - r_soap, col = "skyblue", lwd = 2)
lines(r_tmpf, Lr_tmpf - r_tmpf, col = "purple", lwd = 2)  

# Add a custom legend
legend(
  "topright", 
  legend = c("Cox Springs", "Emigration Canyon", "Franklin Basin", "Monte Cristo", "Mount Timpanogos", "Smith and Morehouse", "Soapstone Basin", "Temple Fork", "Simulation Envelope"),
  col = c("black", "blue", "green", "yellow", "orange", "violet", "skyblue", "purple", "gray"),
  lwd = c(2, 2, 2, 2, 2, 2, 2, 2, 1),
  lty = c(1, 1, 1, 1, 1, 1, 1, 1, 1),
  cex = 0.4
)
dev.off()

```

# No small trees

```{r}

cxsp.no.small <- cxsp %>% dplyr::filter(!is.na(dbh))
emig.no.small <- emig %>% dplyr::filter(!is.na(dbh))
frnk.no.small <- frnk %>% dplyr::filter(!is.na(dbh))
mocr.no.small <- mocr %>% dplyr::filter(!is.na(dbh))
mttp.no.small <- mttp %>% dplyr::filter(!is.na(dbh))
smmh.no.small <- smmh %>% dplyr::filter(!is.na(dbh))
soap.no.small <- soap %>% dplyr::filter(!is.na(dbh))
tmpf.no.small <- tmpf %>% dplyr::filter(!is.na(dbh))

ppp_pcf(cxsp.no.small)
ppp_pcf(emig.no.small)
ppp_pcf(frnk.no.small)
ppp_pcf(mocr.no.small)
ppp_pcf(mttp.no.small)
ppp_pcf(smmh.no.small)
ppp_pcf(soap.no.small)
ppp_pcf(tmpf.no.small)

```

# Tests with no small

## CXSP

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
cxsp_inf_all_big <- envelope(Y=cxsp.no.small.ppp[cxsp.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.no.small.ppp)[rlabel(cxsp.no.small.ppp)$marks=='N']),
                                          (rlabel(cxsp.no.small.ppp)[rlabel(cxsp.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

# Infested - Non-infested
cxsp_inf_non_big <- envelope(Y=cxsp.no.small.ppp[cxsp.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.no.small.ppp)[rlabel(cxsp.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(cxsp_inf_all_big)
plot(cxsp_inf_all)
plot(cxsp_inf_non)
plot(cxsp_inf_non_big)

```
## EMIG

```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
emig_inf_all_big <- envelope(Y=emig.no.small.ppp[emig.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.no.small.ppp)[rlabel(emig.no.small.ppp)$marks=='N']),
                                          (rlabel(emig.no.small.ppp)[rlabel(emig.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(emig_inf_all)
plot(emig_inf_all_big)

# Infested - Non-infested
emig_inf_non_big <- envelope(Y=emig.no.small.ppp[emig.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.no.small.ppp)[rlabel(emig.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(emig_inf_non)
plot(emig_inf_non_big)

```

## FRNK
```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
frnk_inf_all_big <- envelope(Y=frnk.no.small.ppp[frnk.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.no.small.ppp)[rlabel(frnk.no.small.ppp)$marks=='N']),
                                          (rlabel(frnk.no.small.ppp)[rlabel(frnk.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

# Infested - Non-infested
frnk_inf_non_big <- envelope(Y=frnk.no.small.ppp[frnk.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.no.small.ppp)[rlabel(frnk.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(frnk_inf_all_big)
plot(frnk_inf_all)
plot(frnk_inf_non_big)
plot(frnk_inf_non)

```

## MOCR
```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
mocr_inf_all_big <- envelope(Y=mocr.no.small.ppp[mocr.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.no.small.ppp)[rlabel(mocr.no.small.ppp)$marks=='N']),
                                          (rlabel(mocr.no.small.ppp)[rlabel(mocr.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mocr_inf_all_big)
plot(mocr_inf_all)

# Infested - Non-infested
mocr_inf_non_big <- envelope(Y=mocr.no.small.ppp[mocr.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.no.small.ppp)[rlabel(mocr.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(mocr_inf_non_big)
plot(mocr_inf_non)

```

## MTTP
```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
mttp_inf_all_big <- envelope(Y=mttp.no.small.ppp[mttp.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.no.small.ppp)[rlabel(mttp.no.small.ppp)$marks=='N']),
                                          (rlabel(mttp.no.small.ppp)[rlabel(mttp.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

# Infested - Non-infested
mttp_inf_non_big <- envelope(Y=mttp.no.small.ppp[mttp.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.no.small.ppp)[rlabel(mttp.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(mttp_inf_all_big)
plot(mttp_inf_all)
plot(mttp_inf_non_big)
plot(mttp_inf_non)

```

## SMMH
```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
smmh_inf_all_big <- envelope(Y=smmh.no.small.ppp[smmh.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.no.small.ppp)[rlabel(smmh.no.small.ppp)$marks=='N']),
                                          (rlabel(smmh.no.small.ppp)[rlabel(smmh.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

# Infested - Non-infested
smmh_inf_non_big <- envelope(Y=smmh.no.small.ppp[smmh.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.no.small.ppp)[rlabel(smmh.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(smmh_inf_all_big)
plot(smmh_inf_all)
plot(smmh_inf_non_big)
plot(smmh_inf_non)

```

## SOAP
```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
soap_inf_all_big <- envelope(Y=soap.no.small.ppp[soap.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.no.small.ppp)[rlabel(soap.no.small.ppp)$marks=='N']),
                                          (rlabel(soap.no.small.ppp)[rlabel(soap.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(soap_inf_all_big)
plot(soap_inf_all)

# Infested - Non-infested
soap_inf_non_big <- envelope(Y=soap.no.small.ppp[soap.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.no.small.ppp)[rlabel(soap.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(soap_inf_non_big)
plot(soap_inf_non)

```

## TMPF
```{r, warning=FALSE}

#Tests - Time 1
#Infested - All
tmpf_inf_all_big <- envelope(Y=tmpf.no.small.ppp[tmpf.no.small.ppp$marks=='I'], fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.no.small.ppp)[rlabel(tmpf.no.small.ppp)$marks=='N']),
                                          (rlabel(tmpf.no.small.ppp)[rlabel(tmpf.no.small.ppp)$marks=='I'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(tmpf_inf_all_big)
plot(tmpf_inf_all)

# Infested - Non-infested
tmpf_inf_non_big <- envelope(Y=tmpf.no.small.ppp[tmpf.no.small.ppp$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.no.small.ppp)[rlabel(tmpf.no.small.ppp)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')
plot(tmpf_inf_non_big)
plot(tmpf_inf_non)

```

# Do this again with all visits data

```{r, warning=FALSE, echo=FALSE}

all_trees <- read_xlsx("data_with_nearest_info.xlsx")

all_trees <- all_trees %>% mutate_at(c("NL_v1", "NM_v1", "NH_v1", "SL_v1", "SM_v1", "SH_v1", "NL_v2", "NM_v2", "NH_v2", "SL_v2", "SM_v2", "SH_v2", "NL_v3", "NM_v3", "NH_v3", "SL_v3", "SM_v3", "SH_v3", "NL_v4", "NM_v4", "NH_v4", "SL_v4", "SM_v4", "SH_v4"), ~ replace_na(., 0))

all_trees <- all_trees %>%
  mutate(infested_v1 = ifelse(NL_v1 + NM_v1 + NH_v1 + SL_v1 + SM_v1 + SH_v1 != 0, 1, 0),
         infested_v2 = ifelse(NL_v2 + NM_v2 + NH_v2 + SL_v2 + SM_v2 + SH_v2 != 0, 1, 0),
         infested_v3 = ifelse(NL_v3 + NM_v3 + NH_v3 + SL_v3 + SM_v3 + SH_v3 != 0, 1, 0),
         infested_v4 = ifelse(NL_v4 + NM_v4 + NH_v4 + SL_v4 + SM_v4 + SH_v4 != 0, 1, 0))


all_trees <- all_trees[!is.na(all_trees$Longitude.focal),] # get rid of NAs


# convert to UTM to get nice "square" plots
xy = data.frame(all_trees$Longitude.focal, all_trees$Latitude.focal)
#xy <- xy[!is.na(xy$all_trees.Longitude),] # Get rid of NAs
coordinates(xy) <- c("all_trees.Longitude.focal", "all_trees.Latitude.focal")
proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")
NE <- spTransform(xy, CRS("+proj=utm +zone=12 ellps=WGS84"))
NE <- as.data.frame(NE)

all_trees$Longitude.focal <- NE$coords.x1
all_trees$Latitude.focal <- NE$coords.x2

#Marks for status - we'll do marks for non-hosts and then hosts with and w/o infestation
all_trees$infested_v1 = as.factor(all_trees$infested_v1)
all_trees$infested_v2 = as.factor(all_trees$infested_v2)
all_trees$infested_v3 = as.factor(all_trees$infested_v3)
all_trees$infested_v4 = as.factor(all_trees$infested_v4)
all_trees <- all_trees %>% mutate(host = ifelse(species == "abla" | species == "abco", 1, 0))
all_trees$host <- as.factor(all_trees$host)

all_trees <- all_trees %>%
  mutate(marks_v1 = case_when(
    infested_v1 == 1 ~ factor("I"),
    infested_v1 == 0 ~ factor("N")),
    
    marks_v2 = case_when(
    infested_v2 == 1 ~ factor("I"),
    infested_v2 == 0 ~ factor("N")),
    
    marks_v3 = case_when(
    infested_v3 == 1 ~ factor("I"),
    infested_v3 == 0 ~ factor("N")),
    
    marks_v4 = case_when(
    infested_v4 == 1 ~ factor("I"),
    infested_v4 == 0 ~ factor("N")))


cxsp <- all_trees %>% dplyr::filter(site == "CXSP")
emig <- all_trees %>% dplyr::filter(site == "EMIG")
frnk <- all_trees %>% dplyr::filter(site == "FRNK")
mocr <- all_trees %>% dplyr::filter(site == "MOCR")
mttp <- all_trees %>% dplyr::filter(site == "MTTP")
smmh <- all_trees %>% dplyr::filter(site == "SMMH")
soap <- all_trees %>% dplyr::filter(site == "SOAP")
tmpf <- all_trees %>% dplyr::filter(site == "TMPF")



# repeat for all plots
cxsp$plot_x <- scale_coordinates(cxsp$Longitude.focal, 0, 50)
cxsp$plot_y <- scale_coordinates(cxsp$Latitude.focal, 0, 50)

emig$plot_x <- scale_coordinates(emig$Longitude.focal, 0, 50)
emig$plot_y <- scale_coordinates(emig$Latitude.focal, 0, 50)

frnk$plot_x <- scale_coordinates(frnk$Longitude.focal, 0, 50)
frnk$plot_y <- scale_coordinates(frnk$Latitude.focal, 0, 50)

mocr$plot_x <- scale_coordinates(mocr$Longitude.focal, 0, 50)
mocr$plot_y <- scale_coordinates(mocr$Latitude.focal, 0, 50)

mttp$plot_x <- scale_coordinates(mttp$Longitude.focal, 0, 50)
mttp$plot_y <- scale_coordinates(mttp$Latitude.focal, 0, 50)

smmh$plot_x <- scale_coordinates(smmh$Longitude.focal, 0, 50)
smmh$plot_y <- scale_coordinates(smmh$Latitude.focal, 0, 50)

soap$plot_x <- scale_coordinates(soap$Longitude.focal, 0, 50)
soap$plot_y <- scale_coordinates(soap$Latitude.focal, 0, 50)

tmpf$plot_x <- scale_coordinates(tmpf$Longitude.focal, 0, 50)
tmpf$plot_y <- scale_coordinates(tmpf$Latitude.focal, 0, 50)

```

# Cox Srings - Infested vs NOn-infested - All Visits

```{r}
# ppp objects for each visit
cxsp.ppp.v1 <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v1)
cxsp.ppp.v2 <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v2)
cxsp.ppp.v3 <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v3)
cxsp.ppp.v4 <- ppp(cxsp$plot_x, cxsp$plot_y, window = window, marks = cxsp$marks_v4)

# Infested - Non-infested
cxsp_inf_non_v1 <- envelope(Y=cxsp.ppp.v1[cxsp.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp.v1)[rlabel(cxsp.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
cxsp_inf_non_v2 <- envelope(Y=cxsp.ppp.v2[cxsp.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp.v2)[rlabel(cxsp.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
cxsp_inf_non_v3 <- envelope(Y=cxsp.ppp.v3[cxsp.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp.v3)[rlabel(cxsp.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
cxsp_inf_non_v4 <- envelope(Y=cxsp.ppp.v4[cxsp.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(cxsp.ppp.v4)[rlabel(cxsp.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(cxsp_inf_non_v1)
plot(cxsp_inf_non_v2)
plot(cxsp_inf_non_v3)
plot(cxsp_inf_non_v4)

# Bivariate PCF

Lmulti.cxsp.v1 <- envelope(cxsp.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v1)))

Lmulti.cxsp.v2 <- envelope(cxsp.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v2)))

Lmulti.cxsp.v3 <- envelope(cxsp.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v3)))

Lmulti.cxsp.v4 <- envelope(cxsp.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v4)))

plot(Lmulti.cxsp.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 1")
plot(Lmulti.cxsp.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 2")
plot(Lmulti.cxsp.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 3")
plot(Lmulti.cxsp.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 4")

```

# Emigration Canyon - all visits

```{r}
# ppp objects for each visit
emig.ppp.v1 <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v1)
emig.ppp.v2 <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v2)
emig.ppp.v3 <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v3)
emig.ppp.v4 <- ppp(emig$plot_x, emig$plot_y, window = window, marks = emig$marks_v4)

# Infested - Non-infested
emig_inf_non_v1 <- envelope(Y=emig.ppp.v1[emig.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp.v1)[rlabel(emig.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
emig_inf_non_v2 <- envelope(Y=emig.ppp.v2[emig.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp.v2)[rlabel(emig.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
emig_inf_non_v3 <- envelope(Y=emig.ppp.v3[emig.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp.v3)[rlabel(emig.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
emig_inf_non_v4 <- envelope(Y=emig.ppp.v4[emig.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(emig.ppp.v4)[rlabel(emig.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(emig_inf_non_v1)
plot(emig_inf_non_v2)
plot(emig_inf_non_v3)
plot(emig_inf_non_v4)

# Bivariate PCF

plot(Lmulti.emig.min,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF")

Lmulti.emig.v1 <- envelope(emig.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v1)))

Lmulti.emig.v2 <- envelope(emig.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v2)))

Lmulti.emig.v3 <- envelope(emig.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v3)))

Lmulti.emig.v4 <- envelope(emig.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v4)))

plot(Lmulti.emig.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 1")
plot(Lmulti.emig.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 2")
plot(Lmulti.emig.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 3")
plot(Lmulti.emig.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 4")

```

# Franklin Basin - all visits

```{r}
# ppp objects for each visit
frnk.ppp.v1 <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v1)
frnk.ppp.v2 <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v2)
frnk.ppp.v3 <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v3)
frnk.ppp.v4 <- ppp(frnk$plot_x, frnk$plot_y, window = window, marks = frnk$marks_v4)

# Infested - Non-infested
frnk_inf_non_v1 <- envelope(Y=frnk.ppp.v1[frnk.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp.v1)[rlabel(frnk.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
frnk_inf_non_v2 <- envelope(Y=frnk.ppp.v2[frnk.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp.v2)[rlabel(frnk.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
frnk_inf_non_v3 <- envelope(Y=frnk.ppp.v3[frnk.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp.v3)[rlabel(frnk.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
frnk_inf_non_v4 <- envelope(Y=frnk.ppp.v4[frnk.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(frnk.ppp.v4)[rlabel(frnk.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(frnk_inf_non_v1)
plot(frnk_inf_non_v2)
plot(frnk_inf_non_v3)
plot(frnk_inf_non_v4)

# Bivariate PCF

Lmulti.frnk.v1 <- envelope(frnk.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v1)))

Lmulti.frnk.v2 <- envelope(frnk.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v2)))

Lmulti.frnk.v3 <- envelope(frnk.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v3)))

Lmulti.frnk.v4 <- envelope(frnk.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v4)))

plot(Lmulti.frnk.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 1")
plot(Lmulti.frnk.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 2")
plot(Lmulti.frnk.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 3")
plot(Lmulti.frnk.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 4")

```

# Monte Cristo - all visits - need to re-do coordinates with new coordinates!!!

```{r}
# ppp objects for each visit
mocr.ppp.v1 <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v1)
mocr.ppp.v2 <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v2)
mocr.ppp.v3 <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v3)
mocr.ppp.v4 <- ppp(mocr$plot_x, mocr$plot_y, window = window, marks = mocr$marks_v4)

# Infested - Non-infested
mocr_inf_non_v1 <- envelope(Y=mocr.ppp.v1[mocr.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp.v1)[rlabel(mocr.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
mocr_inf_non_v2 <- envelope(Y=mocr.ppp.v2[mocr.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp.v2)[rlabel(mocr.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
mocr_inf_non_v3 <- envelope(Y=mocr.ppp.v3[mocr.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp.v3)[rlabel(mocr.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
mocr_inf_non_v4 <- envelope(Y=mocr.ppp.v4[mocr.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mocr.ppp.v4)[rlabel(mocr.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(mocr_inf_non_v1)
plot(mocr_inf_non_v2)
plot(mocr_inf_non_v3)
plot(mocr_inf_non_v4)

# Bivariate PCF

Lmulti.mocr.v1 <- envelope(mocr.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v1)))

Lmulti.mocr.v2 <- envelope(mocr.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v2)))

Lmulti.mocr.v3 <- envelope(mocr.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v3)))

Lmulti.mocr.v4 <- envelope(mocr.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v4)))

plot(Lmulti.mocr.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 1")
plot(Lmulti.mocr.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 2")
plot(Lmulti.mocr.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 3")
plot(Lmulti.mocr.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 4")

```

# Mount Timp - all visits

```{r}
# ppp objects for each visit
mttp.ppp.v1 <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v1)
mttp.ppp.v2 <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v2)
mttp.ppp.v3 <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v3)
mttp.ppp.v4 <- ppp(mttp$plot_x, mttp$plot_y, window = window, marks = mttp$marks_v4)

# Infested - Non-infested
mttp_inf_non_v1 <- envelope(Y=mttp.ppp.v1[mttp.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp.v1)[rlabel(mttp.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
mttp_inf_non_v2 <- envelope(Y=mttp.ppp.v2[mttp.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp.v2)[rlabel(mttp.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
mttp_inf_non_v3 <- envelope(Y=mttp.ppp.v3[mttp.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp.v3)[rlabel(mttp.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
mttp_inf_non_v4 <- envelope(Y=mttp.ppp.v4[mttp.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(mttp.ppp.v4)[rlabel(mttp.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(mttp_inf_non_v1)
plot(mttp_inf_non_v2)
plot(mttp_inf_non_v3)
plot(mttp_inf_non_v4)

# Bivariate PCF

Lmulti.mttp.v1 <- envelope(mttp.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v1)))

Lmulti.mttp.v2 <- envelope(mttp.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v2)))

Lmulti.mttp.v3 <- envelope(mttp.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v3)))

Lmulti.mttp.v4 <- envelope(mttp.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v4)))

plot(Lmulti.mttp.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 1")
plot(Lmulti.mttp.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 2")
plot(Lmulti.mttp.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 3")
plot(Lmulti.mttp.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 4")

```

# Soapstone - all visits

```{r}
# ppp objects for each visit
soap.ppp.v1 <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v1)
soap.ppp.v2 <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v2)
soap.ppp.v3 <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v3)
soap.ppp.v4 <- ppp(soap$plot_x, soap$plot_y, window = window, marks = soap$marks_v4)

# Infested - Non-infested
soap_inf_non_v1 <- envelope(Y=soap.ppp.v1[soap.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp.v1)[rlabel(soap.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
soap_inf_non_v2 <- envelope(Y=soap.ppp.v2[soap.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp.v2)[rlabel(soap.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
soap_inf_non_v3 <- envelope(Y=soap.ppp.v3[soap.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp.v3)[rlabel(soap.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
soap_inf_non_v4 <- envelope(Y=soap.ppp.v4[soap.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(soap.ppp.v4)[rlabel(soap.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(soap_inf_non_v1)
plot(soap_inf_non_v2)
plot(soap_inf_non_v3)
plot(soap_inf_non_v4)

# Bivariate PCF

Lmulti.soap.v1 <- envelope(soap.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v1)))

Lmulti.soap.v2 <- envelope(soap.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v2)))

Lmulti.soap.v3 <- envelope(soap.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v3)))

Lmulti.soap.v4 <- envelope(soap.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v4)))

plot(Lmulti.soap.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 1")
plot(Lmulti.soap.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 2")
plot(Lmulti.soap.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 3")
plot(Lmulti.soap.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 4")

```

# Smith and Morehouse - all visits

```{r}
# ppp objects for each visit
smmh.ppp.v1 <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v1)
smmh.ppp.v2 <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v2)
smmh.ppp.v3 <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v3)
smmh.ppp.v4 <- ppp(smmh$plot_x, smmh$plot_y, window = window, marks = smmh$marks_v4)

# Infested - Non-infested
smmh_inf_non_v1 <- envelope(Y=smmh.ppp.v1[smmh.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp.v1)[rlabel(smmh.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
smmh_inf_non_v2 <- envelope(Y=smmh.ppp.v2[smmh.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp.v2)[rlabel(smmh.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
smmh_inf_non_v3 <- envelope(Y=smmh.ppp.v3[smmh.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp.v3)[rlabel(smmh.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
smmh_inf_non_v4 <- envelope(Y=smmh.ppp.v4[smmh.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(smmh.ppp.v4)[rlabel(smmh.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(smmh_inf_non_v1)
plot(smmh_inf_non_v2)
plot(smmh_inf_non_v3)
plot(smmh_inf_non_v4)

# Bivariate PCF

Lmulti.smmh.v1 <- envelope(smmh.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v1)))

Lmulti.smmh.v2 <- envelope(smmh.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v2)))

Lmulti.smmh.v3 <- envelope(smmh.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v3)))

Lmulti.smmh.v4 <- envelope(smmh.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v4)))

plot(Lmulti.smmh.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 1")
plot(Lmulti.smmh.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 2")
plot(Lmulti.smmh.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 3")
plot(Lmulti.smmh.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 4")

```

# Temple Fork - all visits

```{r}
# ppp objects for each visit
tmpf.ppp.v1 <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v1)
tmpf.ppp.v2 <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v2)
tmpf.ppp.v3 <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v3)
tmpf.ppp.v4 <- ppp(tmpf$plot_x, tmpf$plot_y, window = window, marks = tmpf$marks_v4)

# Infested - Non-infested
tmpf_inf_non_v1 <- envelope(Y=tmpf.ppp.v1[tmpf.ppp.v1$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp.v1)[rlabel(tmpf.ppp.v1)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
tmpf_inf_non_v2 <- envelope(Y=tmpf.ppp.v2[tmpf.ppp.v2$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp.v2)[rlabel(tmpf.ppp.v2)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
tmpf_inf_non_v3 <- envelope(Y=tmpf.ppp.v3[tmpf.ppp.v3$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp.v3)[rlabel(tmpf.ppp.v3)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')


# Infested - Non-infested
tmpf_inf_non_v4 <- envelope(Y=tmpf.ppp.v4[tmpf.ppp.v4$marks=='I'],fun=pcfinhom, r=seq(0,50,.5), nsim=99,
                      simulate=expression((rlabel(tmpf.ppp.v4)[rlabel(tmpf.ppp.v4)$marks=='N'])),
                      savepatterns = TRUE, savefuns = TRUE,divisor = 'd')

plot(tmpf_inf_non_v1)
plot(tmpf_inf_non_v2)
plot(tmpf_inf_non_v3)
plot(tmpf_inf_non_v4)

# Bivariate PCF

Lmulti.tmpf.v1 <- envelope(tmpf.ppp.v1, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v1)))

Lmulti.tmpf.v2 <- envelope(tmpf.ppp.v2, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v2)))

Lmulti.tmpf.v3 <- envelope(tmpf.ppp.v3, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v3)))

Lmulti.tmpf.v4 <- envelope(tmpf.ppp.v4, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v4)))

plot(Lmulti.tmpf.v1,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 1")
plot(Lmulti.tmpf.v2,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 2")
plot(Lmulti.tmpf.v3,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 3")
plot(Lmulti.tmpf.v4,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 4")

```

# New infested at visits 2,3 and 4

```{r}
ppp_pcf_v2 <- function(i) {
obj_name <- deparse(substitute(i))
i.ppp <- ppp(i$plot_x, i$plot_y, window = window, marks = i$marks_new_v2)
i.pcf <- pcfinhom(i.ppp, r = 0:20, nsim =3)
assign(paste0(obj_name, ".ppp.v2.new"), i.ppp, envir = .GlobalEnv)
assign(paste0(obj_name, ".pcf.v2.new"), i.pcf, envir = .GlobalEnv)
plot(i.pcf)
}
ppp_pcf_v3 <- function(i) {
obj_name <- deparse(substitute(i))
i.ppp <- ppp(i$plot_x, i$plot_y, window = window, marks = i$marks_new_v3)
i.pcf <- pcfinhom(i.ppp, r = 0:20, nsim =3)
assign(paste0(obj_name, ".ppp.v3.new"), i.ppp, envir = .GlobalEnv)
assign(paste0(obj_name, ".pcf.v3.new"), i.pcf, envir = .GlobalEnv)
plot(i.pcf)
}
ppp_pcf_v4 <- function(i) {
obj_name <- deparse(substitute(i))
i.ppp <- ppp(i$plot_x, i$plot_y, window = window, marks = i$marks_new_v4)
i.pcf <- pcfinhom(i.ppp, r = 0:20, nsim =3)
assign(paste0(obj_name, ".ppp.v4.new"), i.ppp, envir = .GlobalEnv)
assign(paste0(obj_name, ".pcf.v4.new"), i.pcf, envir = .GlobalEnv)
plot(i.pcf)
}

ppp_pcf_v2(cxsp)
ppp_pcf_v2(emig)
ppp_pcf_v2(frnk)
ppp_pcf_v2(mocr)
ppp_pcf_v2(mttp)
ppp_pcf_v2(smmh)
ppp_pcf_v2(soap)
ppp_pcf_v2(tmpf)
ppp_pcf_v3(cxsp)
ppp_pcf_v3(emig)
ppp_pcf_v3(frnk)
ppp_pcf_v3(mocr)
ppp_pcf_v3(mttp)
ppp_pcf_v3(smmh)
ppp_pcf_v3(soap)
ppp_pcf_v3(tmpf)
ppp_pcf_v4(cxsp)
ppp_pcf_v4(emig)
ppp_pcf_v4(frnk)
ppp_pcf_v4(mocr)
ppp_pcf_v4(mttp)
ppp_pcf_v4(smmh)
ppp_pcf_v4(soap)
ppp_pcf_v4(tmpf)

```

# Cox Springs - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.cxsp.v2.new <- envelope(cxsp.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v2.new)))

Lmulti.cxsp.v3.new <- envelope(cxsp.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v3.new)))

Lmulti.cxsp.v4.new <- envelope(cxsp.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(cxsp.ppp.v4.new)))


plot(Lmulti.cxsp.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 2 - new infestations vs all other trees")
plot(Lmulti.cxsp.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.cxsp.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Cox Springs Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Emigration Canyon - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.emig.v2.new <- envelope(emig.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v2.new)))

#Lmulti.emig.v3.new <- envelope(emig.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v3.new))) - no new infestations on this visit

Lmulti.emig.v4.new <- envelope(emig.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(emig.ppp.v4.new)))


plot(Lmulti.emig.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 2 - new infestations vs all other trees")
#plot(Lmulti.emig.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.emig.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Emigration Canyon Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Franklin Basin - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.frnk.v2.new <- envelope(frnk.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v2.new)))

#Lmulti.frnk.v3.new <- envelope(frnk.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v3.new))) - no new infestations

Lmulti.frnk.v4.new <- envelope(frnk.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(frnk.ppp.v4.new)))


plot(Lmulti.frnk.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 2 - new infestations vs all other trees")
#plot(Lmulti.frnk.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.frnk.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Franklin Basin Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Monte Cristo - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.mocr.v2.new <- envelope(mocr.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v2.new)))

Lmulti.mocr.v3.new <- envelope(mocr.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v3.new)))

Lmulti.mocr.v4.new <- envelope(mocr.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mocr.ppp.v4.new)))


plot(Lmulti.mocr.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 2 - new infestations vs all other trees")
plot(Lmulti.mocr.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.mocr.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Monte Cristo Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Mount Timpanogos - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.mttp.v2.new <- envelope(mttp.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v2.new)))

Lmulti.mttp.v3.new <- envelope(mttp.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v3.new)))

Lmulti.mttp.v4.new <- envelope(mttp.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(mttp.ppp.v4.new)))


plot(Lmulti.mttp.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 2 - new infestations vs all other trees")
plot(Lmulti.mttp.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.mttp.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Mount Timpanogos Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Smith and Morehouse - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.smmh.v2.new <- envelope(smmh.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v2.new)))

Lmulti.smmh.v3.new <- envelope(smmh.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v3.new)))

Lmulti.smmh.v4.new <- envelope(smmh.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(smmh.ppp.v4.new)))


plot(Lmulti.smmh.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 2 - new infestations vs all other trees")
plot(Lmulti.smmh.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.smmh.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Smith and Morehouse Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Soapstone Basin - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.soap.v2.new <- envelope(soap.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v2.new)))

Lmulti.soap.v3.new <- envelope(soap.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v3.new)))

Lmulti.soap.v4.new <- envelope(soap.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(soap.ppp.v4.new)))


plot(Lmulti.soap.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 2 - new infestations vs all other trees")
plot(Lmulti.soap.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.soap.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Soapstone Basin Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Temple Fork - New infestations vs trees that stayed uninfested

```{r}
# Bivariate PCF

Lmulti.tmpf.v2.new <- envelope(tmpf.ppp.v2.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v2.new)))

Lmulti.tmpf.v3.new <- envelope(tmpf.ppp.v3.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v3.new)))

Lmulti.tmpf.v4.new <- envelope(tmpf.ppp.v4.new, Lcross, nsim = 999, rank = 1, i = "I", global = FALSE, simulate = expression(rlabel(tmpf.ppp.v4.new)))


plot(Lmulti.tmpf.v2.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 2 - new infestations vs all other trees")
plot(Lmulti.tmpf.v3.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 3 - new infestations vs all other trees")
plot(Lmulti.tmpf.v4.new,.-r~r,  shade = c("hi", "lo"), legend = FALSE, main = "Temple Fork Bivariate PCF - visit 4 - new infestations vs all other trees")

```

# Visualizations for each visit

```{r}

data_long <- all_trees %>%
  pivot_longer(
    cols = ends_with("_v1") | ends_with("_v2") | ends_with("_v3") | ends_with("_v4"),  # columns with visit data
    names_to = c(".value", "visit"),       # use ".value" to keep the original variable name
    names_pattern = "(.*)_(v[0-9])"        # regex to separate variable name and visit
  ) %>%
  mutate(visit = as.numeric(gsub("v", "", visit))) 

trees <- data_long

trees <- trees %>%
  arrange(plot_x, plot_y, visit) %>%  # Ensure the data is ordered by tree coordinates and visit
  group_by(plot_x, plot_y) %>%        # Group by unique trees (using coordinates)
  mutate(
    ever_infestation = cumsum(marks == "I") > 0,  # Tracks if the tree was ever infested
    new_infestation = (marks == "I" & lag(ever_infestation, default = FALSE) == FALSE)  # First time "I"
  ) %>%
  ungroup()

cxsp_long <- trees %>% filter(site == "CXSP")

# Filter for first-time infested trees
first_time_data <- cxsp_long %>% filter(new_infestation == TRUE)

# Previously infested trees
prev_data <- cxsp_long %>% filter(marks == "I" & new_infestation == FALSE)

ggplot() +
  # Previously infested trees in light red
  geom_point(
    data = prev_data,
    aes(x = plot_x, y = plot_y),
    color = "lightcoral",
    alpha = 0.5, # Light transparency
    size = 2
  ) +
  # First-time infested trees in red
  geom_point(
    data = first_time_data,
    aes(x = plot_x, y = plot_y),
    color = "red",
    size = 2
  ) +
  labs(
    title = "First-Time Infestation by Visit",
    x = "X Coordinate",
    y = "Y Coordinate"
  ) +
  facet_wrap(~visit) + # Separate plots for each visit
  theme_minimal()

# Cox Springs - visit 1
ggplot(cxsp, aes(x = plot_x, y = plot_y, color = marks_v1)) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("I" = "red", "N" = "green")
  ) +
  labs(
    title = "Cox Springs - Visit 1",
    x = "plot_x",
    y = "plot_y",
    color = "Infestation and Host status"
  ) +
  theme_minimal()

# Cox Springs - visit 2

# make new category with infestations from visit one and new infestations on visit 2
cxsp$new_v2 <- ifelse(cxsp$marks_v1 == "I", "I_v1",
                             ifelse(cxsp$marks_new_v2 == "I", "I_v2", 
                              ifelse(cxsp$marks_new_v2 == "N", "N", NA)))

ggplot(cxsp, aes(x = plot_x, y = plot_y, color = new_v2)) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("I_v1" = "lightcoral", "I_v1" = "red", "N" = "green")
  ) +
  labs(
    title = "Cox Springs - Visit 2",
    x = "plot_x",
    y = "plot_y",
    color = "Infestation and Host status"
  ) +
  theme_minimal()
```